---
format: revealjs
---

# Today's Agenda {background-image="Images/background-data_blue_v4.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
```

<br>

::: {.r-fit-text}

**Section 1: Measurement is Hard**

1. TBD

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2026)
:::

::: notes
Prep for Class

1. Review Canvas submissions

:::


## Questions on the Syllabus? {background-image="Images/background-data_blue_v4.png" .center}

<br>

```{r, fig.align='center'}
tibble(
  x1 = c("Attendance & Participation", "Assignments", "- Report: Analyze the Outcome Variable(s)", "- Report: Analyze the Predictor Variable(s)", "- Report: Test the Causal Mechanism"),
  x2 = c("50%", "50%", "", "", "")
) |>
  kbl(align = c("l", "c"), col.names = c("", "")) |>
  kableExtra::kable_styling(font_size = 38) |>
  column_spec(1:2, background = "white")
```

::: notes

**Any questions about the class from your review of the syllabus?**

<br>

Our job this semester is to design and execute a research project
- Your grade is heavy on being in class and collaborating to get this done

<br>

### Is the attendance policy clear?
1. Attendance cliff: > 4 unexcused absences = 'C'
2. YOUR JOB to request make-up assignments from excused absences

<br>

### Does everybody understand how to earn the participation points?
- Come to class on-time,
- Be engaged while in class (e.g., bring all needed materials, don’t distract others, etc.), and
- Submit all required daily assignments BEFORE class begins.

<br>

### Does everybody see where the readings for each class are assigned and know how to access them?

<br>

### Any questions on using Canvas?

:::







## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}

**For Next Class**

1. Brians et al (2011) Chapter 5

2. Canvas Assignment

:::

::: notes

Everybody take a look at the assignment for next class and let me know if you have any questions!

<br>

Canvas ASSIGNMENT: Submit to Canvas a news event from the last few years that involves a government, anywhere in the world, infringing on the civil rights or liberties of its citizens. You may not submit an event already claimed by someone else in class.

1. APA citation and web link to the event

2. Describe the event

3. Explain the specific right or liberty being infringed upon



:::
