---
output:
  xaringan::moon_reader:
    css: ["default", "extra.css"]
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
##xaringan::inf_mr()
## For offline work: https://bookdown.org/yihui/rmarkdown/some-tips.html#working-offline
## Images not appearing? Put images folder inside the libs folder as that is the main data directory

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
##library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

class: slideblue

.size80[**Today's Agenda**]

<br>

.size40[

1. Installing and configuring the software

2. Choose our first data project

]

<br>
<br>

.center[.size40[
  Justin Leinaweaver (Fall 2021)
]]



---

class: middle, slideblue

.size50[**Let's Install R!**]

.size30[
1. http://www.r-project.org/

2. Click on “CRAN.”

3. Select a site near you or “0-Cloud,”
]

.pull-left[
.size30[
**Windows**

+ "Download R for Windows"
+ "Download and Install R"
+ Select "base"
+ Download the .exe and run it
]]

.pull-right[
.size30[
**macOS**

+ "Download R for (Mac) OS X"
+ Click .pkg under "Latest release"
+ Run the .pkg file
]]



---

class: middle, slideblue

.size70[**Let's Install RStudio!**]

<br>
<br>

.size40[https://www.rstudio.com/products/rstudio/download/]

<br>

.size50[1) Scroll down to "All Installers"]

<br>

.size50[2) Download and run the file for your OS]



---

class: slidegreen, middle

.size60[**RStudio Cloud: A Backup Tool**]

<br>
<br>

.size50[https://rstudio.cloud/

1) Click "sign up"

2) Register for a "Cloud Free" account
]


---

class: middle, slidepurple

.size60[.center[**Our Aim this Semester**]]

.size60[.center[**Design and answer a research question with data**]]

<br>

```{r, fig.align='center', fig.height=2, out.width='85%'}
## Manual DAG
d1 <- tibble(
  x = c(-3, 3),
  y = c(1, 1),
  labels = c("Predictor\n Variable", "Outcome\n Variable")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = -1.9, xend = 1.9, y = 1, yend = 1, arrow = arrow())
```


???

Share the data project options spreadsheet



---

class: center, middle, slidegreen

.size60[**Assignment for Tuesday**]

<br>

.size50[Review the Data Project 1 Codebook and the data for it using Excel]

