---
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    df_print: kable
fontsize: 14pt
classoption: "aspectratio=169"
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "./03_1-Univariate_Viz_v2.Rmd", output_file = "./03_1-Univariate_Viz_v2.pdf")

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```


## Today's Agenda

\Large
Connecting descriptive statistics to univariate visualizations

\vspace{.5in}

\begin{center}
Justin Leinaweaver (Fall 2021)
\end{center}

```{r, echo=FALSE, eval=TRUE}
options("width" = 120)
d <- read_excel("./Data/All_data_FIW_2013-2021-tidy.xlsx", sheet = 2)
```


## {.plain}
\center
\includegraphics[keepaspectratio=true,width=.95\paperwidth]{Images/00-Pizza_Politics1.png}



## Practice Questions
\large

Calculate the minimum, mean and maximum values of the `Total` score for:

1. The years 2013 and 2021 (filter).

2. Each year in the dataset (group_by).

3. Each region in the dataset (group_by).

4. Both year and region (group_by).


## Practice Questions: Baseline Code
```{r, echo=TRUE}
d |>
   summarize(
     Minimum = min(Total),
     Mean = mean(Total),
     Maximum = max(Total)
)
```


## Practice Questions: Baseline Code (v2)
```{r, echo=TRUE}
d |>
   summarize(
     Minimum = min(Total),
     Mean = round(mean(Total), 1),
     Maximum = max(Total)
)
```


## Practice Questions: 2013 and 2021
```{r, echo=TRUE}
d |>
   filter(Edition == 2013) |>
   summarize(
     Minimum = min(Total),
     Mean = round(mean(Total), 1),
     Maximum = max(Total)
)
```


## Practice Questions: By Year
::: columns
:::: column

\small
```{r, echo=TRUE, eval=FALSE}
d |>
   group_by(Edition) |>
   summarize(
     Minimum = min(Total),
     Mean = round(mean(Total), 1),
     Maximum = max(Total)
)
```

::::
:::: column

\small
```{r, echo=FALSE, eval=TRUE}
d |>
   group_by(Edition) |>
   summarize(
     Minimum = min(Total),
     Mean = round(mean(Total), 1),
     Maximum = max(Total)
   )
```
::::
:::



## Practice Questions: By Region
::: columns
:::: column

\small
```{r, echo=TRUE, eval=FALSE}
d |>
   group_by(Region) |>
   summarize(
     Minimum = min(Total),
     Mean = round(mean(Total), 1),
     Maximum = max(Total)
)
```

::::
:::: column

\small
```{r, echo=FALSE, eval=TRUE}
d |>
   group_by(Region) |>
   summarize(
     Minimum = min(Total),
     Mean = round(mean(Total), 1),
     Maximum = max(Total)
   )
```
::::
:::



## Practice Questions: By Region and Year
::: columns
:::: column

\small
```{r, echo=TRUE, eval=FALSE}
d |>
   group_by(Region, Edition) |>
   summarize(
     Minimum = min(Total),
     Mean = round(mean(Total), 1),
     Maximum = max(Total)
)
```

::::
:::: column

\scriptsize
```{r, echo=FALSE, eval=TRUE}
d |>
   group_by(Region, Edition) |>
   summarize(
     Minimum = min(Total),
     Mean = round(mean(Total), 1),
     Maximum = max(Total)
   ) |>
ungroup() |>
slice(1:13)
```
::::
:::



## {.plain}
\center
\includegraphics[keepaspectratio=true,width=.6\paperwidth]{Images/03_1-data_viz.jpg}

\vspace{.25in}

**A method for summarizing data in more compelling, easier to understand and more informative methods than descriptive statistics.**



## Principle 3. Variable Type Determines Tool {.plain}
\large 
**Univariate Visualizations**

+ If a **categorical** variable, make a **bar plot**

+ If a **numeric** variable, make a **box plot** or **histogram**

+ If a **numeric variable across time**, make a **line plot**



## Why draw visualizations by hand?
::: columns
:::: column

\begin{center}
\includegraphics[keepaspectratio=true,width=.45\paperwidth]{Images/03_1-By_hand.jpg}
\end{center}

::::
:::: column
\large

\begin{enumerate}[<+->]
\item Principle 3 in action,

\vspace{.25in}

\item Learn to think like a programmer
\end{enumerate}

::::
:::



## Why draw visualizations by hand?
::: columns
:::: column
\footnotesize

```{r, echo=FALSE}
d <- read_excel("./Data/All_data_FIW_2013-2021-tidy.xlsx", sheet = 2)

d |>
pivot_longer(cols = c(`CL rating`, CL, `PR rating`, PR, Total), names_to = "Variable", values_to = "Values") |>
group_by(Variable) |>
summarize(
    N = n(),
    Mean = round(mean(Values), 1),
    StdDev = round(sd(Values), 1),
    Min = min(Values),
    Max = max(Values)
    )
```

::::
:::: column

\large

1. Principle 3 in action,

\vspace{.25in}

2. Learn to think like a programmer

\vspace{.25in}

3. Learn to "see" descriptive stats tables

::::
:::




## Categorical Variable: Make a Bar Plot {.plain}
\begin{columns}
\begin{column}{0.4\textwidth}

\LARGE
\textbf{Bar Plots}
\vspace{.2in}
\begin{enumerate}
\item Counts, or
\vspace{.2in}
\item Proportions
\end{enumerate}

\end{column}
\begin{column}{0.6\textwidth}

\begin{center}
\includegraphics[keepaspectratio=true,width=.42\paperwidth]{Images/03_1-Johnson-Bar_plot.png}
\end{center}

\end{column}
\end{columns}



## Categorical Variable: Make a Bar Plot {.plain}
\LARGE
\center

\vspace{.2in}

**Draw a bar plot of `CL Rating` counts in 2021 by hand.**

\vspace{.5in}

```{r, echo=TRUE, eval=FALSE}
table(d$"CL rating")
```




## {.plain}
```{r, fig.align='center', fig.asp=0.618, fig.width=6, out.width='95%'}
d |>
filter(Edition == 2021) |>
    ggplot(aes(x = `CL rating`)) +
    geom_bar() +
    theme_bw() +
    labs(x = "CL Rating", y = "") +
    scale_x_continuous(breaks = 1:7)
```


## Categorical Variable: Make a Bar Plot {.plain}
\LARGE
\center

\vspace{.2in}

**Draw a bar plot of "PR Rating" counts in 2021 by hand.**

\vspace{.5in}

```{r, echo=TRUE, eval=FALSE}
table(d$"PR rating")
```



## {.plain}
```{r, fig.align='center', fig.asp=0.618, fig.width=6, out.width='95%'}
d |>
filter(Edition == 2021) |>
    ggplot(aes(x = `PR rating`)) +
    geom_bar() +
    theme_bw() +
    labs(x = "PR Rating", y = "") +
    scale_x_continuous(breaks = 1:7)
```


## Bar Plots vs Descriptive Stats Table
::: columns
:::: column

```{r, fig.align='center', fig.asp=0.7, fig.width=4, out.width='95%'}
d |>
filter(Edition == 2021) |>
    ggplot(aes(x = `PR rating`)) +
    geom_bar() +
    theme_bw() +
    labs(x = "PR Rating", y = "") +
    scale_x_continuous(breaks = 1:7)
```
::::
:::: column
```{r, echo=TRUE}
table(d$`PR rating`)
```
::::
:::


## Numerical Variable: Make a Box Plot {.plain}
\begin{columns}
\begin{column}{0.4\textwidth}

\Large
\textbf{Box Plots}
\vspace{.1in}
\begin{itemize}
\item Minimum
\item 25th percentile
\item Median
\item 75th percentile
\item Maximum
\item IQR x 1.5
\end{itemize}

\end{column}
\begin{column}{0.6\textwidth}

\begin{center}
\includegraphics[keepaspectratio=true,width=.5\paperwidth]{Images/03_1-Johnson-Box_plot.png}
\end{center}

\end{column}
\end{columns}



## Numerical Variable: Make a Box Plot
\LARGE
\center

\vspace{.2in}

**Draw a box plot of `CL` score in 2021 by hand.**

\vspace{.4in}
\large
**Note:** Johnson (2012), p388-389 has the steps...



## {.plain}
```{r, fig.align='center', fig.asp=0.618, fig.width=6, out.width='95%'}
d |>
filter(Edition == 2021) |>
    ggplot(aes(x = CL, y = factor(1))) +
    geom_boxplot() +
    theme_bw() +
    labs(x = "CL Score (0 - 60)", y = "") +
        scale_y_discrete(label = "")
```


## Numerical Variable: Make a Box Plot
\LARGE
\center

\vspace{.2in}

**Draw a box plot of `PR` score in 2021 by hand.**

\vspace{.4in}
\large
**Note:** Johnson (2012), p388-389 has the steps...



## {.plain}
```{r, fig.align='center', fig.asp=0.618, fig.width=6, out.width='95%'}
d |>
filter(Edition == 2021) |>
    ggplot(aes(x = PR, y = factor(1))) +
    geom_boxplot() +
    theme_bw() +
    labs(x = "PR Score (-3 - 40)", y = "") +
        scale_y_discrete(label = "")
```


## Box Plot vs Descriptive Stats Table {.plain}
```{r, fig.align='center', fig.asp=0.618, fig.width=5, out.width='50%'}
d |>
filter(Edition == 2021) |>
    ggplot(aes(x = CL, y = factor(1))) +
    geom_boxplot() +
    theme_bw() +
    labs(x = "CL Score (0 - 60)", y = "") +
        scale_y_discrete(label = "")
```

\small
```{r, echo=TRUE}
summary(d$CL)
```


## Numerical Variable: Make a Histogram {.plain}
\begin{center}
\includegraphics[keepaspectratio=true,width=.7\paperwidth]{Images/03_1-Johnson-Histogram.png}
\end{center}



## Numerical Variable: Make a Histogram {.plain}
\center
\includegraphics[keepaspectratio=true,width=.45\paperwidth]{Images/03_1-Johnson-Histogram.png}

\Large
**Why can't we just draw a bar for every possible value? (e.g. make a bar plot)**


## Histograms are bar plots using bins
\Large

1. Make a table of `CL` scores for 2021 only

2. Convert this table into bins of 10 and re-sum the counts
   + \large Bin 1 is all values from 1 - 10
   + Bin 2 is all values from 11 - 20
   + Bin 3...
   
3. Draw the bar plot of your bins



## Numerical Variable: Make a Histogram {.plain}
```{r, fig.asp=0.618, fig.align='center', out.width='80%', fig.width=5.5}
d |>
filter(Edition == 2021) |>
    ggplot(aes(x = CL)) +
    geom_histogram(bins = 10, color = "white") +
    theme_bw() +
    labs(x = "Civil Liberties Score (2021)", y = "")
```


## Histograms are bar plots using bins
\Large

1. Make a table of `PR` scores for 2021 only

2. Convert this table into bins of 10 and re-sum the counts
   + \large Bin 1 is all values from -10 - 0
   + Bin 2 is all values from 1 - 10
   + Bin 3...
   
3. Draw the bar plot of your bins



## Numerical Variable: Make a Histogram {.plain}
```{r, fig.asp=0.618, fig.align='center', out.width='80%', fig.width=5.5}
d |>
filter(Edition == 2021) |>
    ggplot(aes(x = PR)) +
    geom_histogram(bins = 10, color = "white") +
    theme_bw() +
    labs(x = "Political Rights Score (2021)", y = "")
```


## Numeric Variables: Box vs Histogram
::: columns
:::: column

```{r, fig.width=4, fig.asp=0.75}
d |>
filter(Edition == 2021) |>
    ggplot(aes(x = CL, y = factor(1))) +
    geom_boxplot() +
    theme_bw() +
    labs(x = "Civil Liberties Score (2021)", y = "") +
    scale_y_discrete(label = "")
```

::::
:::: column

```{r, fig.width=4, fig.asp=0.75}
d |>
filter(Edition == 2021) |>
    ggplot(aes(x = CL)) +
    geom_histogram(bins = 10, color = "white") +
    theme_bw() +
    labs(x = "Civil Liberties Score (2021)", y = "")
```
::::
:::


## Box plots make comparing distributions easier {.plain}
```{r, fig.align='center', fig.asp=0.618, fig.width=6, out.width='80%'}
d |>
filter(Edition == 2021) |>
pivot_longer(cols = c(PR, CL), names_to = "Variable", values_to = "Values") |>
    ggplot(aes(x = Values, y = Variable)) +
    geom_boxplot() +
    theme_bw() +
    labs(x = "Freedom House Scores", y = "")
```


## Numeric Variables Across Time {.plain}
\begin{center}
\includegraphics[keepaspectratio=true,width=.7\paperwidth]{Images/03_1-Johnson-Line_plot.png}
\end{center}


## 
\LARGE
\center
**Make a line plot of the average 'PR' and 'CL' scores per year.**

(e.g. two separate lines)

\vspace{.25in}
\large
### R Functions
+ group_by()
+ summarize()
+ mean()



## {.plain}
\LARGE
\center
**Make a line plot of the average 'PR' and 'CL' scores per year.**

\vspace{.3in}
\normalsize
```{r, echo=TRUE, eval=FALSE}
d |>
   group_by(Edition) |>
   summarize(
       MeanPR = mean(PR),
       MeanCL = mean(CL)
   )
```


## Which is the better visualization? {.plain}
::: columns
:::: column

\vspace{.6in}

```{r, fig.align='center', fig.asp=.8, fig.width=4, out.width='95%'}
d |>
   group_by(Edition) |>
   summarize(
       MeanPR = mean(PR),
       MeanCL = mean(CL)
   ) |>
pivot_longer(cols = MeanPR:MeanCL, names_to = "Variable", values_to = "Values") |>
ggplot(aes(x = Edition, y = Values, color = Variable)) +
    geom_line(size = 1.3) +
    theme_bw() +
    labs(x = "", y = "") +
    scale_x_continuous(breaks = 2013:2021) +
  guides(color = FALSE)
```

::::
:::: column

```{r, fig.align='center', fig.asp=1.4, fig.width=4, out.width='95%'}
d |>
   group_by(Edition) |>
   summarize(
       MeanPR = mean(PR),
       MeanCL = mean(CL)
   ) |>
pivot_longer(cols = MeanPR:MeanCL, names_to = "Variable", values_to = "Values") |>
ggplot(aes(x = Edition, y = Values, color = Variable)) +
    geom_line(size = 1.3) +
    theme_bw() +
    labs(x = "", y = "") +
    scale_x_continuous(breaks = 2013:2021) +
    facet_wrap(~ Variable, ncol = 1, scales = "free_y") +
  guides(color = FALSE)
```

::::
:::


## Wrapping Up: Drawing by hand is good!
::: columns
:::: column

\begin{center}
\includegraphics[keepaspectratio=true,width=.45\paperwidth]{Images/03_1-By_hand.jpg}
\end{center}

::::
:::: column
\large

1. Principle 3 in action,

\vspace{.25in}

2. Learn to think like a programmer

\vspace{.25in}

3. Learn to "see" descriptive stats tables

::::
:::


## Principle 3. Variable Type Determines Tool
\large 
**Univariate Visualizations**

+ If a **categorical** variable, make a **bar plot**

+ If a **numeric** variable, make a **box plot** or **histogram**

+ If a **numeric variable across time**, make a **line plot**




## For Thursday
\LARGE
Use Healy (3.1 - 3.4) and the Chang recipes to **try** to recreate all of our by-hand plots using R



