---
output:
  xaringan::moon_reader:
    css: ["default", "extra.css"]
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
##xaringan::inf_mr()
## For offline work: https://bookdown.org/yihui/rmarkdown/some-tips.html#working-offline
## Images not appearing? Put images folder inside the libs folder as that is the main data directory

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
##library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

class: slideblue

.size80[**Today's Agenda**]

<br>

.size40[

1. Review descriptive statistics concepts

2. Explore new ways to manipulate data in R

]

<br>
<br>

.center[.size40[
  Justin Leinaweaver (Fall 2021)
]]



## Wheelan ch2 and Johnson 2012 introduces:

\large

### Measures of Central Tendency 
\begin{itemize}
\item Mean, Median and Mode
\end{itemize}

### Measures of Variability
\begin{itemize}
\item Range and the Interquartile Range
\end{itemize}

### Deviations from Central Tendency
\begin{itemize}
\item Standard deviation
\end{itemize}


## Assignment for Thursday {.plain}
\large
1. Read Wheelan ch 3 and review the recipes in the syllabus

2. Calculate and interpret the descriptive statistics for the variables in our dataset.

\begin{center}
\includegraphics[keepaspectratio=true,width=.65\paperwidth]{Images/02_1-summary.png}
\end{center}



## Quick Review: RStudio Basics
\center
\includegraphics[keepaspectratio=true,width=.95\paperwidth]{Images/02_2-moodle_data1.png}


## Quick Review: RStudio Basics
\center
\includegraphics[keepaspectratio=true,width=.95\paperwidth]{Images/02_2-moodle_data2.png}


## Quick Review: RStudio Basics {.plain}
\begin{center}
\includegraphics[keepaspectratio=true,width=.6\paperwidth]{Images/02_1-RStudio_setup.png}
\end{center}


## Quick Review: Rmarkdown Basics {.plain}
\begin{center}
\includegraphics[keepaspectratio=true,width=.6\paperwidth]{Images/02_2-chunk_refresher.png}
\end{center}



## Class Practice!
\LARGE
\center
**Line up in order by height (again)**

\vspace{.2in}

### Identify the:

\vspace{.15in}

+ Median person,

\vspace{.15in}

+ Modal height (in),

\vspace{.15in}

+ Range (in), and

\vspace{.15in}

+ Interquartile Range (IQR) (in)



## Let's check our work! {.plain}
\LARGE
```{r, echo=TRUE, eval=FALSE}
## Save class heights as an object
heights <- c(68, 62, 58, 53, ...)

## Check the summary function
summary(heights)

## Standard Deviation
sd(heights)
```


## {.plain}
\LARGE
\center
**If you could only choose one or two of these measures to summarize our class heights, which would you select? Why?**

\vspace{.2in}

\Large
### Measures of Central Tendency 
\begin{itemize}
\item Mean, Median and Mode
\end{itemize}

### Measures of Variability
\begin{itemize}
\item Range and the Interquartile Range
\end{itemize}

### Deviations from Central Tendency
\begin{itemize}
\item Standard deviation
\end{itemize}



## Standard Deviation: The context we need for the mean  {.plain}
```{r, fig.align='center', fig.width=6, fig.asp=0.5, out.width='95%'}
set.seed(25)

x <- tibble(
    Low = rnorm(10000, mean = 0, sd = 3),
    High = rnorm(10000, mean = 0, sd = 10)
    )


x |>
pivot_longer(cols = Low:High, names_to = "SD", values_to = "Values") |>
mutate(
    SD = if_else(SD == "High", "High Std Deviation", "Low Std Deviation")
    ) |>
ggplot(aes(x = Values)) +
    geom_histogram(bins = 25, color = "white") +
    facet_wrap(~ SD) +
    theme_bw() +
    geom_vline(xintercept = 0, color = "pink", size = .8) +
    labs(x = "", y = "")
```


## Let's add to our `Getting_Started.Rmd` report
\huge

1. Individual Summary Statistics

\vspace{.2in}

2. Learning to build code using the pipe

\vspace{.2in}

3. Key functions for manipulating datasets



## Packages make R even more powerful
\Large
```{r, echo=TRUE, eval=FALSE}
## Tidying and manipulating data with the tidyverse

## Installation only necessary one time
install.packages("tidyverse")

## Run the library command when you 
## want to use the package
library(tidyverse)
```


## R Built-in Dataset: mtcars
\Large 
```{r}
## Load package
library(tidyverse)

## R has a series of built-in datasets
mtcars[1:4, 1:8]
```

The data was extracted from the 1974 \textit{Motor Trend} US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).


## The Summary Function: Powerful but overwhelming {.plain}
\footnotesize
```{r, echo=FALSE}
options("width" = 100)
summary(mtcars)
```


## Zooming In: The summary function plus '$'
\Large
```{r, echo = TRUE}
summary(mtcars$mpg)

summary(mtcars$cyl)
```


## Zooming In: Individual Summary Statistics Using '$' {.plain}
::: columns
:::: column

\Large
```{r, echo = TRUE, eval=FALSE}
## Mean
mean(mtcars$mpg)
## Median
median(mtcars$mpg)
## Mode
table(mtcars$mpg)
```

::::
:::: column

\Large
```{r, echo = TRUE, eval=FALSE}
## Minimum
min(mtcars$mpg)
## Maximum
max(mtcars$mpg)
## Standard deviation
sd(mtcars$mpg)
```

::::
:::

\vspace{.2in}
\Large
```{r, echo = TRUE, eval=FALSE}
# 1st and 3rd quartiles
quantile(mtcars$mpg, probs = .25)
quantile(mtcars$mpg, probs = .75)
```


## Zooming In: Individual Summary Statistics Using the Pipe
::: columns
:::: column
\Large
\vspace{.3in}

```{r, echo = TRUE, eval = FALSE}
mtcars
```

\vspace{1in}

**Note:** 

Output doesn't fit on the slide.

::::
:::: column
\tiny
```{r, echo = FALSE}
mtcars
```
::::
:::


## Zooming In: Individual Summary Statistics Using the Pipe
::: columns
:::: column
\Large
\vspace{.3in}

```{r, echo = TRUE, eval = FALSE}
mtcars |>
   summary()
```

\vspace{.8in}

\normalsize
```{r, echo = TRUE, eval = FALSE}
## Some old resources may
## still use (%>%):
mtcars %>%
    summary()
```
::::
:::: column
\tiny
```{r, echo = FALSE}
options("width" = 60)
mtcars |>
   summary()
```
::::
:::



## Zooming In: Individual Summary Statistics Using the Pipe
::: columns
:::: column
\Large
\vspace{.3in}

```{r, echo = TRUE, eval = FALSE}
## select variables
mtcars |>
   select(mpg) |>
   summary()
```
::::
:::: column
\LARGE
```{r, echo = FALSE}
options("width" = 60)
mtcars |>
   select(mpg) |>
   summary()
```
::::
:::


## Why use the pipe? You can select multiple variables {.plain}
\Large
```{r, echo = TRUE}
mtcars |>
   select(mpg, cyl) |>
   summary()
```


## Why use the pipe? You can answer more interesting questions
\Large
::: columns
:::: column
\begin{center}
\includegraphics[keepaspectratio=true,width=.3\paperwidth]{Images/02_2-4_cylinders.jpg}
\end{center}
::::
:::: column
\begin{center}
\includegraphics[keepaspectratio=true,width=.3\paperwidth]{Images/02_2-6_cylinders.jpg}
\end{center}
::::
:::
\LARGE
\begin{center}
\textbf{How does average MPG vary across different engine sizes?}
\end{center}


## arrange: Arrange rows by column values
::: columns
:::: column

\LARGE
```{r, echo=TRUE, eval=FALSE}
mtcars |>
   select(mpg, cyl) |>
   arrange(mpg)
```

::::
:::: column

\large
```{r, echo=FALSE, eval=TRUE}
mtcars |>
   select(mpg, cyl) |>
arrange(mpg) |>
slice(1:10)
```

::::
:::



## arrange: Arrange rows by column values
::: columns
:::: column

\LARGE
```{r, echo=TRUE, eval=FALSE}
mtcars |>
   select(mpg, cyl) |>
   arrange(desc(mpg))
```

::::
:::: column

\large
```{r, echo=FALSE, eval=TRUE}
mtcars |>
   select(mpg, cyl) |>
arrange(desc(mpg)) |>
slice(1:10)
```

::::
:::


## arrange: Arrange rows by MULTIPLE column values
::: columns
:::: column

\LARGE
```{r, echo=TRUE, eval=FALSE}
mtcars |>
   select(mpg, cyl) |>
   arrange(cyl, mpg)
```

::::
:::: column

\large
```{r, echo=FALSE, eval=TRUE}
mtcars |>
   select(mpg, cyl) |>
arrange(cyl, mpg) |>
slice(1:10)
```

::::
:::


## filter: Pick observations by their values {.plain}
::: columns
:::: column
\Large
```{r, echo = TRUE, eval=FALSE}
mtcars |>
    filter(cyl == 4) |>
    select(mpg) |>
    summary()
```
::::
:::: column
\Large
```{r, echo=FALSE}
mtcars |>
    filter(cyl == 4) |>
    select(mpg) |>
    summary()
```
::::
:::



## filter: Pick observations by their values {.plain}
::: columns
:::: column
\Large
```{r, echo = TRUE}
mtcars |>
    filter(cyl == 4) |>
    select(mpg) |>
    summary()
```
::::
:::: column
\Large
```{r, echo = TRUE}
mtcars |>
    filter(cyl == 8) |>
    select(mpg) |>
    summary()
```
::::
:::



## summarize: Collapse many values down to a single value
::: columns
:::: column

\Large 

```{r, echo = TRUE}
mean(mtcars$mpg)
```

::::
:::: column

\large 

```{r, echo = TRUE}
mtcars |>
    summarize(Mean = mean(mpg))
```
::::
:::


## Summarize produces a nice table
::: columns
:::: column

\Large 

```{r, echo = TRUE}
mean(mtcars$mpg)

sd(mtcars$mpg)

median(mtcars$mpg)
```

::::
:::: column

\large 

```{r, echo = TRUE}
mtcars |>
   summarize(
     Mean = mean(mpg),
     StdDev = sd(mpg),
     Median = median(mpg)
     )
```
::::
:::



## Summarize can be combined with other functions
::: columns
:::: column

\Large 

```{r, echo = TRUE}
mean(mtcars$mpg)

sd(mtcars$mpg)

median(mtcars$mpg)
```

::::
:::: column

\large 

```{r, echo = TRUE}
mtcars |>
   filter(cyl == 4) |>
   summarize(
     Mean = mean(mpg),
     StdDev = sd(mpg),
     Median = median(mpg)
     )
```
::::
:::



## Summarize and group_by is awesome
::: columns
:::: column

\Large 

```{r, echo = TRUE}
mean(mtcars$mpg)

sd(mtcars$mpg)

median(mtcars$mpg)
```

::::
:::: column

\large 

```{r, echo = TRUE}
mtcars |>
   group_by(cyl) |>
   summarize(
     Mean = mean(mpg),
     StdDev = sd(mpg),
     Median = median(mpg)
     )
```
::::
:::



## Answering more interesting questions with descriptive statistics
\LARGE

1. select: Pick variables by their names

\vspace{.1in}

2. arrange: Arrange rows by column values

\vspace{.1in}

3. filter: Pick observations by their values

\vspace{.1in}

4. summarize: Collapse many values down to a single value

\vspace{.1in}

5. group_by: Summarize for each level of a variable


## 
\huge

\begin{center}
\textbf{What did you find when calculating descriptive statistics for the main indices of Data Project 1?}
\end{center}

\vspace{.2in}

+ Civil Liberties (`CL` and `CL rating`)

+ Political Rights (`PR` and `PR rating`)

+ Total FH Scores (`Total`)



## {.plain}
```{r, echo=FALSE, eval=TRUE}
options("width" = 120)
d <- read_excel("./Data/All_data_FIW_2013-2021-tidy.xlsx", sheet = 2)
```
\Large
```{r, echo=TRUE, eval=FALSE}
## Summarize key measures
d |>
   select(`CL rating`, CL, `PR rating`, PR, Total) |>
   summary()
```

\small
```{r, echo=FALSE, eval=TRUE}
d |>
   select(`CL rating`, CL, `PR rating`, PR, Total) |>
   summary()
```



## {.plain}
::: columns
:::: column
\footnotesize
```{r, echo=FALSE}
options("width" = 80)

d <- read_excel("./Data/All_data_FIW_2013-2021-tidy.xlsx", sheet = 2)

x1 <- d |>
   select(Status:Total) |>
   select(-Status, -starts_with("Add")) |>
   pivot_longer(`PR rating`:Total, names_to = "Variables", values_to = "Values") |>
   group_by(Variables) |>
   summarize(
       mean = round(mean(Values), 1),
       StdDev = round(sd(Values), 1),
       min = min(Values),
       max = max(Values)
    )

x1[1:20,]
```
::::
:::: column
\footnotesize
```{r, echo=FALSE}
x1[21:37,]
```
::::
:::


## Analyzing the Freedom House Data
\Huge
\center
Use the pipe and summarize to calculate the minimum, mean and maximum values of the `Total` score overall


## 
\Large

```{r, echo=TRUE}
d |>
   summarize(
     Minimum = min(Total),
     Mean = mean(Total),
     Maximum = max(Total)
)
```


## Analyzing the Freedom House Data
\Huge
\center
Use the pipe, filter and summarize to calculate the minimum, mean and maximum values of the `Total` score in 2021 only


## 
\Large

```{r, echo=TRUE}
d |>
   filter(Edition == 2021) |>
   summarize(
     Minimum = min(Total),
     Mean = mean(Total),
     Maximum = max(Total)
)
```


## Analyzing the Freedom House Data
\Huge
\center
Use the pipe, group_by and summarize to calculate the minimum, mean and maximum values of the `Total` score for each year in the dataset.


## 
::: columns
:::: column

\Large

```{r, echo=TRUE, eval=FALSE}
d |>
   group_by(Edition) |>
   summarize(
     Minimum = min(Total),
     Mean = mean(Total),
     Maximum = max(Total)
)
```

::::
:::: column

```{r, echo=FALSE, eval=TRUE}
d |>
   group_by(Edition) |>
   summarize(
     Minimum = min(Total),
     Mean = mean(Total),
     Maximum = max(Total)
)
```

::::
:::



## Analyzing the Freedom House Data
\Huge
\center
Calculate the minimum, mean and maximum values of the `Total` score for each region in the dataset (overall).


## 
::: columns
:::: column

\Large

```{r, echo=TRUE, eval=FALSE}
d |>
   group_by(Region) |>
   summarize(
     Minimum = min(Total),
     Mean = mean(Total),
     Maximum = max(Total)
)
```

::::
:::: column

```{r, echo=FALSE, eval=TRUE}
d |>
   group_by(Region) |>
   summarize(
     Minimum = min(Total),
     Mean = mean(Total),
     Maximum = max(Total)
)
```

::::
:::


## Analyzing the Freedom House Data
\Huge
\center
Calculate the minimum, mean and maximum values of the `Total` score for each region per year in the dataset.


## 
::: columns
:::: column

\large

```{r, echo=TRUE, eval=FALSE}
d |>
   group_by(Region, Edition) |>
   summarize(
     Minimum = min(Total),
     Mean = mean(Total),
     Maximum = max(Total)
)
```

::::
:::: column

\footnotesize
```{r, echo=FALSE, eval=TRUE}
d |>
   group_by(Region, Edition) |>
   summarize(
     Minimum = min(Total),
     Mean = round(mean(Total), 1),
     Maximum = max(Total)
)
```

::::
:::



## Assignment 1
\huge

\begin{center}
Prepare a report \textbf{introducing and analyzing} our first data project. What are the \textbf{strengths and weaknesses} of this project for measuring regime type? Support all claims with evidence.
\end{center}


## Getting Started on Assignment 1
\LARGE

If you were introducing someone to the strengths and weaknesses of this data project, what would you want to include in your discussion? 

1. Strengths and weaknesses of the methodology?

2. Strengths and weaknesses demonstrated using the descriptive statistics?


