---
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    df_print: kable
fontsize: 14pt
classoption: "aspectratio=169"
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "./03_2-Univariate_Viz2.Rmd", output_file = "./03_2-Univariate_Viz2.pdf")

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
##library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```


## Today's Agenda

\Large
Building univariate visualizations in R

+ Save your notes in `Getting_Started.Rmd`

+ Section: # Visualizing Data in R

\vspace{.5in}

\begin{center}
Justin Leinaweaver (Fall 2021)
\end{center}


## ggplot() for making visualizations
\begin{center}
\includegraphics[keepaspectratio=true,width=.9\paperwidth]{Images/03_2-ggplot_logo.png}
\end{center}

\center
\large

**Note: You have to load the tidyverse to use ggplot**

```{r, echo=TRUE}
library(tidyverse)
```


## Healy (2021) Introduces ggplot() {.plain}
```{r, echo = TRUE, eval = FALSE}
## 1. Tidy and 2. map your data
p <- ggplot(data = data source, aes(x = variable name))

## 3. Variable type determines geom (choose one)
p + geom_bar()
p + geom_col()
p + geom_histogram()
p + geom_boxplot()
p + geom_line()
```

\center
**This code is on Discord now!**



## The GGplot Function: Two Arguments {.plain}

```{r, echo = TRUE, eval = FALSE}
## 1. Tidy and 2. map your data
p <- ggplot(data = data source, aes(x = variable name))
```

\vspace{.25in}
\Large
1. Identify the source of your tidy data, and
\vspace{.25in}
2. Select your variable(s) using the `aes()` function




## {.plain}
```{r, echo = TRUE, eval = FALSE}
## 1. Tidy and 2. map your data
p <- ggplot(data = data source, aes(x = x axis variable))
```

\vspace{.1in}

\begin{center}
\includegraphics[keepaspectratio=true,width=.35\paperwidth]{Images/03_2-Axes.png}
\end{center}

+ The `x` here refers to the x axis.

+ Rotate the plot by changing `x` to `y`





## Healy (2021) Introduces ggplot() {.plain}

```{r, echo = TRUE, eval = FALSE}
## 1. Tidy and 2. map your data
p <- ggplot(data = data source,
            aes(x = x axis variable, y = y axis variable))

## 3. Variable type determines geom (Chang recipes)
p + geom_bar() ## 3.3
p + geom_col() ## 3.1
p + geom_histogram() ## 6.1
p + geom_boxplot() ## 6.6
p + geom_line() ## 4.1
```


## Built-in Dataset: chickwts
::: columns
:::: column
\vspace{.4in}
\begin{center}
\large
An experiment was conducted to measure and compare the effectiveness of various feed supplements on the growth rate of chickens.
\end{center}
::::
:::: column
\small
```{r, echo = FALSE}
chickwts |> slice(1:9)
```
::::
:::


## Built-in Dataset: chickwts
::: columns
:::: column
\vspace{.25in}
\begin{center}
\large
An experiment was conducted to measure and compare the effectiveness of various feed supplements on the growth rate of chickens.
\end{center}
::::
:::: column
\small
```{r, echo = FALSE}
chickwts |> slice(1:7)
```
::::
:::
\vspace{.25in}
\center
\Large
**How do we visualize the `feed` variable?**



## Built-in Dataset: chickwts {.plain}
::: columns
:::: column
\begin{center}
\vspace{.25in}
\large
An experiment was conducted to measure and compare the effectiveness of various feed supplements on the growth rate of chickens.
\end{center}
::::
:::: column
\small
```{r, echo = FALSE}
chickwts |> slice(1:7)
```
::::
:::
\vspace{.25in}
\large
```{r, echo = TRUE}
## 1. Tidy and 2. map your data
p <- ggplot(data = chickwts, aes(x = feed))
```


## {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='60%'}
## 1. Tidy and 2. map your data
p <- ggplot(data = chickwts, aes(x = feed))
## View the visualization
p
```


## {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='60%'}
## 1. Tidy and 2. map your data
p <- ggplot(data = chickwts, aes(x = feed))
## View the visualization
p + geom_bar()
```


## {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='60%'}
## 1. Tidy and 2. map your data
p <- ggplot(data = chickwts, aes(y = feed))
## View the visualization
p + geom_bar()
```


## Learn to Decipher the Error Text! {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='60%', eval=FALSE}
## 1. Tidy and 2. map your data
p <- ggplot(data = chickwts, aes(x = feed))
## View the visualization
p + geom_histogram()
```
\vspace{.25in}
\large
\textcolor{red}{Error: } StatBin requires a continuous x variable: the x variable is discrete. Perhaps you want stat="count"?



## Know what you are making BEFORE you make it {.plain}

```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='50%', eval=TRUE}
## View the visualization
p + geom_boxplot()
```

## Bar Plots: Counts (Chang 3.3) {.plain}
So, if your variable looks like this (categorical):

```{r, echo = FALSE}
chickwts |> select(feed) |> slice(1:5)
```

```{r, echo = TRUE, eval = FALSE}
## geom_bar counts the levels for you
p + geom_bar()
```


## 
\center
\large
**How do we make a bar plot if the levels have already been counted?**

\vspace{.1in}

::: columns
:::: column
```{r, echo = TRUE, eval = FALSE}
## Function to count levels
chickwts |>
   count(feed)
```
::::
:::: column
\normalsize
```{r, echo = FALSE}
chickwts |>
    count(feed)
```
::::
:::



## {.plain}
\large
::: columns
:::: column
```{r, echo = TRUE, eval = FALSE}
## Save the new data
new1 <- chickwts |>
           count(feed)
```
::::
:::: column
\normalsize
```{r, echo = FALSE}
chickwts |>
    count(feed)
```
::::
:::
\normalsize
\vspace{.3in}
```{r, echo = TRUE}
## 1. Tidy and 2. map your data
p <- ggplot(data = new1, aes(x = feed, y = n))
```


## {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='60%'}
## 1. Tidy and 2. map your data
p <- ggplot(data = new1, aes(x = feed, y = n))
## View the visualization
p
```


## {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='60%'}
## 1. Tidy and 2. map your data
p <- ggplot(data = new1, aes(x = feed, y = n))
## View the visualization
p + geom_col()
```


## geom_bar vs geom_col {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='50%', eval=FALSE}
## 1. Tidy and 2. map your data
p <- ggplot(data = new1, aes(x = feed, y = n))
## View the visualization
p + geom_bar()
```
\vspace{.25in}
\large
\textcolor{red}{Error: } stat_count() can only have an x or y aesthetic.


## {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='55%'}
## All in one line (with pipes)
chickwts |>
   count(feed) |>
   ggplot(aes(x = feed, y = n)) +
      geom_col()
```


## Built-in Dataset: mtcars
::: columns
:::: column
\center
\large
The data was extracted from the 1974 \textit{Motor Trend} US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).
::::
:::: column
\small
```{r, echo = FALSE}
mtcars |> slice(1:9) |> select(mpg, cyl)
```
::::
:::


## Built-in Dataset: mtcars
\small
```{r, echo = FALSE}
mtcars |> slice(1:6) |> select(mpg, cyl)
```
\center
\Large
**How do we visualize the `mpg` variable?**



## Built-in Dataset: mtcars
\small
```{r, echo = FALSE}
mtcars |> slice(1:6) |> select(mpg, cyl)
```
\normalsize
```{r, echo = TRUE}
## 1. Tidy and 2. map your data
p <- ggplot(data = mtcars, aes(x = mpg))
```


## {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='55%'}
## 1. Tidy and 2. map your data
p <- ggplot(data = mtcars, aes(x = mpg))
## View the visualization
p + geom_boxplot()
```


## {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='55%'}
## 1. Tidy and 2. map your data
p <- ggplot(data = mtcars, aes(x = mpg))
## View the visualization
p + geom_histogram()
```


## {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='55%'}
## 1. Tidy and 2. map your data
p <- ggplot(data = mtcars, aes(x = mpg))
## Adjust bin numbers
p + geom_histogram(bins = 8)
```

## {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='55%'}
## 1. Tidy and 2. map your data
p <- ggplot(data = mtcars, aes(x = mpg))
## Adjust bin numbers
p + geom_histogram(bins = 15)
```


## Comparing Histograms to Box Plots
::: columns
:::: column
```{r, echo = FALSE, fig.asp = 0.9, fig.width = 4}
p <- ggplot(data = mtcars,
            aes(x = mpg))

p + geom_boxplot()
```
::::
:::: column
```{r, echo = FALSE, fig.asp = 0.9, fig.width = 4}
p <- ggplot(data = mtcars,
            aes(x = mpg))

p + geom_histogram(bins = 8)
```
::::
:::



## R Built-in Dataset: economics
This dataset was produced from US economic time series data available from \url{https://fred.stlouisfed.org/}.

\small
```{r}
economics |> slice(1:8)
```


## Line Plots (Chang 4.1)
\small
```{r}
economics |> slice(1:5) |> select(date, pop)
```
\normalsize
```{r, echo = TRUE}
## 1. Tidy and 2. map your data
p <- ggplot(data = economics, aes(x = date, y = pop))
```


## {.plain}
```{r, echo = TRUE, fig.asp = 0.618, fig.width = 4, fig.align='center', out.width='55%'}
## 1. Tidy and 2. map your data
p <- ggplot(data = economics, aes(x = date, y = pop))
## Line plot (Chang 4.1)
p + geom_line()
```


## Re-make the visualizations from Tuesday
\Large
1. Bar Plots: 1. `CL Rate` and 2. `PR Rate`

\vspace{.15in}

2. Box Plots: 1. `CL`, 2. `PR`, 3. `Total`

\vspace{.15in}

3. Histogram of the `Total` score

\vspace{.15in}

4. Line plot of the **average** `Total` score per year



## Controlling Figures in Reports
::: columns
:::: column

\begin{center}
\includegraphics[keepaspectratio=true,width=.47\paperwidth]{Images/03_2-Chunk_Options.png}
\end{center}

::::
:::: column

\begin{center}
\includegraphics[keepaspectratio=true,width=.47\paperwidth]{Images/03_2-Chunk_Sizes.png}
\end{center}

::::
:::


## Analyze a Data Project: Regime Type
\Large

\begin{center}
Prepare a report \textbf{introducing and analyzing} our first data project. What are the \textbf{strengths and weaknesses} of this project for measuring regime type? 
\end{center}

+ Support all claims with evidence.




