---
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    df_print: kable
fontsize: 14pt
classoption: "aspectratio=169"
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "./04_2-Data_Project1.Rmd", output_file = "./04_2-Data_Project1.pdf")

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
##library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```


## Today's Agenda
\LARGE
\begin{center}
Data Project 1: Due Sunday (Sep 19)
\end{center}

\large
\vspace{.25in}

Get ready for class:

+ install.packages("rworldmap")

\vspace{.25in}
\center
Justin Leinaweaver (Fall 2021)


## Polishing Visualizations
::: columns
:::: column
\Large
\vspace{.1in}
1. Informative titles
\vspace{.15in}
2. Figure labels
\vspace{.15in}
3. Clean axis labels
\vspace{.15in}
4. Source info
\vspace{.15in}
5. No chart junk

::::
:::: column

\vspace{.2in}

```{r, echo = FALSE, fig.asp = .8, fig.width = 4, out.width = '95%'}
ggplot(data = chickwts, aes(x = feed)) +
    geom_bar(fill = "gold3", width = .6) +
    ggthemes::theme_tufte() +
    labs(x = "", y = "Counts", caption = "Source: Anonymous (1948) Biometrika",
         title = "Which feed builds chicken weights the fastest?") +
    geom_hline(yintercept = c(5, 10), color = "white")
```

::::
:::


## {.plain}
\footnotesize
```{r, echo = TRUE, eval = TRUE, fig.align='center', fig.asp=0.618, fig.width=4, out.width='60%'}
# Map the tidy data
p <- ggplot(data = chickwts, aes(x = feed))

# Visualize the default plot
p +
    geom_bar()
```


## {.plain}
\small
```{r, echo = TRUE, eval = TRUE, fig.align='center', fig.asp=0.618, fig.width=4, out.width='60%'}
# Color in the bars and narrow the width
p +
    geom_bar(fill = "gold3", width = .6)
```



## {.plain}
::: columns
:::: column
\scriptsize
```{r, echo = TRUE, eval = TRUE, fig.align='center', fig.asp=0.618, fig.width=4, out.width='65%'}
p +
    geom_bar(fill = "gold3", width = .6) +
    theme_gray()

p +
    geom_bar(fill = "gold3", width = .6) +
    theme_bw()
```
::::
:::: column
\scriptsize
```{r, echo = TRUE, eval = TRUE, fig.align='center', fig.asp=0.618, fig.width=4, out.width='65%'}
p +
    geom_bar(fill = "gold3", width = .6) +
    theme_classic()

p +
    geom_bar(fill = "gold3", width = .6) +
    theme_minimal()
```
::::
:::


## {.plain}
\footnotesize
```{r, echo = TRUE, eval = TRUE, fig.align='center', fig.asp=0.618, fig.width=5, out.width='60%'}
# Let's try theme bw
p +
    geom_bar(fill = "gold3", width = .6) +
    theme_bw()
```


## {.plain}
\footnotesize
```{r, echo = TRUE, eval = TRUE, fig.align='center', fig.asp=0.618, fig.width=5, out.width='60%'}
# Modify the axes, captions and titles
p +
    geom_bar(fill = "gold3", width = .6) +
    theme_bw() +
    labs(x = "", y = "Counts", caption = "Source: Anonymous (1948) Biometrika",
         title = "Which feed builds chicken weights the fastest?")
```


##

\center
\huge

For the brave, we can get crazier...


## {.plain}
::: columns
:::: column
\scriptsize
```{r, echo = TRUE, eval = TRUE, fig.align='center', fig.asp=0.618, fig.width=5.5, out.width='65%'}
p +
    geom_bar(fill = "gold3", width = .6) +
    ggthemes::theme_economist()

p +
    geom_bar(fill = "gold3", width = .6) +
    ggthemes::theme_wsj()
```
::::
:::: column
\scriptsize
```{r, echo = TRUE, eval = TRUE, fig.align='center', fig.asp=0.618, fig.width=4.5, out.width='65%'}
p +
    geom_bar(fill = "gold3", width = .6) +
    ggthemes::theme_fivethirtyeight()

p +
    geom_bar(fill = "gold3", width = .6) +
    ggthemes::theme_tufte()
```
::::
:::


## {.plain}
\footnotesize
```{r, echo = TRUE, eval = TRUE, fig.align='center', fig.asp=0.618, fig.width=5, out.width='60%'}
## Fancier themes
p +
    geom_bar(fill = "gold3", width = .6) +
    ggthemes::theme_tufte() +
    labs(x = "", y = "Counts", caption = "Source: Anonymous (1948) Biometrika",
         title = "Which feed builds chicken weights the fastest?")
```


## {.plain}
\footnotesize
```{r, echo = TRUE, eval = TRUE, fig.align='center', fig.asp=0.618, fig.width=5, out.width='60%'}
p +
    geom_bar(fill = "gold3", width = .6) +
    ggthemes::theme_tufte() +
    labs(x = "", y = "Counts", caption = "Source: Anonymous (1948) Biometrika",
         title = "Which feed builds chicken weights the fastest?") +
    geom_hline(yintercept = c(5, 10), color = "white")
```


## {.plain}
\small
```{r, echo = TRUE, eval = FALSE, fig.align='center', fig.asp=0.618, fig.width=4, out.width='40%'}
## Customizing a line
## Line types: 'solid', 'dashed', 'dotted', 'dotdash', 'longdash', 'twodash'
ggplot(data = economics, aes(x = date, y = pop)) +
    geom_line(size = 1.3, color = "blue", linetype = "dashed")
```
::: columns
:::: column

```{r, echo = FALSE, eval = TRUE, fig.align='center', fig.asp=0.618, fig.width=4, out.width='60%'}
## Customizing a line
## Line types: 'solid', 'dashed', 'dotted', 'dotdash', 'longdash', 'twodash'
ggplot(data = economics, aes(x = date, y = pop)) +
    geom_line(size = 1.3, color = "blue", linetype = "dashed") +
    theme_classic()

ggplot(data = economics, aes(x = date, y = pop)) +
    geom_line(size = 1.3, color = "red", linetype = "dotted") +
    theme_classic()
```

::::
:::: column

```{r, echo = FALSE, eval = TRUE, fig.align='center', fig.asp=0.618, fig.width=4, out.width='60%'}
## Customizing a line
## Line types: 'solid', 'dashed', 'dotted', 'dotdash',
##             'longdash', 'twodash'
ggplot(data = economics, aes(x = date, y = pop)) +
    geom_line(size = 1.3, color = "green3", linetype = "dotdash") +
    theme_classic()

ggplot(data = economics, aes(x = date, y = pop)) +
    geom_line(size = 1.3, color = "purple", linetype = "twodash") +
    theme_classic()
```

::::
:::



## End of Class: Let's make a map! {.plain}
\center
\includegraphics[keepaspectratio=true,width=.8\paperwidth]{Images/04_2-world_map.png}



## Tuesday: Working on Paper 1 {.plain}
\large
Prepare a report \textbf{introducing and analyzing} our first data project. What are the \textbf{strengths and weaknesses} of this project for measuring regime type? Support all claims with evidence.

+ Section 1: Introducing the Freedom House Data Project
+ Section 2: Analyzing the Methodology
+ Section 3: Analyzing the World in 2021



## Brainstorming Section 4: Ideas to Test?
\Large
\center
Use the data to present and analyze the trends in the world across time

::: columns
:::: column
### Categorical
+ `PR rating`
+ `CL rating`
+ `Status`
::::
:::: column
### Numerical
+ `Total`
+ `PR`
+ `CL`
::::
:::


## 5) Conclusions
\LARGE
**Key takeaways**

Based on your analyses how can we best use this data project to understand the world?



## Let's make a map! {.plain}
\center
\includegraphics[keepaspectratio=true,width=.8\paperwidth]{Images/04_2-world_map.png}



## Let's make a map!
\Large

1. Load the `rworldmap` package

2. Pick your per country data

3. Link your per country data to the country names in the map

4. Output the map



## Let's make a map!
\large

```{r, echo=TRUE, eval=FALSE}
## Install the package (one time)
install.packages("rworldmap") # One time
```

\vspace{.25in}

```{r, echo=TRUE, eval=FALSE}
## 1. Load the package
library(rworldmap)
```


## Let's make a map!

```{r, echo=TRUE, eval=FALSE}
## 2. Pick your per country data (FH in 2021)
d_2021 <- d |>
            filter(Edition == 2021) |>
            select("Country/Territory", "CL rating")
```

\scriptsize
```{r}
d |> filter(Edition == 2021) |> select("Country/Territory", "CL rating") |> slice(1:8)
```



## {.plain}
\small

```{r, echo=TRUE, eval=FALSE}
## 1. Load the package
library(rworldmap)

## 2. Pick your per country data (FH in 2021)
d_2021 <- d |>
             filter(Edition == 2021) |>
             select("Country/Territory", "CL rating")

## 3. Link your data to the map
map1 <- joinCountryData2Map(dF = d_2021,
                            joinCode = "NAME",
                            nameJoinColumn = "Country/Territory")

## 4. Ouput your map
mapCountryData(map1, nameColumnToPlot = "CL rating")
```


## {.plain}
```{r, echo=TRUE, eval=FALSE}
## Output a categorical variable and customize colors
mapCountryData(map1,
               nameColumnToPlot = "CL rating",
               mapTitle="Civil Liberties in 2021 (FH)",
               catMethod = "categorical",
               colourPalette = "heat")
```

\vspace{.2in}

### Color Options

\vspace{.1in}

1. a vector of valid colours, e.g. =c(’red’,’white’,’blue’)

\vspace{.1in}

2. one of "heat", "diverging", "white2Black", "black2White", "topo", "rainbow", "terrain", "negpos8", "negpos9

