---
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    df_print: kable
fontsize: 14pt
classoption: "aspectratio=169"
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "./05_1-Polish_Project_1.Rmd", output_file = "./05_1-Polish_Project_1.pdf")

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```


## Today's Agenda
\LARGE
\begin{center}
Data Project 1: Due Today (Sep 21)
\end{center}

\large
\vspace{.15in}

**For Class Thursday:**

1. install.packages("tinytex")
2. tinytex::install_tinytex()

\vspace{.15in}
\center

Justin Leinaweaver (Fall 2021)


## Report 1 {.plain}
\large
Prepare a report \textbf{introducing and analyzing} our first data project. What are the \textbf{strengths and weaknesses} of this project for measuring regime type? Support all claims with evidence.

+ Section 1: Introducing the Freedom House Data Project
+ Section 2: Analyzing the Methodology
+ Section 3: Analyzing the World in 2021
+ Section 4: Analyzing Trends in the World
+ Section 5: Conclusions


## Hide the Tidyverse Messages
\center
\includegraphics[keepaspectratio=true,width=.85\paperwidth]{Images/05_1-hide_tidy_messages.png}


## Outputting Tables

```{r, echo=TRUE, eval=FALSE}
## Install the package
install.packages("kableExtra")

## Load the package
library(kableExtra)
```

"Create Awesome HTML Table with knitr::kable and kableExtra"



