---
fontsize: 14pt
classoption: "aspectratio=169"
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    df_print: kable
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "./05_2-Building_Beamer_Presentations.Rmd", output_file = "./05_2-Building_Beamer_Presentations.pdf")

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```


## Today's Agenda
\Large

\begin{center}
Conference presentations in R/RStudio with Rmarkdown
\end{center}

\large
\vspace{.15in}

**Get Ready for Class by:**

1. install.packages("tinytex")
2. tinytex::install_tinytex()

\vspace{.15in}
\center

Justin Leinaweaver (Fall 2021)



## Goal: Design a Conference Presentation
\center
\includegraphics[keepaspectratio=true,width=.7\paperwidth]{Images/05_1-Conference_Presentation.jpg}



## For Your Notes: New Presentation Slides
::: columns
:::: column

\large
\vspace{.3in}
\begin{itemize}
\item New RMarkdown document
\vspace{.25in}
\item "Presentation"
\vspace{.25in}
\item "PDF (Beamer)"
\end{itemize}

::::
:::: column

\center
\includegraphics[keepaspectratio=true,width=.4\paperwidth]{Images/05_1-New_Beamer.png}

::::
:::



## Install a LaTex Distribution (once per install)
\Large

```{r, echo = TRUE, eval = FALSE}
## 1. The tinytex package will
##     install a mini-LaTex
install.packages("tinytex")

## 2. Run the following function
tinytex::install_tinytex()
```



## New Template File: Presentation\_Notes.Rmd
::: columns
:::: column

\center
\includegraphics[keepaspectratio=true,width=.4\paperwidth]{Images/05_1-moodle1.png}

::::
:::: column

\vspace{.3in}

1. Download new files to your "Getting Started" folder

\vspace{.25in}

2. Open `Presentation_Notes.Rmd`

\vspace{.25in}

3. `Knit` the slides

::::
:::



## Let's go through the Notes {.plain}
\Large
1. Hidden chunk for library and loading data
2. New slides '##'
3. \**italics*\* and \*\***bold**\*\*
4. Bullet points
5. R code chunks with the echo argument
6. Slides with figures
7. Two column layouts
8. Include external images



## New File: Data_Project\_1-Presentation.Rmd
\Large

+ Hypothetical Presentation: \textcolor{red}{8 minutes}
  + \Large Report sections 1, 2 \& 5: 1 slide each
  + Report sections 3 \& 4: 1-2 slides each

\vspace{.15in}

+ \textcolor{red}{Show}, don't tell (e.g. viz > words)

\vspace{.15in}

+ \textcolor{red}{`Knit`} as you go!



## For Tuesday
\Large

1. Finish your slides with a \textcolor{blue}{\textbf{theme}} and a \textcolor{blue}{\textbf{color}}

\vspace{.25in}

2. \textcolor{blue}{\textbf{Upload}} your slides to the Moodle workshop ('Assignments')





## {.plain}
::: columns
:::: column

\-\-\-

title: "Untitled"

author: "Justin Leinaweaver"

date: "9/13/2021"

output: beamer_presentation

\-\-\-

::::
:::: column

\-\-\-

title: "Untitled"

author: "Justin Leinaweaver"

date: "9/13/2021"

   output: 
   
      beamer_presentation:
	  theme: "?"
	  colortheme: "?"

\-\-\-

::::
:::

\vspace{.15in}
\center
\large
1. https://bookdown.org/yihui/rmarkdown/beamer-presentation.html#themes

2. https://deic-web.uab.cat/~iblanes/beamer_gallery/



## Custom YAML: Themes and Colors {.plain}
::: columns
:::: column

\vspace{.3in}

\-\-\-

title: "Untitled"

author: "Justin Leinaweaver"

date: "9/13/2021"

   output: 
   
      beamer_presentation:
	  theme: "CambridgeUS"
	  colortheme: "dolphin"

\-\-\-

::::
:::: column

\center
\includegraphics[keepaspectratio=true,width=.44\paperwidth]{Images/05_1-cambridge_dolphin.png}

::::
:::

\vspace{.15in}
\center
\large
https://deic-web.uab.cat/~iblanes/beamer_gallery/
