---
format: revealjs
---

## Today's Agenda {background-image="Images/background-data_blue_v4.png" .center}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)

# Data for today
#library(carData)

d <- read_excel("../../DATA/2026-Spring/WDI/Week05-Univariate_Practice/SP2026-Wk05-WDI_Data.xlsx", na = "NA")
```

<br>

::: {.r-fit-text}

**Creating, Transforming and Indexing Variables**

:::

- New Script: "Creating and Transforming Variables.R"

- Add `library(tidyverse)` to the top of the new script

<br>

::: r-stack
Justin Leinaweaver (Spring 2026)
:::

::: notes
Prep for Class

1. Review Canvas submissions

<br>

**SLIDE**: Aims for this week

:::



## {background-image="Images/background-data_blue_v4.png" .center}

**Univariate Analyses**

- Descriptive statistics (means, medians, modes, counts, SDs, etc)

- Visualizations (bar plots, box plots, histograms)

<br>

::: {.fragment}
**This Week**
:::

::: {.incremental}
1. Modifying variables to help us analyze and present them

2. Creating new variables to better measure our concepts
:::

::: notes

Thanks to the work you've done in univariate analysis you now have an array of tools for summarizing data

- e.g. descriptive statistics, frequency tables, bar plots, histograms and box plots

<br>

To really get the most out of those univariate tools we need to add some new skills this week

<br>

**REVEAL** x 2: Today we'll focus on modifying variables to make them easier to analyze and understand

- e.g. rescaling variables, transforming scales and converting numbers into categories

<br>

**REVEAL**: Wednesday and Friday we'll focus on creating new variables to better measure our concepts

- e.g. Combining variables to build an index like how Freedom house builds its measures of global freedom

<br>

Our work today is designed to help you solve specific problems you will run into when doing data analysis work

<br>

**SLIDE**: The first problem

:::



## {background-image="Images/background-data_blue_v4.png" .center}

**Problem**

- Labels (and statistics) are hard to read and understand

:::: {.columns}
::: {.column width="50%"}
```{r, fig.align='center', fig.retina=3, fig.asp=.85, fig.width=5, out.width="70%"}
midwest |>
  ggplot(aes(x = poptotal)) +
  geom_histogram(color = "white") +
  theme_bw() +
  labs(x = "Total Population (midwest counties)", y = "Count")
```
:::

::: {.column width="50%"}
```{r}
midwest |>
  summarize(
    N = sum(!is.na(poptotal)),
    Mean = mean(poptotal),
    SD = sd(poptotal),
    Min = min(poptotal),
    Max = max(poptotal)
  ) |>
  t() |>
  kable()
```
:::
::::

**Solution**

- Rescale the variable: Move the decimal

::: notes

*Read slide*

<br>

**SLIDE**: Ok, how do we do this?

:::




## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.retina=3, fig.asp=.618, fig.width=7}
# Create a new variable: Population in Thousands
midwest$poptotal_k <- midwest$poptotal / 1000

# Univariate analyses
summary(midwest$poptotal_k)

hist(midwest$poptotal_k)
```

::: notes

At this point everyone should understand:

- That you can access specific variables in a dataset with the "$" sign, and

- That you can modify variables using calculator type functions

- Now we're just adding in that we can save this new variable in the same dataset we started with

- **Does this make sense?**

<br>

Now, let's look at the results

- These summary stats are WAY easier to read, especially around the middle

- BUT I don't love these figure axis labels

    - Thinking about population as 1,000 thousands is confusing

<br>

In text I would definitely discuss the summary stats in thousands for this class, but I'd check rescaling into millions for the figure

- Everybody redo this code BUT convert poptotal into millions

- (**SLIDE**)

:::



## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.retina=3, fig.asp=.618, fig.width=7}
# Create a new variable: Population in Millions
midwest$poptotal_m <- midwest$poptotal / 1e6

# Univariate analyses
summary(midwest$poptotal_m)

hist(midwest$poptotal_m)
```

::: notes

**Everybody get this?**

<br>

I really like the histogram labels now.

- It clearly communicates to me that most counties are below 500,000 and there are a collection of HUGE outliers

<br>

Bottom line: Don't be afraid to move the decimal place on your variable if it makes the visualizations easier to read!

<br>

**SLIDE**: Problem 2

:::



## {background-image="Images/background-data_blue_v4.png" .center}

**Problem**

- Distribution is so right-skewed we can't see the actual variation

```{r, fig.align='center', fig.retina=3, fig.asp=.618, fig.width=5}
midwest |>
  ggplot(aes(x = poptotal_m)) +
  geom_histogram(color = "white") +
  theme_bw() +
  labs(x = "Total Population (midwest counties, millions)", y = "Count")
```

**Solution**

- Rescale the variable: Change the scale (log)

::: notes

Problem 2: The distribution is so right-skewed we can't see the differences between the bulk of the sample

<br>

Sticking with the poptotal variable

- **Remind me, what does it mean to say this variable is right skewed? What's the problem here?** 

- (The outliers are hiding the actual variation!)

- The counties at the low end aren't all the same and we want to see that variation too!

<br>

**SLIDE**: Let's fix it

:::



## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.retina=3, fig.asp=.618, fig.width=7}
# Create a new variable: Population in natural logs
midwest$poptotal_log <- log(midwest$poptotal)

# Univariate analyses
summary(midwest$poptotal_log)

hist(midwest$poptotal_log)
```

::: notes

**Did everybody get this?**

<br>

**SLIDE**: Let's briefly unpack this

:::



## {background-image="Images/background-data_blue_v4.png" .center .smaller}

:::: {.columns}
::: {.column width="50%"}
**Original Scale**

```{r, echo=FALSE, fig.align='center', fig.retina=3, fig.asp=.8, fig.width=5}
summary(midwest$poptotal)

#hist(midwest$poptotal)

ggplot(midwest, aes(x = poptotal)) +
  geom_histogram(color = "white") +
  theme_bw()
```
:::

::: {.column width="50%"}
**Log Scale**

```{r, echo=FALSE, fig.align='center', fig.retina=3, fig.asp=.8, fig.width=5}
summary(midwest$poptotal_log)

#hist(midwest$poptotal_log)

ggplot(midwest, aes(x = poptotal_log)) +
  geom_histogram(color = "white", bins = 10) +
  theme_bw() +
  scale_x_continuous(breaks = 8:15)
```
:::
::::

::: {.fragment}
```{r, echo=TRUE}
exp(8)

exp(9)
```
:::

::: notes

The key here is that both of these distributions describe the same data

- Taking the log has not changed the order of the counties at all

- We are simply examining the differences between them on the log scale instead of the normal counting scale

<br>

The log scale expands differences at the low end and shrinks the differences at the high end

- So this tall bar of small counties gets unpacked so we can see how much they differ from each other

- And the huge outliers are still there but get pulled back towards the rest of the distribution

<br>

**REVEAL**: We use exponentiation to undo the log conversion

- Here I'm showing you the population associated with a log value of 8 and a log value of 9

- Moving from log 8 to log 9 means your county size increases from 3k people to 8k people

- So, at the low end of the log scale small changes move you up the distirbution

<br>

**SLIDE**: At the other end of the scale

:::



## {background-image="Images/background-data_blue_v4.png" .center .smaller}

:::: {.columns}
::: {.column width="50%"}
**Original Scale**

```{r, echo=FALSE, fig.align='center', fig.retina=3, fig.asp=.8, fig.width=5}
summary(midwest$poptotal)

#hist(midwest$poptotal)

ggplot(midwest, aes(x = poptotal)) +
  geom_histogram(color = "white") +
  theme_bw()
```

:::

::: {.column width="50%"}
**Log Scale**

```{r, echo=FALSE, fig.align='center', fig.retina=3, fig.asp=.8, fig.width=5}
summary(midwest$poptotal_log)

#hist(midwest$poptotal_log)

ggplot(midwest, aes(x = poptotal_log)) +
  geom_histogram(color = "white", bins = 10) +
  theme_bw() +
  scale_x_continuous(breaks = 8:15)
```
:::
::::

```{r, echo=TRUE}
exp(13)

exp(14)
```

::: notes

At the high end your population size must grow massively to move you further up the scale

- Moving from log 13 to log 14 means your county size increases from 440k people to 1.2 million people!

- **Make sense?**

<br>

The cool thing here is you can use a simple rule of thumb to interpret changes on the log scale

- Each one unit move on the log scale is approximately a tripling of the underlying score

- Technically a natural log relies on Euler's constant of 2.72... but tripling is a useful approximation

<br>

So, in sum, we take the log if the distribution is right-skewed and this puts all of the values on the log scale

- We read increases on the log scale as tripling the score

- This will be something we do A LOT when we get to bivariate and multivariate analyses especially with regression

<br>

**Big picture questions on the log transformation?**

<br>

**SLIDE**: Problem 3

:::



## {background-image="Images/background-data_blue_v4.png" .center}

**Problem**

- You want to focus on one level or region of a distribution

```{r, fig.align='center', fig.retina=3, fig.asp=.618, fig.width=5}
mpg |>
  ggplot(aes(x = drv)) +
  geom_bar() +
  theme_bw() +
  labs(x = "Drive Trains", y = "Count")
```

**Solution**

- Create a dummy variable with ifelse()

::: notes

PROBLEM: You are interested in investigating one part of a distribution only

<br>

For example here, let's say we want to focus our analysis on front wheel drive cars only

- SOLUTION: to address this we'll create a dummy variable with ifelse()

<br>

**SLIDE**: Let's start by talking about the if_else function

:::



## Using if_else to evaluate your data {background-image="Images/background-data_blue_v4.png" .center}

<br>

```{r, echo=TRUE}
# Create a vector of numbers
x <- c(1, 3, 5, 7, 9)
```

<br>

```{r, echo=TRUE}
# Evaluate the numbers in the vector
x > 5
```

<br>

```{r, echo=TRUE}
# Use if_else to evaluate AND re-label
if_else(x > 5, "Above median", "Below median")
```

::: notes

*Step through the code*

<br>

The if_else function is a super handy piece of code to learn

- It basically has three elements

- Element 1 is a test with a true or false answer

- Element 2 is what you want to happen if the answer is true

- Element 3 is what you want to happen if the answer is false

<br>

Essentially if_else gives us a way to evaluate the observations in a variable and to recode them as something else

- **Make sense?**

<br>

**SLIDE**: Let's apply this to our problem

:::




## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.retina=3, fig.asp=.618, fig.width=7}
# Focus on one level of a nominal or ordinal variable 
mpg$front_wheel <- if_else(mpg$drv == "f", "Front Wheel", "Other")

# Univariate analyses
table(mpg$front_wheel)

barplot(table(mpg$front_wheel))
```

::: notes

So, this code creates a new variable

- The if_else test says that when the drive train value in the mpg dataset is "f" the observation should be marked "front wheel"

- When it is false, it should be marked as other

<br>

**Everybody get this code to work?**

<br>

**If you wanted to focus on rear wheel drive cars, what would you change the if_else test to?**

- ("r")

<br>

**SLIDE**: You can also convert an interval variable into a dummy

:::



## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.retina=3, fig.asp=.618, fig.width=7}
# Focus on a region of an interval variable 
mpg$biggest_engines <- if_else(mpg$displ >= 4.6, 
                               "Top 25%", "Bottom 75%")

# Univariate analyses
table(mpg$biggest_engines)

barplot(table(mpg$biggest_engines))
```

::: notes

We can also convert interval variables into a dummy variable

- Instead of the double equals sign we use the relational symbols we practiced a few weeks ago

- e.g. less than, greater than, etc

<br>

**Everybody get this code to work?**

<br>

**SLIDE**: Ok, last problem for today

:::




## {background-image="Images/background-data_blue_v4.png" .center}

**Problem**

- You want to convert an interval variable into an ordinal variable

```{r, fig.align='center', fig.retina=3, fig.asp=.618, fig.width=5}
mpg |>
  ggplot(aes(x = displ)) +
  geom_histogram(color = "white", bins = 10) +
  theme_bw() +
  labs(x = "Engine Displacement", y = "Count")
```

**Solution**

- Cut the interval variable into labelled bins

::: notes

PROBLEM: You want to collapse an interval variable into an ordinal variable

- This is essentially what you were asked to do in textbook exercises 2 and 4

<br>

Let's say we want to break engine sizes into four equally sized groups

- The quartiles do exactly this, but we want to label each car by its quartile

- **SLIDE**: The code

:::




## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE}
# Calculate the quartiles
summary(mpg$displ)
```

<br>

```{r, echo=TRUE}
# Cut the interval
cut(x = mpg$displ, breaks = c(1.5, 2.4, 3.3, 4.6, 7.2))
```

::: notes

First we use summary to find the quartiles, then we use cut to split the interval into bins

- IMPORTANT 1: Go a bit below the minimum and above the maximum to capture those values

- IMPORTANT 2: You must have one more "break" than you have groups!

    - The observations are being put into the bins between the values you have specified
    
    - SO, an engine with a 2 litre engine goes between 1.5 and 2.4
    
<br>

**Does this process make sense?**

<br>

**SLIDE**: Add labels

:::




## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.retina=3, fig.asp=.618, fig.width=7}
# Save the new ordinal variable
mpg$engine_size <- cut(x = mpg$displ, 
                       breaks = c(1.5, 2.4, 3.3, 4.6, 7.2),
                       labels = c("1st Q", "2nd Q", "3rd Q", "4th Q"))

# Univariate analyses
table(mpg$engine_size)

barplot(table(mpg$engine_size))
```

::: notes

**Everybody get this code to work?**

<br>

The cut function takes some practice, so let's do that

- Everybody cut the `poptotal` variable from the midwest dataset into its four quartiles

- Go!

<br>

(**SLIDE**)

:::



## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.retina=3, fig.asp=.618, fig.width=7}
# Save the new ordinal variable
midwest$pop_size <- cut(x = midwest$poptotal, 
    breaks = c(1700, 18840, 35324, 75651, 5105067),
    labels = c("1st Q", "2nd Q", "3rd Q", "4th Q"))

# Univariate analyses
table(midwest$pop_size)

barplot(table(midwest$pop_size))
```

::: notes

**Did everybody get this code?**

<br>

Most of our tools and techniques for today are about setting the table for the analyses to come.

- Once we start analyzing relationships we'll be doing quite a bit of variable transformations.

<br>

**SLIDE**: For next class

:::




## For Next Class {background-image="Images/background-data_blue_v4.png" .center}

<br>

1. Practice today's code on our Canvas assignment

    - Data: SP2026-Wk05-WDI_Data.xlsx
    
2. Read the background document on the Annual National Election Survey (ANES)

3. Read through the ANES variables list on Canvas

::: notes

Canvas Assignment: Practice modifying variables for better analyses

All figures submitted in this exercise must be made using ggplot and polished in line with class requirements (clear axis labels and a complete figure caption with informative title, key takeaway and data sourcing)

1. Compare a histogram of "GNI per capita, Atlas method (current US$)" in thousands of dollars to a second histogram using the logged version of the variable. Submit both histograms and explain what we learn about income in the world from these two views of the data.
2. Make a dummy variable that splits the WDI sample into two groups by level of measles immunization. Label all countries at or above 95% on "Immunization, measles (% of children ages 12-23 months)" as "Herd immunity" and the rest as "At risk". Submit a bar plot of your new variable and explain what we learn from this distribution.
3. Replicate the "wb_income" variable in our WDI sample using the cut function and the data in "GNI per capita, Atlas method (current US$)". Submit the code you used to make the new variable and make an argument that you believe this ordinal variable is (or isn't) a "better" way to think about global development.

World Bank Gross National Income Categories for 2023

- Low income (L) <= 1,145
- Lower middle income (LM) 1,146 - 4,515
- Upper middle income (UM) 4,516 - 14,005
- High income (H) > 14,005


:::