---
format: revealjs
---

## Today's Agenda {background-image="Images/background-data_blue_v4.png" .center}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)

load(file = "../../DATA/2026-Spring/ANES/ANES2020_Data_Extract.RData")

anes$engagement_index <- 
  if_else(anes$polact.givecand == "1. Yes", 1, 0) +
  if_else(anes$polact.giveparty == "1. Yes", 1, 0) +
  if_else(anes$polact.giveoth == "1. Yes", 1, 0) +
  if_else(anes$polact.postsign == "1. Yes", 1, 0) +
  if_else(anes$polact.meetings == "1. Yes", 1, 0) +
  if_else(anes$polact.onlinemeet == "1. Yes", 1, 0) +
  if_else(anes$polact.othwork == "1. Yes", 1, 0) +
  if_else(anes$polact.talkvote == "1. Yes", 1, 0) +
  if_else(anes$polact.talkpol == "1. Yes", 1, 0)

anes$negative_racial_attitudes <- as.numeric(anes$oth.race.concern) +
  as.numeric(anes$oth.race.feel) +
  as.numeric(anes$oth.race.persp) +
  as.numeric(anes$oth.race.protect)

load(file = "../../DATA/2026-Spring/Miller_Saunders_Farhart2016/Miller_Saunders_Farhart2016_v2.RData")
```

<br>

::: {.r-fit-text}

**Bivariate Analyses**

- Practice generating cross-tabs and group statistics

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2026)
:::

::: notes
Prep for Class

1. Review Canvas submissions

<br>

Last class we learned the tools we need to perform basic bivariate analyses in R

- Specifically we generated cross-tabs for analyzing the relationship between two categorical variables, and

- Group descriptive statistics to analyze the relationship between an interval outcome and a categorical predictor

<br>

**SLIDE**: Assignment for today

:::




## For Today {background-image="Images/background-data_blue_v4.png" .center .smaller}

<br>

**Analyze our Additive Index of Americans' Political Engagement (ANES)**

:::: {.columns}

::: {.column width="50%"}
**Required Steps**

1. Choose a predictor

2. Propose a theory

3. Test your hypothesis

4. Explain your findings!
:::

::: {.column width="50%"}
**Predictor Options**

- party identification
- relig.how.often
- social.class
- trust.media
- respondent's marital status
- active.duty.mil
- allow.refugees
- divided.govt
:::

::::
::: notes

**How did this go?**

- **SLIDE**: Let's talk findings!

<br>

Canvas ASSIGNMENT: Practicing Bivariate Descriptive Statistics

Last week we constructed and analyzed an additive index using ANES data that represents Americans' political engagement (political activities in the prio 12 months). I want you to perform a bivariate analysis that helps us understand the variation in this index across ONE of the predictors on the following list:

- party identification (partyid3)
- relig.how.often (5 point scale)
- social.class (4 point scale)
- trust.media (5 point scale)
- respondent's marital status (marital) 6 point scale
- active.duty.mil (3 point scale)
- allow.refugees (7 point)
- divided.govt (3 points)

I want you to do the following steps IN ORDER:

1. Propose a theory AND an associated hypothesis for how this predictor causes changes in our political engagement index as the outcome (e.g. increases or decreases). You must complete this step BEFORE analyzing the data
2. Run the analysis to test the relationship using bivariate descriptive statistics and report the results as a polished table
3. Explain what you have learned from this analysis. Is your theory supported by the data? Why or why not?

Please bear in mind that finding your theory is unsupported is as GOOD an outcome as finding it is supported. Our aim is to to generate knowledge and evaluate our theories, NOT to confirm our intuitions are correct!

:::



## {background-image="Images/background-data_blue_v4.png" .center .smaller}

```{r}
options(width = 110)
```

::: {.panel-tabset}

### engagement_index
```{r, fig.asp=.618, fig.width=7, fig.align='center'}
ggplot(data = anes, aes(x = engagement_index)) +
  geom_bar() +
  theme_bw() +
  scale_x_continuous(breaks = 0:9) +
  labs(x = "Americans' Political Engagement Index", y = "Count")

```


### partyid3
```{r}
df1 <- aggregate(data = anes, engagement_index ~ partyid3, summary)
labs1 <- levels(df1$partyid3)

df1 |> 
  select(-partyid3) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 30)
```

### relig.how.often
```{r}
df1 <- aggregate(data = anes, engagement_index ~ relig.how.often, summary)
labs1 <- levels(df1$relig.how.often)

df1 |> 
  select(-relig.how.often) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 25)
```


### social.class
```{r}
df1 <- aggregate(data = anes, engagement_index ~ social.class, summary)
labs1 <- levels(df1$social.class)

df1 |> 
  select(-social.class) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 28)
```


### trust.media
```{r}
df1 <- aggregate(data = anes, engagement_index ~ trust.media, summary)
labs1 <- levels(df1$trust.media)

df1 |> 
  select(-trust.media) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 25)
```

### marital
```{r}
df1 <- aggregate(data = anes, engagement_index ~ marital, summary)
labs1 <- levels(df1$marital)

df1 |> 
  select(-marital) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 22)
```


### active.duty.mil
```{r}
df1 <- aggregate(data = anes, engagement_index ~ active.duty.mil, summary)
labs1 <- levels(df1$active.duty.mil)

df1 |> 
  select(-active.duty.mil) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 28)
```


### allow.refugees
```{r}
df1 <- aggregate(data = anes, engagement_index ~ allow.refugees, summary)
labs1 <- levels(df1$allow.refugees)

df1 |> 
  select(-allow.refugees) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 22)
```


### divided.govt
```{r}
df1 <- aggregate(data = anes, engagement_index ~ divided.govt, summary)
labs1 <- levels(df1$divided.govt)

df1 |> 
  select(-divided.govt) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 28)
```

:::

::: notes

Ok, here's the univariate histogram

- This represents the variation to be explained!

- Before we look at results pitch us on your theory and hypothesis!

- *PRESENT and DISCUSS each*

<br>

**Bottom line, which are the most useful predictors of political engagement?**

- The differences are pretty small across these predictors

- Consistent with the lit: Independents less politically engaged than partisans

- Social class: Wealthier people more engaged

- Highest levels of trust

- Any religious attendance vs none

- Strong partisans (prefer one party control) larger engagement than those who don't care

<br>

**Questions on the assignment?**

- **The set-up, the code or the interpretations?**

<br>

Let's replicate this work but focused on our second index

- Americans' racial attitudes

- **SLIDE**: But first let's make some code tweaks

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Analyze our Additive Index of Americans' Racial Attitudes**
:::

```{r, echo=TRUE, eval=FALSE}
# Transform 1: Start the scale at 0
anes$negative_racial_attitudes0 <- anes$negative_racial_attitudes - 4

# Transform 2: Convert to % of total
anes$negative_racial_attitudes_pct <- anes$negative_racial_attitudes0 / 16

# Check new results
hist(anes$negative_racial_attitudes_pct)
```

```{r, echo=FALSE, eval=TRUE, fig.align='center', fig.width=7, out.width="50%", fig.asp=.618}
anes$negative_racial_attitudes0 <- anes$negative_racial_attitudes - 4
anes$negative_racial_attitudes_pct <- anes$negative_racial_attitudes0 / 16

ggplot(anes, aes(x = negative_racial_attitudes_pct)) +
  geom_histogram(bins = 8, color = "white") +
  theme_bw() +
  scale_x_continuous(breaks = seq(0,1,.2), labels = scales::percent_format()) +
  labs(x = "Americans' Negative Racial Attitudes", y = "Count")
```

::: notes

The political engagement index makes sense to me as a count

- e.g. each additional thing you report doing increases your score

<br>

The racial attitudes index isn't tracking concrete activities

- This represents an effort to capture the variation in racial animus a respondent holds

- And higher numbers on this scale mean greater negativity to other races

<br>

So, let's transform the index before we analyze it

- *step through code*

- **Everybody have this?**

<br>

**SLIDE**: Let's analyze it!

:::



## {background-image="Images/background-data_blue_v4.png" .center .smaller}

::: {.r-fit-text}
**Analyze our Additive Index of Americans' Racial Attitudes**
:::

<br>

:::: {.columns}

::: {.column width="50%"}
**Required Steps**

1. Choose a predictor

2. Propose a theory

3. Test your hypothesis

4. Explain your findings!
:::

::: {.column width="50%"}
**Predictor Options**

- party identification
- relig.how.often
- social.class
- trust.media
- respondent's marital status
- active.duty.mil
- allow.refugees
- divided.govt
:::

::::
::: notes

Let's do this again but with a different outcome to explain

- Work in PAIRS, choose a predictor, discuss a theory AND THEN test it!

- Get ready to report back on your findings

<br>

(**SLIDE**)

:::



## {background-image="Images/background-data_blue_v4.png" .center .smaller}

```{r}
options(width = 110)
```

::: {.panel-tabset}

### racial attitudes
```{r, fig.asp=.618, fig.width=7, fig.align='center'}
ggplot(anes, aes(x = negative_racial_attitudes_pct)) +
  geom_histogram(bins = 8, color = "white") +
  theme_bw() +
  scale_x_continuous(breaks = seq(0,1,.2), labels = scales::percent_format()) +
  labs(x = "Americans' Negative Racial Attitudes", y = "Count")

```


### partyid3
```{r}
df1 <- aggregate(data = anes, negative_racial_attitudes_pct ~ partyid3, summary)
labs1 <- levels(df1$partyid3)

df1 |> 
  select(-partyid3) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 30)
```

### relig.how.often
```{r}
df1 <- aggregate(data = anes, negative_racial_attitudes_pct ~ relig.how.often, summary)
labs1 <- levels(df1$relig.how.often)

df1 |> 
  select(-relig.how.often) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 25)
```


### social.class
```{r}
df1 <- aggregate(data = anes, negative_racial_attitudes_pct ~ social.class, summary)
labs1 <- levels(df1$social.class)

df1 |> 
  select(-social.class) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 28)
```


### trust.media
```{r}
df1 <- aggregate(data = anes, negative_racial_attitudes_pct ~ trust.media, summary)
labs1 <- levels(df1$trust.media)

df1 |> 
  select(-trust.media) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 25)
```

### marital
```{r}
df1 <- aggregate(data = anes, negative_racial_attitudes_pct ~ marital, summary)
labs1 <- levels(df1$marital)

df1 |> 
  select(-marital) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 22)
```


### active.duty.mil
```{r}
df1 <- aggregate(data = anes, negative_racial_attitudes_pct ~ active.duty.mil, summary)
labs1 <- levels(df1$active.duty.mil)

df1 |> 
  select(-active.duty.mil) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 28)
```


### allow.refugees
```{r}
df1 <- aggregate(data = anes, negative_racial_attitudes_pct ~ allow.refugees, summary)
labs1 <- levels(df1$allow.refugees)

df1 |> 
  select(-allow.refugees) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 22)
```


### divided.govt
```{r}
df1 <- aggregate(data = anes, negative_racial_attitudes_pct ~ divided.govt, summary)
labs1 <- levels(df1$divided.govt)

df1 |> 
  select(-divided.govt) |>
  t() |>
  kable(digits = 3, col.names = labs1) |>
  kable_styling(font_size = 28)
```

:::

::: notes

Before we look at results pitch us on your theory and hypothesis!

- *PRESENT and DISCUSS each*

<br>

**Bottom line, which are the most useful predictors of negative racial attitudes?**

- Party ID

- More religious attendance more tolerance?

- Wealthier is more tolerant?

- Media trust is strong positive effect on tolerance

- Allow refugees has an obviously strong correlation

<br>

**SLIDE**: Time remaining let's take a look at the conspiracy theory data!

:::



## Miller, Saunders & Farhart (2016) {background-image="Images/background-data_blue_v4.png" .center}

Data: Miller_Saunders_Farhart2016_v2.RData

<br>

**Hypotheses**

1. Belief in conspiracy theories increases as trust in government falls

2. Belief in conspiracy theories increases when they are tailored to a respondent's ideological perspective (motivated reasoning)

::: notes

The Miller, Saunders & Farhart (2016) proposes a fascinating and complicated theory for understanding conspiracy theory belief

- Over the course of the semester we will build up to replicating their main findings

- NOTE: Version 2 of the data file has two versions of party ID, use this one!

<br>

As a first stab at this I want us to consider these two hypotheses

- *Read slide*

- **Do these make sense?**

<br>

*Split class into FOUR groups*

- Go sit with your group!

- **SLIDE**: Group tasks

:::




## Miller, Saunders & Farhart (2016) {background-image="Images/background-data_blue_v4.png" .center .smaller}

Data: Miller_Saunders_Farhart2016.RData

<br>

**Outcomes**

- Liberal: CT_Bush_WMDs, CT_Govt_911, CT_Katrina_Levees, CT_2004_election

- Conservative: CT_Obama, CT_DeathPanels, CT_Climate_Hoax,  CT_Hussein_911

**Hypotheses**

1. Belief in liberal CTs increases as trust in government falls (Trust_Govt)

2. Belief in conservative CTs increases as trust in government falls (Trust_Govt)

3. Democrats are more likely to believe liberal coded conspiracies (Party3)

4. Republicans are more likely to believe conservative coded conspiracies (Party3)

::: notes

Groups, here are your jobs

- **Questions?**

- Go!

<br>

(**SLIDE**)

:::



## 1. Govt Trust and Liberal CTs {background-image="Images/background-data_blue_v4.png" .center .smaller}

```{r}
proportions(table("CT_Bush_WMDs" = d$CT_Bush_WMDs, d$Trust_Govt), margin = 2)
proportions(table("CT_Govt_911" = d$CT_Govt_911, d$Trust_Govt), margin = 2)
proportions(table("CT_Katrina_Levees" = d$CT_Katrina_Levees, d$Trust_Govt), margin = 2)
proportions(table("CT_2004_election" = d$CT_2004_election, d$Trust_Govt), margin = 2)
```



## 2. Govt Trust and Conservative CTs {background-image="Images/background-data_blue_v4.png" .center .smaller}

```{r}
proportions(table("CT_Obama" = d$CT_Obama, d$Trust_Govt), margin = 2)
proportions(table("CT_DeathPanels" = d$CT_DeathPanels, d$Trust_Govt), margin = 2)
proportions(table("CT_Climate_Hoax" = d$CT_Climate_Hoax, d$Trust_Govt), margin = 2)
proportions(table("CT_Hussein_911" = d$CT_Hussein_911, d$Trust_Govt), margin = 2)
```

::: notes

1. CT belief increases as trust in government falls


:::


## 3. Dems and Liberal CTs {background-image="Images/background-data_blue_v4.png" .center .smaller}

```{r}
proportions(table("CT_Bush_WMDs" = d$CT_Bush_WMDs, d$Party3), margin = 2)
proportions(table("CT_Govt_911" = d$CT_Govt_911, d$Party3), margin = 2)
proportions(table("CT_Katrina_Levees" = d$CT_Katrina_Levees, d$Party3), margin = 2)
proportions(table("CT_2004_election" = d$CT_2004_election, d$Party3), margin = 2)
```


::: notes

2. Democrats are more likely to believe liberal coded conspiracies


:::


## 4. Repubs and Conservative CTs {background-image="Images/background-data_blue_v4.png" .center .smaller}

```{r}
proportions(table("CT_Obama" = d$CT_Obama, d$Party3), margin = 2)
proportions(table("CT_DeathPanels" = d$CT_DeathPanels, d$Party3), margin = 2)
proportions(table("CT_Climate_Hoax" = d$CT_Climate_Hoax, d$Party3), margin = 2)
proportions(table("CT_Hussein_911" = d$CT_Hussein_911, d$Party3), margin = 2)
```

::: notes

3. Republicans are more likely to believe conservative coded conspiracies


:::




## For Next Class {background-image="Images/background-data_blue_v4.png" .center}

<br>

::: {.r-fit-text}

1. Read Chapter 5 in Pollock and Edwards

2. Test our CT hypotheses with the researchers' indexes!

    - Liberal CTs: Lib_Conspiracy_Index
    
    - Conservative CTs: Cons_Conspiracy_Index

    - Details on Canvas

:::

::: notes

Canvas ASSIGNMENT: Bivariate Tests of Conspiracy Theory Belief

1. Read Chapter 5 in Pollock and Edwards textbook

2. In class we tested four hypotheses inspired by the Muller, Saunders and Farhart (2016) research. For next class I want you to retest ALL FOUR hypotheses using indicies of conspiracy theory belief instead of individual questions. Submit polished tables of results PLUS your analysis: Does the data support EACH hypothesis or not?

    - NOTE: The researchers have already built the two indicies you need for these tests. `Lib_Conspiracy_Index` is an index of CT belief focused on the four liberal coded conspiracies. `Cons_Conspiracy_Index` is an index of CT belief focused on the four conservative coded conspiracies.

Hypotheses:

- H1. Belief in liberal CTs increases as trust in government falls
- H2. Belief in conservative CTs increases as trust in government falls
- H3. Democrats are more likely to believe liberal coded conspiracies
- H4. Republicans are more likely to believe conservative coded conspiracies


:::

