---
format: revealjs
---

# Today's Agenda {background-image="libs/Images/background-data_blue_v3.png"}

```{r}
#  background-size="1920px 1080px"
library(tidyverse)
library(readxl)
```

<br>

Critically Analyze the Research Design of Data Project 2

<br>

<br>

::: r-stack
Justin Leinaweaver (Spring 2024)
:::

::: notes
Prep for Class

1. Update slides BEFORE class
    - Insert research question agreed upon last class and update DAG
    - Add measures to evaluate
:::




## Do more powerful states tend to undermine the human rights of their citizens? {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.width = 10, fig.height=2}
## Manual DAG
d1 <- tibble(
  x = c(-3, 3),
  y = c(1, 1),
  labels = c("Military\nPower\n(CoW NMC)", "Human rights\nand\nthe rule of law\n(FFP)")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = -1.9, xend = 1.85, y = 1, yend = 1, arrow = arrow())
```

::: notes

I think this is coming along nicely!

<br>

**SLIDE**: Before we can test this relationship we need to go deep on our three chosen measures of government expenditures.

:::


## Report 2 {background-image="libs/Images/background-slate_v2.png"}

<br>

**Analyzing our Predictor Variable(s)**

<br>

1. Why is this project important?

2. How confident should we be in the methodology?

3. What do the measures currently show us?

4. How are these measures changing across time?

::: notes

Next week you write your second report.

- A focused univariate analysis of the predictor data set.

<br>

Our work today and Friday should help you make progress on the first two sections of the report.
:::



## Report 2 {background-image="libs/Images/background-slate_v2.png"}

<br>

:::: {.columns}
::: {.column width="50%"}
**The Measures**

1. TBD

2. TBD

3. TBD
:::

::: {.column width="50%"}
**Analyze the Methodology:**

- Source of the Data

- Operationalization

- Instrumentation

- Measurement Process

- Data Validation
:::
::::

::: notes

Here's our job for today.

<br>

### Any questions on the methodology elements we reviewed for the last report?

<br>

- Source: Where does the raw data come from?

- Operationalization: Defining the concepts

- Instrumentation: Designing the tool

- Measurement Process: Using the tool

- Validation: Checking the data

<br>

*Split class into three groups, one per measure*

- Groups, take 20 minutes to evaluate your measure across these four areas.

- Build us a pros and cons list for each item ON THE BOARD!

<br>

*Report back and discuss*
:::


## For Next Class {background-image="libs/Images/background-slate_v2.png"}

<br>

**Data Project 2**

1. **Audit**: How is the data organized? Are there missing data problems (time or region)?

2. **Explore**: Submit **THREE** interesting, puzzling or surprising things about the world of this dataset.

::: notes

From Canvas

- Audit the Dataset: Make sure you understand how the data is organized (observations and variables), and have spent time auditing it for missing data (time or region)

- Explore the Dataset: Submit THREE interesting, puzzling or surprising things about the world of this dataset. Fully explain each of your findings and how you can support your claims using the data.

:::

