---
format: revealjs
---

## Today's Agenda {background-image="Images/background-data_blue_v4.png" .center}

```{r}
#  background-size="1920px 1080px"
library(tidyverse)
```

<br>

::: {.r-fit-text}

Critically analyze the second data project's codebook

:::

<br>

<br>

::: r-stack
Justin Leinaweaver (Spring 2025)
:::

::: notes
Prep for Class

1. Review Canvas submissions

2. Update slides BEFORE class
    - Insert research question agreed upon last class and update DAG
    - Add measures to evaluate

3. Bring markers for every group (x6 in SP24)

4. Make sure the codebook and dataset are on Canvas

<br>

**SLIDE**: Our developing research project

:::




## Our Big Research Question {background-image="Images/background-data_blue_v4.png" .center}

<br>

```{r, fig.width = 10, fig.height=2}
## Manual DAG
d1 <- tibble(
  x = c(-3, 3),
  y = c(1, 1),
  labels = c("TBD", "Press Freedom")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = -1.9, xend = 1.85, y = 1, yend = 1, arrow = arrow())
```

::: notes
Research question brainstorming from Canvas forum

- ...

<br>

Your Brainstorming

- ...

<br>

**SLIDE**: Before we work to answer this research question, we need to deepen our expertise in the new data project!

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Report 2**

**Analyzing our Second Data Project**
:::

1. FINAL paper due Mar 16th

2. PDFs only

3. Support ALL claims with evidence
    - APA formatted in-text citations
    - APA formatted bibliography
    
4. Report should be concise and focused on your arguments
    
::: notes

Same basic requirements as the first report

- HOWEVER, no draft this time, just the real deal!

- We've already done this once so this should be much easier the second time

<br>

**SLIDE**: Big picture aim

:::


## Report 2 {background-image="Images/background-data_blue_v4.png" .center}

<br>

**What do we learn about the world from analyzing the XX produced by YY?**

<br>

The reader of your report is a non-expert in data analysis and is not familiar with this particular project, but is very interested in the topic. 

::: notes

**Any questions on the big picture framing of your report?**

<br>

**SLIDE**: Report components

:::



## Report 2 {background-image="Images/background-data_blue_v4.png" .center}

<br>

**What do we learn about the world from analyzing the XX produced by YY?**

1. The importance of the project in the real-world,
2. The key contributors of uncertainty in the project's data,
3. What we learn about the current world from analyzing the most recently available data, and
4. What we learn about the trajectory of XX in the world from analyzing the data across time

::: notes

In terms of structure and expectations it mirrors the work you did for the first report

- **Any questions on the report?**

<br>

Our plan this week

- Today we map out Premise 2 by analyzing the codebook 

- Wednesday we work on Premise 1, audit the dataset and brainstorm the analyses to consider

- Friday we work to build those analyses

<br>

**SLIDE**: Let's get to work on Premise 2

:::



## {background-image="Images/background-data_blue_v4.png" .center}

**What do we learn about the world from analyzing the XX produced by YY?**

<br>

**Premise 2: The key contributors of uncertainty?**

- **Source:** Where does the raw data come from?

- **Operationalization:** Defining the concepts

- **Instrumentation:** Designing the tool

- **Measurement Process:** Using the tool

- **Validation:** Checking the data

::: notes

**Everybody clear on these ways to analyze the quality of a measurement?**

- **Need any clarifications before we get to work?**

<br>

**OPTION 1: Overall Analyses**

**YOU PUT ON BOARD distributed around the room the five categories (Source, Operationalization, Instrument, Measurement, Validation) AND a header for Pro vs Con below each one*

- *Split class into five groups and assign to each category (to start!)*

- Groups, you have FIVE minutes to add ONE pro and ONE con under your category. Go!

- *Rotate groups through each other category: They must add one pro and one con to each category as they go (gets WAY harder as they go)*

- *PRESENT and DISCUSS each*

<br>

**OPTION 2: Per Indicator Analyses**

*ON BOARD: X headers (one per indicator in the dataset) with space below each for two lists: Pro vs Con*

- *Split class into groups per indicator and have them build pro and con lists for theirs*

- *PRESENT and DISCUSS each*

<br>

**SLIDE**: Next class

:::



## For Next Class {background-image="Images/background-data_blue_v4.png" .center}

<br>

::: {.r-fit-text}

1. Submit evidence for Premise 1

2. Audit the new dataset(s)

:::

::: notes

Canvas Assignment

1. Submit to Canvas one piece of evidence we could use to make the argument that this data project is important (Premise 1 of the paper).

    - Provide an APA citation, and 
    
    - A brief explanation
    
    - No overlapping evidence!

2. Audit the Dataset

    - Make sure you understand how the second dataset is organized (observations and variables), and 
    
    - Spend some time auditing it for missing data
    
    - e.g. do we have any concerns with data missing in each variable, in any particular regions of the world, or in any particular years?

<br>

**Questions on the assignment?**

:::

