---
format: revealjs
---

## Today's Agenda {background-image="Images/background-data_blue_v4.png" .center}

```{r}
#  background-size="1920px 1080px"
library(tidyverse)
```

<br>

::: {.r-fit-text}

**Critically analyze the second data project's codebook**

- The Yale Environmental Performance Index (EPI)

:::

<br>

<br>

::: r-stack
Justin Leinaweaver (Spring 2025)
:::

::: notes
Prep for Class

1. Review Canvas submissions

2. Update slides BEFORE class
    - Insert research question agreed upon last class and update DAG
    - Add measures to evaluate

3. Bring markers for every group (x6 in SP24)

4. Make sure the codebook and dataset are on Canvas

<br>

**SLIDE**: Our developing research project

:::




## Our Research Question {background-image="Images/background-data_blue_v4.png" .center}

<br>

**Does a free press promote better environmental quality around the world?**

```{r, fig.align='center', fig.width = 10, fig.height=2}
## Manual DAG
d1 <- tibble(
  x = c(-3, 3),
  y = c(1, 1),
  labels = c("Press\nFreedom", "Environmental\nQuality")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = -1.9, xend = 1.85, y = 1, yend = 1, arrow = arrow())
```

::: notes

**Is everybody happy with this research question?**

- **Want to propose any changes?**

<br>

**SLIDE**: Before we work to answer this research question, we need to deepen our expertise in the new data project!

<br>

Other ideas you brainstormed

- Does government transparency and accountability translate into better environmental quality?

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Report 2**

**Analyzing our Second Data Project**
:::

1. FINAL paper due Mar 16th

2. PDFs only

3. Support ALL claims with evidence
    - APA formatted in-text citations
    - APA formatted bibliography
    
4. Report should be concise and focused on your arguments
    
::: notes

Same basic requirements as the first report

- HOWEVER, no draft this time, just the real deal!

- We've already done this once so this should be much easier the second time

<br>

**SLIDE**: Big picture aim

:::


## Report 2 {background-image="Images/background-data_blue_v4.png" .center}

<br>

**What do we learn about the world from analyzing the Environmental Performance Index (EPI) produced by the Yale Center for Environmental Law & Policy?**

<br>

The reader of your report is a non-expert in data analysis and is not familiar with this particular project, but is very interested in the topic. 

::: notes

**Any questions on the big picture framing of your report?**

- The reader is a busy policy principal so you have to be concise!

<br>

**SLIDE**: Report components

:::



## Report 2 {background-image="Images/background-data_blue_v4.png" .center}

<br>

**What do we learn about the world from analyzing the Environmental Performance Index (EPI) produced by the Yale Center for Environmental Law & Policy?**

1. The importance of the project in the real-world,
2. The key contributors of uncertainty in the project's data,
3. What we learn about the current world from analyzing the most recently available data, and
4. What we learn about the trajectory of environmental quality in the world from analyzing the data across time

::: notes

In terms of structure and expectations it mirrors the work you did for the first report

- **Any questions on the report?**

<br>

Our plan this week

- Today we map out Premise 2 by analyzing the codebook 

- Wednesday we work on Premise 1, audit the dataset and brainstorm the analyses to consider

- Friday we work to build those analyses

<br>

**SLIDE**: Let's get to work on Premise 2

:::



## {background-image="Images/background-data_blue_v4.png" .center}

**What do we learn about the world from analyzing the Environmental Performance Index (EPI) produced by the Yale Center for Environmental Law & Policy?**

<br>

**Premise 2: The key contributors of uncertainty?**

- **Source:** Where does the raw data come from?

- **Operationalization:** Defining the concepts

- **Instrumentation:** Designing the tool

- **Measurement Process:** Using the tool

- **Validation:** Checking the data

::: notes

**Everybody clear on these ways to analyze the quality of a measurement?**

- **Need any clarifications before we get to work?**

<br>

*Split class into small groups (2-3 max)*

- PAIRS/GROUPS, you have TEN minutes to build a list of STRENGTHS ONLY

- What are the biggest methodological reasons you believe someone should use this dataset?

<br>

*ON BOARD: Report back*

<br>

- PAIRS/GROUPS, now you know the baseline stuff, back to the strengths!

- Five more minutes to deepen this list! I want a STRONG argument that represents the best of this project!

<br>

- *ON BOARD*: Report back - build list

<br>

**SLIDE**: Plan of attack for this section

:::



## {background-image="Images/background-data_blue_v4.png" .center}

**What do we learn about the world from analyzing the Environmental Performance Index (EPI) produced by the Yale Center for Environmental Law & Policy?**

<br>

**Premise 2: Proposed Paragraph Structure**

- Big idea: The first reason we can use this dataset to learn about the world is...

- Flesh it out: Explain the idea and provide evidence

- Critical analysis: Add a note of caution (e.g. a weakness)

::: notes

<br>

**SLIDE**: Next class

<br>

**Other Options to Consider in Future**

**OPTION 2: Round-robin**

- **YOU PUT ON BOARD distributed around the room the five categories (Source, Operationalization, Instrument, Measurement, Validation) AND a header for Pro vs Con below each one*

- *Split class into five groups and assign to each category (to start!)*

- Groups, you have FIVE minutes to add ONE pro and ONE con under your category. Go!

- *Rotate groups through each other category: They must add one pro and one con to each category as they go (gets WAY harder as they go)*

- *PRESENT and DISCUSS each*

<br>

**OPTION 3: Per Indicator Analyses**

- *ON BOARD: X headers (one per indicator in the dataset) with space below each for two lists: Pro vs Con*

- *Split class into groups per indicator and have them build pro and con lists for theirs*

- *PRESENT and DISCUSS each*

:::



## For Next Class {background-image="Images/background-data_blue_v4.png" .center}

<br>

::: {.r-fit-text}

1. Submit evidence for Premise 1

2. Audit the new dataset(s)

:::

::: notes

Canvas Assignment

1. Submit to Canvas one piece of evidence we could use to make the argument that this data project is important (Premise 1 of the paper).

    - Provide an APA citation, and 
    
    - A brief explanation
    
    - No overlapping evidence!

2. Audit the Dataset

    - Make sure you understand how the second dataset is organized (observations and variables), and 
    
    - Spend some time auditing it for missing data
    
    - e.g. do we have any concerns with data missing in each variable, in any particular regions of the world, or in any particular years?

<br>

**Questions on the assignment?**

:::

