---
format: revealjs
---

# Today's Agenda {background-image="Images/background-data_blue_v4.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)

load("../../DATA/2026-Spring/Miller_Saunders_Farhart2016/Miller_Saunders_Farhart2016_v2.RData")

# Build the index
d$ct_index <- as.numeric(d$CT_Obama) + as.numeric(d$CT_DeathPanels) + as.numeric(d$CT_Govt_911) + as.numeric(d$CT_Katrina_Levees) + as.numeric(d$CT_Climate_Hoax) + as.numeric(d$CT_Bush_WMDs) + as.numeric(d$CT_Hussein_911) + as.numeric(d$CT_2004_election)

# Rescale the index to start at zero
d$ct_index0 <- d$ct_index - 8

# Convert to percentage out of 24
d$ct_index_pct <- d$ct_index0 / 24
```

<br>

::: {.r-fit-text}

**Bivariate Analyses**

- Visualizing the relationship between two variables

- NEW R Script: Bivariate_Visualizations.R

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2026)
:::

::: notes
Prep for Class

1. Check Canvas submissions

<br>

Last week we began performing bivariate analyses with cross-tabs and descriptive statistics by group

- **SLIDE**: Let's start by reviewing your assignment results for today

:::



## Miller, Saunders & Farhart (2016) {background-image="Images/background-data_blue_v4.png" .center}

**H1. Belief in liberal CTs increases as trust in government falls**

<br>

```{r, echo=FALSE}
d |>
  group_by(Trust_Govt) |>
  summarize(
    Mean = mean(Lib_Conspiracy_Index, na.rm = T),
    #StdDev = sd(Lib_Conspiracy_Index, na.rm = T),
    "1st quartile" = quantile(Lib_Conspiracy_Index, probs = .25, na.rm = T),
    Median = median(Lib_Conspiracy_Index, na.rm = T),
    "3rd quartile" = quantile(Lib_Conspiracy_Index, probs = .75, na.rm = T)
  ) |>
  filter(!is.na(Trust_Govt)) |>
  kable(digits = 2) |>
  kable_styling(font_size = 30)
```

<br>

```{r, echo=TRUE, eval=FALSE}
# Interval x Ordinal - Descriptive Stats by Group
aggregate(data = d, Lib_Conspiracy_Index ~ Trust_Govt, summary)
```

::: notes

The results using the indicies are WAY easier to read than the cross-tabs for individual ordinal questions, right?

<br>

Ideally, the more variation in your outcome variable the more tools at your disposal

- Never willingly throw away variation!

<br>

**So, does the data support the first hypothesis?**

:::



## Miller, Saunders & Farhart (2016) {background-image="Images/background-data_blue_v4.png" .center}

**H2. Belief in conservative CTs increases as trust in government falls**

<br>

```{r, echo=FALSE}
d |>
  group_by(Trust_Govt) |>
  summarize(
    Mean = mean(Cons_Conspiracy_Index, na.rm = T),
    "1st quartile" = quantile(Cons_Conspiracy_Index, probs = .25, na.rm = T),
    Median = median(Cons_Conspiracy_Index, na.rm = T),
    "3rd quartile" = quantile(Cons_Conspiracy_Index, probs = .75, na.rm = T)
  ) |>
  filter(!is.na(Trust_Govt)) |>
  kable(digits = 2) |>
  kable_styling(font_size = 30)
```

<br>

```{r, echo=TRUE, eval=FALSE}
# Interval x Ordinal - Descriptive Stats by Group
aggregate(data = d, Cons_Conspiracy_Index ~ Trust_Govt, summary)
```

::: notes

**Does the data support the second hypothesis?**

- (Not really... Very slight U-shape here...

:::



## Miller, Saunders & Farhart (2016) {background-image="Images/background-data_blue_v4.png" .center}

**H3. Democrats are more likely to believe liberal coded conspiracies**

<br>

```{r, echo=FALSE}
d |>
  group_by(Party3) |>
  summarize(
    Mean = mean(Lib_Conspiracy_Index, na.rm = T),
    "1st quartile" = quantile(Lib_Conspiracy_Index, probs = .25, na.rm = T),
    Median = median(Lib_Conspiracy_Index, na.rm = T),
    "3rd quartile" = quantile(Lib_Conspiracy_Index, probs = .75, na.rm = T)
  ) |>
  filter(!is.na(Party3)) |>
  kable(digits = 2) |>
  kable_styling(font_size = 30)
```

<br>

```{r, echo=TRUE, eval=FALSE}
# Interval x Ordinal - Descriptive Stats by Group
aggregate(data = d, Lib_Conspiracy_Index ~ Party3, summary)
```

::: notes

**Does the data support the third hypothesis?**

:::



## Miller, Saunders & Farhart (2016) {background-image="Images/background-data_blue_v4.png" .center}

**H4. Republicans are more likely to believe conservative coded conspiracies**

<br>

```{r, echo=FALSE}
d |>
  group_by(Party3) |>
  summarize(
    Mean = mean(Cons_Conspiracy_Index, na.rm = T),
    "1st quartile" = quantile(Cons_Conspiracy_Index, probs = .25, na.rm = T),
    Median = median(Cons_Conspiracy_Index, na.rm = T),
    "3rd quartile" = quantile(Cons_Conspiracy_Index, probs = .75, na.rm = T)
  ) |>
  filter(!is.na(Party3)) |>
  kable(digits = 2) |>
  kable_styling(font_size = 30)
```

<br>

```{r, echo=TRUE, eval=FALSE}
# Interval x Ordinal - Descriptive Stats by Group
aggregate(data = d, Cons_Conspiracy_Index ~ Party3, summary)
```

::: notes

**Does the data support the final hypothesis?**

<br>

**Ok, how are we doing with bivariate analysis using cross-tabs and descriptive stats by group?**

<br>

This week we extend our bivariate analyses visualizations

- **SLIDE**: In ALMOST all case, visualizations are preferred to tables and group means

:::



## {background-image="Images/background-data_blue_v4.png"}

::: {.r-fit-text}
**Bivariate Analyses: Visualization**
:::

<br>

Add the following to your notes:

```{r, echo=TRUE}
# Bivariate Visualizations

# Categorical x Categorical
# Bar Plots: Side-by-side, Stacked bars, Proportions

# Interval x Categorical
# Facet wraps, Box plots

# Interval x Interval
# Scatter plots

# Load the tidyverse for practice data
library(tidyverse)
```

::: notes

Add to your notes!

<br>

**SLIDE**: Like last week we'll learn the tools using built-in data

:::



## {background-image="Images/background-data_blue_v4.png" .center}

### **What explains why some diamonds are more expensive than others?**

<br>

```{r, echo=TRUE}
# Create an ordinal version of prices by quartile
diamonds$price2 <- cut(diamonds$price,
    breaks = c(0, 950, 2401, 5324, 19000),
    labels = c("25th", "50th", "75th", "100th"))

# Check the new variable
table(diamonds$price2)
```

::: notes

To use the diamonds data to learn our bivariate tools we'll need a version of the outcome, price, in ordinal form

- Same code as last week

<br>

**Everybody got this?**

<br>

**SLIDE**: Let's start our bivariate analyses with two ordinal variables

:::



## Categorical x Categorical {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=8, fig.asp=.618, cache=FALSE}
ggplot(diamonds, aes(x = cut)) +
  geom_bar()
```

::: notes

We'll begin our bivariate analysis with a univariate step

- Our interest is in seeing if the types of diamond cuts explain variations in diamond prices

- For this we want to begin by putting the predictor, or independent variable, on the x-axis

<br>

I will now step you through THREE different types of bar plot for visualizing this relationship

- **SLIDE**: The first type of bivariate bar plot is what we call a side-by-side plot

:::






## Categorical x Categorical {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=8, fig.asp=.618, cache=FALSE}
# Side-by-side bar plots
ggplot(diamonds, aes(x = cut, fill = price2)) +
  geom_bar(position = "dodge")
```

::: notes

To get to the side-by-side bar plot we add two things to the code

- We "fill" in the outcome variable which defines the legend, and

- We add a position argument to tell ggplot what kind of bar plot to make

<br>

In terms of representing a relationship remember:

- The independent variables goes on the x axis

- The dependent variable is the legend

- This way we compare the price distribution across each level of the independent variable

<br>

Let's now clean this visualization up

- **SLIDE**: First needed fix: the order of values in the legend

:::



## Categorical x Categorical {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=8, fig.asp=.618, `code-line-numbers`="1,4"}
# Side-by-side bar plots
ggplot(diamonds, aes(x = cut, fill = price2)) +
  geom_bar(position = "dodge") +
  scale_fill_discrete(limits = c("100th", "75th", "50th", "25th"))
```

::: notes

Call me crazy, but number in a plot should go in descending order

- Our brains are not wired to think higher is lower

<br>

Moving forward I am using code highlighting to draw you to the only thing that changed from the last run of the code

- SO, the code is identical to the last plot you made but now we've added one more function

- You will need to add the "+" on line 3 to let ggplot know there is a new function to add

<br>

**Everybody got this?**

<br>

**SLIDE**: One more polish: Clean up the axis labels

:::



## Categorical x Categorical {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=8, fig.asp=.618, `code-line-numbers`="1,5"}
# Side-by-side bar plots
ggplot(diamonds, aes(x = cut, fill = price2)) +
  geom_bar(position = "dodge") +
  scale_fill_discrete(limits = c("100th", "75th", "50th", "25th")) +
  labs(x = "Diamond Cuts", y = "Count", fill = "Prices")
```

::: notes

NOTE: I've added a labs function to show you how to change the label on the legend

- "fill" is the label on the legend

<br>

Ok, this is a side-by-side bar plot

- **What do we learn from this?**

- **Is the cut of a diamond a good predictor of the price of that diamond?**

- (1. Staggering differences in the numbers in each category; Ideal cut appears to be the modal value!)

- (2. Premium cut is tilted to the highest prices and ideal cut the opposite!)

<br>

**Questions on making a side-by-side bar plot two categorical variables?**

<br>

**SLIDE**: It can be easier to interpret your bivariate bar plots if you stack the bars

:::



## Categorical x Categorical {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=8, fig.asp=.618, `code-line-numbers`="1,3"}
# Stacked bar plots
ggplot(diamonds, aes(x = cut, fill = price2)) +
  geom_bar(position = "stack") +
  scale_fill_discrete(limits = c("100th", "75th", "50th", "25th")) +
  labs(x = "Diamond Cuts", y = "Count", fill = "Prices")
```

::: notes

Once again, just copy the exact code chunk you just wrote to a new space

- Re-label it as we are now making a stacked bar plot

- Then modify this one line of code

- All you have to do is change the position argument from "dodge" to "stack"

- **Everybody got this?**

<br>

**SLIDE**: Before we interpret it, let's fix the bars to match the order of the legend

:::



## Categorical x Categorical {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=8, fig.asp=.618, `code-line-numbers`="1,3"}
# Stacked bar plots
ggplot(diamonds, aes(x = cut, fill = price2)) +
  geom_bar(position = position_stack(reverse = TRUE)) +
  scale_fill_discrete(limits = c("100th", "75th", "50th", "25th")) +
  labs(x = "Diamond Cuts", y = "Count", fill = "Prices")
```

::: notes

Ok, one more tweak in the position argument to reverse the order of the levels in each bar

- **Everybody get this?**

<br>

This is a stacked bar plot

- It simply takes the side-by-side bars and puts them on top of each other

<br>

**Does everybody see how this helps us compare the distributions of prices across each color category?**

- This makes it SO clear that most diamonds are ideal cut

- And the counts go up with quality!

- e.g. diamonds are a rigged market!

<br>

**SLIDE**: Given the uneven bars, let's consider one final approach to making a bivariate bar plot

:::



## Categorical x Categorical {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=6, fig.asp=.618, `code-line-numbers`="1,3,6"}
# Proportionate bar plot
ggplot(diamonds, aes(x = cut, fill = price2)) +
  geom_bar(position = "fill") +
  scale_fill_discrete(limits = c("100th", "75th", "50th", "25th")) +
  labs(x = "Diamond Cuts", y = "Count", fill = "Prices") +
  scale_y_continuous(labels = scales::percent)
```

::: notes

Last style of common bar plot is a proportional stacked bar plot.

- This only requires a change to the position argument, BUT

- You will definitely want to change the axis labels to %s so let's say it takes two changes to the code

<br>

Proportional stacked bar plots are a very cool approach

- They transform each stacked bar to a proportion of 100% per x variable level.

- So, the y-axis marks off the proportion of the total represented by each level of the fill variable.

<br>

**SLIDE**: Annoyingly, we need one more piece of code to fix the order within each bar

:::




## Categorical x Categorical {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=6, fig.asp=.618, `code-line-numbers`="1,2,4"}
# Reverse the order of the levels
diamonds$price3 <- fct_rev(diamonds$price2)

ggplot(diamonds, aes(x = cut, fill = price3)) +
  geom_bar(position = "fill") +
  scale_fill_discrete(limits = c("100th", "75th", "50th", "25th")) +
  labs(x = "Diamond Cuts", y = "Count", fill = "Prices") +
  scale_y_continuous(labels = scales::percent)
```

::: notes

R already sees our ordinal price variable as having levels, but the bar plot needs us to reverse them

- Hence, we create a NEW version of the price variable but with the levels backwards

- **Did everybody get this?**

<br>

I want to be clear, stacked bar plots are useful and VERY frequently used

- BUT they are VERY dangerous

<br>

In our data, there are over 21k ideal cut diamonds and 1k fair cut diamonds

- Small changes in the value of fair cut diamonds will have a HUGE impact on the %s

- e.g. In a room of 10 people, losing 2 is a 20% decrease

- WHEREAS small changes in the values of ideal cut diamonds will have basically no effect on the %s

- **Does this danger make sense?**

<br>

Bottom line, be VERY careful with comparing proportions across groups

- If the group sizes are equal, go nuts

- If not, beware

- THIS is why we use quartiles so often, they create equal sized groups for comparisons!

<br>

Ok, that covers bivariate visualizations for the relationship between two categorical variables

- **Any questions on our three versions of the bar plot?**

<br>

**SLIDE**: Next up!

:::



## Interval x Categorical {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=6, fig.asp=.75}
ggplot(diamonds, aes(x = price)) +
  geom_histogram(bins = 20, color = "white")
```

::: notes

Next visualization challenge!

- We have an outcome variable that is interval (diamond prices), and

- We have a predictor variable that is ordinal (diamond cuts)

<br>

Last class we calculated summary statistics of the interval variable for each level of the categorical variable using the aggregate function

- Today I'll show you TWO options for visualizing this relationship

<br>

**SLIDE**: Our first option is simply to make separate histograms for each type of diamond cut

:::



## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=6, fig.asp=.75, `code-line-numbers`="1,4"}
# Interval x Categorical Option 1. Facet Wrap
ggplot(diamonds, aes(x = price)) +
  geom_histogram(bins = 20, color = "white") +
  facet_wrap(~ cut)
```

::: notes

Ok, your first option is to use the facet_wrap function to extend ANY univariate visualization into the levels of the categorical variable

- Use a tilde and then the categorical variable

- **Did everybody get this code to run?**

<br>

**SLIDE**: When using a facet wrap you should choose a layout that makes comparisons easier for the viewer

:::




## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=6, fig.asp=1.5, `code-line-numbers`="1,4"}
# Interval x Categorical Option 1. Facet Wrap
ggplot(diamonds, aes(x = price)) +
  geom_histogram(bins = 20, color = "white") +
  facet_wrap(~ cut, ncol = 1)
```

::: notes

It is simply easier to compare distributions of this shape when they are stacked on top of each other

- Other types of plot will lead you to different layouts

<br>

**SLIDE**: One more important tweak when using facet wraps

:::



## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=6, fig.asp=1.5, `code-line-numbers`="1,4"}
# Interval x Categorical Option 1. Facet Wrap
ggplot(diamonds, aes(x = price)) +
  geom_histogram(bins = 20, color = "white") +
  facet_wrap(~ cut, ncol = 1, scale = "free_y")
```

::: notes

Facet wrap forces all of the plots to be on the same x and y axis range

- Clearly, this makes the "fair" distribution impossible to view

<br>

scale = "free_y" tells the facet wrap to let the y-axis ranges vary in each level of the categorical variable

- This way we can view the distribution of "Fair" on its own terms

- This helps us compare the distributions across the levels

<br>

**Any questions on the facet wrap approach?**

<br>

**SLIDE**: Our second option for interval x categorical is to go back to our box plots!

:::



## Interval x Categorical {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=8, fig.asp=.618}
ggplot(diamonds, aes(y = price)) +
  geom_boxplot()
```

::: notes

Alright, here's our old friend the box plot

- You mocked it for being "terrible" (thanks Sara)

- NOTE: We are putting the outcome variable on the y axis

    - Outcomes (DVs) ALWAYS go on the y-axis!

<br>

**SLIDE**: Now let's unlock the power!

:::



## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=8, fig.asp=.618, `code-line-numbers`="1,2"}
# Interval x Categorical Variable: 2. Box plots
ggplot(diamonds, aes(x = cut, y = price)) +
  geom_boxplot()
```

::: notes

This code is basically the same as what we've done before but with more than one variable in the aes function

- **Everybody got this code working?**

<br>

Box plots are an excellent way to compare IQRs across the levels of a categorical variable

- Rather than comparing group means which can be skewed by outliers

- Or medians which tell us nothing about the spread of the data

- The boxes give us a wide version of the "middle" we can compare!

<br>

**What do we learn about diamond prices from comparing the middles of each type of cut?**

- The diamond market is a scam!!!

<br>

**Ok, any questions on the facet wraps or box plots for examining an interval variable across levels of a categorical variable?**

<br>

**SLIDE**: Last bivariate relationship

:::



## Interval x Interval {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, fig.align='center', fig.width=6, fig.asp=.75, cache=FALSE}
# Scatter Plot
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point()
```

::: notes

Our final new tool: Scatter Plots

- For visualizing the relationship between two interval variables

- Your outcome variable ALWAYS goes on the y-axis, predictor ALWAYS on the x-axis

<br>

We read the plot from left to right

- As you increase the size of the diamonds what happens to their prices?

<br>

**What do we learn from this?**

- (The only one of the four "C's" that actually matters)

- Color, cut, clarity all matter on the margins, but diamond prices are driven by their size!

<br>

**SLIDE**: Wrapping up our new code

:::



## {background-image="Images/background-data_blue_v4.png"}

::: {.r-fit-text}
**Bivariate Analyses: Visualization**
:::

<br>

**Categorical x Categorical**

- Bar Plots: Side-by-side, Stacked bars, Proportions

**Interval x Categorical**

- Facet wraps, Box plots

**Interval x Interval**

- Scatter plots

::: notes

**Any questions on the tools we learned today?**

<br>

**SLIDE**: Practice for next class

:::



## For Next Class {background-image="Images/background-data_blue_v4.png" .center}

<br>

::: {.r-fit-text}

**Practicing Bivariate Visualizations Using ANES Data**

- Analyzing political engagement across levels of:

    1. Social class
    
    2. Party identification

:::
::: notes

Everybody take a look at the details on Canvas

<br>

**Any questions on the assignment?**

<br>

Canvas ASSIGNMENT: Practicing Bivariate Visualizations Using ANES Data

For this assignment we will practice the bivariate visualizations tools while analyzing the political engagement of respondents to the 2020 ANES survey.

1. Does self reported social class (social.class) predict the likelihood a respondent donates money to political candidates (polact.givecand)?
    - Visualize this relationship with a side-by-side, a stacked and a proportionate bar plot. THEN submit all three AND make an argument that ONE of these is the "best" one to describe the relationship between these two variables.

2. Does party identification (partyid3) predict the likelihood a respondent is highly engaged in American politics (engagement_index)? (Int x Cat)
    - Use the nine point political engagement index we built and explored last week as your outcome variable
    - Visualize this relationship with a box plot and a histogram with a facet wrap. THEN submit both AND make an argument that ONE of these is the "best" one to describe the relationship between these two variables.

:::

