---
format: revealjs
---

# Today's Agenda {background-image="libs/Images/background-data_blue_v3.png"}

```{r}
#  background-size="1920px 1080px"
library(tidyverse)
library(haven)

d <- read_csv("../Data_in_Class-SP24/CoW-National_Material_Capabailities/CoW-NMC-Data.csv", na = "-9")
```

<br>

Working on Report 2

- Importance Arguments (Section 1)

- Auditing the Data

<br>

::: r-stack
Justin Leinaweaver (Spring 2024)
:::

::: notes
Prep for Class

1. Check the Canvas submissions

2. Update slides BEFORE class
    - Add data view for discussing observations
    - Add code checking for missing data by variable, region and year
    - Is it time to introduce rworldmap for checking regional missing data

<br>

For today I asked each of you to come to class with material for Section 1 and having audited the dataset.

<br>

**Everybody ready to go with these?**

<br>

**SLIDE**: Let's start with the Section 1 evidence
:::



## Section 1 {background-image="libs/Images/background-slate_v2.png" .center}

**Why is this project important?**

<br>

- What problem are the researchers trying to solve?

- How does the problem impact actors, institutions or interactions in the real world?

- Why should we take these researchers seriously? (e.g. credentials, funding, transparency, etc.) 
 
- Is there evidence that the project has had concrete impacts on the world?

::: notes
Here is the general advice I offered you in week 3 for the kinds of elements that help construct an argument about importance.

- My hope is the evidence you found for today will help everyone in class make this a compelling and well supported argument!

<br>

Let's go around the room so you can each briefly introduce the evidence you found and how you think it helps us make this argument.

- *PRESENT and DISCUSS each*

<br>

An excellent jump-start on Section 1.

- **Any questions before we jump to auditing the data?**
:::



## Section 2: Auditing the Data  {background-image="libs/Images/background-slate_v2.png" .smaller .center}

```{r}
d |>
  select(stateabb, year:cinc) |>
  slice_sample(n=10) |>
  kableExtra::kbl(digits = 2)
```

::: notes
Let's go back to our work on Section 2 of the report

- The codebook gives us vital information about how the researchers actually measure the world

<br>

BUT, that's only ONE source of uncertainty

- Data sources and weaknesses in the measurement instrument can often only be spotted once you begin to explore the data itself

<br>

Everybody import the data and get ready to explore!

- Don't forget the missing data code is -9 for the CoW NMC data project
:::



## Audit: Missing Data (Indicators) {background-image="libs/Images/background-slate_v2.png"}

<br>

::: {.fragment}
```{r}
d |>
  pivot_longer(cols = milex:upop, names_to = "Indicators", values_to = "Values") |>
  group_by(Indicators) |>
  reframe(
    NAs = sum(is.na(Values))
  ) |>
  mutate(
    Proportion = str_c(round(NAs/15951, 3)*100, "%")
  ) |>
  kableExtra::kbl(align = c("l", "c", "c"))
```
:::

::: notes
Let's start by looking for missing data by variable

- Everybody use the summary function to check for NAs in each of the six indicators

<br>

**SLIDE: What do we learn from these results?**
:::



## Audit: Missing Data (Years) {background-image="libs/Images/background-slate_v2.png"}

```{r, fig.asp=.618, out.width='95%', fig.align='center', fig.width=8, cache=TRUE}
d |>
  pivot_longer(cols = milex:upop, names_to = "Indicators", values_to = "Values") |>
  group_by(year) |>
  reframe(
    NAs = sum(is.na(Values))
  ) |>
  ggplot(aes(x = year, y = NAs)) +
  geom_line() +
  theme_bw() +
  labs(x = "", y = "Missing Data (count)") +
  scale_x_continuous(breaks = seq(1820, 2020, 20))
```

::: notes
Here we see missing data by year

- Estimating missing data overall can be a little tricky so I'm just showing you the results

<br>

The spikes coincide with WWI, WWII, the collapse of the USSR and the year 2014

- I get the first three of those, but am not sure about the 2014 spike

<br>

**SLIDE: What do we learn from these results?**

<br>

**SLIDE**: Let's split this out by variable
:::



## Audit: Missing Data (Years) {background-image="libs/Images/background-slate_v2.png"}

```{r, fig.asp=.618, out.width='95%', fig.align='center', fig.width=8, cache=TRUE}
d |>
  pivot_longer(cols = milex:upop, names_to = "Indicators", values_to = "Values") |>
  group_by(year, Indicators) |>
  reframe(
    NAs = sum(is.na(Values))
  ) |>
  ggplot(aes(x = year, y = NAs, color = Indicators)) +
  geom_line() +
  theme_bw() +
  labs(x = "", y = "Missing Data (count)") +
  scale_color_brewer(type = "qual", palette = 3) +
  scale_x_continuous(breaks = seq(1820, 2020, 20))
```

::: notes
**What do we learn from splitting out the data?**

<br>

It makes a ton of sense to me that military spending would be the hardest data to acquire!

<br>

Clearly, the sample size changes across the years so let's now examine by how much.

- Everybody make us a bar plot of countries across time (e.g. count the years!)

<br>

**SLIDE**: Results
:::



## Audit: How does the sample change? {background-image="libs/Images/background-slate_v2.png"}

```{r, fig.asp=.618, out.width='95%', fig.align='center', fig.width=8, cache=TRUE}
ggplot(d, aes(x = year)) +
  geom_bar() +
  theme_bw() +
  labs(x = "", y = "Number of Countries") +
  scale_x_continuous(breaks = seq(1820, 2020, 20))
```

::: notes
**What do we learn from this visualization?**

<br>

If we are going to understand "proportion as a whole" to be meaningful, then we better be confident the sample has the whole!

<br>

**So, what does this count represent?**

- **Does the codebook say this is a count of ALL countries or just those relevant to power?**

<br>

**SLIDE**: Last audit step for us today, let's try to verify the calculation of the `cinc` scores
:::



## Audit 2: Verifying CINC scores  {background-image="libs/Images/background-slate_v2.png" .smaller}

<br>

::: {.fragment}
**1. Filter the Data**
```{r, echo=TRUE}
d1910 <- filter(d, year == 1910) 
```

<br>

```{r, echo=TRUE}
# View the top of the dataset
d1910
```
:::

::: notes
Remember, the CINC score for each country-year is an average of the proportions in the six indicator measures

- That's dense, so let's unpack it as an audit check on the data

- In other words, can we replicate their index? If not that should cause us concerns!

<br>

As practice, let's verify the US cinc score for the year 1910

- I'm picking a spot four years before the outbreak of WW1.

<br>

Let's attack this with our programmer hats on.

- That simply means breaking a complex question down into simpler steps

<br>

**SLIDE**: I've input the NMC dataset as 'd' and my first step is to extract only the observations from 1910

<br>

**What's our next step in this process?**

- (**SLIDE**: We need to calculate the sum of each variable in this year)
:::



## Audit 2: Verifying CINC scores  {background-image="libs/Images/background-slate_v2.png"}

<br>

**2. Sum together all observations for each variable in one year**

```{r, echo=TRUE}
sum(d1910$milex, na.rm = TRUE)
sum(d1910$milper, na.rm = TRUE)
sum(d1910$irst, na.rm = TRUE)
sum(d1910$pec, na.rm = TRUE)
sum(d1910$tpop, na.rm = TRUE)
sum(d1910$upop, na.rm = TRUE)
```

::: notes
Note here that you have to include the argument "na.rm = TRUE" so that R knows what to do with missing data.

- Otherwise, adding together all values in a column with missing data results in an "NA."

<br>

This is a good thing!

- We don't want to produce results that we don't understand

- If there is missing data and you CHOOSE to ignore it, you have to make that choice EXPLICITLY in the code.

<br>

**Make sense?**
:::



## Audit 2: Verifying CINC scores  {background-image="libs/Images/background-slate_v2.png"}

<br>

**3. Identify the US amounts by the totals**

```{r, echo=TRUE}
d1910_USA <- filter(d1910, stateabb == "USA")
```

<br>

```{r, echo=TRUE}
d1910_USA
```

::: notes
**Everybody understand this step and how we got these results?**
:::



## Audit 2: Verifying CINC scores  {background-image="libs/Images/background-slate_v2.png"}

<br>

**4. Calculate the US proportions**

```{r, echo=TRUE}
d1910_USA$milex / sum(d1910$milex, na.rm = TRUE)

d1910_USA$milper / sum(d1910$milper, na.rm = TRUE)

d1910_USA$irst / sum(d1910$irst, na.rm = TRUE)

d1910_USA$pec / sum(d1910$pec, na.rm = TRUE)

d1910_USA$tpop / sum(d1910$tpop, na.rm = TRUE)

d1910_USA$upop / sum(d1910$upop, na.rm = TRUE)
```


::: notes
**Everybody understand this step and how we got these results?**

<br>

Everybody calculate the average of these six proportions

- Add together the six values and divide them by six

- Try it!
:::



## Audit 2: Verifying CINC scores  {background-image="libs/Images/background-slate_v2.png"}

<br>

**5. Calculate the CINC**

```{r, echo=TRUE}
(.133 + .0218 + .4527 + .4581 + .0778 + .2153) / 6
```

<br>

```{r, echo=TRUE}
d1910_USA$cinc
```

::: notes
**Everybody understand this step and how we got these results?**

<br>

**Alright, what have we learned about the cinc score?**

- **Are we more or less confident in the measure?**

<br>

**SLIDE**: Let's talk about Sections 3 and 4 for the second report
:::


## {background-image="libs/Images/background-slate_v2.png"}

**Section 3: Current Snapshot (2016)**

1. Analyze the current distribution of the `cinc` score

2. Analyze the current distribution of one of the indicators in the dataset (assigned)

::: {.fragment}
**Section 4: Across Time (since WWII)**

1. Analyze the `cinc` score **AND** your assigned indicator across time (1946, 1969, 1993, 2016)

2. Select a country that has experienced variation in its `cinc` score and your assigned indicator across time (e.g. avoid the stable cases).
:::

::: notes
Just like last time!

<br>

**Questions on the assignment?**

- Get to it!
:::




