---
fontsize: 14pt
classoption: "aspectratio=169"
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    df_print: kable
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "./08_2-Project2_Presentations.Rmd", output_file = "./08_2-Project2_Presentations.pdf")

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```


## Today's Agenda
\LARGE

\begin{center}
Planning our class research project
\end{center}

\large
\center

\vspace{.65in}

Justin Leinaweaver (Fall 2021)



## Assignments 1 and 2
\center
\Large

**What do we learn about the world by analyzing individual data projects?**

\vspace{.2in}

+ Dataset 1: Freedom in the World (FH)

\vspace{.1in}

+ Dataset 2: National Material Capabilities (CoW)



## Assignment 3: Research Project
\center
\Large

**What do we learn about the world by combining the data projects?**

\vspace{.2in}

+ Dataset 1: Freedom in the World (FH)

\vspace{.1in}

+ Dataset 2: National Material Capabilities (CoW)



## Assignment 3: Research Project
\Large

What we need:

1. A research question that can be answered with our data

    + \large \textit{Do states with greater levels of X have more Y?}

    + *What is the impact of increasing X on Y?*




## Assignment 3: Research Project
\Large

What we need:

1. A research question that can be answered with our data, and

2. A model (e.g. the story explaining the arrow)

```{r, echo = FALSE, fig.asp = .4, fig.width = 5, fig.align = 'center', out.width = '80%'}
tribble(
    ~Name, ~x, ~y,
    "Predictor", 1, 1,
    "Outcome", 3, 1
) %>%
    ggplot(aes(x = x, y = y)) +
    geom_point(size = .1) +
    geom_label(aes(label = Name), size = 6) +
    coord_cartesian(xlim = c(0.8, 3.3), ylim = c(.93, 1.04)) +
    ggthemes::theme_tufte() +
    annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
    scale_x_continuous(breaks = NULL) +
    scale_y_continuous(breaks = NULL) +
    labs(x = "", y = "")
```



## Group Presentations on Tuesday {.plain}
\large

Your presentation should:

+ Be 8 minutes in length (with Rmarkdown slides),

+ Propose a research question, 

+ Propose a model answering that question, and

+ Provide supporting UNIVARIATE analyses that show sufficient variation exists in the predictor and the outcome


## Custom YAML: Themes and Colors {.plain}
::: columns
:::: column

\vspace{.3in}

\-\-\-

title: "Untitled"

author: "Justin Leinaweaver"

date: "9/13/2021"

   output: 
   
      beamer_presentation:
	  theme: "CambridgeUS"
	  colortheme: "dolphin"

\-\-\-

::::
:::: column

\center
\includegraphics[keepaspectratio=true,width=.44\paperwidth]{Images/05_1-cambridge_dolphin.png}

::::
:::

\vspace{.15in}
\center
\large
https://deic-web.uab.cat/~iblanes/beamer_gallery/
