---
format: revealjs
---

## Today's Agenda {background-image="Images/background-data_blue_v4.png" .center}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)

load("../../DATA/2026-Spring/ANES/ANES2020_Data_Extract.RData")

anes$engagement_index <- 
  if_else(anes$polact.givecand == "1. Yes", 1, 0) +
  if_else(anes$polact.giveparty == "1. Yes", 1, 0) +
  if_else(anes$polact.giveoth == "1. Yes", 1, 0) +
  if_else(anes$polact.postsign == "1. Yes", 1, 0) +
  if_else(anes$polact.meetings == "1. Yes", 1, 0) +
  if_else(anes$polact.onlinemeet == "1. Yes", 1, 0) +
  if_else(anes$polact.othwork == "1. Yes", 1, 0) +
  if_else(anes$polact.talkvote == "1. Yes", 1, 0) +
  if_else(anes$polact.talkpol == "1. Yes", 1, 0)

anes$racial_negativity <- (as.numeric(anes$oth.race.concern) +
  as.numeric(anes$oth.race.feel) +
  as.numeric(anes$oth.race.persp) +
  as.numeric(anes$oth.race.protect) - 4) / 16
```

<br>

::: {.r-fit-text}

**Bivariate Analyses**

- Practice visualizing the relationship between two variables

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2026)
:::

::: notes
Prep for Class

1. Review Canvas submissions

<br>

**SLIDE**: Let's kick things off with your assignment for today

:::



## {background-image="Images/background-data_blue_v4.png" .center}

```{r, fig.align='center', fig.asp=.618, out.width="50%", fig.width=6.5}
#RColorBrewer::brewer.pal(2, name = "Accent")

ggplot(data = anes, aes(x = social.class, fill = polact.givecand)) +
  geom_bar(position = "dodge") +
  labs(x = "Social Class (ANES 2020)", y = "Count", fill = "Donated to a\n Candidate") +
  scale_x_discrete(limits = c("1. Lower class", "2. Working class", "3. Middle class", "4. Upper class")) +
  scale_fill_manual(values = c("#BEAED4", "#FDC086"))
```

:::: {.columns}
::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.618, fig.width=6.5}
ggplot(data = anes, aes(x = social.class, fill = polact.givecand)) +
  geom_bar(position = "stack") +
  labs(x = "Social Class (ANES 2020)", y = "Count", fill = "Donated to a\n Candidate") +
  scale_x_discrete(limits = c("1. Lower class", "2. Working class", "3. Middle class", "4. Upper class")) +
  scale_fill_manual(values = c("#BEAED4", "#FDC086"))
```
:::

::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.618, fig.width=6.5}
ggplot(data = anes, aes(x = social.class, fill = polact.givecand)) +
  geom_bar(position = "fill") +
  labs(x = "Social Class (ANES 2020)", y = "", fill = "Donated to a\n Candidate") +
  scale_x_discrete(limits = c("1. Lower class", "2. Working class", "3. Middle class", "4. Upper class")) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("#BEAED4", "#FDC086"))
```
:::
::::

::: notes

**Ok, does self reported social class predict the likelihood a respondent donates money to political candidates?**

- Make your case!

<br>

**And which type of bar plot "best" answers this question?**

- I think you can make the case for all three!

<br>

**SLIDE**: Code (if needed)

:::



## {background-image="Images/background-data_blue_v4.png" .center .smaller}

```{r, echo=TRUE, eval=FALSE}
# side-by-side
ggplot(data = anes, aes(x = social.class, fill = polact.givecand)) +
  geom_bar(position = "dodge") +
  labs(x = "Social Class (ANES 2020)", y = "Count", fill = "Donated to a\n Candidate") +
  scale_x_discrete(limits = c("1. Lower class", "2. Working class", 
                              "3. Middle class", "4. Upper class")) +
  scale_fill_manual(values = c("#BEAED4", "#FDC086"))

# stacked
ggplot(data = anes, aes(x = social.class, fill = polact.givecand)) +
  geom_bar(position = "stack") +
  labs(x = "Social Class (ANES 2020)", y = "Count", fill = "Donated to a\n Candidate") +
  scale_x_discrete(limits = c("1. Lower class", "2. Working class", 
                              "3. Middle class", "4. Upper class")) +
  scale_fill_manual(values = c("#BEAED4", "#FDC086"))

# proportionate
ggplot(data = anes, aes(x = social.class, fill = polact.givecand)) +
  geom_bar(position = "fill") +
  labs(x = "Social Class (ANES 2020)", y = "", fill = "Donated to a\n Candidate") +
  scale_x_discrete(limits = c("1. Lower class", "2. Working class", 
                              "3. Middle class", "4. Upper class")) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("#BEAED4", "#FDC086"))
```




## {background-image="Images/background-data_blue_v4.png" .center}

:::: {.columns}
::: {.column width="50%"}

```{r, fig.align='center', fig.asp=1.4, fig.width=6.5}
ggplot(data = anes, aes(y = partyid3, x = engagement_index)) +
  geom_boxplot(fill = "darkslateblue") +
  theme_bw() +
  labs(x = "Americans' Political Engagement Index", y = "") +
  scale_y_discrete(limits = c("3. Republican", "2. Independent", "1. Democrat")) +
  scale_x_continuous(breaks = 0:9)
```
:::

::: {.column width="50%"}
```{r, fig.align='center', fig.asp=1.4, fig.width=6.5}
anes2 <- select(.data = anes, partyid3, engagement_index)
anes3 <- na.omit(anes2)

ggplot(data = anes3, aes(x = engagement_index)) +
  geom_bar(fill = "darkslateblue") +
  facet_wrap(~ partyid3, ncol = 1, scales = "free_y") +
  theme_bw() +
  labs(x = "Americans' Political Engagement Index", y = "") +
  scale_x_continuous(breaks = 0:9)
```
:::
::::

::: notes

**Ok, does party identification predict the likelihood a respondent is highly engaged in American politics?**

- Make your case!

<br>

**And which of these two visualization approaches "best" answers this question?**

<br>

**SLIDE**: Code (if needed)

:::



## {background-image="Images/background-data_blue_v4.png" .center .smaller}

```{r, echo=TRUE, eval=FALSE}
# Box plot (rotated)
ggplot(data = anes, aes(y = partyid3, x = engagement_index)) +
  geom_boxplot(fill = "darkslateblue") +
  theme_bw() +
  labs(x = "Americans' Political Engagement Index", y = "") +
  scale_y_discrete(limits = c("3. Republican", "2. Independent", "1. Democrat")) +
  scale_x_continuous(breaks = 0:9)

## Histogram with facets

# To Remove the NAs: 1. Create data subset with two variables
anes2 <- select(.data = anes, partyid3, engagement_index)

# To Remove the NAs: 2. Only keep rows with complete data
anes3 <- na.omit(anes2)

ggplot(data = anes3, aes(x = engagement_index)) +
  geom_bar(fill = "darkslateblue") +
  facet_wrap(~ partyid3, ncol = 1, scales = "free_y") +
  theme_bw() +
  labs(x = "Americans' Political Engagement Index", y = "") +
  scale_x_continuous(breaks = 0:9)
```



## Bivariate Analyses {background-image="Images/background-data_blue_v4.png" .center .smaller}

<br>

**Categorical x Categorical**

- Cross-tabs, Proportions (with margins) and Bar plots (stacked, side-by-side, proportions)

**Numeric x Categorical**

- Descriptive statistics by group (aggregate across summary, mean, median, percentiles, standard deviations), facet wraps, and box plots

**Numeric x Numeric**

- Descriptive statistics by group (cut numeric into categorical), and scatter plots

::: notes

**Ok, how are we doing with bivariate analyses?**

- **Questions on the concepts of each tool? e.g. when to use it and what to use it for?**

- **Questions on the R code to execute these tools?**

<br>

**SLIDE**: Today I want us to keep practicing your work thinking in terms of causal argument and testing those arguments with data

:::



## {background-image="Images/background-data_blue_v4.png" .center}

**Applied Bivariate Analyses: ANES (2020)**

::: {.incremental}
1. Choose your Outcome: Political Engagement OR Negative Racial Attitudes

2. Choose your Predictor: Select a NEW ordinal variable from ANES

3. Propose a theory and hypothesis

4. Test 1: Bivariate analysis using the index version of the outcome

5. Test 2: Bivariate analysis using ONE of the questions in the index

6. Present your findings to the class!

:::

::: notes

*Split class into groups of THREE*

- Sit with your group!

<br>

Ok, groups time to practice using bivariate analyses to test causal arguments!

- **REVEAL** x 5

<br>

Test 1 is an interval outcome x a categorical predictor

- You need to do this analysis with BOTH a visualization AND descriptive stats by group level

<br>

Test 2 is a categorical outcome x a categorical predictor

- You need to do this analysis with BOTH a visualization AND a cross-tab

<br>

**Questions on the assignment?**

<br>

We'll come back together with 15 minutes left to see what we've learned about the world!

- Go!

<br>

**SLIDE**: For next class

:::



## {background-image="Images/background-data_blue_v4.png" .center}

```{r}
# 1. Been Arrested (been.arrested) -> Political engagement (engagement_index)
ggplot(anes, aes(x = been.arrested, y = engagement_index)) +
  geom_boxplot() +
  theme_bw() +
  scale_y_continuous(breaks = 0:9)

ggplot(anes, aes(x = been.arrested, fill = polact.talkvote)) +
  geom_bar() +
  theme_bw() +
  labs(fill = "")

# Multi?
ggplot(anes, aes(x = partyid3, y = engagement_index, fill = been.arrested)) +
  geom_boxplot() +
  theme_bw() +
  scale_y_continuous(breaks = 0:9)


# 2. Ethnicity (race.ethnicity) -> Negative racial attitudes (racial_negativity)
ggplot(anes, aes(y = race.ethnicity, x = racial_negativity)) +
  geom_boxplot() +
  theme_bw() +
  scale_x_continuous(labels = scales::percent_format())



# 3. American nationalism (american.id.import) -> Political engagement (engagement_index)

ggplot(anes, aes(x = american.id.import, y = engagement_index)) +
  geom_boxplot() +
  theme_bw() +
  scale_y_continuous(breaks = 0:9)

ggplot(anes, aes(x = american.id.import, fill = polact.talkvote)) +
  geom_bar(position = "fill") +
  theme_bw() +
  labs(fill = "")


# 4. American nationalism (american.id.import) -> Negative racial attitudes (racial_negativity)
ggplot(anes, aes(x = american.id.import, y = racial_negativity)) +
  geom_boxplot() +
  theme_bw() +
  scale_y_continuous(labels = scales::percent_format())




# 5. Age (age) -> Negative racial attitudes (racial_negativity)
ggplot(anes, aes(y = age, x = racial_negativity)) +
  geom_point() +
  theme_bw()










```


## For Next Class {background-image="Images/background-data_blue_v4.png" .center}

<br>

::: {.r-fit-text}

**Visualizing Americans' Beliefs in Conspiracy Theories**

1. Read textbook excerpt (Ch6, p129-146)

2. Re-test our conspiracy theory hypotheses with visualizations

:::

::: notes

Everybody take a look at the assignment on Canvas

<br>

**Questions on the assignment?**

<br>

Canvas ASSIGNMENT: Visualizing Americans' Beliefs in Conspiracy Theories 

1. Read the excerpt from Chapter 6 in Pollock and Edwards' textbook (p129-146)

2. On Monday you used bivariate descriptive statistics to analyze Americans' beliefs in conspiracy theories. For next class I want you to retest ALL FOUR hypotheses using bivariate visualizations. Submit polished visualizations of the results with complete figure captions.

Hypotheses:

- H1. Belief in liberal CTs (Lib_Conspiracy_Index) increases as trust in government falls
- H2. Belief in conservative CTs (Cons_Conspiracy_Index) increases as trust in government falls
- H3. Democrats are more likely to believe liberal coded conspiracies
- H4. Republicans are more likely to believe conservative coded conspiracies

:::