---
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
fontsize: 14pt
classoption: "aspectratio=169"
header-includes:
- \usepackage{caption, xcolor}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "./09_1-Models.Rmd", output_file = "./09_1-Models.pdf")

library(tidyverse)
##library(readxl)
library(stargazer)
##library(kableExtra)
##library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```


## Today's Agenda
\large
**Planning our class research project**

+ Select a research question,

+ Describe our model, and

+ Merge the data.

\vspace{.5in}

\begin{center}
Justin Leinaweaver (Fall 2021)
\end{center}



## Group Presentations
\large

Your presentation should:

+ Be 8 minutes in length (with Rmarkdown slides),

+ Propose a research question, 

+ Propose a model answering that question, and

+ Provide supporting UNIVARIATE analyses that show sufficient variation exists in the predictor and the outcome



## Decision Time!
\Large
\vspace{.25in}

1. What is our research question?

\vspace{.15in}

2. What is our model?

```{r, echo = FALSE, fig.asp = .4, fig.width = 5, fig.align = 'center', out.width = '90%'}
tribble(
    ~Name, ~x, ~y,
    "Predictor", 1, 1,
    "Outcome", 3, 1
) %>%
    ggplot(aes(x = x, y = y)) +
    geom_point(size = .1) +
    geom_label(aes(label = Name), size = 6) +
    coord_cartesian(xlim = c(0.8, 3.3), ylim = c(.93, 1.04)) +
    ggthemes::theme_tufte() +
    annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
    scale_x_continuous(breaks = NULL) +
    scale_y_continuous(breaks = NULL) +
    labs(x = "", y = "")
```




## Assignment 3: The Research Report {.plain}

1. Introduction
   + What is the question, why is it important and what is your argument?

\vspace{.1in}

2. Bivariate Analyses
   + What is the answer to the question using only bivariate visualizations and simple OLS regressions?

\vspace{.1in}

3. Multivariate Analyses
   + What is the answer to the question using multivariate visualizations and multiple OLS regressions?

\vspace{.1in}

4. Conclusion
   + What did you find and how confident should we be in your results?
   
   
  
## Assignment 3: The Research Report {.plain}
\LARGE

1. Introduction

	+ \Large What is the question?

	+ \Large Why is it important?

	+ \Large What is your argument?


## For Tuesday
\Large

1. Read Wheelan ch 4 on correlation

\vspace{.25in}

2. Take note of the Chang (2018) recipes on bivariate analyses in the syllabus.
    + \Large (3.2, 3.3, 3.7, 3.8, 5, 6.6, 15.14)





