---
format: revealjs
---

# Today's Agenda {background-image="libs/Images/background-data_blue_v3.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
```

<br>

<br>

**Apply bivariate analyses to our research project**

<br>

<br>

::: r-stack
Justin Leinaweaver (Spring 2024)
:::

::: notes
Prep for Class

1. ?
:::



## For Today: Present your "best" visualizations {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. 'cut' and 'clarity'

2. 'cut' and 'price'

3. 'carat' and 'price'

<br>

**Source:** 'diamonds' dataset in tidyverse

::: notes
Ok, here's what I asked you to do for class today

- My hope is that this gave you a chance to practice all of the visualizations from last class, PLUS

- A chance to practice our polishing skills again
:::



## Diamonds: 'cut' and 'clarity' {background-image="libs/Images/background-slate_v2.png" .center}

<br>

::: {.fragment}
```{r, echo=FALSE, fig.align='center', fig.asp=.618, fig.width=8, cache=TRUE}
ggplot(diamonds, aes(x = clarity, fill = forcats::fct_rev(cut))) +
  geom_bar(position = "fill") +
  theme_bw() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(type = "seq", palette = 4) +
  labs(x = "Diamond Clarity", fill = "Cut Type",
       caption = "Source: The Loose Diamonds Search Engine",
       title = "Jewelers appear to use higher quality cuts with higher quality diamonds")
```
:::

::: notes
Alright, let's see what everybody made!

- *Encourage everyone to walk around and see all the plots*

<br>

**What are we seeing? What do we like?**

<br>

**REVEAL**: Here's my stab
:::



## Diamonds: 'cut' and 'price' {background-image="libs/Images/background-slate_v2.png" .center}

::::: {.fragment}
:::: {.columns}
::: {.column width="45%"}

<br>

```{r, echo=FALSE, fig.align='center', fig.asp=1, fig.width=6, cache=TRUE}
ggplot(diamonds, aes(x = price, y = forcats::fct_rev(cut))) +
  geom_boxplot(fill = "seagreen2") +
  theme_bw() +
  labs(x = "", y = "", title = "Does the quality of the a diamond's cut influence its price?", caption = "Source: The Loose Diamonds Search Engine") +
  scale_x_continuous(labels = scales::dollar_format()) 
```
:::

::: {.column width="55%"}
```{r, echo=FALSE, fig.align='center', fig.asp=1.1, fig.width=7, cache=TRUE}
ggplot(diamonds, aes(x = price)) +
  geom_histogram(bins = 15) +
  facet_wrap(~ cut, ncol = 1) +
  theme_bw() +
  scale_x_continuous(labels = scales::dollar_format()) 
```
:::
::::
:::::

::: notes
The distributions overlap considerably.

- Doesn't look like cut and price are highly correlated!
:::



## Diamonds: 'carat' and 'price' {background-image="libs/Images/background-slate_v2.png" .center}

::: {.fragment}
```{r, fig.align='center', fig.width=9, fig.asp=.65, cache=TRUE}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = .1) +
  theme_bw() +
  labs(x = "Size of the Diamond (carats)", y = "Prices",
       title = "The larger the diamond the more expensive!") +
  scale_y_continuous(labels = scales::dollar_format())
```
:::

::: notes

:::



## Bivariate Analyses {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. Descriptive Statistics by Group
    - Counts, Proportions, Means, Medians, Standard Deviations, Percentiles, Ranges, Interquartile Ranges (IQR) and Correlations

2. Visualizations
    - Bar plots (stacked, side-by-side, proportions), box plots and scatter plots

::: notes
**Any questions on our bivariate analyses tools?**
:::



## Are powerful states more likely to respect the human rights of their citizens? {background-image="libs/Images/background-slate_v2.png" .center .smaller}

:::: {.columns}
::: {.column width="20%"}
- `cinc`
- `milex`
- `milper`
- `irst`
- `pec`
- `tpop`
- `upop`
:::

::: {.column width="60%"}

<br>

<br>

```{r, fig.width = 9, fig.height=2}
## Manual DAG
d1 <- tibble(
  x = c(-3, 3),
  y = c(1, 1),
  labels = c("Military\nPower\n(CoW NMC)", "Human rights\nand\nthe rule of law\n(FFP)")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = -1.9, xend = 1.85, y = 1, yend = 1, arrow = arrow())
```
:::

::: {.column width="20%"}
- `FSI_Total`
- `P3_Human_Rts`
:::
::::
















## Old Slides

## Rescaling Variables: log10 (Price) {background-image="libs/Images/background-slate_v2.png" .center}

:::: {.columns}
::: {.column width="45%"}

<br>

```{r, echo=FALSE, fig.align='center', fig.asp=1, fig.width=6, cache=TRUE}
ggplot(diamonds, aes(x = price, y = forcats::fct_rev(cut))) +
  geom_boxplot(fill = "seagreen2") +
  theme_bw() +
  labs(x = "", y = "", title = "Does the quality of the a diamond's cut influence its price?", caption = "Source: The Loose Diamonds Search Engine") +
  scale_x_log10(labels = scales::dollar_format()) 
```
:::

::: {.column width="55%"}
```{r, echo=FALSE, fig.align='center', fig.asp=1.1, fig.width=7, cache=TRUE}
ggplot(diamonds, aes(x = price)) +
  geom_histogram(bins = 15) +
  facet_wrap(~ cut, ncol = 1) +
  theme_bw() +
  scale_x_log10(labels = scales::dollar_format())
```
:::
::::




## Rescaling Variables: The Log Scale {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.align='center', fig.asp=.618, fig.width=9}
tibble(
  x1 = 1:10,
  x2 = log10(x1)
) |>
  ggplot(aes(x = x1, y = x2)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(x = "Normal Scale", y = "Log 10 Scale",
       title = "The Common Logarithmic Scale (base 10)") +
  scale_x_continuous(breaks = 1:20) +
  scale_y_continuous(breaks = seq(0, 2, .1))
```

::: notes
While the most common transformations focus on moving a decimal place (e.g. making thousands into millions), there are a number of other options

<br>

One very useful tool is a family of transformations under the log scale

- On this plot I'm showing you how the numbers from 1-10 translate into a common log  scale (base 10).
:::



## Rescaling Variables: The Log Scale {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.align='center', fig.asp=.618, fig.width=9}
tibble(
  x1 = 1:10,
  x2 = log10(x1)
) |>
  ggplot(aes(x = x1, y = x2)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(x = "Normal Scale", y = "Log 10 Scale",
       title = "The Common Logarithmic Scale (base 10)") +
  scale_x_continuous(breaks = 1:20) +
  scale_y_continuous(breaks = seq(0, 2, .1)) +
  annotate("point", x = 4, y = .6, size = 4, color = "red") +
  annotate("point", x = 8, y = .9, size = 4, color = "red")
```

::: notes
For examples

- The log base 10 value of 5 is approximately .6, and

- The log 10 value of 6 is approximately .78
:::



## Rescaling Variables: The Log Scale {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.align='center', fig.asp=.618, fig.width=9}
tibble(
  x1 = 1:10,
  x2 = log10(x1)
) |>
  ggplot(aes(x = x1, y = x2)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(x = "Normal Scale", y = "Log 10 Scale",
       title = "The Common Logarithmic Scale (base 10)") +
  scale_x_continuous(breaks = 1:20) +
  scale_y_continuous(breaks = seq(0, 2, .1)) +
  annotate("point", x = 4, y = .6, size = 4, color = "red") +
  annotate("segment", x = 1, xend = 4, y = .6, yend = .6, linetype = "dashed", color = "red") +
  annotate("segment", x = 4, xend = 4, y = 0, yend = .6, linetype = "dashed", color = "red") +
  annotate("point", x = 8, y = .9, size = 4, color = "red") +
  annotate("segment", x = 1, xend = 8, y = .9, yend = .9, linetype = "dashed", color = "red") +
  annotate("segment", x = 8, xend = 8, y = 0, yend = .9, linetype = "dashed", color = "red")
```

::: notes
Moving from a value of 4 to 8 on our normal scale represents a difference of 4 units

<br>

BUT, on the log10 scale this 4 unit difference is reduced!

- The difference is now from .6 to .9 or .3

<br>

Moving to the log 10 scale makes differences on the normal scale much smaller.

<br>

**SLIDE**: The effect is even more stark as we get into much bigger numbers
:::



## Rescaling Variables: The Log Scale {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.align='center', fig.asp=.618, fig.width=9}
tibble(
  x1 = 1:10000,
  x2 = log10(x1)
) |>
  ggplot(aes(x = x1, y = x2)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(x = "Normal Scale", y = "Log 10 Scale",
       title = "The Common Logarithmic Scale (base 10)") +
  #scale_x_continuous(breaks = 1:20) +
  #scale_y_continuous(breaks = seq(0, 2, .1)) +
  annotate("point", x = 5000, y = 3.7, size = 4, color = "red") +
  annotate("segment", x = 1, xend = 5000, y = 3.7, yend = 3.7, linetype = "dashed", color = "red") +
  annotate("segment", x = 5000, xend = 5000, y = 0, yend = 3.7, linetype = "dashed", color = "red") +
  annotate("point", x = 7500, y = 3.88, size = 4, color = "red") +
  annotate("segment", x = 1, xend = 7500, y = 3.88, yend = 3.88, linetype = "dashed", color = "red") +
  annotate("segment", x = 7500, xend = 7500, y = 0, yend = 3.88, linetype = "dashed", color = "red")
```

::: notes
Let's move into the thousands range and rexamine the log 10 scale

- A value of 5,000 on the normal scale corresponds to a log 10 value of approximately 3.7

- A value of 7,500, which is an increase of 2.5k, corresponds to a log 10 value of approximately 3.88

<br>

That means an increase of 2.5k is shrunk to only a difference of .18 on the log 10 scale!

<br>

**Anybody know why we might do this?**

- **What is the value of the log scale transformation?**

<br>

In short, the log scale helps us examine distributions where many observations are compressed in the smaller values OR where you have extreme outliers

- At the small end the log scale helps us distinguish between cases with similar values on the normal scale

- At the big end the log scale pulls outliers back towards the rest of the pack

<br>

**SLIDE**: Illustrate with GDP
:::


