---
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    df_print: kable
fontsize: 14pt
classoption: "aspectratio=169"
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "", output_file = "")

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)

d <- read_csv("../Data_Projects/04-Research_Project/P296-FA21-Research_Project_Data.csv", na = c(-9, "NA"))

d2 <- d |>
  arrange(Country, year) |>
  group_by(Country) |>
  mutate(
    milper_prop = (milper / tpop)*100,
    milper_prop_lag1 = lag(milper_prop, n = 1),
    milper_prop_lag3 = lag(milper_prop, n = 3)
  ) |>
  ungroup()
```


## Today's Agenda
\large

Build univariate and bivariate visualizations for your research project report.

\vspace{.5in}

\center
Justin Leinaweaver (Fall 2021)



## Visualizations
\large
1. Histogram: The new predictor variable (milper_prop)
2. Stacked bars: FH status by year
3. Line: Mean FH total across time (+/- 1 Std Dev)  
4. Boxes: milper_prop x FH status
5. Scatter (with correlation): milper_prop x FH total
6. Scatter (with correlation): milper_prop x pr
7. Scatter (with correlation): milper_prop x cl


## 
```{r, fig.width=7, out.width='95%', fig.asp=0.5}
## Histogram: The new predictor variable (milper_prop)
d2 |>
  pivot_longer(cols = c(milper, milper_prop), names_to = "Variables", values_to = "Values") |>
  mutate(
    Variables = if_else(Variables == "milper", "Military Personnel (k)", "Military Personnel (% Population)")
  ) |>
  ggplot(aes(x = Values)) +
  geom_histogram(color = "white", fill = "skyblue2") +
  facet_wrap(~ Variables, scales = "free") +
  labs(x = "", y = "Counts", 
       title = "Most countries enlist well below 1% of their populations in the military",
       subtitle = "(Almost all below 2%)") +
  theme_bw()
```



## {.plain}
```{r, fig.width=7, out.width='92%', fig.asp=0.618}
## Bar plot: status x year
d2 |>
  filter(!is.na(status)) |>
  mutate(
    status2 = case_when(
      status == "NF" ~ "Not Free",
      status == "PF" ~ "Partly Free",
      status == "F" ~ "Free"
    ),
    status2 = factor(status2, levels = c("Not Free", "Partly Free", "Free"))
  ) |>
  ggplot(aes(x = year, fill = status2)) +
  geom_bar() +
  labs(x = "", y = "Counts", fill = "",
       title = "Freedom was on the march around the world (until 9/11)!") +
  theme_bw() +
  guides(fill = FALSE) +
  annotate("label", x = 2011, y = 175, label = "Not Free", size = 5.5) +
  annotate("label", x = 2011, y = 120, label = "Partly Free", size = 5.5) +
 annotate("label", x = 2011, y = 40, label = "Free", size = 5.5) +
  scale_fill_brewer(palette = 1)

```



## {.plain}
```{r, fig.width=7, out.width='95%', fig.asp=0.618}
## Line: Mean FH total across time (+/- 1 Std Dev)  
d2 |>
  group_by(year) |>
  summarize(
    mean = mean(fh_total, na.rm = TRUE),
    low = mean - sd(fh_total, na.rm = TRUE),
    high = mean + sd(fh_total, na.rm = TRUE)
  ) |>
  ggplot(aes(x = year, y = mean)) +
  geom_ribbon(aes(ymin = low, ymax = high), fill = "lightblue") +
  geom_line(color = "black") +
  theme_bw() +
  labs(x = "", y = "FH Total Score",
       title = "The average FH score has been falling since the 1970s", subtitle = "(Although things may have stabilized since 2005)",
       caption = "Line and ribbon represents the mean +/- 1 Std Dev") +
  scale_y_continuous(breaks = seq(2, 14, 2), limits = c(1,14))
```


## {.plain}
```{r, fig.width=7, out.width='95%', fig.asp=0.618}
## Boxes: milper_prop x FH status
d2 |>
  filter(!is.na(status)) |>
  mutate(
    status2 = case_when(
      status == "NF" ~ "Not Free",
      status == "PF" ~ "Partly Free",
      status == "F" ~ "Free"
    ),
    status2 = factor(status2, levels = c("Not Free", "Partly Free", "Free"))
  ) |>
  filter(!is.na(status)) |>
  ggplot(aes(x = status2, y = milper_prop)) +
  geom_boxplot(fill = "lightblue") +
  theme_bw() +
  labs(x = "", y = "Military Personnel (% Population)",
       title = "Freedom appears to correlate with smaller enlistments")
```


## {.plain}
```{r, fig.width=7, out.width='95%', fig.asp=0.5}
# 5. Scatter: milper_prop x FH total
# 6. Scatter: milper_prop x pr
# 7. Scatter: milper_prop x cl
# cor(d2$milper_prop, d2$fh_total, use = "complete")
# cor(d2$milper_prop, d2$pr, use = "complete")
# cor(d2$milper_prop, d2$cl, use = "complete")

d2 |>
  pivot_longer(cols = pr:fh_total, names_to = "FH", values_to = "Values") |>
  mutate(
    FH = case_when(
      FH == "cl" ~ "Civil Liberties (0.29)",
      FH == "pr" ~ "Political Rights (0.26)",
      FH == "fh_total" ~ "FH Total (0.28)"
    )
  ) |>
  ggplot(aes(y = Values, x = milper_prop)) +
  #geom_point(alpha = .1) +
  geom_jitter(alpha = .05, width = .25) +
  facet_wrap(~ FH,scales = "free") +
  theme_bw() #+
  labs(x = "", y = "Military Personnel (% Population)",
       title = "Freedoms appear to decline as military enlistments increase in size")

```


## For Tuesday
\Large

1. Wheelan chapter 11 "Regression"

\vspace{.3in}

2. *Linear Regression in R* on Moodle, "Simple OLS" section only





