---
format: revealjs
---

# Today's Agenda {background-image="libs/Images/background-data_blue_v3.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
library(ggeffects)
```

<br>

<br>

**Build the Third Report**

- Bivariate Analyses: Simple OLS

<br>

<br>

::: r-stack
Justin Leinaweaver (Spring 2024)
:::

::: notes
Prep for Class

1. Make sure merged data available on Canvas

2. Post assignment details on Canvas

<br>

**SLIDE**: Before we start on the third report, let's review your work for today evaluating the fit of our practice OLS regressions.
:::



## Evaluate the fit of both regressions: {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. Regress city fuel economy ('cty') on engine displacement ('displ') in the 'mpg' datatset

2. Regress percent college educated ('percollege') on the percent of people below the  poverty line ('percbelowpoverty') in the 'midwest' datatset

:::notes
**How did this go?**
:::



## City Fuel Economy and Engine Size {.center .smaller}

<br>

:::: {.columns}
::: {.column width="50%"}
```{r}
model1 <- lm(data = mpg, cty ~ displ)

modelsummary(list("Fuel Economy (City)" = model1), fmt = 2, stars = c('*' = .05), gof_omit = "IC|Log|F",
             coef_map = c("displ"= "Engine Size", "(Intercept)" = "Constant")) |>
  column_spec(1, width = "9em") |>
  column_spec(2, background = c("orange", rep("white", 3), "orange", "orange", "white", "white")) |>
  kable_styling(font_size = 24)
```
:::

::: {.column width="50%"}
```{r, echo=TRUE}
nrow(mpg)
```

<br>

```{r, echo = FALSE, fig.retina=3, fig.asp=0.7, fig.align = 'center', out.width = '90%', fig.width=7, cache=TRUE}
# Check residuals
plot(model1, which = 1)
```
:::
::::

::: notes

1. Missing data problems?

2. Are the coefficients significant?

3. What does the R<sup>2</sup> indicate? 

4. Any problems in the residuals plot?
:::



## College Education and Poverty  {.center .smaller}

<br>

:::: {.columns}
::: {.column width="50%"}
```{r}
model2 <- lm(data = midwest, percollege ~ percbelowpoverty)

modelsummary(list("College (%)" = model2), fmt = 2, stars = c('*' = .05), gof_omit = "IC|Log|F",
             coef_map = c("percbelowpoverty"= "Below Poverty (%)", "(Intercept)" = "Constant")) |>
  column_spec(1, width = "10em") |>
  column_spec(2, background = c("orange", rep("white", 3), "orange", "orange", "white", "white")) |>
  kable_styling(font_size = 24)
```
:::

::: {.column width="50%"}
```{r, echo=TRUE}
nrow(midwest)
```

<br>

```{r, echo = FALSE, fig.retina=3, fig.asp=0.7, fig.align = 'center', out.width = '90%', fig.width=7, cache=TRUE}
# Check residuals
plot(model2, which = 1)
```
:::
::::

::: notes

1. Missing data problems?

2. Are the coefficients significant?

3. What does the R<sup>2</sup> indicate? 

4. Any problems in the residuals plot?
:::



## Report 3 {background-image="libs/Images/background-slate_v2.png" .center}

**Submission Requirements**

<br>

- Submit a **pdf** to **Canvas** (due **Mar 26th**)

- Include an **APA** formatted bibliography

- Polish **all** included visualizations and tables

::: notes

Ok, this week we work on your second report.

- The basic requirements same as last time

- Late penalty (-10) points per day
:::



## Report 3 {background-image="libs/Images/background-slate_v2.png" .center}

**Bivariate Test of your Hypotheses**

<br>

1. Introduction

2. Test hypotheses with bivariate visualizations and statistics

3. Test hypotheses with simple OLS regressions

::: notes
Next report has three required sections.
:::



## Report 3 {background-image="libs/Images/background-slate_v2.png" .center}

**Bivariate Test of your Hypotheses**

<br>

1. Introduction
    - Research Question?
    - Why is it important?
    - What are your three hypotheses?

2. Test hypotheses with bivariate visualizations and statistics

3. Test hypotheses with simple OLS regressions

::: notes

Section 1: Introduction
- What is the research question?
- Make an argument: Why should we care?
- What are your hypotheses? List all three
:::



## Report 3 {background-image="libs/Images/background-slate_v2.png" .center .smaller}

**Bivariate Test of your Hypotheses**

<br>

1. Introduction

2. Test hypotheses with bivariate visualizations and statistics
    - 2019 only
    - Scatterplots and correlations
        - GII x military
        - GII x education
        - GII x health

3. Test hypotheses with simple OLS regressions

::: notes

This is what we worked on in class last Friday.

<br>

Section 2: Test your hypothesis using bivariate visualizations and statistics

- Make a scatterplot AND calculate the correlation for each (2019 only):
    - GII x military
    - GII x education
    - GII x health
:::
    
    
    
## Report 3 {background-image="libs/Images/background-slate_v2.png" .center .smaller}

**Bivariate Test of your Hypotheses**

<br>

1. Introduction

2. Test hypotheses with bivariate visualizations and statistics

3. Test hypotheses with simple OLS regressions
    - 2019 only
    - Fit, interpret and evaluate each regression
        - GII ~ military
        - GII ~ education
        - GII ~ health

::: notes

Today's work!

Section 3: Test your hypothesis using simple OLS regressions

- Fit, interpret and evaluate three simple OLS regressions (2019 only)
::: 
