---
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    df_print: kable
fontsize: 14pt
classoption: "aspectratio=169"
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "./09_1-Simple_OLS.Rmd", output_file = "./09_1-Simple_OLS.pdf")

library(tidyverse)
##library(readxl)
library(stargazer)
##library(kableExtra)
library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```


## Today's Agenda

\Large
Fitting, interpreting and analyzing simple OLS regressions

\vspace{.5in}

\begin{center}
Justin Leinaweaver (Fall 2021)
\end{center}



##
\center
\huge
**Part 1**

\vspace{.25in}

**Why do we need regressions?**



## {.plain}
```{r, fig.align = 'center', fig.asp=.75, out.width = '60%', fig.width = 4.5, cache=TRUE}
diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)",
       title = str_c("The correlation of diamond weight to price is:\n", round(cor(diamonds$carat, diamonds$price), 2))) +
  coord_cartesian(xlim = c(0,4))
```

\center
**What conclusions can we draw based on the visualization and the correlation?**



## {.plain}
```{r, fig.align = 'center', fig.asp=.75, out.width = '60%', fig.width = 4.5, cache=TRUE}
diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  #geom_vline(xintercept = 1.5, color = "red2", linetype = "dashed", size = 1.3) +
  #geom_vline(xintercept = 2.5, color = "red2", linetype = "dashed", size = 1.3) +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)",
       title = str_c("The correlation of diamond weight to price is:\n", round(cor(diamonds$carat, diamonds$price), 2))) +
  coord_cartesian(xlim = c(0,4))
```

\center
**What is your most precise estimate of the price difference between a 1.5 carat and a 2.5 carat diamond?**



## 
\large
\begin{center}
Regression is a technique for estimating the relationship between predictor variables (X) and an outcome (Y) using the formula for a line.

\vspace{.2in}

\Huge

Y = $\alpha$ + $\beta$ X

\vspace{.2in}

\large

\textbf{What clarifications do you need from the readings?}

\end{center}



## Why 'Regression'?
```{r, fig.show="hold", out.width="48%", fig.width=5, fig.height=3, cache=TRUE}
diamonds |>
  slice_sample(prop = .4) |>
  ggplot(aes(x = carat)) +
  geom_histogram(bins = 25, color = "white") +
  theme_bw() +
  labs(x = "Carats", y = "")

diamonds |>
  slice_sample(prop = .4) |>
  ggplot(aes(x = price)) +
  geom_histogram(bins = 25, color = "white") +
  theme_bw() +
  labs(x = "Price ($)", y = "")
```

\center
**If you pull a single diamond from this collection at random, what is the most likely price and size it will be?**



## Why 'Regression'?
```{r, fig.show="hold", out.width="48%", fig.width=5, fig.height=3, cache=TRUE}
diamonds |>
  slice_sample(prop = .4) |>
  ggplot(aes(x = carat)) +
  geom_histogram(bins = 25, color = "white") +
  theme_bw() +
  labs(x = "Carats", y = "") +
  geom_vline(xintercept = mean(diamonds$carat), color = "red", size=1.5)

diamonds |>
  slice_sample(prop = .4) |>
  ggplot(aes(x = price)) +
  geom_histogram(bins = 25, color = "white") +
  theme_bw() +
  labs(x = "Price ($)", y = "") +
  geom_vline(xintercept = mean(diamonds$price), color = "red", size=1.5)
```

\center
**Pros and cons of guessing the average?**




## Why 'Regression'?
```{r, fig.show="hold", out.width="48%", fig.width=5, fig.height=3, cache=TRUE}
diamonds |>
  slice_sample(prop = .4) |>
  ggplot(aes(x = carat)) +
  geom_histogram(bins = 25, color = "white") +
  theme_bw() +
  labs(x = "Carats", y = "")

diamonds |>
  slice_sample(prop = .4) |>
  ggplot(aes(x = price)) +
  geom_histogram(bins = 25, color = "white") +
  theme_bw() +
  labs(x = "Price ($)", y = "")
```

\center
**If you first pull a massive diamond from the collection, what is the most likely price and size of the next one you draw?**




## Regression aka 'regression to the mean'
```{r, fig.show="hold", out.width="48%", fig.width=5, fig.height=3, cache=TRUE}
diamonds |>
  slice_sample(prop = .4) |>
  ggplot(aes(x = carat)) +
  geom_histogram(bins = 25, color = "white") +
  theme_bw() +
  labs(x = "Carats", y = "") +
  geom_vline(xintercept = mean(diamonds$carat), color = "red", size=1.5)

diamonds |>
  slice_sample(prop = .4) |>
  ggplot(aes(x = price)) +
  geom_histogram(bins = 25, color = "white") +
  theme_bw() +
  labs(x = "Price ($)", y = "") +
  geom_vline(xintercept = mean(diamonds$price), color = "red", size=1.5)
```

\center
**A random draw of an outlier is more likely to be followed by an average value than another outlier.**





## 
::: columns
:::: column
```{r, echo=FALSE, fig.align = 'center', fig.asp=.8, out.width = '95%', fig.width = 4, cache=TRUE}
diamonds |>
  slice_sample(prop = .4) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  annotate("point", x = mean(diamonds$carat), y = mean(diamonds$price), color = "red", size = 3) +
  theme_bw() +
  labs(x = "Carats", y = "Price") +
  coord_cartesian(xlim = c(0,4))

```
::::
:::: column
\vspace{.5in}
```{r, echo=FALSE}
diamonds |>
  pivot_longer(cols = c(carat, price), names_to = "Variable", values_to = "Values") |>
  group_by(Variable) |>
  summarize(
    Mean = round(mean(Values), 3)
  )
```
::::
:::

\center
**If data tends towards its 'central tendency' then let's start our model using the sample means.**




## 
::: columns
:::: column
```{r, echo=FALSE, fig.align = 'center', fig.asp=.8, out.width = '95%', fig.width = 4, cache=TRUE}
## Add line guesses
x1 <- .798
y1 <- 3933
x2 <- 0
y2 <- 0
x3 <- .25
y3 <- 500
x4 <- .5
y4 <- 500

b1 <- (y2-y1)/(x2-x1)
a1 <- y1 - b1*x1

b2 <- (y3-y1)/(x3-x1)
a2 <- y1 - b2*x1

b3 <- (y4-y1)/(x4-x1)
a3 <- y1 - b3*x1

diamonds |>
  slice_sample(prop = .4) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  geom_abline(slope = b1, intercept = a1, color = "green", size = 1.1) +
  geom_abline(slope = b2, intercept = a2, color = "orange", size = 1.1) +
  geom_abline(slope = b3, intercept = a3, color = "purple", size = 1.1) +
  annotate("point", x = mean(diamonds$carat), y = mean(diamonds$price), color = "red", size = 3) +
  theme_bw() +
  labs(x = "Carats", y = "Price") +
  coord_cartesian(xlim = c(0,4))

```
::::
:::: column
\vspace{.5in}
```{r, echo=FALSE}
diamonds |>
  pivot_longer(cols = c(carat, price), names_to = "Variable", values_to = "Values") |>
  group_by(Variable) |>
  summarize(
    Mean = round(mean(Values), 3)
  )
```
::::
:::

\center
**How do we pick the second point we need to draw the "best" line?**



##
::: columns
:::: column

```{r, echo=FALSE, fig.align = 'center', fig.asp=.8, out.width = '95%', fig.width = 4, cache=TRUE}
diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  annotate("point", x = mean(diamonds$carat), y = mean(diamonds$price), color = "red", size = 2) +
  geom_vline(xintercept = mean(diamonds$carat), color = "red", linetype = "dashed") +
  geom_hline(yintercept = mean(diamonds$price), color = "red", linetype = "dashed") +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)") +
  annotate("text", x = 0.1, y = 3000, label = "3", size = 5.5, color = "red") +
    annotate("text", x = 3, y = 3000, label = "4", size = 5.5, color = "red") +
  annotate("text", x = .1, y = 15000, label = "1", size = 5.5, color = "red") +
  annotate("text", x = 3, y = 15000, label = "2", size = 5.5, color = "red") +
  coord_cartesian(xlim = c(0,4))
```

::::
:::: column

\vspace{.2in}
```{r, echo=FALSE}
diamonds |>
  mutate(
    Quadrants = case_when(
      carat < mean(diamonds$carat) & price > mean(diamonds$price) ~ "1",
      carat < mean(diamonds$carat) & price < mean(diamonds$price) ~ "3",
      carat > mean(diamonds$carat) & price > mean(diamonds$price) ~ "2",
      carat > mean(diamonds$carat) & price < mean(diamonds$price) ~ "4"
    )
  ) |> 
  count(Quadrants)
```

::::
:::

\center
**The line should follow the data, so let's divide the plot into quadrants and focus on the one with the most observations.**



## 
::: columns
:::: column

```{r, echo=FALSE, fig.align = 'center', fig.asp=.8, out.width = '95%', fig.width = 4, cache=TRUE}
diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  annotate("point", x = mean(diamonds$carat), y = mean(diamonds$price), color = "red", size = 2) +
  geom_vline(xintercept = mean(diamonds$carat), color = "red", linetype = "dashed") +
  geom_hline(yintercept = mean(diamonds$price), color = "red", linetype = "dashed") +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)") +
  annotate("text", x = 0.1, y = 3000, label = "3", size = 5.5, color = "red") +
    annotate("text", x = 3, y = 3000, label = "4", size = 5.5, color = "red") +
  annotate("text", x = .1, y = 15000, label = "1", size = 5.5, color = "red") +
  annotate("text", x = 3, y = 15000, label = "2", size = 5.5, color = "red") +
  annotate("point", x = .4631, y = 1358, color = "red", size = 2) +
  coord_cartesian(xlim = c(0,4))
```

::::
:::: column

\vspace{.5in}
```{r, echo=FALSE}
diamonds |>
  filter(carat < mean(diamonds$carat), price < mean(diamonds$price)) |>
  pivot_longer(cols = c(carat, price), names_to = "Variable", values_to = "Values") |>
  group_by(Variable) |>
  summarize(
    Mean = round(mean(Values), 3)
  ) |>
  mutate(
    Group = "Quadrant 3"
  ) |>
  select(Group, everything())
```

::::
:::

\vspace{.25in}
\center
**The mean values in quadrant three give us a second point!**



## {.plain}
\center
**Estimated Line: Price = -2,203 + 7,689 (Carats)**

```{r, fig.align = 'center', fig.asp=.7, out.width = '55%', fig.width = 4.5, cache=TRUE}
diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  annotate("point", x = mean(diamonds$carat), y = mean(diamonds$price), color = "red", size = 2) +
  annotate("point", x = .4631, y = 1358, color = "red", size = 2) +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)") +
  geom_abline(slope = 7689.65, intercept = -2203.0767779328, color = "red") +
  coord_cartesian(xlim = c(0,4))
```

\center
**How do we interpret this estimated formula for a line?**



## {.plain}
\center
**Estimated Line: Price = -2,203 + 7,689 (Carats)**

```{r, fig.align = 'center', fig.asp=.7, out.width = '55%', fig.width = 4.5, cache=TRUE}
diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  annotate("point", x = mean(diamonds$carat), y = mean(diamonds$price), color = "red", size = 2) +
  annotate("point", x = .4631, y = 1358, color = "red", size = 2) +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)") +
  geom_abline(slope = 7689.65, intercept = -2203.0767779328, color = "red") +
  coord_cartesian(xlim = c(0,4))
```

\center
**What are the strengths and weaknesses of this two-point estimator of the 'general tendency' line?**




## {.plain}
\center
\footnotesize
Two-Point Line: Price = -2,203 + 7,689 (Carats)

\normalsize
**OLS Line: Price = -2,256 + 7,756 (Carats)**

```{r, fig.align = 'center', fig.asp=.7, out.width = '55%', fig.width = 4.5, cache=TRUE}
model1 <- lm(data = diamonds, price ~ carat)
preds1 <- ggeffects::ggpredict(model1, "carat")

diamonds |>
  ggplot(aes(x = carat)) +
  geom_point(alpha = .05, aes(y = price)) +
  #geom_abline(slope = 7689.65, intercept = -2203.0767779328, color = "red", size = .8) +
  geom_ribbon(data = preds1, aes(x = x, ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line(data = preds1, aes(x = x, y = predicted), color = "royalblue1", size = 1.2) +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)") +
  coord_cartesian(ylim = c(0, 19000), xlim = c(0,4))
```

\center
**Per the readings, how does OLS fit a line to the data?**





## {.plain}
\center
**Price = -2,256 + 7,756 (Carats)**

```{r, fig.align = 'center', fig.asp=.7, out.width = '55%', fig.width = 4.5, cache=TRUE}

set.seed(277)
d2 <- diamonds |>
  slice_sample(prop = .002)

d2 |>
  modelr::add_predictions(model1) |> 
  ggplot(aes(x = carat)) +
  geom_point(aes(y = price)) +
  geom_ribbon(data = preds1, aes(x = x, ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line(data = preds1, aes(x = x, y = predicted), color = "royalblue1", size = 1.2) +
  geom_segment(aes(xend = carat, y = price, yend = pred), color = "red") +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)") +
  coord_cartesian(ylim = c(0, 19000), xlim = c(0,4))
```

\center
**OLS finds the line of best fit that minimizes the squared residuals of ALL THE POINTS.**



## {.plain}
\center
What is the predicted price of a diamond that weighs 1.5 carats?

```{r, fig.align = 'center', fig.asp=.618, out.width = '60%', fig.width = 4, cache=TRUE}
diamonds |>
  ggplot(aes(x = carat)) +
  geom_point(alpha = .05, aes(y = price)) +
  geom_ribbon(data = preds1, aes(x = x, ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line(data = preds1, aes(x = x, y = predicted), color = "royalblue1", size = 1.2) +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)") +
  coord_cartesian(ylim = c(0, 19000), xlim = c(0,4))
```

\large
\center
Price = -2,256.36 + 7,756.43 x Carats



## {.plain}
\center
What is the predicted price of a diamond that weighs 1.5 carats?

```{r, fig.align = 'center', fig.asp=.618, out.width = '60%', fig.width = 4, cache=TRUE}
diamonds |>
  ggplot(aes(x = carat)) +
  geom_point(alpha = .05, aes(y = price)) +
  geom_ribbon(data = preds1, aes(x = x, ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line(data = preds1, aes(x = x, y = predicted), color = "royalblue1", size = 1.2) +
  annotate("point", x = 1.5, y = 9378.285, color = "red", size = 3) +
  annotate("segment", x = 1.5, xend = 1.5, y = 0, yend = 9378, color = "red", linetype = "dashed", size = 1) +
  annotate("segment", x = 0, xend = 1.5, y = 9378, yend = 9378, color = "red", linetype = "dashed", size = 1) +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)") +
  coord_cartesian(ylim = c(0, 19000), xlim = c(0,4))

##predict.lm(model1, newdata = data.frame("carat" = 1.5), interval = "confidence")
```

\center
Price = -2,256.36 + 7,756.43 x 1.5 = \$9,378.29

(95\% confidence interval: \$9,354 - \$9,401)



## {.plain}
\center
What is the predicted price of a diamond that weighs 2.5 carats?

```{r, fig.align = 'center', fig.asp=.618, out.width = '60%', fig.width = 4, cache=TRUE}
diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)") +
  coord_cartesian(ylim = c(0, 18000), xlim = c(0,4))
```

\large
\center
Price = -2,256.36 + 7,756.43 x Carats



## {.plain}
\center
What is the predicted price of a diamond that weighs 2.5 carats?

```{r, fig.align = 'center', fig.asp=.618, out.width = '60%', fig.width = 4, cache=TRUE}
diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  geom_smooth(method = "lm") +
  annotate("point", x = 2.5, y = 17134.72, color = "red", size = 3) +
  annotate("segment", x = 2.5, xend = 2.5, y = 0, yend = 17134.72, color = "red", linetype = "dashed", size = 1) +
  annotate("segment", x = 0, xend = 2.5, y = 17134.72, yend = 17134.72, color = "red", linetype = "dashed", size = 1) +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)") +
  coord_cartesian(ylim = c(0, 18000), xlim = c(0,4))

##predict.lm(model1, newdata = data.frame("carat" = 2.5), interval = "confidence")
```

\center
Price = -2,256.36 + 7,756.43 x 2.5 = \$17,134.72

(95\% confidence interval: \$17,085 - \$17,183)



## {.plain}
\center
Regression lets us answer the question: What is the average price difference between 1.5 and 2.5 carat diamonds?

```{r, fig.align = 'center', fig.asp=.618, out.width = '60%', fig.width = 4, cache=TRUE}
diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  geom_smooth(method = "lm") +
  annotate("point", x = 2.5, y = 17134.72, color = "red", size = 3) +
  annotate("point", x = 1.5, y = 9378.29, color = "red", size = 3) +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)") +
  coord_cartesian(ylim = c(0, 18000), xlim = c(0,4))
```

\center
\large

\$17,134.72 - \$9,378.29 = \$7,756.43 (on average)



## An OLS Estimate of Diamond Prices {.plain}
::: columns
:::: column
\vspace{.25in}

```{r, fig.align = 'center', fig.asp=.75, out.width = '92%', fig.width = 4, cache=TRUE}
diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)") +
  coord_cartesian(ylim = c(0, 18000), xlim = c(0,4))
```

::::
:::: column

```{r, results='asis'}
model1 <- lm(data = diamonds, price ~ carat)

stargazer(model1, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", dep.var.labels = "Price", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser"), font.size = "footnotesize", float = FALSE, covariate.labels = "Carats")
```

::::
:::
\begin{center}
Y = $\alpha$ + $\beta$ X

Price = -2,256.36 + 7,756.43 x Carats
\end{center}



## Why use OLS regressions?
\large

### Basics
\vspace{.15in}
+ Quantifies the relationship between variables
\vspace{.15in}
+ Can be used to make predictions
\vspace{.15in}
+ Provides estimates of uncertainty



## Why use OLS regressions?
\large

### Basics
+ Quantifies the relationship between variables

+ Can be used to make predictions

+ Provides estimates of uncertainty

### Future Weeks
+ Can be adjusted for confounders (e.g. control variables), nonlinear relationships and different data structures



##
\center
\huge
**Part 2**

\vspace{.25in}

**How do we fit a regression and make predictions in R?**



## Fitting OLS Regressions in R
```{r, echo = TRUE, eval = FALSE}
## Fitting an OLS model
## lm(data = dataset, outcome ~ predictor)

## Regress `price` on `carat` in the `diamonds` dataset
model1 <- lm(data = diamonds, price ~ carat)

## Check the Results
summary(model1)

## Fancy Table (use type = "html" in our reports)
stargazer::stargazer(model1, type = "text")
```


## Making Predictions (Approach 1)
```{r, echo=TRUE}
## Create dataset with values of the predictors
new1 <- tibble(carat = 1.5)

## Prediction Function
predict(model1, newdata = new1, interval = "confidence")
```


## Making Predictions (Approach 1)
```{r, echo=TRUE, eval=FALSE}
## Create dataset with values of the predictors
new1 <- tibble(carat = 1.5)

## Prediction Function
predict(model1, newdata = new1, interval = "confidence")
```
\vspace{.25in}

\begin{center}
\large
\textbf{Redo the above but calculate the predictions for carat values of 1, 2, 3 and 4 simultaneously.}
\end{center}




## Making Predictions (Approach 1)
\small
```{r, echo=TRUE}
## Create dataset with values of the predictors
new2 <- tibble(carat = c(1, 2, 3, 4))

## Prediction Function
predict(model1, newdata = new2, interval = "confidence")
```


## 
```{r, echo=TRUE}
## Making Predictions (Approach 2) 
ggeffects::ggpredict(model1, terms = "carat")
```


## {.plain}
::: columns
:::: column
\footnotesize

```{r, echo=TRUE, eval=FALSE}
## Visualize the Predictions (Approach 2) 
ggpredict(model1, terms = "carat") |>
  ggplot(aes(x = x, y = predicted)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line() +
  theme_bw() +
  labs(x = "Carats", y = "Predicted Price")
```

::::
:::: column

\vspace{1.25in}

```{r, echo=FALSE, fig.align = 'center', fig.asp=.75, out.width = '100%', fig.width = 4, cache=TRUE}
## Visualize the Predictions (Approach 2) 
ggpredict(model1, terms = "carat") |>
  ggplot(aes(x = x, y = predicted)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line() +
  theme_bw() +
  labs(x = "Carats", y = "Predicted Price")
```

::::
:::


##
\center
\huge
**Part 3**

\vspace{.25in}

**How do we evaluate the fit of a regression model?**


## 1. Is it logical?
::: columns
:::: column
\vspace{.1in}

\begin{center}
```{r, echo = FALSE, results = "asis", eval=TRUE}
stargazer(model1, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", dep.var.labels = "Price", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser"), font.size = "footnotesize", float = FALSE, covariate.labels = "Carats")
```
\end{center}

::::
:::: column

\center
\includegraphics[keepaspectratio=true,width=.31\paperwidth]{Images/11_1-Diamonds_carats.jpg}

::::
:::



## 2. Are the coefficients significant?

\begin{center}
```{r, echo = FALSE, results = "asis", eval=TRUE}
stargazer(model1, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", dep.var.labels = "Price", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser"), font.size = "small", float = FALSE, covariate.labels = "Carats")
```
\end{center}



## What is Statistical Significance?
\large

$H_{A}$ = Larger diamonds are more expensive than smaller ones

\center
\includegraphics[keepaspectratio=true,width=.75\paperwidth]{Images/11_1-Diamonds_2.jpg}

$H_{0}$ = Diamond size has no impact on price


## What is Statistical Significance? {.plain}
```{r, fig.show="hold", out.width="48%", fig.width=4.5, fig.height=4.5, cache=TRUE}
diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)",
       title = "Alternative Hypothesis",
       subtitle = str_c("Sum of Squared Errors: ", round(sum((model1$residuals)^2)/1e9, 1), " billion")) +
  coord_cartesian(ylim = c(0, 18000), xlim = c(0,4))

model0 <- lm(data = diamonds, price ~ 1)

diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = .05) +
  geom_hline(yintercept = mean(diamonds$price), color = "red") +
  theme_bw() +
  labs(x = "Carats", y = "Price ($)",
       title = "Null Hypothesis",
       subtitle = str_c("Sum of Squared Errors: ", round(sum((model0$residuals)^2)/1e9, 1), " billion")) +
  coord_cartesian(ylim = c(0, 18000), xlim = c(0,4))
```


## Are deeper diamonds more expensive? {.plain}
```{r, fig.show="hold", out.width="48%", fig.width=4.5, fig.height=4.5, cache=TRUE}
diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = depth, y = price)) +
  geom_point(alpha = .05) +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(x = "Depth", y = "Price ($)",
       title = "Alternative Hypothesis")

#model0 <- lm(data = diamonds, price ~ 1)

diamonds |>
  slice_sample(prop = .1) |>
  ggplot(aes(x = depth, y = price)) +
  geom_point(alpha = .05) +
  geom_hline(yintercept = mean(diamonds$price), color = "red") +
  theme_bw() +
  labs(x = "Depth", y = "Price ($)",
       title = "Null Hypothesis")
```



## 3. How high is an adjusted R$^2$ of 0.85?
```{r, fig.align = 'center', fig.asp = 0.618, fig.width = 7, out.height = '85%', eval=TRUE, cache=TRUE}
set.seed(50)
d1 <- tibble(
    x = rnorm(50, 12, 3),
    y1 = x + rnorm(50, 15, 8),
    y2 = x + rnorm(50, 15, 5),
    y3 = x + rnorm(50, 15, 1.5),
    y4 = x + rnorm(50, 15, .5)    
)

d2 <- d1 |>
pivot_longer(cols = y1:y4, names_to = "Version", values_to = "Values") |>
mutate(
    Version = case_when(
        Version == "y1" ~ "R2 = 0",
        Version == "y2" ~ "R2 = 0.24",
        Version == "y3" ~ "R2 = 0.81",
        Version == "y4" ~ "R2 = 0.97")
    )

ggplot(data = d2, aes(x = x, y = Values)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    theme_bw() +
    facet_wrap(~ Version)

```


## {.plain}
\vspace{.2in}

```{r, fig.align = 'center', echo=TRUE, out.width='66%', fig.width=6, fig.asp=0.7}
## 4. Any problems in the residuals plot?
plot(model1, which = 1)
```


## Heteroskedastic Errors are Bad
\center
\includegraphics[keepaspectratio=true,width=.85\paperwidth]{Images/09_1-heteroskedastic_errors.png}



## Non-Linear Errors are Bad
\center
\includegraphics[keepaspectratio=true,width=.7\paperwidth]{Images/09_1-nonlinear_errors.png}



## {.plain}
\vspace{.2in}

```{r, fig.align = 'center', echo=TRUE, out.width='66%', fig.width=6, fig.asp=0.7}
## 4. Any problems in the residuals plot?
plot(model1, which = 1)
```


## Evaluating the Fit of an OLS Regression
\Large
\begin{enumerate}
\item Is it logical?
\vspace{.1in}
\item Are the coefficients significant?
\vspace{.1in}
\item How high is the adjusted R$^2$
\vspace{.1in}
\item Any problems in the residuals plot?
\end{enumerate}




## Let's practice using the `mpg` dataset
\large

**Does engine size (`displ`) explain fuel economy (`cty`)?**

1. Scatter plot: `cty` x `displ`

2. Regress `cty` on `displ`

3. Use ggpredict() to draw the marginal effect of engine size on fuel economy

4. Evaluate the fit of the model



## Explaining Fuel Economy
::: columns
:::: column

```{r, fig.align = 'center', fig.asp=.75, out.width = '90%', fig.width = 4, cache=TRUE}
mpg |>
  ggplot(aes(x = displ, y = cty)) +
  geom_point() +
  #geom_smooth(method = "lm") +
  theme_bw() +
  labs(x = "Engine Displacement", y = "Fuel Economy (city mpg)")
```

::::
:::: column

```{r, results='asis'}
model2 <- lm(data = mpg, cty ~ displ)

stargazer(model2, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", dep.var.labels = "Fuel Economy", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser"), font.size = "footnotesize", float = FALSE, covariate.labels = "Engine Displacement")
```
::::
:::



## {.plain}
\scriptsize
```{r, fig.align = 'center', fig.asp=.618, out.width = '60%', fig.width = 5, cache=FALSE, echo=TRUE}
ggeffects::ggpredict(model2, terms = "displ") |>
  ggplot(aes(x = x, y = predicted)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line() +
  theme_bw() +
  labs(x = "Engine Size (litres)", y = "Predicted Fuel Economy")
```


## 1. Is it logical?
::: columns
:::: column
\vspace{.1in}

\begin{center}
```{r, echo = FALSE, results = "asis", eval=TRUE}
stargazer(model2, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", dep.var.labels = "Fuel Economy", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser"), font.size = "footnotesize", float = FALSE, covariate.labels = "Engine Displacement") 
```
\end{center}

::::
:::: column

\center
\includegraphics[keepaspectratio=true,width=.3\paperwidth]{Images/11_1-diesel-engines.jpg}

::::
:::



## 2. Are the coefficients significant?

\begin{center}
```{r, echo = FALSE, results = "asis", eval=TRUE}
stargazer(model2, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", dep.var.labels = "Fuel Economy", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser"), font.size = "small", float = FALSE, covariate.labels = "Engine Displacement") 
```
\end{center}


## 3. How high is an adjusted R$^2$ of 0.64?
```{r, fig.align = 'center', fig.asp = 0.618, fig.width = 7, out.height = '85%', eval=TRUE, cache=TRUE}
set.seed(50)
d1 <- tibble(
    x = rnorm(50, 12, 3),
    y1 = x + rnorm(50, 15, 8),
    y2 = x + rnorm(50, 15, 5),
    y3 = x + rnorm(50, 15, 1.5),
    y4 = x + rnorm(50, 15, .5)    
)

d2 <- d1 |>
pivot_longer(cols = y1:y4, names_to = "Version", values_to = "Values") |>
mutate(
    Version = case_when(
        Version == "y1" ~ "R2 = 0",
        Version == "y2" ~ "R2 = 0.24",
        Version == "y3" ~ "R2 = 0.81",
        Version == "y4" ~ "R2 = 0.97")
    )

ggplot(data = d2, aes(x = x, y = Values)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    theme_bw() +
    facet_wrap(~ Version)

```


## {.plain}
\vspace{.2in}

```{r, fig.align = 'center', echo=TRUE, out.width='66%', fig.width=6, fig.asp=0.7}
## 4. Any problems in the residuals plot?
plot(model2, which = 1)
```


## 
\huge

\begin{center}
Redo the regression but using the highway fuel economy variable (`hwy`) as the outcome.
\end{center}


##
\begin{center}
```{r, echo = FALSE, results = "asis", eval=TRUE}
model3 <- lm(data = mpg, hwy ~ displ)

stargazer(model2, model3, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", dep.var.labels = c("Fuel Economy (City)", "Fuel Economy (Highway)"), star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser"), font.size = "footnotesize", float = FALSE, covariate.labels = "Engine Displacement") 
```
\end{center}


## OLS Regression Tools in R
\large

+ lm(): Fit the regression

+ tibble(): Make a new dataset

+ predict(): Make predictions using the regression model

+ ggeffects::ggpredict(): Same as above

+ stargazer::stargazer(): Make a fancy table

+ plot(model1, which = 1): Check the residuals