---
format: revealjs
---

## Today's Agenda {background-image="Images/background-data_blue_v4.png" .center}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
library(ggeffects)

# Input practice WDI data
wdi <- read_excel("../Data_in_Class/Class11-1-Simple_OLS-WDI_LifeExpectancy/WDI-Practice_Simple_OLS-2024-04-04.xlsx", na = "NA")
```

<br>

::: {.r-fit-text}

**Ordinary Least Squares (OLS) Regression**

- Practice fitting, interpreting and evaluating OLS regressions

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2025)
:::

::: notes
Prep for Class

1. Check Canvas submissions

<br>

**SLIDE**: Why do we use regressions?

:::



## Why use OLS regressions? {background-image="Images/background-data_blue_v4.png" .center}

<br>

- Quantifies the relationship between variables

- Uses **ALL** of the data

- Makes predictions with estimates of uncertainty

- Gives us criteria for evaluating the fit of the line

**Future Weeks**

- Can be adjusted for confounders (e.g. control variables), nonlinear relationships and different data structures

::: notes

**Any questions on our work last week on regression?**

- **e.g. the intuitions, interpretations, evaluations or methods in R?**

<br>

**SLIDE**: Let's review your assignments

:::



## Assignment For Today {background-image="Images/background-data_blue_v4.png" .center}

<br>

What is the effect of college completion on poverty rates in midwest cities?

1. Fit the regression

2. Submit a polished regression table, residuals plot and predictions plot

3. Answer the question

4. Evaluate the regression

::: notes

**How did this go?**

<br>

**Toughest part?**

<br>

*Save your conclusion for next slide*

- **Ok, what's your answer to the question: What is the effect of college completion on poverty rates in midwest cities?**

<br>

**SLIDE**: My results and the evaluation tools

:::



## {background-image="Images/background-data_blue_v4.png"}

::::: {.columns}

:::: {.column width="35%"}

<br>

```{r, echo=FALSE}
res1 <- lm(data = midwest, percbelowpoverty ~ percollege)

modelsummary(list("Poverty (%)" = res1), 
             output = "gt", fmt = 2, 
             stars = c('*' = .05), gof_omit = "IC|Log|F",
             coef_rename = c("percollege"= "College Degree (%)")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

::::

:::: {.column width="65%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=6}
plot(ggpredict(res1, terms = "percollege")) +
  labs(x = "College Degree (%)",
       y = "Poverty (%)",
       title = "Investments in college access and affordability may directly\nreduce poverty rates")
```

### Residuals

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=6}
plot(res1, which = 1)
```

:::

::::

:::::

::: notes

**And, based on your evaluation of the regression, how confident are you in that answer?**

```{r, echo=TRUE}
nrow(midwest)
```

<br>

*Make sure to discuss the seemingly small magnitude of the effect*

- **Are these moves in poverty a big deal or not?**

- **How hard would it be to increase college completion rates in meaningful ways?**

<br>

**SLIDE**: Our plan for today is to keep practicing all of these skills!

:::




## Global Life Expectancy {background-image="Images/background-data_blue_v4.png" .center}

```{r, fig.align='center', fig.asp=.618, fig.width=9}
wdi |>
  ggplot(aes(x = life_expectancy_total)) +
  geom_histogram(color = "white") +
  theme_bw() +
  labs(x = "Life Expectancy (Years)", y = "Count of Observations", caption = "Source: WDI 2020 data") +
  annotate("text", x = 82.3, y = .5, label = "Ireland", color = "white", srt = 90, hjust = 0) +
  annotate("text", x = 75.6, y = .5, label = "United States", color = "white", srt = 90, hjust = 0) +
  annotate("text", x = 71, y = .5, label = "Russia", color = "white", srt = 90, hjust = 0) +
  annotate("text", x = 62, y = .5, label = "Afghanistan", color = "white", srt = 90, hjust = 0) +
  scale_x_continuous(breaks = seq(50,90,5))

#View(d |> select(`Country Name`, life_expectancy_total))
```

::: notes

I've uploaded World Bank data to our Canvas modules for today

- Everybody grab that data!

<br>

Countries around the world vary dramatically in terms of their life expectancies

- Ireland: 82
- USA: 76
- Russia: 71
- Afghanistan: 62
- Nigeria: 53

<br>

Our job today is to use regressions to identify and evaluate possible answers to the question of why people in some countries tend to live longer than those in others

:::



## Regress life expectancy on... {background-image="Images/background-data_blue_v4.png" .center}

<br>

::::: {.columns}
:::: {.column width="50%"}
```{r, fig.align='left', fig.asp=.8, fig.width=6}
wdi |>
  ggplot(aes(x = life_expectancy_total)) +
  geom_histogram(color = "white") +
  theme_bw() +
  labs(x = "Life Expectancy (Years)", y = "Count of Observations", caption = "Source: WDI 2020 data") +
  scale_x_continuous(breaks = seq(50,90,5))
```
::::

:::: {.column width="50%"}

<br>

::: {.r-fit-text}
1. fertility_rate_per_woman

2. compulsory_education_yrs

3. unemployment_pct
:::
::::
:::::

::: notes
Work with the people around you to fit and evaluate three regressions of life expectancy

- Model 1 regresses life expectancy on the fertility rate per woman

- Model 2 regresses life expectancy on the number of years of education you are required to complete

- Model 3 regresses life expectancy on the proportion of adults who are unemployed

<br>

Your job:

1. Fit each model,

2. Interpret the coefficient, 

3. Make predictions using the model, and then 

4. Evaluate each model with our four steps

<br>

**Questions?**

- Get to it!

<br>

**SLIDE**: Fertility rate models

:::



## {background-image="Images/background-data_blue_v4.png"}

::::: {.columns}

:::: {.column width="40%"}

<br>

```{r, echo=FALSE}
res1 <- lm(data = wdi, life_expectancy_total ~ fertility_rate_per_woman)

modelsummary(list("Life Expectancy" = res1),
             out = "gt",
             fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log",
             coef_rename = c("fertility_rate_per_woman" = "Fertility (births per woman)", "tobacco_use_pct" = "Tobacco Use (%)", "compulsory_education_yrs" = "Compulsory Education (years)")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

::::

:::: {.column width="60%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(ggpredict(res1, terms = "fertility_rate_per_woman")) +
  labs(x = "Fertility (births per woman)",
       y = "Life Expectancy (years)")
```

### Residuals

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(res1, which = 1)
```

:::

::::

:::::

::: notes

**Did everybody get these results for the fertility rate regression?**

<br>

**Interpret the coefficient for me**

- As average births per woman increases by one, life expectancy decreases by almost 5 years!

<br>

**What is obvious from this set of predictions?**

- The effect is DRAMATICALLY larger

- ONE additional child is a 5 year decrease in life expectancy (on average)!

- The effect of fertility rate on life expectancy is staggeringly large!

<br>

**So, how well does the regression line fit the data?**

- **e.g. missing data, significance, R2, residuals**

<br>

**Any concerns with heteroskedasticity in the fertility rate regression?**

- Slight non-linearity here but the effect is rather minor

- Does appear that we have fewer cases at the low end of the scale (e.g. very high fertility rates) and so the model fits this extreme less well

<br>

**SLIDE**: Let's talk education

:::



## {background-image="Images/background-data_blue_v4.png"}

::::: {.columns}

:::: {.column width="40%"}

<br>

```{r, echo=FALSE}
res2 <- lm(data = wdi, life_expectancy_total ~ compulsory_education_yrs)

modelsummary(list("Life Expectancy" = res2),
             out = "gt",
             fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log",
             coef_rename = c("fertility_rate_per_woman" = "Fertility (births per woman)", "tobacco_use_pct" = "Tobacco Use (%)", "compulsory_education_yrs" = "Compulsory Education (years)")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

::::

:::: {.column width="60%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(ggpredict(res2, terms = "compulsory_education_yrs")) +
  labs(x = "Compulsory Education (years)",
       y = "Life Expectancy (years)")
```

### Residuals

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(res2, which = 1)
```

:::

::::

:::::

::: notes

**Did everybody get these results for the education regression?**

<br>

**Interpret the coefficient for me**

- Add one year of compulsory education and see an average increase of life expectancies by almost one year

<br>

**What is obvious from this set of predictions?**

- A country with ZERO required school that implements a primary education requirement will see, on average, an improvement of 4 years of life expectancy
    - 0 to 6 yrs education = 69.64-65.31 = 4.33
    
- A country moving from a primary only requirement to a secondary ed requirement will see, on average, an improvement of 4 years of life expectancy

- Do both and add 8 years!

<br>

**So, how well does the regression line fit the data?**

- **e.g. missing data, significance, R2, residuals**

<br>

**Any concerns with heteroskedasticity in the education regression?**

- This residual plot shows the flaw in just relying on the red line because it's tracking changes across the levels that are probably just scattered randomly

- Some evidence of heteroskedasticity: At the low end error is high, at the high end it's the opposite

- Probably not a deal breaker but not a clear case of homoscedasticity

<br>

**So, if you ran a development agency would you get more bang from your buck focusing on women's rights and healthcare or increasing compulsory education?**

<br>

**SLIDE**: Let's talk unemployment

:::



## {background-image="Images/background-data_blue_v4.png"}

::::: {.columns}

:::: {.column width="40%"}

<br>

```{r, echo=FALSE}
res3 <- lm(data = wdi, life_expectancy_total ~ unemployment_pct)

modelsummary(list("Life Expectancy" = res3),
             out = "gt",
             fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log",
             coef_rename = c("fertility_rate_per_woman" = "Fertility (births per woman)", "unemployment_pct" = "Unemployment Rate (%)", "compulsory_education_yrs" = "Compulsory Education (years)")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

::::

:::: {.column width="60%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(ggpredict(res3, terms = "unemployment_pct")) +
  labs(x = "Unemployment Rate (%)",
       y = "Life Expectancy (years)")
```

### Residuals

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(res3, which = 1)
```

:::

::::

:::::

::: notes

**Did everybody get these results for the unemployment regression?**

<br>

**Interpret the coefficient for me**

- As the unemployment rate increases by 1%, average life expectancy decreases .12 years
    
    - So, add 8% unemployment and shave one year off of average life expectancies!

<br>

**What is obvious from this set of predictions?**

- Negative but the effect magnitude is TINY

<br>

**So, how well does the regression line fit the data?**

- **e.g. missing data, significance, R2, residuals**

<br>

- Some missing data, insignificant coefficient, some problems in the residuals plot

<br>

**SLIDE**: Put them all together!

:::



## Models of Global Life Expectancy {background-image="Images/background-data_blue_v4.png" .center .smaller}

```{r}
modelsummary(models = list(res1, res2, res3),
             out = "gt",
             fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log",
             coef_rename = c("fertility_rate_per_woman" = "Fertility (births per woman)", "unemployment_pct" = "Unemployment Rate (%)", "compulsory_education_yrs" = "Compulsory Education (years)")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "large")
  ), locations = gt::cells_body())

```

::: notes

**Ok, where does this leave us?**

- **Is one model more useful than the others? Why?**

<br>

Canvas assignment for next class

- What is the effect of tobacco use on average life expectancies around the world?

- Details on Canvas

:::







