---
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    df_print: kable
fontsize: 14pt
classoption: "aspectratio=169"
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "", output_file = "")

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)

d <- read_csv("../Data_Projects/04-Research_Project/P296-FA21-Research_Project_Data.csv", na = c(-9, "NA"))

d2 <- d |>
  arrange(Country, year) |>
  group_by(Country) |>
  mutate(
    milper_prop = (milper / tpop)*100,
    milper_prop_lag1 = lag(milper_prop, n = 1),
    milper_prop_lag3 = lag(milper_prop, n = 3)
  ) |>
  ungroup()
```


## Today's Agenda
\Large

Build simple OLS regressions for your research project report.

\vspace{.5in}

\center
Justin Leinaweaver (Fall 2021)



## Assignment 3: The Research Report {.plain}

1. Introduction
   + What is the question, why is it important and what is your argument?

\vspace{.1in}

2. **Bivariate Analyses**
   + **What is the answer to the question using only bivariate visualizations and simple OLS regressions?**

\vspace{.1in}

3. Multivariate Analyses
   + What is the answer to the question using multivariate visualizations and multiple OLS regressions?

\vspace{.1in}

4. Conclusion
   + What did you find and how confident should we be in your results?
   
   

## Research Project: Simple OLS Regressions
\large

Fit, evaluate and make predictions for the following regressions:

1. Regress `fh_total` on `milper_prop`

2. Regress `pr` on `milper_prop`

3. Regress `cl` on `milper_prop`

\vspace{.25in}

Then, **rebuild each model** with a one period time lag on `milper_prop` (`milper_prop_lag1`)




## Adding a Time Lag to a Variable
\footnotesize
```{r, echo=TRUE, eval=FALSE}
## This code is already in the project Rmd file
d2 |> 
  mutate(
    milper_prop_lag1 = lag(milper_prop, n = 1)
  )
```

```{r, echo=FALSE}
d2 |> 
  slice(6:11) |>
  select(Country, year, fh_total, milper_prop) |>
  mutate(
    milper_prop_lag1 = lag(milper_prop, n = 1)
  )
```



## To Start
\Large

Everybody fit, evaluate and make a marginal effects plot for the first model.

\vspace{.3in}

1. Regress `fh_total` on `milper_prop`



##
::: columns
:::: column

\begin{center}
```{r, results='asis'}
res1a <- lm(data = d2, fh_total ~ milper_prop)

stargazer(res1a, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser", "f"), font.size = "footnotesize", float = FALSE, covariate.labels = "Military Personnel (\\% Pop)", dep.var.labels = c("FH Total"))
```
\end{center}

::::
:::: column
\large

\vspace{.25in}

\begin{enumerate}
\item Logical?
\vspace{.2in}
\item Coefficients significant?
\vspace{.2in}
\item Adjusted R$^2$
\vspace{.2in}
\item Residuals?
\end{enumerate}

::::
:::



##
```{r, fig.asp=.9, fig.width=6, out.width='55%', fig.align='center'}
plot(res1a, which = 1)
```





##
::: columns
:::: column

\begin{center}
```{r, results='asis'}
stargazer(res1a, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser", "f"), font.size = "footnotesize", float = FALSE, covariate.labels = "Military Personnel (\\% Pop)", dep.var.labels = c("FH Total"))
```
\end{center}

::::
:::: column

```{r, fig.asp=0.8, fig.width=4, out.width='85%', fig.align='center'}
preds1a <- ggeffects::ggpredict(res1a, terms = "milper_prop")

preds1a |>
  ggplot(aes(x = x, y = predicted)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line() +
  theme_bw() +
  labs(x = "Military Personnel (% Population)", y = "Predicted Freedom House Score") 
```

::::
:::



##
::: columns
:::: column

\begin{center}
```{r, results='asis'}
stargazer(res1a, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser", "f"), font.size = "footnotesize", float = FALSE, covariate.labels = "Military Personnel (\\% Pop)", dep.var.labels = c("FH Total"))
```
\end{center}

::::
:::: column

```{r, fig.asp=0.8, fig.width=4, out.width='85%', fig.align='center'}
preds1a |>
  ggplot(aes(x = x, y = predicted)) +
  geom_hline(yintercept = 14, color = "red", size = 2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line() +
  theme_bw() +
  labs(x = "Military Personnel (% Population)", y = "Predicted Freedom House Score") 

## What proportion of cases above 5%?
# d2 |>
#   filter(milper_prop > 5)
# 
# nrow(d2)
# 
# table(d2$milper_prop > 5)
```

::::
:::



##
::: columns
:::: column

\begin{center}
```{r, results='asis'}
stargazer(res1a, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser", "f"), font.size = "footnotesize", float = FALSE, covariate.labels = "Military Personnel (\\% Pop)", dep.var.labels = c("FH Total"))
```
\end{center}

::::
:::: column

```{r, fig.asp=0.8, fig.width=4, out.width='85%', fig.align='center'}
ggplot(data = d2, aes(x = milper_prop, y = fh_total)) +
  geom_point() +
  theme_bw() +
  geom_ribbon(data = preds1a, aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line(data = preds1a, aes(x = x, y = predicted)) +
  theme_bw() +
  labs(x = "Military Personnel (% Population)", y = "Predicted Freedom House Score") 
```

::::
:::

\vspace{.2in}
\center
Fortunately, only 23 / 9k cases have milper_prop values above 5%



## Research Project: Simple OLS Regressions
\large

Fit, evaluate and make predictions for the following regressions:

1. Regress `fh_total` on `milper_prop`

2. Regress `pr` on `milper_prop`

3. Regress `cl` on `milper_prop`

\vspace{.25in}

Then, **rebuild each model** with a one period time lag on `milper_prop` (`milper_prop_lag1`)



##
\begin{center}
```{r, results='asis'}
res1a <- lm(data = d2, fh_total ~ milper_prop)
res1b <- lm(data = d2, pr ~ milper_prop)
res1c <- lm(data = d2, cl ~ milper_prop)

stargazer(res1a, res1b, res1c, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser"), font.size = "small", float = FALSE, covariate.labels = "Military Personnel (\\% Pop)", dep.var.labels = c("FH Total", "Political Rights", "Civil Liberties"))
```
\end{center}



## 
\begin{center}
```{r, results='asis'}
res2a <- lm(data = d2, fh_total ~ milper_prop_lag1)
res2b <- lm(data = d2, pr ~ milper_prop_lag1)
res2c <- lm(data = d2, cl ~ milper_prop_lag1)

stargazer(res1a, res2a, res1b, res2b, res1c, res2c, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser", "f"), font.size = "footnotesize", float = FALSE, covariate.labels = c("Military Personnel (\\% Pop)", "Time lag = 1"), dep.var.labels = c("FH Total", "Political Rights", "Civil Liberties"))

#stargazer(res1a, res1b, res1c, type = "latex", digits = 2, header = FALSE, dep.var.caption = "", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, omit.stat = c("rsq", "ser"), font.size = "small", float = FALSE, covariate.labels = "Military Personnel (\\% Pop)", dep.var.labels = c("FH Total", "Political Rights", "Civil Liberties"))
```
\end{center}


## {.plain}
```{r, fig.asp=0.618, fig.width=5, out.width='35%', fig.align='center'}
preds1a <- ggeffects::ggpredict(res1a, terms = "milper_prop")

preds1a |>
  ggplot(aes(x = x, y = predicted)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line() +
  theme_bw() +
  labs(x = "Military Personnel (% Population)", y = "Predicted Freedom House Score") 
```

```{r, echo=FALSE, fig.show="hold", out.width="35%", fig.width=4, fig.height=3, fig.align='center'}
preds1b <- ggeffects::ggpredict(res1b, terms = "milper_prop")

preds1b |>
  ggplot(aes(x = x, y = predicted)) +
  #geom_hline(yintercept = 7, color = "red", size = 2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line() +
  theme_bw() +
  labs(x = "Military Personnel (% Population)", y = "Predicted Political Rights") 


preds1c <- ggeffects::ggpredict(res1c, terms = "milper_prop")

preds1c |>
  ggplot(aes(x = x, y = predicted)) +
  #geom_hline(yintercept = 7, color = "red", size = 2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightblue") +
  geom_line() +
  theme_bw() +
  labs(x = "Military Personnel (% Population)", y = "Predicted Civil Liberties") 
```



## {.plain}
```{r, echo=FALSE, fig.show="hold", out.width="38%", fig.width=5, fig.height=4, fig.align='center'}
plot(res1a, which = 1)

plot(res1b, which = 1)
```


```{r, echo=FALSE, fig.align='center', out.width="34%", fig.height=3.5, fig.width=4.5}
plot(res1c, which = 1)
```



## For Tuesday
\Large

1. Wheelan chapter 5 "Probability" 

\vspace{.3in}

2. Wheelan chapter 6 "Problems with Probability




