---
format: revealjs
---

# Today's Agenda {background-image="libs/Images/background-data_blue_v3.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
library(ggeffects)
```

<br>

<br>

**Working on Report 3**

- Scatterplots, correlations and simple OLS regressions

<br>

<br>

::: r-stack
Justin Leinaweaver (Spring 2024)
:::

::: notes
Prep for Class

1. TBD
:::



## Report 3 {background-image="libs/Images/background-slate_v2.png" .center}

**Bivariate Test of your Hypotheses**

<br>

**Submission Requirements**

- Submit a **pdf** to **Canvas** (due **Apr 5th**)

- Include an **APA** formatted bibliography

- Polish **all** included visualizations and tables

- Lead with **arguments** NOT evidence

::: notes

Ok, this week we work on your second report.

- The basic requirements same as last time

- Late penalty (-10) points per day
:::



## Report 3 {background-image="libs/Images/background-slate_v2.png" .center}

**Bivariate Test of your Hypotheses**

<br>

**Are powerful states more likely to respect the human rights of their citizens?**

1. Analyze the relationship between a country's `cinc` score (CoW NMC) and its measured level of respect for human rights and the rule of law (FFP) **using a scatter plot, a correlation and a simple OLS regression**. 

2. Repeat the above using the CoW NMC **indicator you were assigned in Report 2** instead of the `cinc` score.

::: notes
Here's what we worked on last class

- **Does everybody have these basic requirements completed (e.g. two scatter plots, correlations and OLS regressions)?**

<br>

In order to make an argument that answers the research question in this report you will need to turn to analyzing and interpreting the work you did last class.

<br>

**SLIDE**: Let's briefly start off today by discussing the elements of an effective quantitative argument
:::



## Making a Quantitative Argument {background-image="libs/Images/background-slate_v2.png" .center}

<br>

::: {.r-fit-text}
Write your report so that any interested reader can both:

1. Understand your argument, and

2. Get something of value from **EVERY** statistical tool you use
:::

::: notes

Point 1 means it is your job to lead EVERY paragraph with your ideas, not the data

- Topic sentences that explain conclusions are KEY to this exercise

<br>

Point 2 means it is your job, as the data analyst, to help the reader understand why these statistical tools are useful

<br>

**SLIDE**: Let's talk about how to do point 2 well
:::




## Making a Quantitative Argument {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. You must explain the tools before using them

```{r, fig.align='center', fig.asp=.7}
tibble(
  x1 = rnorm(1000, 0, 5)
) |>
  ggplot(aes(x = x1)) +
  geom_histogram(color = "white") +
  theme_bw() +
  labs(x = "", y = "Count of Observations")
```

::: notes
Include brief explanations of the tools you are using to support your argument BEFORE you use them to analyze the data

<br>

If I'm going to include a histogram then I need to BRIEFLY explain to the reader what it does and why it helps us.

- **Speaking of, why do we need histograms? How do they help us?**

<br>

A histogram lets us see how a variable is distributed across the observations and that helps us understand what the world looks like!

- Do all the observations have similar central values? Is this a case where the average is a useful summary of the data?

- How spread out are the observations? Is this a variable that captures substantial variation in the world?

- Any evidence of outliers?
:::



## Making a Quantitative Argument {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. You must explain the tools before using them

```{r, fig.align='center', fig.asp=.7}
tibble(
  x1 = rnorm(1000, 25, 5),
  y1 = x1*2 + rnorm(1000, 0, 9)
) |>
  ggplot(aes(x = x1,y = y1)) +
  geom_point() +
  theme_bw() +
  #geom_smooth(method = "lm") +
  labs(x = "Predictor", y = "Outcome")
  
```

::: notes

If I'm going to include a scatterplot then I need to BRIEFLY explain to the reader what they do and why they help us.

- **Speaking of, why do we need scatter plots? How do they help us?**

<br>

- See the relationship

- Identify outliers

- Look for trends and correlations

:::



## Making a Quantitative Argument {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. You must explain the tools before using them

:::: {.columns}
::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.7}
set.seed(72)
x1 <- rnorm(1000, 25, 5)
set.seed(120)
y1 <- x1*3 + rnorm(1000, 0, 9)

res1 <- lm(y1 ~ x1)

modelsummary(res1, fmt = 2, stars = c("*" = .05), gof_omit = "Lik|IC|F", output = "gt") |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```
:::

::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.7, fig.width=6}
plot(ggpredict(res1))
```
:::
::::

::: notes
If I'm going to include a regression then I need to BRIEFLY explain to the reader what they do and why they help us.

- **Why do we need OLS regressions? How do they help us?**

<br>

- Quantify the relationship

- Create estimates of uncertainty

- Make predictions about the future

<br>

Bottom line on this first point, you must explain to the reader why you are doing what you are doing in the paper.

- **Make sense?**

<br>

**SLIDE**: Second piece of advice
:::



## Making a Quantitative Argument {background-image="libs/Images/background-slate_v2.png" .center}

<br>

2. Use specific examples to illustrate statistical tools

```{r, fig.align='center', fig.asp=.7}
set.seed(72)
x1 <- rnorm(1000, 25, 5)
set.seed(120)
y1 <- x1*3 + rnorm(1000, 0, 9)

tibble(
  x1 = x1,
  y1 = y1
) |>
  ggplot(aes(x = x1,y = y1)) +
  geom_point() +
  theme_bw() +
  labs(x = "Predictor", y = "Outcome") +
  #geom_smooth(method = "lm") +
  annotate("text", x = 38, y = 140, label = "United\nStates", color = "blue2") +
  annotate("point", x = 38.5, y = 132, color = "blue", size = 3.5) +
  annotate("text", x = 10, y = 23, label = "Russia", color = "red") +
  annotate("point", x = 10.3, y = 30, color = "red", size = 3.5)
```

::: notes
Related to the idea of explaining the tools you are using is the necessity of translating your findings into more digestible terms.

<br>

For those trying to get their heads around a scatterplot, identifying specific observations can help immeasurably

- Reminds the reader what your observations are and provides context for each axis scale

<br>

Do this whenever possible.

- Try to ensure that the reader can quickly get to grips with what is being shown to them and why

- This is also why I require informative figure titles and captions!


<br>

**Make sense?**
:::



## Making a Quantitative Argument {background-image="libs/Images/background-slate_v2.png" .center}

<br>

2. Use specific examples to illustrate statistical tools

```{r, fig.align='center', fig.asp=.7}
set.seed(72)
x1 <- rnorm(1000, 25, 5)
set.seed(120)
y1 <- x1*3 + rnorm(1000, 0, 9)

tibble(
  x1 = x1,
  y1 = y1
) |>
  ggplot(aes(x = x1,y = y1)) +
  geom_point() +
  theme_bw() +
  labs(x = "Predictor", y = "Outcome") +
  geom_smooth(method = "lm") +
  annotate("text", x = 38, y = 140, label = "United\nStates", color = "blue2") +
  annotate("point", x = 38.5, y = 132, color = "blue", size = 3.5) +
  annotate("text", x = 10, y = 23, label = "Russia", color = "red") +
  annotate("point", x = 10.3, y = 30, color = "red", size = 3.5)
```

::: notes
This goes doubly for understanding regression coefficients!

- You should definitely report the slope of the line as that's vital, BUT

- You should also present that in the terms of the actual variables, e.g. as the predictor increases by one the outcome increases by three!

<br>

Even better, use observation examples to help the reader understand

- If Russia here at the bottom of the plot increased its predictor to the average (approximately 25), they would increase their outcome by X

- Always present stat results in human terms!

<br>

**Make sense?**

<br>

**SLIDE**: Last one...
:::




## Making a Quantitative Argument {background-image="libs/Images/background-slate_v2.png" .center}

<br>

3. Help the reader understand the uncertainty

:::: {.columns}
::: {.column width="40%"}
```{r, fig.align='center', fig.asp=.7}
res1 <- lm(y1 ~ x1)

modelsummary(res1, fmt = 2, stars = c("*" = .05), gof_omit = "Lik|IC|F", output = "gt") |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```
:::

::: {.column width="60%"}
```{r, fig.align='center', fig.asp=.7}
plot(res1, which = 1)
```
:::
::::

::: notes

Remember, a big part of your job as the analyst is to help the reader understand the uncertainty in the data.

<br>

In other words, how confident are you in the answer you are providing to the big question in the report? 

- What aspects of the research design and the sample you are analyzing give you confidence in the answer? 

- What aspects make you hesitant?

<br>

This is where all the good strengths and weaknesses work you did in the first two reports comes into play, AND

- where you discuss the strengths and weaknesses of the statistics you are providing as evidence

<br>

**Any questions on any of this advice?**
:::



## Report 3 {background-image="libs/Images/background-slate_v2.png" .center}

**Bivariate Test of your Hypotheses**

<br>

**Are powerful states more likely to respect the human rights of their citizens?**

1. Analyze the relationship between a country's `cinc` score (CoW NMC) and its measured level of respect for human rights and the rule of law (FFP) **using a scatter plot, a correlation and a simple OLS regression**. 

2. Repeat the above using the CoW NMC **indicator you were assigned in Report 2** instead of the `cinc` score.

::: notes

Alright, let's get back to work!

:::




























