---
format: revealjs
---

# Today's Agenda {background-image="Images/background-data_blue_v4.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
library(modelr)
library(ggeffects)

# Input merged data
d <- read_excel("../Data_in_Class/Data-2024-EPI_RSF-Merged.xlsx", na = "NA")

d2 <- read_excel("../Data_in_Class/Data-2015-2022-EPI_RSF-Merged.xlsx", na = "NA")
```

<br>

::: {.r-fit-text}

**Ordinary Least Squares (OLS) Regression**

- Applying OLS to our research project

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2025)
:::

::: notes
Prep for Class

1. Check Canvas submissions

<br>

**SLIDE**: Review assignments

:::



## For Today {background-image="Images/background-data_blue_v4.png" .center}

<br>

**Working on our Research Project (2024 Only)**

1. Regress Yale's EPI on RSF's Press Freedom Index

2. Regress one other EPI indicator on RSF's Press Freedom Index

<br>

Submit a polished regression table, residuals plot, predictions plot, and your analysis of the findings.

::: notes

**How did this go?**

- **Any big picture problems to discuss before we report back on your findings?**

<br>

**SLIDE**: Overall results

:::



## {background-image="Images/background-data_blue_v4.png"}

::::: {.columns}

:::: {.column width="40%"}

<br>

```{r, echo=FALSE}
res1 <- lm(data = d, EPI.new ~ Press_Global)

modelsummary(list("Environmental Quality" = res1),
             out = "gt",
             fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log",
             coef_rename = c("Press_Global" = "Press Freedom Index")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

::::

:::: {.column width="60%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7, cache=TRUE}
plot(ggpredict(res1, terms = "Press_Global")) +
  labs(x = "Press Freedom Index",
       y = "Environmental Quality")
```

### Residuals

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7, cache=TRUE}
plot(res1, which = 1)
```

:::
::::
:::::

::: notes

**Did everybody get these results?**

<br>

**What do we learn from these regression results?**

- *YOU be quiet, it's on them*

<br>

**Did anybody check to see if a squared term gives a better fit here?**

- (**SLIDE**: I hope so!!!!!!)

:::



## {background-image="Images/background-data_blue_v4.png"}

::::: {.columns}

:::: {.column width="40%"}

<br>

```{r, echo=FALSE}
res2 <- lm(data = d, EPI.new ~ Press_Global + I(Press_Global^2))

modelsummary(list("Environmental Quality" = res2),
             out = "gt",
             fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log",
             coef_rename = c("Press_Global" = "Press Freedom Index",
                             "I(Press_Global^2)" = "Press Freedom Index (Squared)")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

::::

:::: {.column width="60%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7, cache=TRUE}
plot(ggpredict(res2, terms = "Press_Global")) +
  labs(x = "Press Freedom Index",
       y = "Environmental Quality")
```

### Residuals

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7, cache=TRUE}
plot(res2, which = 1)
```

:::
::::
:::::

::: notes

Hot damn, that's cool.

<br>

Ok, next job, go check out the indicator regressions your classmates made for today

- Make note of any you think tell an interesting story!

- And call out great work!

- Go!

<br>

Ok, now I want everyone to remake their indicator regression BUT this time using the across time data

- e.g. the dataset with observations from 2015 - 2022

- Go!

<br>

**Ok, are the results of your across time analysis consistent with the 2024 results or not?** 

- **Examples to share?**

- *(Save controlling for year for multiple regression*)

:::



## Next Class {background-image="Images/background-data_blue_v4.png" .center}

<br>

**Causal Inference Week**

- Huntington-Klein 2022 chapter 5 "The Challenge of Identification"

::: notes

Next week we tackle some big, important ideas.

- Specifically, how do we make an argument about causality if all our methods so far have just been aimed at describing associations?

- Important material for social scientists AND for those heading off to Research Design and Senior Seminar!


:::
