---
output:
  xaringan::moon_reader:
    css: ["default", "extra.css"]
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
##xaringan::inf_mr()
## For offline work: https://bookdown.org/yihui/rmarkdown/some-tips.html#working-offline
##chakra: libs/remark-latest.min.js

## Images not appearing? Put images folder inside the libs folder as that is the main data directory

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
##library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

```{r setup, include=FALSE}
## Include this to render DT tables correctly until fix is released
options(htmltools.dir.version = FALSE, htmltools.preserve.raw = FALSE)
```

class: slideblue

.size60[**Today's Agenda**]

<br>

.size40[
**Multivariate Analyses**

1. Descriptive statistics by groups

2. Using facets to extend any visualization

3. Mapping 3rd variables with colors or shapes
]

.center[.size30[
  Justin Leinaweaver (Fall 2021)
]]

???

Last week: Research design, causal inference and identification strategies using DAGs

ANY QUESTIONS ON THIS MATERIAL BEFORE WE EXPLORE ADAPTING OUR ANALYSES TO THESE METHODS?




---

class: middle

.size50[**1. Descriptive statistics by groups**]

<br>

.code150[
```{r, echo = TRUE, eval = FALSE}
dataset |>
  group_by(Cat_Predictor1, Cat_Predictor2, ...) |> #<<
  summarize(
    N = n(),
    Mean = mean(Outcome),
    StdDev = sd(Outcome)
  )
```
]

???

You can always add more categorical variables to the group_by function



---

.size40[**1. Descriptive statistics by groups**]

.pull-left[
.code130[
```{r, echo = TRUE, eval = FALSE}
## Column order with group_by
diamonds |>
  group_by(color) |> #<<
  summarize(
    N = n(),
    Mean = mean(price),
    StdDev = sd(price)
  )
```
]]

.pull-right[

<br>

.size30[
```{r, echo = FALSE, eval = TRUE}
diamonds |>
  group_by(color) |>
  summarize(
    N = n(),
    Mean = mean(price),
    StdDev = sd(price),
  ) |>
  ungroup() |>
  slice(1:14) |>
  knitr::kable(digits=0)
```
]]



---

.size40[**1. Descriptive statistics by groups**]

.pull-left[
.code125[
```{r, echo = TRUE, eval = FALSE}
## Column order with group_by
diamonds |>
  group_by(color, cut) |> #<<
  summarize(
    N = n(),
    Mean = mean(price),
    StdDev = sd(price)
  )
```
]]

.pull-right[
```{r, echo = FALSE, eval = TRUE}
diamonds |>
  group_by(color, cut) |>
  summarize(
    N = n(),
    Mean = mean(price),
    StdDev = sd(price),
  ) |>
  ungroup() |>
  slice(1:14) |>
  knitr::kable(digits=0)
```
]



---

.size40[**1. Descriptive statistics by groups**]

.pull-left[
.code125[
```{r, echo = TRUE, eval = FALSE}
## Column order with group_by
diamonds |>
  group_by(cut, color) |> #<<
  summarize(
    N = n(),
    Mean = mean(price),
    StdDev = sd(price)
  )
```
]]

.pull-right[
```{r, echo = FALSE, eval = TRUE}
diamonds |>
  group_by(cut, color) |>
  summarize(
    N = n(),
    Mean = mean(price),
    StdDev = sd(price),
  ) |>
  ungroup() |>
  slice(1:14) |>
  knitr::kable(digits=0)
```
]



---

.size40[**1. Descriptive statistics by groups**]

.pull-left[
.code125[
```{r, echo = TRUE, eval = FALSE}
## Column order with group_by
diamonds |>
  group_by(cut, color) |> #<<
  summarize(
    N = n(),
    Mean = mean(price),
    StdDev = sd(price)
  )
```
]]

.pull-right[
```{r, echo = FALSE, eval = TRUE}
diamonds |>
  group_by(cut, color) |>
  summarize(
    N = n(),
    Mean = mean(price),
    StdDev = sd(price),
  ) |>
  ungroup() |>
  slice(1:10) |>
  knitr::kable(digits=0)
```
]

<br>

.center[.size40[**How do we include a table like this in a report? (35 rows)**]]



---

class: middle

.size30[.center[**Option 1: Add a dynamic data table to your report**]]

<br>

.code125[
```{r, echo=TRUE, eval=FALSE}
## Load the DT package (install it first!)
library(DT)

## Use the datatable function
diamonds |>
  group_by(cut, color) |>
  summarize(
    N = n(),
    Mean = mean(price),
    StdDev = sd(price),
  ) |>
  datatable(filter = "top", rownames = FALSE) #<<
```
]



---

class: middle

.size30[.center[**Option 1: Add a dynamic data table to your report**]]

<br>

.code125[
```{r, echo=TRUE, eval=FALSE}
## Load the DT package (install it first!)
library(DT)

## Use the datatable function
diamonds |>
  group_by(cut, color) |>
  summarize(
    N = n(),
    Mean = mean(price),
    StdDev = sd(price),
  ) |>
  datatable(filter = "top", rownames = FALSE) |> #<<
  formatRound(columns = c("Mean", "StdDev"), digits = 0) #<<
```
]



---

class: middle

```{r, echo=FALSE, eval=TRUE}
## Load the DT package (install it first!)
library(DT)

## Use the datatable function
diamonds |>
  group_by(cut, color) |>
  summarize(
    N = n(),
    Mean = mean(price),
    StdDev = sd(price),
  ) |>
  datatable(rownames = FALSE) |> 
  formatRound(columns = c("Mean", "StdDev"), digits = 0)
```




---

class: middle

.size30[.center[**Option 2: Standard Table (Cat x Cat x Numeric)**]]

.code125[
```{r, echo = TRUE, eval = TRUE}
diamonds |>
  group_by(cut, color) |>
  summarize(
    Mean = round(mean(price), 0)
  ) |>
  pivot_wider(names_from = color, values_from = Mean) #<<
```
]

???

1) Choose 1 stat, 
2) pivot_wider to change rows into columns



---

class: middle

.size40[.center[**Why aren't colorless (D, E, F) diamonds of higher quality more expensive?**]]

<br>

.size30[
```{r, echo = FALSE, eval = TRUE}
diamonds |>
  group_by(cut, color) |>
  summarize(
    Mean = round(mean(price), 0)
  ) |>
  pivot_wider(names_from = color, values_from = Mean) |>
  knitr::kable()
```
]

--

<br>

.size40[.center[How do we ’control’ for diamond size in this table?]]



---

class: middle

.code110[
```{r, echo = TRUE, eval = FALSE}
diamonds |>
  filter(carat < 1) |> #<<
  group_by(cut, color) |>
  summarize(
    Mean = round(mean(price), 0)
  ) |>
  pivot_wider(names_from = color, values_from = Mean) |>
  knitr::kable()
```
]

.size30[
```{r, echo = FALSE, eval = TRUE}
diamonds |>
  filter(carat < 1) |> #<<
  group_by(cut, color) |>
  summarize(
    Mean = round(mean(price), 0)
  ) |>
  pivot_wider(names_from = color, values_from = Mean) |>
  knitr::kable()
```
]

???

Not clear why the colorless diamonds (D, E, F) are not more expensive.

Diamond prices determined by four c's: carat, cut, clarity, color

CLASS: Now redo the table for medium diamonds (between 1 and 2 carats in size).



---

class: middle

.code110[
```{r, echo = TRUE, eval = FALSE}
diamonds |>
  filter(carat > 1, carat < 2) |> #<<
  group_by(cut, color) |>
  summarize(
    Mean = round(mean(price), 0)
  ) |>
  pivot_wider(names_from = color, values_from = Mean) |>
  knitr::kable()
```
]

.size30[
```{r, echo = FALSE, eval = TRUE}
diamonds |>
  filter(carat > 1, carat < 2) |> #<<
  group_by(cut, color) |>
  summarize(
    Mean = round(mean(price), 0)
  ) |>
  pivot_wider(names_from = color, values_from = Mean) |>
  knitr::kable()
```
]

???

Result makes way more sense.

CLASS: Now redo the table for large diamonds (above 2 carats in size).



---

class: middle

.code110[
```{r, echo = TRUE, eval = FALSE}
diamonds |>
  filter(carat > 2) |> #<<
  group_by(cut, color) |>
  summarize(
    Mean = round(mean(price), 0)
  ) |>
  pivot_wider(names_from = color, values_from = Mean) |>
  knitr::kable()
```
]

.size30[
```{r, echo = FALSE, eval = TRUE}
diamonds |>
  filter(carat > 2) |> #<<
  group_by(cut, color) |>
  summarize(
    Mean = round(mean(price), 0)
  ) |>
  pivot_wider(names_from = color, values_from = Mean) |>
  knitr::kable()
```
]



---

class: middle, slideblue

.size50[.center[**Multivariate Analyses**]]

<br>

.size40[
1. .greytext[Descriptive statistics by groups]
  
2. **Using facets to extend any visualization**
    
3. .greytext[Mapping 3rd variables with colors or shapes]
]



---

.size40[1) Start with a univariate visualization]

.code125[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.618, out.width='65%', fig.width=5}
ggplot(data = mpg, aes(x = cty)) +
  geom_histogram(bins = 10, color = "white") 
```
]



---

.size40[2) facet_wrap to repeat plot for data subsets]

.code125[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.5, out.width='85%', fig.width=5}
ggplot(data = mpg, aes(x = cty)) +
  geom_histogram(bins = 10, color = "white") +
  facet_wrap(~ drv) #<<
```
]



---

.size40[2) facet_wrap to repeat plot for data subsets]

.code125[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.5, out.width='85%', fig.width=5}
ggplot(data = mpg, aes(x = cty)) +
  geom_histogram(bins = 10, color = "white") +
  facet_wrap(~ drv, scales = "free_y") #<<
```
]



---

.size40[You can specify the number of columns]

.code125[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.5, out.width='85%', fig.width=5}
ggplot(data = mpg, aes(x = cty)) +
  geom_histogram(bins = 10, color = "white") +
  facet_wrap(~ class, ncol = 4) #<<
```
]

???

CLASS: Try different numbers of columns and see what happens




---

.size40[facet_grid to subset by two cat variables]

.code125[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.5, out.width='85%', fig.width=5}
ggplot(data = mpg, aes(x = cty)) +
  geom_histogram(bins = 10, color = "white") +
  facet_grid(drv ~ class) #<<
```
]



---

.size40[We can use facets with ANY baseline visualization]

.code125[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.618, out.width='65%', fig.width=5}
ggplot(data = mpg, aes(x = displ, y = cty)) +
    geom_point()
```
]



---

.size40[We can use facets with ANY baseline visualization]

.code125[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.45, out.width='80%', fig.width=5}
ggplot(data = mpg, aes(x = displ, y = cty)) +
    geom_point() +
    facet_wrap(~ drv) #<<
```
]




---

class: middle, slideblue

.size50[.center[**Multivariate Analyses**]]

<br>

.size40[
1. .greytext[Descriptive statistics by groups]
  
2. .greytext[Using facets to extend any visualization]
    
3. **Mapping 3rd variables with colors or shapes**
]



---

.size40[Mapping 3rd variables with colors or shapes]

.code125[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.618, out.width='65%', fig.width=5}
## Baseline
ggplot(data = mpg, aes(x = displ, y = cty)) +
    geom_point()
```
]



---

.size40[Mapping 3rd variables with colors or shapes]

.code125[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.65, out.width='70%', fig.width=5}
ggplot(data = mpg, aes(x = displ, y = cty)) + 
    geom_point(color = "blue") #<<
```
]



---

.size40[Mapping 3rd variables with colors or shapes]

.code125[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.65, out.width='70%', fig.width=5}
ggplot(data = mpg, aes(x = displ, y = cty)) + 
    geom_point(aes(color = drv)) #<<
```
]


---

.size40[Mapping 3rd variables with colors or shapes]

.code125[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.65, out.width='70%', fig.width=5}
ggplot(data = mpg, aes(x = displ, y = cty)) + 
    geom_point(aes(shape = drv)) #<<
```
]



---

.size40[Mapping 3rd variables with colors or shapes]

.code125[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.65, out.width='70%', fig.width=5}
ggplot(data = mpg, aes(x = displ, y = cty)) + 
    geom_point(aes(shape = drv), size = 2) #<<
```
]


---

.size40[Mapping 3rd variables with colors or shapes]

.code120[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.65, out.width='70%', fig.width=5}
ggplot(data = mpg, aes(x = displ, y = cty)) + 
    geom_point(aes(color = drv, shape = drv), size = 2) #<<
```
]


---

.size40[Mapping 3rd variables with colors or shapes]

.code120[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.45, out.width='90%', fig.width=6}
ggplot(data = mpg, aes(x = displ, y = cty)) +
    geom_point(aes(color = drv, shape = drv), size = 2) +
    facet_wrap(~ drv) #<<
```
]


---

.size40[Mapping 3rd variables with colors or shapes]

.code120[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.45, out.width='85%', fig.width=6}
ggplot(data = mpg, aes(x = displ, y = cty)) +
  geom_point(aes(color = drv, shape = drv), size = 2) +
  facet_wrap(~ drv) +
  scale_color_manual(values = c("orange", "red", "blue")) #<<
```
]



---

.code120[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.45, out.width='85%', fig.width=6}
ggplot(data = mpg, aes(x = displ, y = cty)) +
  geom_point(aes(color = drv, shape = drv), size = 2) +
  facet_wrap(~ drv) +
  scale_color_manual(values = c("orange", "red", "blue")) +
  theme(legend.position = "top") #<<
```
]



---

.code120[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.45, out.width='85%', fig.width=6}
ggplot(data = mpg, aes(x = displ, y = cty)) +
  geom_point(aes(color = drv, shape = drv), size = 2) +
  facet_wrap(~ drv) +
  scale_color_manual(values = c("orange", "red", "blue")) +
  theme(legend.position = "top")  +
  geom_smooth(method = "lm", se = FALSE) #<<
```
]



---

class: middle, slidegreen

.size50[.center[**Multivariate Analyses**]]

<br>

.size40[
1. Descriptive statistics by groups
  
2. Using facets to extend any visualization
    
3. Mapping 3rd variables with colors or shapes

4. **Bonus: Animate the 3rd/4th variable!**
]



---

class: middle

.size40[.center[**Install the needed packages and load gganimate**]]

<br>

.code150[
```{r, echo=TRUE, eval=FALSE}
install.packages("gifski")

install.packages("png")

install.packages("gganimate")

library(gganimate)
```
]

```{r}
## Load it for you
library(gganimate)
```



---

.center[**Transitions between different subsets of the data**]

.code120[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=0.7, out.width='60%', fig.width=5, cache=TRUE}
ggplot(data = mpg, aes(x = displ, y = cty)) +
  geom_point() +
  transition_states(cyl) + #<<
  labs(title = "Number of Cylinders: {closest_state}") #<<
```
]

???

transition_states is for moving across levels of a categorical variable



---

.center[**Transitions between different subsets of the data**]

.code120[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=0.618, out.width='60%', fig.width=5, cache=TRUE}
ggplot(data = diamonds, aes(x = cut, fill = color)) +
  geom_bar() +
  transition_states(cut) + #<<
  shadow_mark() + #<<
  enter_grow() + #<<
  enter_fade() #<<
```
]

???

Build the plot by subsets with dynamic adding




---

.center[**Transitions across time use a different function**]

.code120[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=0.7, out.width='60%', fig.width=5, cache=TRUE}
ggplot(data = mpg, aes(x = class, y = cty)) +
  geom_boxplot() +
  transition_time(year) + #<<
  labs(title = 'Year: {frame_time}') #<<
```
]


---

.center[**Transitions can also leave the prior data in place**]

.code120[
```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=0.618, out.width='60%', fig.width=5, cache=TRUE}
ggplot(data = economics, aes(x = date, y = pop)) +
  geom_line(color = "darkblue", size = 1.4) +
  transition_reveal(date) + #<<
  coord_cartesian(ylim = c(150000, 350000)) +
  labs(title = "US Population (thousands)")
```
]



---

class: middle

.size50[.center[**Last one, so let's make something awesome**]]

.code135[
```{r, echo=TRUE, eval=FALSE}
## New package needed for the data (install before loading)
install.packages("gapminder")

library(gapminder)
```
]

```{r}
library(gapminder)

gapminder |>
  slice(1:6) |>
  knitr::kable(format = "html")
```



---

class: middle

.code115[
```{r, fig.retina=3, fig.align='center', fig.asp=0.7, out.width='65%', fig.width=6, cache=TRUE, echo=TRUE, eval=TRUE}
gapminder |>
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(alpha = 0.7, show.legend = FALSE)
```
]




---

class: middle

.code115[
```{r, fig.retina=3, fig.align='center', fig.asp=0.7, out.width='65%', fig.width=6, cache=TRUE, echo=TRUE, eval=TRUE}
gapminder |>
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) + #<<
  geom_point(alpha = 0.7, show.legend = FALSE)
```
]




---

class: middle

.code115[
```{r, fig.retina=3, fig.align='center', fig.asp=0.7, out.width='65%', fig.width=6, cache=TRUE, echo=TRUE, eval=TRUE}
gapminder |>
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_x_log10() #<<
```
]



---

class: middle

.code115[
```{r, fig.retina=3, fig.align='center', fig.asp=0.45, out.width='85%', fig.width=8, cache=TRUE, echo=TRUE, eval=TRUE}
gapminder |>
  filter(continent != "Oceania") |> #<<
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_x_log10() +
  facet_wrap(~ continent, ncol = 4) #<<
```
]



---

class: middle

.code115[
```{r, fig.retina=3, fig.align='center', fig.asp=0.45, out.width='85%', fig.width=8, cache=TRUE, echo=TRUE, eval=TRUE}
gapminder |>
  filter(continent != "Oceania") |>
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_x_log10() +
  facet_wrap(~ continent, ncol = 4) +
  scale_colour_manual(values = country_colors) #<<
```
]



---

class: middle

.code115[
```{r, fig.retina=3, fig.align='center', fig.asp=0.45, out.width='80%', fig.width=8, cache=TRUE, echo=TRUE, eval=TRUE}
gapminder |>
  filter(continent != "Oceania") |>
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_x_log10() +
  facet_wrap(~ continent, ncol = 4) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(3, 13)) #<<
```
]



---

class: middle

.code115[
```{r, fig.retina=3, fig.align='center', fig.asp=0.45, out.width='80%', fig.width=8, cache=TRUE, echo=TRUE, eval=FALSE}
gapminder |>
  filter(continent != "Oceania") |>
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_x_log10() +
  facet_wrap(~ continent, ncol = 4) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(3, 13)) +
  transition_time(year) + #<<
  labs(title = 'Year: {frame_time}') #<<
```
]



---

class: middle

```{r, fig.retina=3, fig.align='center', fig.asp=0.4, out.width='100%', fig.width=8, cache=TRUE, echo=FALSE, eval=TRUE}
gapminder |>
  filter(continent != "Oceania") |>
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_x_log10() +
  facet_wrap(~ continent, ncol = 4) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(3, 13)) +
  transition_time(year) + #<<
  labs(title = 'Year: {frame_time}', x = "GDP per capita (log 10)", y = "Life Expectancy") 
```




---

class: center

.size40[**Do countries with larger military enlistments have fewer political freedoms?**]
    
```{r, fig.retina=3, fig.align = 'center', fig.height = 3, fig.width=5, out.width='70%'}
## Manual DAG
tibble(
  x = c(-3, 3, 0),
  y = c(1, 1, 2),
  labels = c("Military\n Enlistment", "Political\n Freedoms", "Confounders")
) |>
  ggplot(aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4), ylim = c(0.8, 2.1)) +
  geom_label(aes(label = labels), size = 7) +
    annotate("segment", x = .4, xend = 2.5, y = 1.9, yend = 1.2, arrow = arrow()) +
  annotate("segment", x = -.4, xend = -2.5, y = 1.9, yend = 1.2, arrow = arrow()) +
  annotate("segment", x = -1.5, xend = 1.6, y = 1, yend = 1, arrow = arrow())
```


