---
format: revealjs
---

# Today's Agenda {background-image="libs/Images/background-data_blue_v3.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
```

<br>

<br>

**Multivariate Analyses**

1. Descriptive statistics by groups

2. Using facets to extend any visualization

3. Mapping 3rd variables with colors or shapes

<br>

::: r-stack
Justin Leinaweaver (Spring 2024)
:::

::: notes
Prep for Class

1. Upload WDI data for class to Canvas: WDI-Life_Expectancy-2015.csv

2. ??
:::


## Building a Causal Argument {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Predictor", "Confounders", "Outcome")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.45, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.1, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.1, arrow = arrow())
```

::: notes

Last week we started exploring the concepts and approaches we can use to make causal arguments

- e.g. arguments that changes in X cause a change in Y

<br>

We discussed:

- the Data Generating Process (DGP), 

- identification strategies, 

- the use of Directed Acyclic Graphs (DAGs), and 

- the need to adjust our analyses for confounders.

<br>

### Any questions on those big ideas?
:::



## Building a Causal Argument {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Government\nSpending", "Confounders", "Gender\nInequality\nIndex")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.55, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

::: notes
In class on Friday we worked to:

1. Draw DAGs for each of our three hypotheses, and

2. Gather data that operationalizes the concepts we identified as important confounders in these relatonships.

<br>

### Any questions on those DAGs or the variables we selected?
:::



## Multivariate Analyses {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. Descriptive statistics by groups
  
2. Using facets to extend any visualization
    
3. Mapping 3rd variables with colors or shapes

::: notes
Today I will introduce you to a series of statistical tools that will allow you to adjust your analyses for a confounder variable.

- These are what we call multivariate analyses

- e.g. Analyses on three or more variables simultaneously

<br>

Note that you can also use these techniques to extend simple descriptive analyses to summarize data across multiple variables

<br>

I'll introduce these tools to you today and then we'll apply them in class on Friday on your research projects.
:::



## Do countries with more hospital capacity have longer life expectancies? {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', fig.width = 7, fig.height=1.3, out.width='90%', cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(-3, 3),
  y = c(1, 1),
  labels = c("Hospital\nBeds", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = -2.1, xend = 1.9, y = 1, yend = 1, arrow = arrow())
```

::: notes
Let's say you were working on a research project exploring the impact of healthcare spending on key outcomes.

- Specifically we are interested in estimating the effect of hospital capacity on life expectancies across the world.
:::


## Do countries with more hospital capacity have longer life expectancies? {background-image="libs/Images/background-slate_v2.png" .center .smaller}

<br>

```{r}
## Prep WDI data for the class
# Input data
# d <- read_excel("../Data_in_Class-SP23/Wk14-Multivariate/WDI-Life_Expectancy-2015-ORIGINAL.xlsx", na = "NA")
# 
# d2 <- d |>
#   mutate(
#     gdp_cats = case_when(
#       gdp < quantile(d$gdp, probs = .25, na.rm = TRUE) ~ "Bottom 25%",
#      gdp < quantile(d$gdp, probs = .75, na.rm = TRUE) ~ "Middle 50%",
#      gdp >= quantile(d$gdp, probs = .75, na.rm = TRUE) ~ "Top 25%"
#     ),
#     hosp_beds_cats = case_when(
#       hospital_beds_per_1k < 1 ~ "< 1 per 1k",
#       hospital_beds_per_1k < 2 ~ "< 2 per 1k",
#       #hospital_beds_per_1k < 3 ~ "< 3 per 1k",
#       hospital_beds_per_1k >= 2 ~ "> 2 per 1k",
#     )
#   )
# 
# d3 <- d2 |>
#   select(-access_electricity_pct_pop, -gdp_pc) |>
#   na.omit()
# 
# write_excel_csv(d3, file = "../Data_in_Class-SP23/Wk14-Multivariate/WDI-Life_Expectancy-2015.csv")

d <- read_csv("../../Course_History/2023-Spring/Data_in_Class-SP23/Wk14-Multivariate/WDI-Life_Expectancy-2015.csv")

d |>
  select(country, life_expectancy, hospital_beds_per_1k, hosp_beds_cats, gdp, gdp_cats) |>
  slice_sample(n = 9) |>
  kbl(digits = 1, align = c('l', rep('c', 5))) |>
  kableExtra::kable_styling(font_size = 20)
```

- Import 'WDI-Life_Expectancy-2015.csv' into R


::: notes
I've pulled the 2015 data from the World Bank's WDI database for us to explore.

- Everybody grab this dataset from Canvas and import it into RStudio.

<br>

Variables

- Life expectancy is total in years

- Hospital beds is per 1,000 people in the population

- GDP is gross domestic product

- I've also given you a categorical version of both hospital beds and GDP for us to explore

<br>

### Ok, what is the first step we need to do to answer this question?

### - e.g. the first thing you need a good sense of before building your causal argument?

- (You MUST first analyze the variation in the outcome variable!)

1. If there is no variation there, then your project cannot proceed, and

2. Your job is then to explain the distribution you find there

<br>

Let's get started!

- Everybody make a histogram of life expectancy and calculate the descriptive statistics for it.
:::


## Do countries with more hospital capacity have longer life expectancies? {background-image="libs/Images/background-slate_v2.png" .center}

:::: {.columns}
::: {.column width="50%"}
```{r, fig.align='center', fig.retina=3, out.width='100%', fig.asp=0.9, fig.width=5, cache=TRUE}
ggplot(data = d, aes(x = life_expectancy)) +
  geom_histogram(bins = 12, color = "white") +
  theme_bw() +
  labs(x = "Life Expectancy (years)", y = "")
```
:::

::: {.column width="50%"}
```{r, eval = TRUE}
d |>
  summarize(
    Min = min(life_expectancy),
    "1st Qu" = quantile(life_expectancy, probs = .25),
    Median = median(life_expectancy),
    Mean = mean(life_expectancy),
    "3rd Qu" = quantile(life_expectancy, probs = .75),
    Max = max(life_expectancy)
  ) |>
  t() |>
  kbl(digits = 1, align = 'c') |>
  column_spec(1:2, background = "white")
#summary(d$life_expectancy)
```
:::
::::

::: notes
**Analyze these univariate analyses for me**

- **What do we learn from this about the outcome we are trying to explain?**
:::


## Do countries with more hospital capacity have longer life expectancies? {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', fig.width = 7, fig.height=1.3, out.width='90%', cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(-3, 3),
  y = c(1, 1),
  labels = c("Hospital\nBeds", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = -2.1, xend = 1.9, y = 1, yend = 1, arrow = arrow())
```

<br>

Calculate the mean life expectancy ('life_expectancy') across the levels of hospital beds ('hosp_beds_cats')

::: notes
Let's now move from a univariate to a bivariate analysis.
:::



## Do countries with more hospital capacity have longer life expectancies? {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, echo=TRUE}
aggregate(data = d, life_expectancy ~ hosp_beds_cats, FUN = mean)
```

::: notes
**Analyze these bivariate analyses for me**

- **What do we learn from this about the relationship we are trying to test?**

<br>

Key here is that we have established a correlation with this analysis, but not causation.

### What do we need to add to make this a causal argument?

- (**SLIDE**: An identification strategy that specifies the confounders!)
:::



## Building a Causal Argument {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Hospital\nBeds", "GDP", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.5, xend = 2.4, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.9, yend = 1.2, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.9, yend = 1.2, arrow = arrow())
```

::: notes
My DAG displays one identification strategy for this argument.

### What does it mean to say that GDP is presented as a confounder?

- Argument that the size of your economy influences your hospital capacity, and

- Argument that the size of your economy influences your average life expectancies

<br>

**SLIDE**: Put another way...
:::



## Building a Causal Argument {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Hospital\nBeds", "GDP", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  annotate("rect", xmin = 1.45, xmax = 2.45, ymin = .9, ymax = 1.1, color = "red", fill = "white", linewidth = 1.3) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.5, xend = 2.4, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.9, yend = 1.2, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.9, yend = 1.2, arrow = arrow())
```

::: notes

This diagram argues that we cannot get a causal estimate of the effect of changing the number of hospital beds on life expectancy without adjusting our analysis for the size of the economy.

<br>

**SLIDE**: This brings us to the tools we will need to adjust our analyses for the confounders.
:::


## Multivariate Analyses {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. **Descriptive statistics by groups**
  
2. Using facets to extend any visualization
    
3. Mapping 3rd variables with colors or shapes

::: notes

First approach extends our descriptive statistics across groups in the data.
:::



## 1. Descriptive Statistics by Group {background-image="libs/Images/background-slate_v2.png" .center}

**Bivariate Analyses: Numerical x Categorical**

<br>

```{r, echo=TRUE}
aggregate(data = d, life_expectancy ~ hosp_beds_cats, FUN = mean)
```

::: notes
Let's go back to the aggregate code you wrote earlier.
:::



## 1. Descriptive Statistics by Group {background-image="libs/Images/background-slate_v2.png" .center .smaller}

**Multivariate Analyses: Numerical x Categorical x Categorical**

<br>

```{r, echo=TRUE}
aggregate(data = d, life_expectancy ~ hosp_beds_cats + gdp_cats, FUN = mean)
```

::: notes
Adding a third variable that is categorical to the aggregate function is very simple.

- Modify the formula argument to add a second predictor variable just like adding two numbers together with the plus sign.

- You can add as many CATEGORICAL variables to the right side of the formula as you want.

<br>

### Does the code make sense?

<br>

### So, when we calculate the averages while adjusting for GDP what happens to our estimate of the effect we are interested in?

### - Does increasing hospital beds increase life expectancies when controlling for GDP?

<br>

**SLIDE**: You can also add a third variable to the aggregate function if it is a numeric variable.
:::



## 1. Descriptive Statistics by Group {background-image="libs/Images/background-slate_v2.png" .center .smaller}

**Multivariate Analyses: Numerical x Categorical x Numerical**

<br>

:::: {.columns}
::: {.column width="50%"}
```{r, echo = TRUE, eval = TRUE, `code-line-numbers`="4"}
aggregate(data = d, 
          life_expectancy ~ hosp_beds_cats, 
          FUN = mean,
          subset = gdp < 8e10) 
```
:::

::: {.column width="50%"}
```{r, echo = TRUE, eval = TRUE, `code-line-numbers`="4"}
aggregate(data = d, 
          life_expectancy ~ hosp_beds_cats, 
          FUN = mean,
          subset = gdp > 8e10) 
```
:::
::::

::: notes
You can also use the subset argument in aggregate to split your sample by a numeric variable.

- The key here is to use evaluation functions like less than, greater than, less than or equal to and greater than or equal to.

<br>

I hope everyone can see now how powerful the aggregate function can be.

- Add multiple categorical variables to the formula, OR

- Use subset to add numeric variables.

<br>

### Questions on using subset in the aggregate function to adjust your results for a numeric variable?
:::



## Multivariate Analyses {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. Descriptive statistics by groups
  
2. **Using facets to extend any visualization**
    
3. Mapping 3rd variables with colors or shapes

::: notes

Let's shift to our second multivariate analysis technique.

- Using facets to extend any visualization

<br>

In short, facets let you take any visualization and break it into separate plots across levels of another variable.

- Think of this like the visual version of what we just did using aggregate

<br>

**SLIDE**: Let's start with a bivariate visualization
:::



## Do countries with more hospital capacity have longer life expectancies? {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', fig.width = 7, fig.height=1.3, out.width='90%', cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(-3, 3),
  y = c(1, 1),
  labels = c("Hospital\nBeds", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = -2.1, xend = 1.9, y = 1, yend = 1, arrow = arrow())
```

<br>

Visualize life expectancy ('life_expectancy') and hospital beds ('hospital_beds_per_1k') with a scatter plot

::: notes
???
:::



## {background-image="libs/Images/background-slate_v2.png" .center}

**2. Using facets to extend a visualization**

```{r, echo=TRUE, fig.retina=3, fig.align = 'center', fig.width = 7, fig.asp = .618, out.width='70%', cache=TRUE}
ggplot(data = d, aes(x = hospital_beds_per_1k, y = life_expectancy)) +
  geom_point() +
  geom_smooth(method = "lm")
```

::: notes
**Analyze this bivariate analysis for me**

- **What do we learn from this about the relationship we are trying to test?**
:::



## {background-image="libs/Images/background-slate_v2.png" .center}

**2. Using facets to extend a visualization**

```{r, echo=TRUE, fig.align ='center', fig.width = 10, fig.asp =.4, `code-line-numbers`="4", cache=TRUE}
ggplot(data = d, aes(x = hospital_beds_per_1k, y = life_expectancy)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ gdp_cats)
```

::: notes

Here we see the facet_wrap function.

- The argument requires a tilde ('~') and then the name of a third categorical variable.

<br>

### Analyze these multivariate analyses for me

### - What do we learn from this about the relationship we are trying to test?

<br>

The slope is positive across all three levels of GDP!

- This means we have more confidence that increasing hospital beds increases life expectancy across all three levels of economy size.

- This is how we build a causal argument by pairing our analyses to our identification strategy.

<br>

### Any questions on using the facet_wrap function to adjust an analysis for a confounder?
:::



## Multivariate Analyses {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. Descriptive statistics by groups
  
2. Using facets to extend any visualization
    
3. **Mapping 3rd variables with colors or shapes**

::: notes
Last multivariate approach for today!

- Mapping 3rd variables with colors or shapes
:::



## {background-image="libs/Images/background-slate_v2.png" .center}

**3. Mapping 3rd variables with colors or shapes**

```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.65, out.width='65%', fig.width=7, cache=TRUE}
ggplot(data = d, aes(x = hospital_beds_per_1k, y = life_expectancy)) +
  geom_point()
```

::: notes
Back to our standby baseline plot.
:::


## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

**3. Mapping 3rd variables with colors or shapes**

```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.65, out.width='60%', fig.width=7, cache=TRUE, `code-line-numbers`="1"}
ggplot(data = d, aes(x = hospital_beds_per_1k, y = life_expectancy, color = gdp_cats)) +
  geom_point()
```

::: notes
Rather than facet wrap, we can tell R to assign colors to the levels of a categorical variable.
:::


## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

**3. Mapping 3rd variables with colors or shapes**

```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.65, out.width='60%', fig.width=7, cache=TRUE, `code-line-numbers`="1"}
ggplot(data = d, aes(x = hospital_beds_per_1k, y = life_expectancy, shape = gdp_cats)) +
  geom_point()
```

::: notes
Printing in black and white?

- Use shapes instead of colors!
:::



## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

**3. Mapping 3rd variables with colors or shapes**

```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.65, out.width='60%', fig.width=7, cache=TRUE, `code-line-numbers`="2"}
ggplot(data = d, aes(x = hospital_beds_per_1k, y = life_expectancy, 
                     color = gdp_cats, shape = gdp_cats)) + 
  geom_point()
```

::: notes

Best approach? Do both!
:::


## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

**3. Mapping 3rd variables with colors or shapes**

```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.45, out.width='80%', fig.width=8, cache=TRUE, `code-line-numbers`="4"}
ggplot(data = d, aes(x = hospital_beds_per_1k, y = life_expectancy, 
                     color = gdp_cats, shape = gdp_cats)) +
  geom_point() +
  facet_wrap(~ gdp_cats)
```

::: notes
If your dots overlap too much then also add a facet wrap!
:::


## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

**3. Mapping 3rd variables with colors or shapes**

```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.45, out.width='80%', fig.width=8, cache=TRUE, `code-line-numbers`="5"}
ggplot(data = d, aes(x = hospital_beds_per_1k, y = life_expectancy, 
                     color = gdp_cats, shape = gdp_cats)) +
  geom_point() +
  facet_wrap(~ gdp_cats) +
  guides(color = "none", shape = "none")
```

::: notes
If doing both then omit the legend as it is redundant (each facet is labeled).

- Because the plot adds a color and a shape you have to remove both of their legends hence the two arguments here
:::



## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

**3. Mapping 3rd variables with colors or shapes**

```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.4, out.width='80%', fig.width=8, cache=TRUE, `code-line-numbers`="6"}
ggplot(data = d, aes(x = hospital_beds_per_1k, y = life_expectancy, 
                     color = gdp_cats, shape = gdp_cats)) +
  geom_point() +
  facet_wrap(~ gdp_cats) +
  guides(color = "none", shape = "none") +
  scale_color_manual(values = c("orange", "red", "blue"))
```

::: notes
You can also pick the colors manually
:::


## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

**3. Mapping 3rd variables with colors or shapes**

```{r, echo=TRUE, fig.retina=3, fig.align='center', fig.asp=.4, out.width='80%', fig.width=8, cache=TRUE, `code-line-numbers`="7"}
ggplot(data = d, aes(x = hospital_beds_per_1k, y = life_expectancy, 
                     color = gdp_cats, shape = gdp_cats)) +
  geom_point() +
  facet_wrap(~ gdp_cats) +
  guides(color = "none", shape = "none") +
  scale_color_manual(values = c("orange", "red", "blue"))  +
  geom_smooth(method = "lm")
```

::: notes

Finally, let's put our OLS lines back on!

<br>

### Any questions on this approach?
:::



## Multivariate Analyses {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. Descriptive statistics by groups
  
2. Using facets to extend any visualization
    
3. Mapping 3rd variables with colors or shapes

::: notes

**Any questions on our multivariate tools from today?**

<br>

**SLIDE**: Descriptive stats practice...
:::



## ggplot2 Dataset: mpg {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r}
mpg |>
  slice_sample(n = 8) |>
  kbl(digits = 2, align = 'c') |>
  column_spec(column = c(1, 4, 8), background = "gold") |>
  kableExtra::kable_styling(font_size = 22)
```

<br>

**Practice**: Which auto manufacturers improved their average city fuel economy when moving from 1999 to 2008 models?

::: notes

:::


## Which auto manufacturers improved their average city fuel economy when moving from 1999 to 2008 models? {background-image="libs/Images/background-slate_v2.png" .center .smaller}

:::: {.columns}
::: {.column width="50%"}
```{r, echo = TRUE, eval = FALSE}
# Save the group means
x1 <- aggregate(data = mpg, 
          cty ~ manufacturer + year, 
          FUN = mean)

# Order the output
arrange(x1, manufacturer)
```
:::

::: {.column width="50%"}
```{r, echo = FALSE, eval = TRUE}
# Save the group means
x1 <- aggregate(data = mpg, 
          cty ~ manufacturer + year, 
          FUN = mean)

# Order the output
arrange(x1, manufacturer)
```
:::
::::

::: notes
Because there are only two levels in year I treated it here as a categorical variable

<br>

### So, what do we learn from this?

<br>

**SLIDE**: Visualization practice...
:::



## ggplot2 Dataset: mpg {background-image="libs/Images/background-slate_v2.png" .center}

```{r}
mpg |>
  slice_sample(n = 7) |>
  kbl(digits = 2, align = 'c') |>
  column_spec(column = c(1,3,5,8), background = "gold") |>
  kableExtra::kable_styling(font_size = 22)
```

<br>

**Practice**: How do various manufacturers design their fleets to meet the EPA requirements for corporate average fuel economy (CAFE)?

::: notes

EPA requires manufacturers to meet a certain level of fuel economy across their entire fleet.

- So, some models can have big, less efficient engines but only if they offset those sales with sales of cleaner cars.

<br>

Let's try developing a visualization to let us see this dynamic in practice!

### How could we visualize these four variables?

- (**SLIDE**)
:::


## ggplot2 Dataset: mpg {background-image="libs/Images/background-slate_v2.png" .center}

```{r}
mpg |>
  slice_sample(n = 3) |>
  kbl(digits = 2, align = 'c') |>
  column_spec(column = c(1,3,5,8), background = "gold") |>
  kableExtra::kable_styling(font_size = 22)
```

<br>

Make a scatter plot of city fuel economy x engine size

- Add colors and shapes for drive train

- Add facet wrapping for the manufacturers

::: notes

:::


## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

```{r, eval=TRUE, echo=TRUE, fig.asp=.65, fig.align='center', out.width='70%', fig.width=9, fig.retina=3}
ggplot(data = mpg, aes(x = displ, y = cty, color = drv, shape = drv)) +
  geom_point() +
  facet_wrap(~ manufacturer, ncol = 5)
```

::: notes

**Lessons from this?**
:::



## Multivariate Analyses {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. Descriptive statistics by groups
  
2. Using facets to extend any visualization
    
3. Mapping 3rd variables with colors or shapes

::: notes

**Any questions on our multivariate tools from today?**

<br>

**SLIDE**: Add assignment for next class
:::



## For Next Class {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. ...

2. ...

::: notes
Practice multivariate...
:::



