---
format: revealjs
---

## Today's Agenda {background-image="Images/background-data_blue_v4.png" .center}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
library(ggeffects)

# Input practice WDI data
wdi <- read_excel("../../DATA/2026-Spring/WDI/Lectures/World_Development_Indicators-2023-Life_Expectancy.xlsx", na = "NA")

# Input class selections from WDI
d <- read_excel("../../DATA/2026-Spring/WDI/Week03-Class_Selections/Spring2026-WDI_Data_Class_Selections-SHORT_LABELS.xlsx", na = "NA")

d_long <- read_excel("../../DATA/2026-Spring/WDI/Week03-Class_Selections/Spring2026-WDI_Data_Class_Selections.xlsx", na = "NA")

# # Increasing modelsummary color and size
# "xx-small", "x-small", "small", "medium", "large", "x-large", or "xx-large"
# modelsummary(models,
#              out = "gt",
#              fmt = 2, 
#              stars = c("*" = .05), 
#              gof_omit = "IC|F|Log") |>
#   gt::tab_style(style = list(
#                   gt::cell_fill(color = 'white'),
#                   gt::cell_text(size = "large")
#   ), locations = gt::cells_body())
```

<br>

::: {.r-fit-text}

**Ordinary Least Squares (OLS) Regression**

- Practice fitting, interpreting and evaluating OLS regressions

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2026)
:::

::: notes
Prep for Class

1. Check Canvas submissions

<br>

**SLIDE**: Let's refresh our memories on regression

:::




## Why use OLS regressions? {background-image="Images/background-data_blue_v4.png" .center}

<br>

- **Quantifies the relationship** between variables

- Uses **ALL** of the data

- Makes predictions with estimates of **uncertainty**

- Gives us **criteria** for evaluating the fit of the line

**Future Weeks**

- Can be adjusted for confounders (e.g. control variables), nonlinear relationships and different data structures

::: notes

**Any questions on our work last week on regression?**

- **e.g. the intuitions, interpretations, evaluations or methods in R?**

<br>

**SLIDE**: For today I asked you to use regressions to answer a big question

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Outcome to Explain: Average Life Expectancy**
:::

```{r, fig.align='center', fig.asp=0.618, fig.width=8}
# wdi |> select(country, life_expectancy) |> View()

wdi |>
  ggplot(aes(x = life_expectancy)) +
  geom_histogram(color = "white", bins = 15) +
  theme_bw() +
  labs(x = "Global Average Life Expectancy", y = "Count") +
  annotate("text", x = 61.25, y = .5, label = "Liberia", color = "white", srt = 90, hjust = 0, size = 6) +
  annotate("text", x = 65.75, y = .5, label = "S. Africa", color = "white", srt = 90, hjust = 0, size = 6) +
  annotate("text", x = 70.4, y = .5, label = "Indonesia", color = "white", srt = 90, hjust = 0, size = 6) +
  annotate("text", x = 75, y = .5, label = "Mexico", color = "white", srt = 90, hjust = 0, size = 6) +
  annotate("text", x = 79.5, y = .5, label = "Germany", color = "white", srt = 90, hjust = 0, size = 6) +
  annotate("text", x = 84, y = .5, label = "Japan", color = "white", srt = 90, hjust = 0, size = 6) +
  scale_x_continuous(breaks = seq(50,90,5))
```

::: notes

What explains the variation in average global life expectancy?

<br>

Here I'm showing you a histogram of our key outcome variable

- THIS is the variation I want to try and explain

- Why do the people in some countries live longer than 80 years and some well below 70?

<br>

**Where is the US is in this distribution?**

- (A score of 78 years which puts us at 62nd globally or the 70th percentile)

- (R code below)

<br>

**SLIDE**: The regression results

<br>

```{r, echo=TRUE}
wdi |> 
  select(country, life_expectancy) |>
  arrange(desc(life_expectancy)) |>
  mutate( rank = 1:222 ) |>
  filter(country == "United States")

quantile(wdi$life_expectancy, na.rm = TRUE, probs = .70)
```
:::




## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**OLS Models of Global Life Expectancy**
:::

::: {.panel-tabset}

### Results

```{r}
#names(wdi)
res1 <- lm(data = wdi, life_expectancy ~ measles_immune_pct)
res2 <- lm(data = wdi, life_expectancy ~ secondary_education_pct)
res3 <- lm(data = wdi, life_expectancy ~ fertility_rate)

modelsummary(models = list(res1, res2, res3), 
             output = "gt", 
             fmt = 2, 
             stars = c('*' = .05), 
             gof_omit = "IC|Log|F",
             coef_map = c("measles_immune_pct"= "Measles Immunizations (%)",
                          "secondary_education_pct" = "Secondary Education (%)",
                          "fertility_rate" = "Fertility Rate",
                          "(Intercept)" = "Constant"))
```

### Code
```{r, echo=TRUE, eval=FALSE}
# Fit the models
res1 <- lm(data = wdi, life_expectancy ~ measles_immune_pct)
res2 <- lm(data = wdi, life_expectancy ~ secondary_education_pct)
res3 <- lm(data = wdi, life_expectancy ~ fertility_rate)

# Output the table
modelsummary(models = list(res1, res2, res3), 
             fmt = 2, 
             stars = c('*' = .05), 
             gof_omit = "IC|Log|F",
             coef_map = 
               c("measles_immune_pct"= "Measles Immunizations (%)",
                 "secondary_education_pct" = "Secondary Education (%)",
                 "fertility_rate" = "Fertility Rate",
                 "(Intercept)" = "Constant"))
```

:::

::: notes

**Did everybody make this table (e.g. three models together with these results)?**

<br>

Let's step through your analyses of this work starting with the statistical significance of the models

- In other words, how well does each model fit the data (e.g. missing data, significance and R2)?

<br>

**Is the measles model a good fit for the data or not? Why?**

<br>

**Is the education model a good fit for the data or not? Why?**

<br>

**Is the fertility rate model a good fit for the data or not? Why?**

<br>

**SLIDE**: Alright, let's shift to the substantive significance of each model

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.panel-tabset}

### Model 1

```{r, echo=FALSE, fig.align='center', fig.asp=.618, fig.width=7, out.width="87%"}
plot(ggpredict(res1, "measles_immune_pct")) +
  labs(x = "Measles Immunity (%)", 
       y = "Average Life Expectancy (years)",
       title = "Predicted Values of Life Expectancy") +
  scale_y_continuous(limits = c(45,88), breaks = seq(45,85,5))
```

### Code

```{r, echo=TRUE, eval=FALSE}
plot(ggpredict(res1, "measles_immune_pct")) +
  labs(x = "Measles Immunity (%)", 
       y = "Average Life Expectancy (years)") +
  scale_y_continuous(limits = c(45,88), breaks = seq(45,85,5))
```

:::

::: notes

**How substantial is the impact of measles immunizations on average life expectancy?**

<br>

**Is focusing on this a "good" policy recommendation for improving life expectancy? Why or why not?**

:::




## {background-image="Images/background-data_blue_v4.png" .center}

```{r, fig.align='center', fig.asp=.618, fig.width=6}
plot(ggpredict(res2, "secondary_education_pct")) +
  labs(x = "Secondary Education (%)", y = "Average Life Expectancy (years)", title = "Predicted Values of Life Expectancy") +
  scale_y_continuous(limits = c(45,88), breaks = seq(45,85,5))
```

::: notes

**How substantial is the impact of secondary education on average life expectancy?**

<br>

**Is focusing on this a "good" policy recommendation for improving life expectancy? Why or why not?**

:::




## {background-image="Images/background-data_blue_v4.png" .center}

```{r, fig.align='center', fig.asp=.618, fig.width=6}
plot(ggpredict(res3, "fertility_rate")) +
  labs(x = "Fertility Rate", y = "Average Life Expectancy (years)", title = "Predicted Values of Life Expectancy") +
  scale_y_continuous(limits = c(45,88), breaks = seq(45,85,5))
```

::: notes

**How substantial is the impact of fertility rates on average life expectancy?**

<br>

**Is focusing on this a "good" policy recommendation for improving life expectancy? Why or why not?**

:::




## {background-image="Images/background-data_blue_v4.png" .center}

```{r, fig.align='center', fig.asp=.618, fig.width=6, out.width="45%"}
plot(ggpredict(res1, "measles_immune_pct")) +
  labs(x = "Measles Immunity (%)", y = "Average Life Expectancy (years)", title = "Predicted Values (Model 1)") +
  scale_y_continuous(limits = c(45,88), breaks = seq(45,85,5))
```

:::: {.columns}
::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.618, fig.width=6}
plot(ggpredict(res2, "secondary_education_pct")) +
  labs(x = "Secondary Education (%)", y = "Average Life Expectancy (years)", title = "Predicted Values (Model 2)") +
  scale_y_continuous(limits = c(45,88), breaks = seq(45,85,5))
```
:::

::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.618, fig.width=6}
plot(ggpredict(res3, "fertility_rate")) +
  labs(x = "Fertility Rate", y = "Average Life Expectancy (years)", title = "Predicted Values (Model 3)") +
  scale_y_continuous(limits = c(45,88), breaks = seq(45,85,5))
```
:::
::::

::: notes

**So, which of these three models has the largest substantive significance?**

<br>

**SLIDE**: Bottom line time

:::




## {background-image="Images/background-data_blue_v4.png" .center}

```{r}
modelsummary(models = list(res1, res2, res3), 
             output = "gt", 
             fmt = 2, 
             stars = c('*' = .05), 
             gof_omit = "IC|Log|F",
             coef_map = c("measles_immune_pct"= "Measles Immunizations (%)",
                          "secondary_education_pct" = "Secondary Education (%)",
                          "fertility_rate" = "Fertility Rate",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-small")
  ), locations = gt::cells_body())

```

:::: {.columns}
::: {.column width="33%"}
```{r, fig.align='center', fig.asp=.85, fig.width=6}
plot(ggpredict(res1, "measles_immune_pct")) +
  labs(x = "Measles Immunity (%)", y = "Average Life Expectancy (years)", title = "Predicted Values (Model 1)") +
  scale_y_continuous(limits = c(45,88), breaks = seq(45,85,5))
```
:::

::: {.column width="33%"}
```{r, fig.align='center', fig.asp=.85, fig.width=6}
plot(ggpredict(res2, "secondary_education_pct")) +
  labs(x = "Secondary Education (%)", y = "Average Life Expectancy (years)", title = "Predicted Values (Model 2)") +
  scale_y_continuous(limits = c(45,88), breaks = seq(45,85,5))
```
:::

::: {.column width="33%"}
```{r, fig.align='center', fig.asp=.85, fig.width=6}
plot(ggpredict(res3, "fertility_rate")) +
  labs(x = "Fertility Rate", y = "Average Life Expectancy (years)", title = "Predicted Values (Model 3)") +
  scale_y_continuous(limits = c(45,88), breaks = seq(45,85,5))
```
:::
::::

::: notes

**Ok, bottom line, which model is the most useful for explaining the variation in global life expectancy? Why?**

<br>

Today I'd like us to keep practicing the exercise of fitting, evaluating and interpreting OLS regressions

- **SLIDE**: Let's go back to the WDI data you selected and evaluated back in Week 3

:::



## {background-image="Images/background-data_blue_v4.png" .center .smaller}

::: {.r-fit-text}
**What explains the variation in maternal mortality ratios around the world?**
:::

- Data: Spring2026-WDI_Data_Class_Selections-SHORT_LABELS.xlsx

::: {.fragment}

```{r, fig.align='center', fig.asp=.618, out.width="80%", fig.width=7}
d |>
  ggplot(aes(x = maternal_mortality_ratio)) +
  geom_histogram(color = "white", bins = 15) +
  theme_bw() +
  labs(x = "Maternal Mortality Ratio (per 100,000 births)", y = "Count") +
  scale_x_continuous(breaks = seq(0, 1000, 200))
```

:::

::: notes

*Read slide*

<br>

The World Health Organization (WHO) defines the maternal mortality ratio as "the number of maternal deaths during a given time period per 100,000 live births during the same time period."

- It essentially captures the risk of death from a maternal cause in a single pregnancy (proxied by a single live birth).

<br>

We'll assume for the present purposes that this data is valid and reliable

- Let's begin with univariate analysis

- Everybody make a histogram and get ready to describe the variation in our measure

- (**REVEAL and DISCUSS**)

<br>

**Where is the US on this scale?**

- (Around 20 which is 59th in the world)

<br>

**What countries do this best?**

- (Belarus, Poland, UAE, Czechia, Israel, Norway)

<br>

**What countries are at the bad end of this distribution?**

- (Nigeria, Central African Republic, Chad, South Sudan, Liberia)

<br>

**SLIDE**: Ok, let's build some models to explain this variation

<br>

```{r, echo=TRUE}
d |> 
  select(country, maternal_mortality_ratio) |>
  arrange(maternal_mortality_ratio) |>
  mutate( rank = 1:217 ) |>
  filter(country == "United States")

d |> 
  select(country, maternal_mortality_ratio) |>
  arrange(maternal_mortality_ratio) |>
  slice_head(n = 7)

d |> 
  select(country, maternal_mortality_ratio) |>
  arrange(maternal_mortality_ratio) |>
  filter(!is.na(maternal_mortality_ratio)) |>
  slice_tail(n = 7)
```

:::




## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Predictors to Explain Maternal Mortality Ratios**
:::

```{r}
d_long |>
  select(-country, -ccode, -year) |>
  names()
```

::: notes

Ok, here's the list of variables you selected with their full informative labels

- The dataset on Canvas for today has shortened the labels for ease of analysis

<br>

Let's brainstorm some relationships to test with OLS regressions!

- *Track proposed models on the board*

- We won't stress causality too much but let's come up with AT LEAST THREE models to test. Hit me!

<br>

Now we have our work, let's get to it!

1. Fit the regressions

2. Evaluate the fit of each one

3. Evaluate the substantive significance of each one

- Go!

<br>

**SLIDE**: Some stuff I found...

```{r, eval=FALSE}
names(d_long)
names(d)
#summary(d)
# model1 <- lm(data = d, ~)
# summary(model1)
# plot(ggpredict(model1, ""))

# run correlations to find associations
d |>
  select(-country, -ccode, -year) |>
  cor(use = "pairwise")
```

:::



## {background-image="Images/background-data_blue_v4.png" .center}

```{r}
# SP26: replaced urban_population with education_spending
model1a <- lm(data = d, maternal_mortality_ratio ~ electricity_pct_pop)
model1b <- lm(data = d, maternal_mortality_ratio ~ basic_drinking_water)
model1c <- lm(data = d, maternal_mortality_ratio ~ education_spending)
model1d <- lm(data = d, maternal_mortality_ratio ~ fertility_rate)
#model1e <- lm(data = d, maternal_mortality_ratio ~ women_in_parliaments)

modelsummary(models = list(model1a, model1b, model1c, model1d), 
             stars = c("*" = .05),
             output = "gt", 
             fmt = 2, 
             gof_omit = "IC|Log|F",
             coef_map = c("electricity_pct_pop"= "Electricty (% of pop)",
                          "basic_drinking_water" = "Drinking Water (% of pop)",
                          "education_spending" = "Education Spending (% GDP)",
                          "urban_population" = "Urban Population (%)",
                          "fertility_rate" = "Fertility Rate",
                          "women_in_parliaments" = "Women in Parliament",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

::: notes

*Match the class selections*



:::



## {background-image="Images/background-data_blue_v4.png" .center .smaller}

::: {.panel-tabset}

### Electricity

```{r, fig.asp=.618, fig.align='center', out.width="80%", fig.width=7}
plot(ggpredict(model1a, "electricity_pct_pop")) +
  labs(x = "Electricity (% of population)", y = "Maternal Mortality Ratio") +
  scale_y_continuous(limits = c(-100,1200))
```

### Water

```{r, fig.asp=.618, fig.align='center', out.width="80%", fig.width=7}
plot(ggpredict(model1b, "basic_drinking_water")) +
  labs(x = "Basic Drinking Water (% of population)", y = "Maternal Mortality Ratio") +
  scale_y_continuous(limits = c(-100,1200))
```

### Education

```{r, fig.asp=.618, fig.align='center', out.width="80%", fig.width=7}
plot(ggpredict(model1c, "education_spending")) +
  labs(x = "Education Spending (% GDP)", y = "Maternal Mortality Ratio") +
  scale_y_continuous(limits = c(-100,1200))
```

### Fertility

```{r, fig.asp=.618, fig.align='center', out.width="80%", fig.width=7}
plot(ggpredict(model1d, "fertility_rate")) +
  labs(x = "Fertility Rate", y = "Maternal Mortality Ratio") +
  scale_y_continuous(limits = c(-100,1200))
```

### Women in Parliament

```{r, fig.asp=.618, fig.align='center', out.width="80%", fig.width=7, eval=FALSE}
plot(ggpredict(model1e, "women_in_parliaments")) +
  labs(x = "Electricty (% of population)", y = "Maternal Mortality Ratio") +
  scale_y_continuous(limits = c(-100,1200))
```

:::

::: notes


:::




## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**For Next Class: Multiple OLS Regression**

<br>

1. Read Chapter 12

2. Submit Exercises 1, 3 (skip part F) and 5

:::





