---
format: revealjs
---

# Today's Agenda {background-image="Images/background-data_blue_v3.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)

d <- read_excel("../Data_in_Class/Class11-1-Simple_OLS-WDI_LifeExpectancy/WDI-Practice_Simple_OLS-2024-04-04.xlsx", na = "NA")
```

<br>

::: {.r-fit-text}
**Multivariate Analyses**

- Descriptive Statistics and Visualizations
:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2025)
:::

::: notes
Prep for Class

1. Keep using WDI-Practice_Simple_OLS-2024-04-04.xlsx on Modules in Week 11

<br>

Everybody make sure they have, or redownload, the WDI data I posted in Week 11 on our Canvas Modules page

- We'll keep working with that data today
:::


## Building a Causal Argument {background-image="Images/background-slate_v2.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Predictor", "Confounders", "Outcome")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.45, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.1, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.1, arrow = arrow())
```

::: notes

Last week we started exploring the concepts and approaches we can use to make causal arguments

- e.g. arguments that changes in X cause a change in Y

<br>

We discussed:

- the Data Generating Process (DGP), 

- identification strategies, 

- the use of Directed Acyclic Graphs (DAGs), and 

- the need to adjust our analyses for confounders.

<br>

### Any questions on those big ideas?
:::



## Building a Causal Argument {background-image="Images/background-slate_v2.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("National Material\nCapabailities", "Confounders", "Human Rights and\nthe Rule of Law")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.6, xend = 2.3, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

::: notes
In class on Friday we worked to:

1. Draw a DAG for our key mechanism, and

2. Gather data that operationalizes the concepts we identified as important confounders in these relatonships.

<br>

### Any questions on those DAGs or the variables we selected?
:::



## Multivariate Analyses {background-image="Images/background-slate_v2.png" .center}

<br>

::: {.r-fit-text}
1. Descriptive statistics by groups
  
2. Using facets to extend any visualization
    
3. Mapping 3rd variables with colors, shapes or sizes
:::

::: notes
Today I will introduce you to a series of multivariate statistical tools

- These allow you to adjust your analyses for a confounder variable (or multiple confounder variables)

<br>

Today we'll focus on adapting to a single confounder, but know that these techniques can be used for multiple confounders (although it does get messy)

:::



## Do countries with higher average fertility rates have shorter life expectancies? {background-image="Images/background-slate_v2.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', fig.width = 7, fig.height=1.3, out.width='90%', cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(-3, 3),
  y = c(1, 1),
  labels = c("Fertility\nRate", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = -2.1, xend = 1.9, y = 1, yend = 1, arrow = arrow())
```

::: notes

Two weeks ago we used simple OLS models to explore global predictors of life expectancy

- One of our strongest findings was that there is an association between higher average fertility rates and considerably lower life expectancies

<br>

**SLIDE**: The question for us is can we build a causal argument that this is more than simple correlation?

:::



## Building a Causal Argument {background-image="Images/background-slate_v2.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

**Possible Confounders**: GDP per capita, unemployment rate, tobacco use, compulsory education

::: notes

Here is a list of the other variables in the dataset that we can consider as confounders in this relationship

<br>

**Remind me, what is the test for a confounder?**

- **When should we adjust our analysis for one of these variables?**

<br>

A confounder is a variable that explains some of the variation in both the treatment and the outcome variables.

1. The confounder MUST cause BOTH variables in the mechanism

2. The direction of effect is FROM the confounder TO the mechanism

<br>

I think it can be useful to think about this in terms of the "gold standard" of running a RCT (randomized controlled experiment)

- Do you believe that, on average, the high and low fertility rate countries are identical?

- If not, are the key differences also things that should influence life expectancy?

- If you believe they are not identical on average and the variable also influences the outcome then it is a confounder!

<br>

**Of these options, are we convinced all of them are likely confounders of this relationship? Why or why not?**

:::



## Multivariate Analyses: Approach 1 {background-image="Images/background-slate_v2.png" .center}

<br>

::: {.r-fit-text}
**Descriptive statistics by groups**

- categorical x categorical x categorical

- categorical x categorical x numeric
:::

::: notes

Data analysis principle 3 remains in effect: Variable type determines tool!

<br>

When your multivariate analysis is composed of mostly categorical variables then descriptive statistics by group is a viable approach

<br>

The WDI data we are working with is all numerical data, but we know how to convert numeric into categorical right?

- **What function do we use?**

- (**SLIDE**: Cut!)

:::



## Multivariate Analyses: Approach 1 {background-image="Images/background-slate_v2.png" .center}

<br>

**Identify the quartiles**
```{r, echo=TRUE}
summary(d$fertility_rate_per_woman)
```

<br>

**Convert numeric var to categorical**
```{r, echo=TRUE}
d$fertility2 <- cut(x = d$fertility_rate_per_woman, 
                    breaks = c(0, 1.558, 2.040, 3.257, 7), 
                    labels = c("Fertility (Q1)", "Fertility (Q2)", 
                               "Fertility (Q3)", "Fertility (Q4)"))
```

::: {.fragment}
**Practice: Create categorical versions of `gdp_per_capita` and `tobacco_use_pct`**
:::

::: notes

Refresh memories on using cut() to convert numeric to categorical

- Remember, your breaks argument needs one more value than labels!

- Defining the bins requires a start and end point for each bin

<br>

**Questions on the code?**

<br>

**REVEAL**: Practice this code by using it to convert gdp and tobacco using the same approach as shown here

- e.g. group the cases into quartiles

- Remember, quartiles mean four equally sized groups which can make it a useful way to describe variation in each group for comparison

<br>

**SLIDE**: Code results

:::



## Multivariate Analyses: Approach 1 {background-image="Images/background-slate_v2.png" .center}

<br>

```{r, echo=TRUE}
d$gdp_pc2 <- cut(x = d$gdp_per_capita, 
                 breaks = c(0, 2262.2, 6370.9, 22805.3, 183000),
                 labels = c("GDP pc (Q1)", "GDP pc (Q2)", 
                            "GDP pc (Q3)", "GDP pc (Q4)"))
```

```{r, echo=TRUE}
d$tobacco2 <- cut(x = d$tobacco_use_pct, 
                  breaks = c(0, 11.80, 20.30, 26.60, 50), 
                  labels = c("Tobacco (Q1)", "Tobacco (Q2)", 
                             "Tobacco (Q3)", "Tobacco (Q4)"))
```

**AUDIT: Check the group sizes**
```{r, echo=TRUE}
table(d$gdp_pc2)
```

::: notes

Always a good idea to audit the data work you do

- In this case checking the size of each category with table()

- Since we made quartiles each groups should be approximately the same size

<br>

**SLIDE**: Now that we have categorical variables we can build our first multivariate analysis

:::


## Multivariate Analyses: Approach 1 {background-image="Images/background-slate_v2.png" .center}

<br>

**Bivariate Descriptive Statistics**
```{r, echo=TRUE}
aggregate(data = d, life_expectancy_total ~ fertility2, FUN = mean)
```

<br>

```{r, echo=TRUE}
aggregate(data = d, life_expectancy_total ~ fertility2, FUN = median)
```


::: notes

**Does everybody remember the aggregate function from our bivariate work a few weeks ago?**

<br>

Uses the formula method for variable selection

- Any descriptive stats function can be used (mean, median, sd, summary, min, max, etc)

<br>

Here we see our main causal mechanism in action

- e.g. how mean life expectancy changes across fertility rate groupings

- In terms of both the means and medians, increasing average fertility rates is associated with a reduction in average life expectancy in 2020.

<br>

There are two ways to add a third variable to the aggregate function

- **SLIDE**: Option 1 is to use the subset argument

:::



## Multivariate Analyses: Approach 1 {background-image="Images/background-slate_v2.png" .center}

**The Subsets Argument**
```{r, echo=TRUE}
aggregate(data = d, life_expectancy_total ~ fertility2, FUN = mean,
          subset = gdp_per_capita < 6370.9)
```

<br>

```{r, echo=TRUE}
aggregate(data = d, life_expectancy_total ~ fertility2, FUN = mean,
          subset = gdp_per_capita > 6370.9)
```

::: notes

Exact same code as before but now include a subset argument

- Subset allows you to define what part of the dataset you want to focus on using relational symbols

- Any relational symbols can be used in subset (e.g. >, <, >=, <=, etc)

<br>

Here I am running the aggregate function one time on countries with GDP pc above the median and one time below the median

- In other words, I am estimating the relationship between fertility and life expectancy one time for countries below median wealth and one time above it

<br>

**Questions on the code?**

<br>

**Remind me, why are we doing this and what do these results indicate?**

<br>

This represents our first attempt to establish a causal argument

- Remember, our interest is on estimating the effect of fertility rate changes on average life expectancies

- Here we have adjusted our analysis for GDP pc as a confounder in the relationship

<br>

Per these results, adjusting our analyses for the wealth of the country does not change our estimated association!

- Put more assertively, IFF GDP pc is the only confounder then this is evidence that increasing fertility rates may reduce average life expectancies

- This is the kind of evidence we would provide to make a causal argument

<br>

**Make sense?**

<br>

Now this is a very weak attempt to adjust for wealth

- Just above and below the median

- **SLIDE**: Let's try another technique to adjust for more levels of wealth

:::



## Multivariate Analyses: Approach 1 {background-image="Images/background-slate_v2.png" .center}

<br>

**Adding Extra Categorical Variables**
```{r, echo=TRUE}
aggregate(data = d, life_expectancy_total ~ fertility2 + gdp_pc2, 
          FUN = mean)
```

::: notes

The second way to modify the aggregate function is to add additional categorical variables to the formula argument

- This creates a long set of results that shows the mean life expectancy for each combination of fertility and gdp quartiles

<br>

**Everybody get this code to work?**

<br>

**Do our results hold up to adjusting for four levels of state wealth?**

- We can see the effect of interest (fertility -> life exp) holds up in all instances except the richest countries!

- Apparently, in rich countries it looks like the negative effects of fertility rate may be reduced!

<br>

**Any questions about using aggregate for multivariate analyses?**

<br>

**SLIDE**: Practice on tobacco use

:::


## Multivariate Analyses: Approach 1 {background-image="Images/background-slate_v2.png" .center}

<br>

**What is the impact of adjusting for tobacco use as a confounder?**

<br>

::: {.fragment}
```{r, echo=FALSE}
aggregate(data = d, life_expectancy_total ~ fertility2 + tobacco2, 
          FUN = mean) |>
  pivot_wider(names_from = fertility2, values_from = life_expectancy_total) |>
  arrange(tobacco2) |>
  kbl(digits = 1, align = c("l", rep("c", 4))) |>
  kableExtra::kable_styling(font_size = 30)
```
:::

::: notes
Now everybody practice this one time using the categorical version of tobacco use

- Check to see if the negative effect of fertility on life expectancy remains when adjusting for the quartiles of tobacco use

<br>

(**REVEAL**: Results in wide format for ease of reading)

- Read results within each row and across the columns, e.g. countries with lowest tobacco use is row 1, how does their life expectancy change as fertility increases?

<br>

**What do we learn from this?**

- IFF tobacco use is the only confounder then this is evidence that increasing fertility rates reduces average life expectancies
:::



## Multivariate Analyses: Approach 2 {background-image="Images/background-slate_v2.png" .center}

<br>

**If you are adding a categorical variable you can use a facet wrap on ANY visualization**


::: notes

Multivariate approach 2 uses facet wraps

- IFF the third variable you want to add to any analysis is categorical, you can use a facet wrap to do it
:::



## Multivariate Analyses: Approach 2 {background-image="Images/background-slate_v2.png" .center}

```{r, echo=TRUE, fig.asp=0.75, fig.align='center', fig.width=9, cache=TRUE}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  theme_bw()
```

::: notes

Here's our bivariate scatter plot of the key association we're analyzing

- I'm going to omit the polish and just focus on the new techniques here.

- I've also added the geom_smooth function to add an OLS line of best fit to our plot

<br>

**Everybody clear on this code?**

<br>

**SLIDE**: Let's now adjust this analysis for GDP pc using a facet wrap

:::



## Multivariate Analyses: Approach 2 {background-image="Images/background-slate_v2.png" .center}

<br>

```{r, echo=TRUE, fig.asp=0.25, fig.align='center', fig.width=14, cache=TRUE, `code-line-numbers`="5"}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  theme_bw() +
  facet_wrap(~ gdp_pc2, ncol = 5)
```

::: notes

We worked with facet wraps a few weeks ago when doing bivariate analyses

- **Questions on the code?**

<br>

**What do we learn from this effort to adjust the relationship for the wealth confounder?**

<br>

- Some cases missing GDP data, but fortunately, the relationship there appears similar to the others

- The negative slope is evident in each facet

- If our identification hinged on adjusting for wealth this represents another step in the right direction

<br>

Now everybody practice this one time using the categorical version of tobacco use

- Check to see if the negative effect of fertility on life expectancy remains when adjusting for the quartiles of tobacco use

- (**SLIDE**)

:::


## Multivariate Analyses: Approach 2 {background-image="Images/background-slate_v2.png" .center}

<br>

```{r, echo=TRUE, fig.asp=0.25, fig.align='center', fig.width=14}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  theme_bw() +
  facet_wrap(~ tobacco2, ncol = 5)
```

::: notes

**What do we learn from this effort to adjust the relationship for tobacco use as a confounder?**

<br>

**SLIDE**: One more multivariate approach for today

:::



## Multivariate Analyses: Approach 3 {background-image="Images/background-slate_v2.png" .center}

<br>

::: {.r-fit-text}
**You can adapt scatter plots to confounders**

- Categorical confounder: Colors and/or shapes

- Numeric confounder: Vary the sizes
:::

::: notes

Multivariate approach 3 focuses on the points in a scatter plot

- IFF the third variable you want to add to any analysis is categorical, you can use colors and shapes

- IFF the third variable you want to add to any analysis is numeric, you can use sizes
:::




## Multivariate Analyses: Approach 3 {background-image="Images/background-slate_v2.png" .center}

**Categorical Confounder: Use Color and/or Shapes**

```{r, echo=TRUE, fig.asp=0.618, fig.align='center', fig.width=8, `code-line-numbers`="3"}
ggplot(d, aes(x = fertility_rate_per_woman, 
              y = life_expectancy_total, 
              color = gdp_pc2)) +
  geom_point() + 
  theme_bw()
```

::: notes

TBD

:::



## Multivariate Analyses: Approach 3 {background-image="Images/background-slate_v2.png" .center}

**Categorical Confounder: Use Color and/or Shapes**

```{r, echo=TRUE, fig.asp=0.618, fig.align='center', fig.width=8, `code-line-numbers`="3"}
ggplot(d, aes(x = fertility_rate_per_woman, 
              y = life_expectancy_total, 
              shape = gdp_pc2)) +
  geom_point() + 
  theme_bw()
```

::: notes

If printing in B&W

:::



## Multivariate Analyses: Approach 3 {background-image="Images/background-slate_v2.png" .center}

**Categorical Confounder: Use Color and/or Shapes**

```{r, echo=TRUE, fig.asp=0.618, fig.align='center', fig.width=8, `code-line-numbers`="3"}
ggplot(d, aes(x = fertility_rate_per_woman, 
              y = life_expectancy_total, 
              color = gdp_pc2, shape = gdp_pc2)) +
  geom_point() + 
  theme_bw()
```

::: notes

Use Both!

<br>

If there aren't too many observations or there is considerable spacing between groups this can be a really effective visualization technique

<br>

Everybody practice this replacing GDP with tobacco again

- (**SLIDE**)

:::



## Multivariate Analyses: Approach 3 {background-image="Images/background-slate_v2.png" .center}

**Categorical Confounder: Use Color and/or Shapes**

```{r, echo=TRUE, fig.asp=0.618, fig.align='center', fig.width=8, `code-line-numbers`="3"}
ggplot(d, aes(x = fertility_rate_per_woman, 
              y = life_expectancy_total, 
              color = tobacco2, shape = tobacco2)) +
  geom_point() + 
  theme_bw()
```

::: notes

I don't think this actually clarifies the test given the overlap in the groups

- **SLIDE**: So combine approaches and re-add the facet wrap!

:::



## Multivariate Analyses: Approach 3 {background-image="Images/background-slate_v2.png" .center}

**Categorical Confounder: Use Color and/or Shapes**

```{r, echo=TRUE, fig.asp=0.25, fig.align='center', fig.width=14, `code-line-numbers`="3,6"}
ggplot(d, aes(x = fertility_rate_per_woman, 
              y = life_expectancy_total, 
              color = tobacco2, shape = tobacco2)) +
  geom_point() + 
  theme_bw() +
  facet_wrap(~ tobacco2, ncol = 5)
```

::: notes

**Any questions on using colors or shapes for categorical confounders?**

<br>

**SLIDE**: Adding numeric confounders is also a bit clunky but I'd like you to be aware of the option

:::



## Multivariate Analyses: Approach 3 {background-image="Images/background-slate_v2.png" .center}

**Numerical Confounder: Use Sizing**

```{r, echo=TRUE, fig.asp=0.618, fig.align='center', fig.width=8, `code-line-numbers`="3"}
ggplot(d, aes(x = fertility_rate_per_woman, 
              y = life_expectancy_total, 
              size = gdp_per_capita)) +
  geom_point() + 
  theme_bw()
```

::: notes

Size will rescale the points per a numeric variable

- Some refer to these as bubble plots

<br>

Shows us there is a clear separation between the rich and poor world in this relationship!

<br>

Everybody practice this on tobacco!

- (**SLIDE**)

:::



## Multivariate Analyses: Approach 3 {background-image="Images/background-slate_v2.png" .center}

**Numerical Confounder: Use Sizing**

```{r, echo=TRUE, fig.asp=0.618, fig.align='center', fig.width=8, `code-line-numbers`="3"}
ggplot(d, aes(x = fertility_rate_per_woman, 
              y = life_expectancy_total, 
              size = tobacco_use_pct)) +
  geom_point() + 
  theme_bw()
```

::: notes

Tobacco use much more spread out across this relationship

- Less distinction rich vs poor on this one
:::



## Multivariate Analyses {background-image="Images/background-slate_v2.png" .center}

<br>

::: {.r-fit-text}
1. Descriptive statistics by groups
  
2. Using facets to extend any visualization
    
3. Mapping 3rd variables with colors, shapes or sizes
:::

::: notes

**Any questions on the multivariate techniques we covered today?**

<br>

**SLIDE**: Practice for next class!

:::


## For Next Class {background-image="Images/background-slate_v2.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

**Possible Confounders**: GDP per capita, tobacco use, unemployment rate, compulsory education

::: notes

We covered GDP and tobacco in class today

- Your job for Wednesday is to replicate our work today using the other two possible confounders

- Does the central mechanism hold up when we adjust the analyses for either unemployment or years of compulsory education?

<br>

**Questions?**

:::

