---
format: revealjs
---

## Today's Agenda {background-image="Images/background-data_blue_v4.png" .center}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)

# Create new dataset with categorical data for class to work with
# d <- read_excel("../Data_in_Class/Class11-1-Simple_OLS-WDI_LifeExpectancy/WDI-Practice_Simple_OLS-2024-04-04.xlsx", na = "NA")
# 
# d$gdp_pc_cat <- cut(x = d$gdp_per_capita, breaks = c(0, 2262.2, 6370.9, 22805.3, 183000), labels = c("GDP pc (Q1)", "GDP pc (Q2)", "GDP pc (Q3)", "GDP pc (Q4)"))
# 
# d$tobacco_cat <- cut(x = d$tobacco_use_pct, breaks = c(0, 11.80, 20.30, 26.60, 50), labels = c("Tobacco (Q1)", "Tobacco (Q2)", "Tobacco (Q3)", "Tobacco (Q4)"))
# 
#write_excel_csv(d, file = "../Data_in_Class/WDI-Multivariate_Viz-2024-04-04.csv")

d <- read_excel("../Data_in_Class/World_Development_Indicators/WDI-Multivariate_Viz-2024-04-04.xlsx", na = "NA")

data1 <- read_excel("../Data_in_Class/Data-2024-EPI_RSF-Merged_w_Confounders.xlsx", na = "NA")
```

<br>

::: {.r-fit-text}
**Multivariate Analyses**

- Descriptive Statistics and Visualizations

- Data: WDI-Multivariate_Viz-2024-04-04.xlsx
:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2025)
:::

::: notes
Prep for Class

1. Review Canvas submissions

2. Update histograms for confounders we selected last class

3. Data: WDI-Multivariate_Viz-2024-04-04.xlsx

<br>

Everybody make sure they have, or redownload, the WDI data I posted in Week 11 on our Canvas Modules page

- We'll keep working with that data today

<br>

**SLIDE**: Last week...

:::



## Building a Causal Argument {background-image="Images/background-data_blue_v4.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Predictor", "Confounders", "Outcome")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.45, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.1, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.1, arrow = arrow())
```

::: notes

Last week we started exploring the concepts and approaches we can use to make causal arguments

- e.g. arguments that changes in X cause a change in Y

<br>

We discussed:

- the Data Generating Process (DGP), 

- identification strategies, 

- the use of Directed Acyclic Graphs (DAGs), and 

- the need to adjust our analyses for confounders.

<br>

**Any questions on those big ideas?**

:::



## Building a Causal Argument {background-image="Images/background-data_blue_v4.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Press\nFreedom", "Confounders", "Environmental\nQuality")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.6, xend = 2.3, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

::: notes

In class on Friday we worked to:

1. Draw a DAG for our key mechanism, and

2. Gather data that operationalizes the concepts we identified as important confounders in these relatonships.

<br>

**Any questions on those DAGs or the variables we selected?**

<br>

For today I asked each of you to perform univariate analyses on our chosen confounders

- **How did that go?**

<br>

**Do we have sufficient variation in each to be useful?**

- **In other words, do the countries of the world actually vary in their levels of these confounders?**

- (**SLIDE**: Histograms of each)

:::



## Analyzing the Confounders {background-image="Images/background-data_blue_v4.png" .center}

```{r, fig.retina=3, fig.align='center'}
data1 |>
  select(gdp_current_usd2023:population2023) |>
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Values") |>
  ggplot(aes(x = Values)) +
  geom_histogram() +
  theme_bw() +
  facet_wrap(~ Variable, scales = "free")
```

::: notes

**What do we learn about the world from these distributions?**

```{r}
data1 |>
  select(gdp_current_usd2023, gdp_pc_current_usd2023, population2023) |>
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Values") |>
  group_by(Variable) |>
  summarize(
    Mean = mean(Values, na.rm = T),
    SD = sd(Values, na.rm = T),
    Min = min(Values, na.rm = T),
    Max = max(Values, na.rm = T),
    Missing = sum(is.na(Values))
  )
```

```{r}
data1 |>
  select(compulsory_ed_years2023, control_corruption_estimate2023, control_corruption_percentile2023) |>
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Values") |>
  group_by(Variable) |>
  summarize(
    Mean = mean(Values, na.rm = T),
    SD = sd(Values, na.rm = T),
    Min = min(Values, na.rm = T),
    Max = max(Values, na.rm = T),
    Missing = sum(is.na(Values))
  )
```



<br>

Quick aside, DO NOT include figures in your reports with exponents in the axis labels OR in the regression results.

- Rescale the variables and replot them

<br>

Let's practice this one time by rescaling GDP into billions of USD

- **SLIDE**: How-to

:::



## Keep those axis labels readable! {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}

```{r, fig.retina=3, fig.align='center', echo=TRUE, out.width="55%", fig.asp=.75, fig.width = 6}
# Rescale GDP into trillions USD
data1$gdp_trillions <- data1$gdp_current_usd2023 / 1e12

# Visualize the new variable
ggplot(data = data1, aes(x = gdp_trillions)) +
  geom_histogram()
```

:::

::: notes

The process is fairly simple

- Create a new variable that is a modification of the current one

- Divide by a million to rescale into millions

- Divide by a trillion to rescale into trillions

<br>

**Questions on rescaling a variable?**

<br>

**SLIDE**: Plan for today

:::



## Multivariate Analyses 1 {background-image="Images/background-data_blue_v4.png" .center}

<br>

::: {.r-fit-text}
**Working with Categorical Variables as Confounders**

1. Descriptive statistics by groups
  
2. Multiple Facets
    
3. Adding colors, shapes and sizes
:::

::: notes
Today I will introduce you to a series of multivariate statistical tools

- These tools allow you to adjust your visualizations and descriptive statistics for a confounder variable

- HOWEVER, these tools, and our work this week, will focus ONLY on adding categorical variables to our analyses

- Adding numeric variables is something we will do using mulitple regression next week

<br>

**SLIDE**: Our practice data

:::



## Does increasing fertility rates directly lower life expectancy around the world? {background-image="Images/background-data_blue_v4.png" .center}

<br>

```{r, fig.retina = 3, fig.align = 'center', fig.width = 7, fig.height=1.3, out.width='90%', cache=TRUE}
# Do countries with higher average fertility rates have shorter life expectancies?

## Manual DAG
d1 <- tibble(
  x = c(-3, 3),
  y = c(1, 1),
  labels = c("Fertility\nRate", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = -2.1, xend = 1.9, y = 1, yend = 1, arrow = arrow())
```

::: notes

Two weeks ago we used simple OLS models to explore global predictors of life expectancy

- One of our strongest findings was that countries with average fertility rates have lower average life expectancies

<br>

**SLIDE**: The question for us is, can we use statistics to establish that this is a causal effect and not just a correlation?

:::



## Building a Causal Argument {background-image="Images/background-data_blue_v4.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

**Possible Confounders**: GDP per capita, unemployment rate, tobacco use, compulsory education

::: notes

Here is a list of the other variables in the dataset that we can consider as confounders in this relationship

<br>

**Remind me, what is the test for a confounder?**

- **When should we adjust our analysis for one of these variables?**

<br>

A confounder is a variable that explains some of the variation in both the treatment and the outcome variables.

1. The confounder MUST cause BOTH variables in the mechanism

2. The direction of effect is FROM the confounder TO the mechanism

<br>

I think it can be useful to think about this in terms of the "gold standard" of running a RCT (randomized controlled experiment)

- Do you believe that, on average, the high and low fertility rate countries are identical?

- If not, are the key differences also things that should influence life expectancy?

- If you believe they are not identical on average and the variable also influences the outcome then it is a confounder!

<br>

**Of these options, are we convinced all of them are likely confounders of this relationship? Why or why not?**

<br>

Now, remember our techniques for today are all designed for when you have categorical variables as confounders

- None of these are currently categorical!

- **SIDE**: So, our first job is to create categorical versions of each of these variables

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Convert Numeric Variables to Categorical**
:::

```{r, echo=TRUE}
# Get the Quartiles
summary(d$fertility_rate_per_woman)
```

<br>

```{r, echo=TRUE}
# Make the categorical version
d$fertility_cat <- cut(x = d$fertility_rate_per_woman, 
                    breaks = c(0, 1.558, 2.040, 3.257, 7), 
                    labels = c("Fertility (Q1)", "Fertility (Q2)", 
                               "Fertility (Q3)", "Fertility (Q4)"))

# Check the result
table(d$fertility_cat)
```

::: notes

We have previously used the cut() function to convert numeric variables into categorical variables

- Let's practice that skill again now!

<br>

Everybody copy this code using the cut function to make a categorical version of fertility

- **Does everybody remember this approach?**

<br>

For ease of our work today I have done this for you for the GDP and tobacco variables

- Both are in today's dataset with "_cat" at the end.

<br>

**SLIDE**: with our categorical confounders in hand we can start on our first multivariate approach...

:::





## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Descriptive Statistics by Groups**

<br>

**Step 1: Start with the bivariate analysis**

```{r, echo=TRUE}
aggregate(data = d, life_expectancy_total ~ fertility_cat, 
          FUN = mean)
```
:::

::: notes

Step 1 to taking your descriptive statistics into multivariate analyses starts with the aggregate function

- Remember, this is what we used in our bivariate analyses.

- Any descriptive stats function can be used (mean, median, sd, summary, min, max, etc)

<br>

Here we see our main causal mechanism in action

- e.g. how mean life expectancy changes across fertility rate groupings

- Life expectancy decreases are associated with increases in the fertility rate

<br>

There are two ways to add a third variable to the aggregate function

- **SLIDE**: Option 1 is to use the subset argument

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Descriptive Statistics by Groups**

<br>

**Step 2a: Add a subset (numeric)**

```{r, echo=TRUE}
aggregate(data = d, life_expectancy_total ~ fertility_cat, 
          FUN = mean,
          subset = gdp_per_capita < 6370.9)
```

:::

::: notes

Exact same code as before but now include a subset argument

- Subset allows you to define what part of the dataset you want to focus on using relational symbols

- Any relational symbols can be used in subset (e.g. >, <, >=, <=, etc)

<br>

I know I told you our focus today would be on categorical confounders, but, hey, this works too!

- **Everybody clear on this code?**

<br>

**Does our association between fertility rate and life expectancy hold amongst the poorer countries in the world?**

<br>

**SLIDE**: Rather than having to run this code repeatedly for different subsets lets use our quartiles variable

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Descriptive Statistics by Groups**

<br>

**Step 2a: Add extra variables (categorical)**

```{r, echo=TRUE}
aggregate(data = d, life_expectancy_total ~ fertility_cat + gdp_pc_cat, 
          FUN = mean)
```

:::

::: notes

The second way to modify the aggregate function is to add additional categorical variables to the formula argument

- This creates a long set of results that shows the mean life expectancy for each combination of fertility and gdp quartiles

<br>

**Everybody get this code to work?**

- Note that there's a catch to these results!

- In the poorest states (GDP pc Q1) there are NO states with the lowest fertility rates!

- In the richest states it's the opposite!

<br>

**So, do our results hold up to adjusting for four levels of state wealth?**

- We can see the effect of interest (fertility -> life exp) holds up in all instances except the richest countries!

- Apparently, in rich countries it looks like the negative effects of fertility rate may be reduced!

<br>

This means IFF our causal argument specified GDP pc as the only confounder then this analysis is evidence of a causal relationship and not just correlation!

- **Does everybody understand this distinction?**

<br>

**Any questions about using aggregate for multivariate analyses?**

<br>

**SLIDE**: Practice on tobacco use

:::


## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Descriptive Statistics by Groups**

<br>

**Replicate but replace GDP with tobacco as a confounder**
:::

::: {.fragment}
```{r, echo=FALSE}
aggregate(data = d, life_expectancy_total ~ fertility_cat + tobacco_cat, 
          FUN = mean) |>
  pivot_wider(names_from = fertility_cat, values_from = life_expectancy_total) |>
  arrange(tobacco_cat) |>
  kbl(digits = 1, align = c("l", rep("c", 4))) |>
  kableExtra::kable_styling(font_size = 30)
```
:::

::: notes
Now everybody practice this one time using the categorical version of tobacco use

- Check to see if the negative effect of fertility on life expectancy remains when adjusting for the quartiles of tobacco use

<br>

(**REVEAL**: Results in wide format for ease of reading)

- Read results within each row and across the columns, e.g. countries with lowest tobacco use is row 1, how does their life expectancy change as fertility increases?

<br>

**What do we learn from this?**

- IFF tobacco use is the only confounder then this is evidence that increasing fertility rates reduces average life expectancies
:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Extend a Visualization with Facets**
:::

<br>

```{r, echo=TRUE, fig.retina=3, fig.asp=0.7, fig.align='center', fig.width=7}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  theme_bw()
```


::: notes

Multivariate approach 2 uses facet wraps

- IFF the third variable you want to add to any analysis is categorical, you can use a facet wrap to do it

<br>

Just like before we begin with our bivariate analysis

- I'm going to omit the polish and just focus on the new techniques here.

- I've also added the geom_smooth function to add an OLS line of best fit to our plot

<br>

**Everybody clear on this code?**

<br>

**SLIDE**: Let's now adjust this analysis for GDP pc using a facet wrap

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Extend a Visualization with Facets**
:::

<br>

```{r, echo=TRUE, fig.asp=0.25, fig.align='center', fig.width=14, cache=TRUE}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  theme_bw() +
  facet_wrap(~ gdp_pc_cat, ncol = 5)
```

::: notes

The ONLY change here is the facet_wrap line (line 5)

- We worked with facet wraps a few weeks ago when doing bivariate analyses

- **Questions on the code?**

<br>

**What do we learn from this effort to adjust the relationship for the wealth confounder?**

<br>

- Some cases missing GDP data, but fortunately, the relationship there appears similar to the others

- The negative slope is evident in each facet

- If our identification hinged on adjusting for wealth this represents another step in the right direction

<br>

Now everybody practice this one time using the categorical version of tobacco use

- Check to see if the negative effect of fertility on life expectancy remains when adjusting for the quartiles of tobacco use

- (**SLIDE**)

:::


## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Extend a Visualization with Facets**
:::

<br>

```{r, echo=TRUE, fig.asp=0.25, fig.align='center', fig.width=14}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  theme_bw() +
  facet_wrap(~ tobacco_cat, ncol = 5)
```

::: notes

**What do we learn from this effort to adjust the relationship for tobacco use as a confounder?**

<br>

**SLIDE**: One more multivariate approach for today

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Adding colors, shapes and sizes**
:::

```{r, echo=TRUE, fig.retina=3, fig.asp=0.618, fig.align='center', fig.width=8}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total)) +
  geom_point() + 
  theme_bw()
```

::: notes

Ok, let's reset our code again!

<br>

Multivariate approach 3 focuses on the points in a scatter plot

- **SLIDE**: IFF the third variable you want to add to any analysis is categorical, you can use colors and shapes

:::




## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Adding colors, shapes and sizes**
:::

```{r, echo=TRUE, fig.retina=3, fig.asp=0.618, fig.align='center', fig.width=8}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total, 
              color = gdp_pc_cat)) +
  geom_point() + 
  theme_bw()
```

::: notes

You can use all of our other tools for adjusting these colors

- scale_color_manual, scale_color_brewer, etc

<br>

**Questions on this code?**

<br>

**SLIDE**: working in B&W

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Adding colors, shapes and sizes**
:::

```{r, echo=TRUE, fig.asp=0.618, fig.align='center', fig.width=8, fig.retina=3}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total, 
              shape = gdp_pc_cat)) +
  geom_point(size = 2) + 
  theme_bw()
```

::: notes

If printing in B&W use shapes

- Usually helpful to increase the size of the points

<br>

**SLIDE**: Use both!

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Adding colors, shapes and sizes**
:::

```{r, echo=TRUE, fig.asp=0.618, fig.align='center', fig.width=8, fig.retina=3}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total, 
              color = gdp_pc_cat, shape = gdp_pc_cat)) +
  geom_point(size = 2) + 
  theme_bw()
```

::: notes

If there aren't too many observations or there is considerable spacing between groups this can be a really effective visualization technique

<br>

Everybody practice this replacing GDP with tobacco again

- (**SLIDE**)

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Adding colors, shapes and sizes**
:::

```{r, echo=TRUE, fig.asp=0.618, fig.align='center', fig.width=8, fig.retina=3}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total, 
              color = tobacco_cat, shape = tobacco_cat)) +
  geom_point(size = 2) + 
  theme_bw()
```

::: notes

I don't think this actually clarifies the test given the overlap in the groups

- **SLIDE**: So combine approaches and re-add the facet wrap!

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Adding colors, shapes and sizes**
:::

```{r, echo=TRUE, fig.asp=0.25, fig.align='center', fig.width=14, fig.retina=3}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total, 
              color = tobacco_cat, shape = tobacco_cat)) +
  geom_point() + 
  theme_bw() +
  facet_wrap(~ tobacco_cat, ncol = 5)
```

::: notes

**Any questions on using colors or shapes for categorical confounders?**

<br>

**SLIDE**: Adding numeric confounders is also a bit clunky but I'd like you to be aware of the option

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Adding colors, shapes and sizes**
:::

```{r, echo=TRUE, fig.asp=0.618, fig.align='center', fig.width=8, fig.retina=3}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total, 
              size = gdp_per_capita)) +
  geom_point() + 
  theme_bw()
```

::: notes

Size will rescale the points per a numeric variable

- Some refer to these as bubble plots

<br>

Shows us there is a clear separation between the rich and poor world in this relationship!

<br>

Everybody practice this on tobacco!

- (**SLIDE**)

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Multivariate Analyses: Adding colors, shapes and sizes**
:::

```{r, echo=TRUE, fig.asp=0.618, fig.align='center', fig.width=8, fig.retina=3}
ggplot(d, aes(x = fertility_rate_per_woman, y = life_expectancy_total, 
              size = tobacco_use_pct)) +
  geom_point() + 
  theme_bw()
```

::: notes

Tobacco use much more spread out across this relationship

- Less distinction rich vs poor on this one
:::



## Multivariate Analyses 1 {background-image="Images/background-data_blue_v4.png" .center}

<br>

::: {.r-fit-text}
**Working with Categorical Variables as Confounders**

1. Descriptive statistics by groups
  
2. Multiple Facets
    
3. Adding colors, shapes and sizes
:::

::: notes

**Any questions on the multivariate techniques we covered today?**

<br>

**SLIDE**: Practice for next class!

:::


## For Next Class {background-image="Images/background-data_blue_v4.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

**Possible Confounders**: unemployment rate, compulsory education

::: notes

We covered GDP and tobacco in class today

- Your job for Wednesday is to replicate our work today using the other two possible confounders

- Does the central mechanism hold up when we adjust the analyses for either unemployment or years of compulsory education?

<br>

**Questions?**

:::

