---
format: revealjs
---

# Today's Agenda {background-image="libs/Images/background-data_blue_v3.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)

d <- read_excel("../Data_in_Class-SP24/World_Development_Indicators/Practice_Simple_OLS/WDI-Practice_Simple_OLS-2024-04-04.xlsx", na = "NA")
```

<br>

::: {.r-fit-text}

**Apply multivariate analyses to our research project**

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2024)
:::

::: notes
Prep for Class

1. Bring a usb with the data file to class so we can live code together

<br>

Last class I introduced you to a series of tools we can use to begin performing multivariate analyses

1. Descriptive statistics by groups
  
2. Using facets to extend any visualization
    
3. Mapping 3rd variables with colors, shapes or sizes

<br>

**SLIDE**: Assignment for today
:::


## For Today {background-image="libs/Images/background-slate_v2.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

**Possible Confounders**: GDP per capita, tobacco use, unemployment rate, compulsory education

::: notes

Your job for today was to use these new tools to examine the other two possible confounders

- Does the central mechanism hold up when we adjust the analyses for either unemployment or years of compulsory education?

<br>

**How did this go?**

:::



## Adjusting for Unemployment {background-image="libs/Images/background-slate_v2.png" .center}

```{r, echo=FALSE, fig.asp=.85, fig.align='center', fig.width=8}
d |>
  mutate(
    unemployment2 = case_when(
      unemployment_pct < quantile(d$unemployment_pct, probs = .25, na.rm = T) ~ "Unemployment (Q1)",
      unemployment_pct < quantile(d$unemployment_pct, probs = .50, na.rm = T) ~ "Unemployment (Q2)",
      unemployment_pct < quantile(d$unemployment_pct, probs = .75, na.rm = T) ~ "Unemployment (Q3)",
      unemployment_pct >= quantile(d$unemployment_pct, probs = .75, na.rm = T) ~ "Unemployment (Q4)"
      )
    ) |>
  select(fertility_rate_per_woman, life_expectancy_total, unemployment2) |>
  na.omit() |>
  ggplot(aes(x = fertility_rate_per_woman, y = life_expectancy_total, color = unemployment2, shape = unemployment2)) +
  geom_point() + 
  geom_smooth(method = "lm", se = T) +
  theme_bw() +
  facet_wrap(~ unemployment2, ncol = 2) +
  guides(color = "none", shape = "none") +
  labs(x = "Fertility Rate (births per woman)", y = "Life Expectancy (years)",
       caption = "Source: WDI (2024) for data from 2020")
```

::: notes

Note: I removed the NAs by filtering the dataset before making the plots

<br>

**Does adjusting for unemployment change the key relationship of interest?**

- **What is our takeaway from this?**

- IFF our identification strategy requires adjusting for unemployment then these correlations may be causal!

<br>

(**SLIDE**: Code to remove missing data on next slide)

:::



## Remove the Missing Data {background-image="libs/Images/background-slate_v2.png" .center}

```{r, echo=TRUE, eval=FALSE}
# Create new variable
d$unemp2 <- cut(d$unemployment_pct, 
            breaks = c(0, 4.191, 6.453, 11.718, 28),
            labels = c("Unemployment (Q1)", "Unemployment (Q2)", 
                       "Unemployment (Q3)", "Unemployment (Q4)"))

# Make a new dataset with only the needed variables
d_subset <- select(.data = d, 
                   fertility_rate_per_woman, 
                   life_expectancy_total, 
                   unemp2)

# Make a new dataset without the missing data
d2 <- na.omit(d_subset)

# Make the visualization with d2
```

::: notes

Two new functions here

<br>

FIRST, select() lets you create a new data object with only some of the variables in the original dataset

- Here we are creating a copy of the dataset with only three variables in it

<br>

SECOND, na.omit() removes any observation from the dataset if it has missing data in ANY of the variables

- Be very careful with this!

- If you apply na.omit without narrowing the variables using select then you may lose a ton of data you didn't mean to remove

:::


## Adjusting for Education {background-image="libs/Images/background-slate_v2.png" .center}

```{r, echo=FALSE, fig.asp=.85, fig.align='center', fig.width=8}
d |>
  mutate(
    education2 = case_when(
      compulsory_education_yrs < quantile(d$compulsory_education_yrs, probs = .25, na.rm = T) ~ "Education (Q1)",
      compulsory_education_yrs < quantile(d$compulsory_education_yrs, probs = .50, na.rm = T) ~ "Education (Q2)",
      compulsory_education_yrs < quantile(d$compulsory_education_yrs, probs = .75, na.rm = T) ~ "Education (Q3)",
      compulsory_education_yrs >= quantile(d$compulsory_education_yrs, probs = .75, na.rm = T) ~ "Education (Q4)"
      )
    ) |>
  select(fertility_rate_per_woman, life_expectancy_total, education2) |>
  na.omit() |>
  ggplot(aes(x = fertility_rate_per_woman, y = life_expectancy_total, color = education2, shape = education2)) +
  geom_point() + 
  geom_smooth(method = "lm", se = T) +
  theme_bw() +
  facet_wrap(~ education2, ncol = 2) +
  guides(color = "none", shape = "none") +
  labs(x = "Fertility Rate (births per woman)", y = "Life Expectancy (years)",
       caption = "Source: WDI (2024) for data from 2020")
```

::: notes

**Does adjusting for education change the key relationship of interest?**

:::



## Multivariate Analyses {background-image="libs/Images/background-slate_v2.png" .center}

<br>

::: {.r-fit-text}
1. Descriptive statistics by groups
  
2. Using facets to extend any visualization
    
3. Mapping 3rd variables with colors, shapes or sizes
:::

::: notes

**Any questions on these techniques for adding confounders to our analyses?**

:::




## Causal Arguments {background-image="libs/Images/background-slate_v2.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("National Material\nCapabailities", "Confounders", "Human Rights and\nthe Rule of Law")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.6, xend = 2.3, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())

## Input project data
d <- read_excel("../Data_in_Class-SP24/Merged_Dataset-2006-2016_with_confounders.xlsx", na = "NA")
```

**Possible Confounders**: GDP, Immunizations DPT, Infant mortality & Compulsory education

::: notes

Today we work to apply our new techniques to our ongoing class project!

- GDP (constant 2015 US$)
    - (Source: World Bank national accounts data, and OECD National Accounts data files.)
    
- Immunization, DPT (% of children ages 12-23 months)
    - (Source: WHO and UNICEF (http://www.who.int/immunization/monitoring_surveillance/en/).)

- Mortality rate, infant (per 1,000 live births)
    - (Source: Estimates developed by the UN Inter-agency Group for Child Mortality Estimation (UNICEF, WHO, World Bank, UN DESA Population Division) at www.childmortality.org.)

- Compulsory education, duration (years)
    - (Source: UNESCO Institute for Statistics (UIS). UIS.Stat Bulk Data Download Service. Accessed September 19, 2023. https://apiportal.uis.unesco.org/bdds.)

<br>

**SLIDE**: Step 1 check for variation in the new measures!

:::



## Adjusting for Confounders {background-image="libs/Images/background-slate_v2.png" .center}
:::: {.columns}
::: {.column width="50%"}
```{r, fig.align='center', fig.asp=1, fig.width=7}
d |>
  select(gdp_2015usd_billions, immunization_dpt, compulsory_education, infant_mortality) |>
  pivot_longer(cols = gdp_2015usd_billions:infant_mortality, names_to = "Confounders", values_to = "Values") |>
  ggplot(aes(x = Values)) +
  geom_histogram(color = "white") +
  theme_bw() +
  facet_wrap(~ Confounders, scales = "free")
```
:::

::: {.column width="50%"}
```{r, fig.align='center', fig.asp=1, fig.width=6}
d |>
  filter(Year == 2016) |>
  ggplot(aes(x = cinc_pct, y = P3_Human_Rts)) +
  geom_point() +
  theme_bw() + 
  geom_smooth(method = "lm") +
  labs(x = "CINC (%)", y = "Human Rights / Rule of Law",
       caption = "Note: Focusing on 2016 data only")
```
:::
::::

::: notes

Today your job is to think about how to adjust your analyses for these confounders

- Let's focus just on 2016 for now (the most recent year available)

<br>

Use our tools from Monday and explore different ways of adjusting the plot on the right for the levels of the variables on the left

<br>

**Questions?**

- Let's get to work!

- Make sure to share what you find with the rest of the class!

:::


## For Next Class {background-image="libs/Images/background-slate_v2.png" .center}

<br>

::: {.r-fit-text}

1. *Linear Regression in R*, "Multiple OLS" section only

2. Wheelan ch 12 "Common Regression Mistakes"

:::

::: notes
Next class I want to introduce you to multiple regression (e.g. OLS regression with more than one predictor variable)
:::


## Old slides

## Check for Variation {background-image="libs/Images/background-slate_v2.png" .center}

```{r, fig.align='center', fig.asp=.75, fig.width=8}
d |>
  mutate(
    logGDP = log(gdp_2015usd)
  ) |>
  ggplot(aes(x = logGDP)) +
  geom_histogram(color = "white") +
  theme_bw() +
  labs(x = "GDP (2015 USD, Log Scale)", y = "") +
  scale_x_continuous(breaks = 19:30)

#scale_x_continuous(breaks = c(21, 24, 27, 30), labels = c("$1.3 billion", "$26.5 billion", "$532.0 billion", "$10.7 trillion"))
```

::: notes

This is a view of GDP in the natural log scale

- Each step up by one on the natural log scale is equivalent to multiplying by euler's number

- approximately 2.7 times

<br>

So, basically, for a country to move from 21 to 22 they have to triple the size of their economy

- This means each step up the scale represents a huge increase in your economy

<br>

**SLIDE**: Let me put this in dollar terms

:::



## Check for Variation {background-image="libs/Images/background-slate_v2.png" .center}

```{r, fig.align='center', fig.asp=.75, fig.width=8}
d |>
  mutate(
    logGDP = log(gdp_2015usd)
  ) |>
  ggplot(aes(x = logGDP)) +
  geom_histogram(color = "white") +
  theme_bw() +
  labs(x = "GDP (2015 USD, Log Scale)", y = "") +
  scale_x_continuous(breaks = c(21, 24, 27, 30), labels = c("$1.3 billion", "$26.5 billion", "$532.0 billion", "$10.7 trillion"))
```

::: notes

I show you this simply to let you know that there is considerable variation in GDP

- Countries differ from each other in GDP a lot, but the US messes up our ability to see it

- e.g. most countries have small GDPs relative to the biggest economies and this scale helps us see that

:::



## Causal Arguments {background-image="libs/Images/background-slate_v2.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("National Material\nCapabailities", "Confounders", "Human Rights and\nthe Rule of Law")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.6, xend = 2.3, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())

## Input project data
d <- read_excel("../Data_in_Class-SP24/Merged_Dataset-2006-2016_with_confounders.xlsx", na = "NA")
```

**Possible Confounders**: GDP, Immunizations DPT, Infant mortality & Compulsory education

::: notes

For today, I would like each of you to make a multivariate visualization for each of the confounders here.

- So, examine the primary relationship between cinc and human rights but adjusted for each of the confounders

- **FOCUS on most recent year only (e.g. 2016)**

:::



## Possible Confounder: Immunizations {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.align='center', fig.asp=.4, fig.width=13}
d |>
  filter(Year == 2016) |>
  mutate(
    # immune2 = case_when(
    #   immunization_dpt < quantile(d$immunization_dpt, probs = .25, na.rm = T) ~ "Immunization (Bottom 25%)",
    #   immunization_dpt < quantile(d$immunization_dpt, probs = .75, na.rm = T) ~ "Immunization (Middle 50%)",
    #   immunization_dpt >= quantile(d$immunization_dpt, probs = .75, na.rm = T) ~ "Immunization (Top 25%)"
    # )
    immune2 = case_when(
      immunization_dpt < quantile(d$immunization_dpt, probs = .75, na.rm = T) ~ "Immunization (Bottom 75%)",
      immunization_dpt >= quantile(d$immunization_dpt, probs = .75, na.rm = T) ~ "Immunization (Top 25%)"
    )
  ) |>
  select(immune2, cinc_pct, P3_Human_Rts) |>
  na.omit() |>
  ggplot(aes(x = cinc_pct, y = P3_Human_Rts)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ immune2, scales = "free_x") +
  scale_y_continuous(breaks = 0:10) +
  coord_cartesian(ylim = c(0,10)) +
  labs(x = "CINC (%)", y = "Human Rights / Rule of Law",
       caption = "Note that the x-axis scales differ across facets") +
  theme_bw()
```

<br>

**facet_wrap(~ gdp_quartiles, scales = "free_x")**

::: notes

As we noted before, not much variation in immunization!

- Maybe a tiny effect in the least healthy states. Maybe.

- Not a super promising confounder

<br>

Important caveat! The x-axis scales differ across the facets

- Code on the slide

- So, be careful interpreting the x-axes here

:::




## Possible Confounder: GDP {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.align='center', fig.asp=.4, fig.width=13}
d |>
  filter(Year == 2016) |>
  mutate(
    gdp2 = case_when(
      gdp_2015usd < quantile(d$gdp_2015usd, probs = .25, na.rm = T) ~ "GDP (Bottom 25%)",
      gdp_2015usd < quantile(d$gdp_2015usd, probs = .75, na.rm = T) ~ "GDP (Middle 50%)",
      gdp_2015usd >= quantile(d$gdp_2015usd, probs = .75, na.rm = T) ~ "GDP (Top 25%)"
    )
  ) |>
  select(gdp2, cinc_pct, P3_Human_Rts) |>
  na.omit() |>
  ggplot(aes(x = cinc_pct, y = P3_Human_Rts)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ gdp2, scales = "free_x") +
  scale_y_continuous(breaks = 0:10) +
  coord_cartesian(ylim = c(0,10)) +
  labs(x = "CINC (%)", y = "Human Rights / Rule of Law",
       caption = "Note that the x-axis scales differ across facets") +
  theme_bw()
```


::: notes

This looks WAY better!

- Those lines go through those data points!!

:::




## Possible Confounder: Infant Mortality {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.align='center', fig.asp=.4, fig.width=13}
d |>
  filter(Year == 2016) |>
  mutate(
    infant2 = case_when(
      infant_mortality < quantile(d$infant_mortality, probs = .25, na.rm = T) ~ "Infant Mortality (Bottom 25%)",
      infant_mortality < quantile(d$infant_mortality, probs = .75, na.rm = T) ~ "Infant Mortality (Middle 50%)",
      infant_mortality >= quantile(d$infant_mortality, probs = .75, na.rm = T) ~ "Infant Mortality (Top 25%)"
    )
  ) |>
  select(infant2, cinc_pct, P3_Human_Rts) |>
  na.omit() |>
  ggplot(aes(x = cinc_pct, y = P3_Human_Rts)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ infant2, scales = "free_x") +
  scale_y_continuous(breaks = 0:10) +
  coord_cartesian(ylim = c(0,10)) +
  labs(x = "CINC (%)", y = "Human Rights / Rule of Law",
       caption = "Note that the x-axis scales differ across facets") +
  theme_bw()
```

::: notes

Remember here that higher infant mortality is BAD!

- Not nearly as compelling as GDP, but maybe...

<br>

Keep in mind, we are only examining the variation in 3 groups

- That might not be fine-grained enough

:::



## Possible Confounder: Education {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, fig.align='center', fig.asp=.4, fig.width=13}
d |>
  filter(Year == 2016) |>
  mutate(
    ed2 = case_when(
      compulsory_education < quantile(d$compulsory_education, probs = .25, na.rm = T) ~ "Education (Bottom 25%)",
      compulsory_education < quantile(d$compulsory_education, probs = .75, na.rm = T) ~ "Education (Middle 50%)",
      compulsory_education >= quantile(d$compulsory_education, probs = .75, na.rm = T) ~ "Education (Top 25%)"
    )
  ) |>
  select(ed2, cinc_pct, P3_Human_Rts) |>
  na.omit() |>
  ggplot(aes(x = cinc_pct, y = P3_Human_Rts)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ ed2, scales = "free_x") +
  scale_y_continuous(breaks = 0:10) +
  coord_cartesian(ylim = c(0,10)) +
  labs(x = "CINC (%)", y = "Human Rights / Rule of Law",
       caption = "Note that the x-axis scales differ across facets") +
  theme_bw()
```

::: notes

Same as last one, I call this a maybe

:::









