---
format: revealjs
---

# Today's Agenda {background-image="Images/background-data_blue_v4.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)

d <- read_excel("../Data_in_Class/World_Development_Indicators/WDI-Multivariate_Viz-2024-04-04.xlsx", na = "NA")
```

<br>

::: {.r-fit-text}

**Apply multivariate analyses to our research project**

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2025)
:::

::: notes
Prep for Class

1. Check Canvas submissions

2. Bring a usb with the data file to class so we can live code together

<br>

Last class I introduced you to a series of tools we can use to begin performing multivariate analyses

1. Descriptive statistics by groups
  
2. Using facets to extend any visualization
    
3. Mapping 3rd variables with colors, shapes or sizes

<br>

**SLIDE**: Assignment for today
:::


## For Today {background-image="Images/background-data_blue_v4.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

**Possible Confounders**: Unemployment rate, Compulsory education

::: notes

Your job for today was to use these new tools to examine the other two possible confounders

- Does the central mechanism hold up when we adjust the analyses for either unemployment or years of compulsory education?

<br>

**How did this go?**

:::



## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=FALSE, fig.asp=.85, fig.align='center', fig.width=8}
# colorbrewer qualitative names: Accent, Dark2, Paired, Pastel1, Pastel2, Set1, Set2, Set3
d |>
  mutate(
    unemployment2 = case_when(
      unemployment_pct < quantile(d$unemployment_pct, probs = .25, na.rm = T) ~ "Unemployment (Q1)",
      unemployment_pct < quantile(d$unemployment_pct, probs = .50, na.rm = T) ~ "Unemployment (Q2)",
      unemployment_pct < quantile(d$unemployment_pct, probs = .75, na.rm = T) ~ "Unemployment (Q3)",
      unemployment_pct >= quantile(d$unemployment_pct, probs = .75, na.rm = T) ~ "Unemployment (Q4)"
      )
    ) |>
  select(fertility_rate_per_woman, life_expectancy_total, unemployment2) |>
  na.omit() |>
  ggplot(aes(x = fertility_rate_per_woman, y = life_expectancy_total, color = unemployment2, shape = unemployment2)) +
  geom_point() + 
  geom_smooth(method = "lm", se = T) +
  theme_bw() +
  facet_wrap(~ unemployment2, ncol = 2) +
  guides(color = "none", shape = "none") +
  labs(x = "Fertility Rate (births per woman)", y = "Life Expectancy (years)",
       caption = "Source: WDI (2024) for data from 2020") +
  scale_color_manual(values = RColorBrewer::brewer.pal(n = 4, name = "Dark2"))
```

::: notes

Note: I removed the NAs by filtering the dataset before making the plots

<br>

**Does adjusting for unemployment change the key relationship of interest?**

- **What is our takeaway from this?**

- IFF our identification strategy requires adjusting for unemployment then these correlations may be causal!

<br>

(**SLIDE**: Code to remove missing data on next slide)

:::



## Remove the Missing Data {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE, eval=FALSE}
# Create new variable
d$unemp2 <- cut(d$unemployment_pct, 
            breaks = c(0, 4.191, 6.453, 11.718, 28),
            labels = c("Unemployment (Q1)", "Unemployment (Q2)", 
                       "Unemployment (Q3)", "Unemployment (Q4)"))

# Make a new dataset with only the needed variables
d_subset <- select(.data = d, 
                   fertility_rate_per_woman, 
                   life_expectancy_total, 
                   unemp2)

# Make a new dataset without the missing data
d2 <- na.omit(d_subset)

# Make the visualization with d2
```

::: notes

Two new functions here

<br>

FIRST, select() lets you create a new data object with only some of the variables in the original dataset

- Here we are creating a copy of the dataset with only three variables in it

<br>

SECOND, na.omit() removes any observation from the dataset if it has missing data in ANY of the variables

- Be very careful with this!

- If you apply na.omit without narrowing the variables using select then you may lose a ton of data you didn't mean to remove

<br>

**Questions on the code?**

<br>

**SLIDE**: Education confounder

:::


## Adjusting for Education {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=FALSE, fig.asp=.85, fig.align='center', fig.width=8}
d |>
  mutate(
    education2 = case_when(
      compulsory_education_yrs < quantile(d$compulsory_education_yrs, probs = .25, na.rm = T) ~ "Education (Q1)",
      compulsory_education_yrs < quantile(d$compulsory_education_yrs, probs = .50, na.rm = T) ~ "Education (Q2)",
      compulsory_education_yrs < quantile(d$compulsory_education_yrs, probs = .75, na.rm = T) ~ "Education (Q3)",
      compulsory_education_yrs >= quantile(d$compulsory_education_yrs, probs = .75, na.rm = T) ~ "Education (Q4)"
      )
    ) |>
  select(fertility_rate_per_woman, life_expectancy_total, education2) |>
  na.omit() |>
  ggplot(aes(x = fertility_rate_per_woman, y = life_expectancy_total, color = education2, shape = education2)) +
  geom_point() + 
  geom_smooth(method = "lm", se = T) +
  theme_bw() +
  facet_wrap(~ education2, ncol = 2) +
  guides(color = "none", shape = "none") +
  labs(x = "Fertility Rate (births per woman)", y = "Life Expectancy (years)",
       caption = "Source: WDI (2024) for data from 2020") +
  scale_color_manual(values = RColorBrewer::brewer.pal(n = 4, name = "Dark2"))
```

::: notes

**Does adjusting for education change the key relationship of interest?**

:::



## Multivariate Analyses {background-image="Images/background-data_blue_v4.png" .center}

<br>

::: {.r-fit-text}
1. Descriptive statistics by groups
  
2. Using facets to extend any visualization
    
3. Mapping 3rd variables with colors, shapes or sizes
:::

::: notes

**Any questions on these techniques for adding confounders to our analyses?**

:::




## Causal Arguments {background-image="Images/background-data_blue_v4.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Press\nFreedom", "Confounders", "Environmental\nQuality")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.6, xend = 2.3, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())

## Input project data
#d <- read_excel("../Data_in_Class-SP24/Merged_Dataset-2006-2016_with_confounders.xlsx", na = "NA")
```

**Possible Confounders**: Size of economy, wealth, population, education, corruption

::: notes

Today we work to apply our new techniques to our ongoing class project!

<br>

Let's focus just on the 2024 data (the most recent year available)

- **Brainstorm time! What should we make?**

- *ON BOARD*

<br>

Ok, that's our plan, let's make some stuff!

- Let's get to work!

- Make sure to share what you find with the rest of the class!

:::




## For Next Class {background-image="Images/background-data_blue_v4.png" .center}

<br>

::: {.r-fit-text}

1. *Multiple Linear Regression in R*

2. Wheelan ch 12 "Common Regression Mistakes"

:::

::: notes
Next class I want to introduce you to multiple regression (e.g. OLS regression with more than one predictor variable)
:::

