---
format: revealjs
---

# Today's Agenda {background-image="libs/Images/background-data_blue_v3.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
```

<br>

<br>

**Multivariate Analyses**

- Using animations

<br>

::: r-stack
Justin Leinaweaver (Spring 2024)
:::

::: notes
Prep for Class


NEED TO ADD PRACTICE after each animation technique

- They tweak the code to answer a new question

:::



## Multivariate Analyses {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. Descriptive statistics by groups
  
2. Using facets to extend any visualization
    
3. Mapping 3rd variables with colors or shapes

4. **Bonus: Animate the 3rd/4th variable!**



## Install the needed packages and load gganimate {background-image="libs/Images/background-slate_v2.png" .center}

<br>

```{r, echo=TRUE, eval=FALSE}
install.packages("gifski")

install.packages("png")

install.packages("gganimate")

library(gganimate)
```

```{r}
## Load it for you
library(gganimate)
```



## Transitions Across Subsets {background-image="libs/Images/background-slate_v2.png" .center .smaller}

```{r, echo=TRUE, fig.align='center', fig.asp=0.7, fig.width=6, cache=TRUE}
ggplot(data = mpg, aes(x = displ, y = cty)) +
  geom_point(size = 1.5, color = "red") +
  transition_states(cyl) + 
  shadow_mark(past = T, future=F, alpha=0.25, color = "darkgrey") +
  labs(title = "Number of Cylinders: {closest_state}") 
```

::: notes
transition_states is for moving across levels of a categorical variable

- The shadow_mark function is cool in that it lets you keep the prior points visible for comparison
:::




## Transitions Across Time {background-image="libs/Images/background-slate_v2.png" .center .smaller}

```{r, echo=TRUE, fig.align='center', fig.asp=0.7, fig.width=7, cache=TRUE}
ggplot(data = mpg, aes(x = class, y = cty)) +
  geom_boxplot() +
  transition_time(year) + 
  labs(title = 'Year: {frame_time}') 
```



## Animating a Line Plot {background-image="libs/Images/background-slate_v2.png" .center .smaller}

```{r, echo=TRUE, fig.align='center', fig.asp=0.7, fig.width=7, cache=TRUE}
ggplot(data = economics, aes(x = date, y = pop)) +
  geom_line(color = "darkblue", linewidth = 1.4) +
  transition_reveal(date) + 
  labs(title = "US Population (thousands)")
```



## Last one... {background-image="libs/Images/background-slate_v2.png" .center}

```{r, echo=TRUE, eval=FALSE}
## New package needed for the data (install before loading)
install.packages("gapminder")

library(gapminder)
```

<br>

```{r}
library(gapminder)

gapminder |>
  slice_sample(n = 8) |>
  knitr::kable(format = "html") |>
  kableExtra::kable_styling(font_size = 22)
```



## {background-image="libs/Images/background-slate_v2.png" .center}

```{r, fig.align='center', fig.asp=0.7, fig.width=9, cache=TRUE, echo=TRUE}
gapminder |>
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(alpha = 0.7)
```

::: notes
Nice and easy

- Just plot countries by their wealth and life expectancy
:::



## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

```{r, fig.align='center', fig.asp=0.7, fig.width=9, cache=TRUE, echo=TRUE}
gapminder |>
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) + 
  geom_point(alpha = 0.7, show.legend = FALSE)
```

::: notes
Resize the points to approximate the size of the country's population and color by country

:::



## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

```{r, fig.align='center', fig.asp=0.7, fig.width=9, `code-line-numbers`="4", cache=TRUE, echo=TRUE}
gapminder |>
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_x_log10(labels = scales::label_comma())
```



## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

```{r, `code-line-numbers`="2,6", fig.align='center', fig.asp=0.5, fig.width=11, cache=TRUE, echo=TRUE}
gapminder |>
  filter(continent != "Oceania") |> 
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_x_log10(labels = scales::label_comma()) +
  facet_wrap(~ continent, ncol = 4) 
```



## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

```{r, `code-line-numbers`="7", fig.align='center', fig.asp=0.5, fig.width=11, cache=TRUE, echo=TRUE}
gapminder |>
  filter(continent != "Oceania") |>
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_x_log10(labels = scales::label_comma()) +
  facet_wrap(~ continent, ncol = 4) +
  scale_colour_manual(values = country_colors) 
```



## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

```{r, `code-line-numbers`="8", fig.align='center', fig.asp=0.5, fig.width=11, cache=TRUE, echo=TRUE}
gapminder |>
  filter(continent != "Oceania") |>
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_x_log10(labels = scales::label_comma()) +
  facet_wrap(~ continent, ncol = 4) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(3, 13)) 
```



## {background-image="libs/Images/background-slate_v2.png" .center .smaller}

```{r, `code-line-numbers`="9,10", fig.align='center', fig.asp=0.5, fig.width=11, cache=TRUE, echo=TRUE}
gapminder |>
  filter(continent != "Oceania") |>
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_x_log10(labels = scales::label_comma()) +
  facet_wrap(~ continent, ncol = 4) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(3, 13)) +
  transition_time(year) + 
  labs(title = 'Year: {frame_time}', x = "GDP per capita (log 10)", y = "Life Expectancy") 
```



## For Next Class {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. ...

::: notes

:::












