---
output:
  xaringan::moon_reader:
    css: ["default", "extra.css"]
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
##xaringan::inf_mr()
## For offline work: https://bookdown.org/yihui/rmarkdown/some-tips.html#working-offline
##chakra: libs/remark-latest.min.js

## Images not appearing? Put images folder inside the libs folder as that is the main data directory

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
##library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

```{r setup, include=FALSE}
## Include this to render DT tables correctly until fix is released
options(htmltools.dir.version = FALSE, htmltools.preserve.raw = FALSE)
```

class: slideblue

.size60[**Today's Agenda**]

<br>

.size40[
Apply multivariate analysis tools to our research project.
]

.center[.size30[
  Justin Leinaweaver (Fall 2021)
]]



---

class: middle, slidegreen

.size60[**Multivariate Analyses Tools**]

.size50[
+ Descriptive statistics by multiple groups

+ Facets

+ Adding third variables with colors, shapes or animations
]



---

class: middle, slidegreen

.size50[.center[**Aim is to take the visualization that represents our main causal effect and modify it for levels of the confounder.**]]

--

<br>

.size50[
What are the levels of a numeric confounder?

+ Visualize the confounder
+ Cut it into a categorical variable
]



---

class: middle, slidegreen

Main relationship

Confounder List

+ ?















