---
format: revealjs
---

# Today's Agenda {background-image="libs/Images/background-data_blue_v3.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
library(ggeffects)

d <- read_excel("../Data_in_Class-SP24/World_Development_Indicators/Practice_Simple_OLS/WDI-Practice_Simple_OLS-2024-04-04.xlsx", na = "NA")
```

<br>

::: {.r-fit-text}
**Multivariate Analyses**

- Adding confounders to OLS regressions
:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2024)
:::

::: notes
Prep for Class

1. Keep using the simple OLS WDI data

<br>

We'll continue using the WDI data from Monday.

- Everybody open up a new script file and import the WDI data so we can work with it again

<br>

**SLIDE**: Kick this off with the research problem

:::



## Building a Causal Argument {background-image="libs/Images/background-slate_v2.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

**Possible Confounders**: GDP per capita, unemployment rate, tobacco use, compulsory education

::: notes

Always want to keep our work rooted in the problem which is making a causal argument using observational data

<br>

This week we've been trying to adjust our analysis for confounders using multivariate data visualization and descriptive statistics tools

<br>

The aim for today is to use OLS regressions to bring a confounder into the adjustment process

<br>

**SLIDE**: Why move back to regression?

:::



## Why use OLS regressions? {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
- Assumes **"regression to the mean"**

- **Quantifies** the relationship between variables

- Uses **ALL** of the data

- Provides estimates of **uncertainty**

- Can be adjusted for **non-linear effects** and **confounders**
:::

::: notes

*Read slide*

<br>

Key to this is to remember that regression is a method for finding conditional means. 

- e.g. drawing a line through the middle of the points, and by middle we mean averages.

<br>

**SLIDE**: And what is a simple regression?

:::


## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Ordinary Least Squares (OLS) Regression**
:::

<br>

OLS regression is a technique for estimating the relationship between an outcome (Y) and a predictor variable (X) by finding the line that minimizes the squared residuals.

$$Y = \alpha + \beta X$$

```{r, echo = TRUE, eval = FALSE}
model <- lm(data, outcome ~ predictor)
```

::: notes

Conceptually, this is what you have been working with

- "Simple" in the OLS case means bivariate analysis: One predictor and one outcome.

<br>

The lm function in R gives us an estimate of the intercept term and the beta coefficient which together give us our "model" using the formula for a line.

<br>

**Everybody still with me?**

:::


## Our Baseline Estimate {background-image="libs/Images/background-slate_v2.png" .center}

```{r, echo=TRUE}
res1 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman)
```

<br>

:::: {.columns}
::: {.column width="50%"}
```{r}
modelsummary(models = list("Baseline" = res1), out = "gt", fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log",
             coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```
:::

::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.85, fig.width=6}
predict1 <- ggpredict(res1)

plot(predict1) +
  labs(x = "Fertility (births per woman)", y = "Predicted Average Life Expectancy (years)",
       title = "Increases in average fertility rates are associated with\n lower life expectancies around the world")
```
:::
::::

::: notes

Here are the results of regressing life expectancy on fertility rates around the world in 2020

- We estimated this during our work in week 11

- Everybody make sure you have loaded the data and can fit this model.

<br>

**How do we interpret this coefficient in real world terms?**

- An increase in the fertility rate by one additional birth per woman is associated, on average, with a decrease in average life expectancy of almost five years.  

<br>

**How well does this model fit the data when evaluated using our four step evaluation process?**

- Missing 8 countries out of 217 in the dataset

- Coefficient is significant

- R2 says fertility rates explain some 70% of the variation in life expectancy

- **SLIDE**: Residuals plot

:::


## Our Baseline Estimate {background-image="libs/Images/background-slate_v2.png" .center}

<br>

:::: {.columns}
::: {.column width="50%"}
```{r}
modelsummary(models = list("Baseline" = res1), out = "gt", fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log",
             coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```
:::

::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.85, fig.width=8}
plot(res1, which = 1)
```
:::
::::

::: notes

**Any concerns raised by the residuals plot?**

- (DEFINITE evidence of non-linearity!)

- Positive error at the extremes but negative in the middle

<br>

If we have time we'll fix this non-linearity before the end of class using a multiple regression.

:::


## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Ordinary Least Squares (OLS) Regression**
:::

<br>

Multiple OLS regression estimates the relationship between an outcome (Y) and a series of predictor variables (X$_i$) by finding the line that minimizes the squared residuals.

$$Y = \alpha + \beta_1 X_1 + \beta_2 X_2 + ... + \beta_k X_k$$

```{r, echo = TRUE, eval = FALSE}
model <- lm(data, outcome ~ predictor1 + predictor2 + ...)
```

::: notes

In mathematical terms, we can add as many predictor variables to the formula for a line as we want and still have it produce a SINGLE line

- One intercept term is now joined by multiple X variables (your predictors)

- Each X is partnered with a beta coefficient that represents the effect of that predictor on the outcome ceteris paribus ('holding all else equal')

<br>

In R terms, it is equally easy to add predictors to the lm function!

- Add the variable as a list after the tilde and separate predictors with plus signs

<br>

For our first multiple regression we are going to add the simplest possible predictor variable, a dummy variable

- **Anybody know what a "dummy" variable is?**

<br>

Think "dummy" like a store mannequin

- e.g. it stands in for something else (plastic person for showing how clothes might fit)

<br>

A dummy variable, also called dichotomous, can only have a value of zero or 1

- If the value is 1 then the characteristic is present, if zero then it is absent

<br>

**SLIDE**: Let's make this idea more concrete!

:::


## {background-image="libs/Images/background-slate_v2.png" .center}

**1. Find the median of wealth**
```{r, echo=TRUE}
summary(d$gdp_per_capita)
```

<br>

**2. Create a dummy variable**
```{r, echo=TRUE}
d$wealthier_median <- if_else(d$gdp_per_capita >= 6370.9,
                            true = 1, false = 0)
```

<br>

**3. Fit a multiple regression using the dummy confounder**
```{r, echo=TRUE}
res2 <- lm(data = d, life_expectancy_total ~ 
             fertility_rate_per_woman + wealthier_median)
```

::: notes 

Let's reduce the variation in GDP per capita to a dummy variable

- Specifically, we'll create a new variable that gets a value of '1' if a state is wealthier than the median and a '0' if not

<br>

We're using the if_else function to do this

- if_else requires a test in the data and then an argument for what to do if the test is true and what to do if the test is false

- Here I am giving if_else the test of greater than or equal to the median GDP per capita in the dataset

- If true then that country gets a '1' and if false a '0'

<br>

**Make sense?**

- Everybody run these lines of code and examine the results!

- (**SLIDE**)
:::


## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Multiple Regression: Add a Dummy Confounder**
:::

<br>

:::: {.columns}
::: {.column width="25%"}

```{r, echo=FALSE}
modelsummary(models = list(res1, res2), out = "gt", fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log", coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", "wealthier_median" = "Wealth > Median", "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "large")
  ), locations = gt::cells_body())
```

:::

::: {.column width="5%"}

:::

::: {.column width="70%"}

**Model 1: Simple**

- $$Y = \alpha + \beta_1 X_1$$

**Model 2: Multiple**

- $$Y = \alpha + \beta_1 X_1 + \beta_2 X_2$$

:::
::::

::: notes

I'm using modelsummary to output the two models side-by-side

- **Did everybody get the results for Model 2 here?**

<br>

Keep in mind, both of these models describe a SINGLE line

- The second simply has two slope coefficients instead of one

<br>

**Focus on Model 1 for a sec, what are the components from the table that go into this formula for a line?**

- (**SLIDE**)

:::



## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Multiple Regression: Add a Dummy Confounder**
:::

<br>

:::: {.columns}
::: {.column width="25%"}

```{r, echo=FALSE}
modelsummary(models = list(res1, res2), out = "gt", fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log", coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", "wealthier_median" = "Wealth > Median", "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "large")
  ), locations = gt::cells_body())
```

:::

::: {.column width="5%"}

:::

::: {.column width="70%"}

**Model 1: Simple**

- Life_Exp = 84.72 + -4.89 (Fertility)

<br>

**Model 2: Multiple**

- $$Y = \alpha + \beta_1 X_1 + \beta_2 X_2$$

:::
::::

::: notes

**And using this line, what does our regression predict is the average life expectancy for a country with a fertility rate of two births per woman on average?**

- (`r predict(res1, newdata = data.frame("fertility_rate_per_woman" = 2))`)

<br>

**Now focus on Model 2, what are the components from the table that go into this formula for a line?**

- (**SLIDE**)


:::



## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Multiple Regression: Add a Dummy Confounder**
:::

<br>

:::: {.columns}
::: {.column width="25%"}

```{r, echo=FALSE}
modelsummary(models = list(res1, res2), out = "gt", fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log", coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", "wealthier_median" = "Wealth > Median", "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "large")
  ), locations = gt::cells_body())
```

:::

::: {.column width="5%"}

:::

::: {.column width="70%"}

**Model 1: Simple**

- Life = 84.7 + -4.9 (Fertility)

<br>

**Model 2: Multiple**

- Life = 79.4 + -3.7 (Fertility) +  4.8 (Wealthy)

:::
::::

::: notes

**Is everybody clear on how we get from the lm function to the table to the formula for a line?**

<br>

Before we use Model 2 to make predictions, let's talk about the fit of the model

- **Does Model 2 fit the data "better" than model 1? Why or why not?**

<br>

- We lose 7 more states from the regression but 202/217 is still pretty darn good

- Two significant coefficients instead of one

- R2 increases; we now explain 76% of the variation in life expectancy

- RMSE error is reduced (e.g. avg residuals are smaller)

<br>

**SLIDE**: Let's check the residuals plot

:::



## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Multiple Regression: Add a Dummy Confounder**
:::

<br>

:::: {.columns}
::: {.column width="50%"}

**Simple OLS Regression**
```{r, fig.align='center', fig.asp=.85, fig.width=8}
plot(res1, which = 1)
```
:::

::: {.column width="50%"}
**Multiple OLS Regression 1**
```{r, fig.align='center', fig.asp=.85, fig.width=8}
plot(res2, which = 1)
```
:::
::::

::: notes

**What happens to the residuals when we move from the simple to the multiple regression in this case?**

<br>

Multiple regression model line simply fits the data better than our simple model line

- Almost completely removes the non-linearity!

- Much more homoscedastic errors

<br>

To be clear, this is not something you should expect simply by moving to a multiple regression

- It just so happens that adjusting our model for wealth has addressed this underlying problem!

<br>

**Make sense?**

:::




## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Multiple Regression: Add a Dummy Confounder**
:::

<br>

:::: {.columns}
::: {.column width="25%"}

```{r, echo=FALSE}
modelsummary(models = list(res1, res2), out = "gt", fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log", coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", "wealthier_median" = "Wealth > Median", "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "large")
  ), locations = gt::cells_body())
```

:::

::: {.column width="5%"}

:::

::: {.column width="70%"}

**Model 1: Simple**

- Life = 84.7 + -4.9 (Fertility)

<br>

**Model 2: Multiple**

- Life = 79.4 + -3.7 (Fertility) +  4.8 (Wealthy)

:::
::::

::: notes

Now let's talk about using the multiple regression model to make predictions

- Remember that we just used Model 1 to predict an average life expectancy of `r predict(res1, newdata = data.frame("fertility_rate_per_woman" = 2))` if a country has a fertility rate of two

<br>

**What does model 2 predict for the same case? e.g. Fertility = 2?**

<br>

Can't do it without making an assumption for rich states!

- That's the trick with a multiple regression!

- The coefficient for fertility can ONLY make predictions if you specify the other predictors in the model

<br>

**So, what does model 2 predict for fertility 2 if the country is rich? e.g. the value of wealthy here is a '1'**

- `r predict(res2, newdata = data.frame("fertility_rate_per_woman" = 2, "wealthier_median" = 1))`

<br>

**And for states below median wealth levels?**

- `r predict(res2, newdata = data.frame("fertility_rate_per_woman" = 2, "wealthier_median" = 0))`

<br>

So, remember, making predictions with a multiple regression requires specifying the values for all the predictors in the model

- **Does that make sense?**

<br>

**SLIDE**: Let's use ggpredict to handle this

:::



## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Multiple Regression: Add a Dummy Confounder**
:::

```{r, echo=TRUE, fig.align='center', fig.asp=.75, fig.width=7}
# Make the predictions
pred2 <- ggpredict(res2, 
                   terms = c("fertility_rate_per_woman", 
                             "wealthier_median"))

# Visualize the predictions
plot(pred2)
```

::: notes

Given that our wealth variable is a dummy, we have to use ggpredict a little differently

- We have to ask it to make predictions using both variables in the terms argument

<br>

**Everybody get the code to work?**

<br>

**Why are there two lines here and not just one?**

- **I keep telling you the regression makes a single line, don't I?**

<br>

(**SLIDE**)

:::


## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Multiple Regression: Add a Dummy Confounder**
:::

<br>

:::: {.columns}
::: {.column width="50%"}

```{r, echo=FALSE}
modelsummary(models = list(res2), out = "gt", fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log", coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", "wealthier_median" = "Wealth > Median", "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

:::

::: {.column width="50%"}

<br>


```{r, echo=F, fig.align='center', fig.asp=.75, fig.width=7}
# Visualize the predictions
plot(pred2)
```
:::
::::

::: notes

The key here is that plugging in different values for the predictors shifts the line up or down, BUT THE LINE REMAINS THE SAME FUNCTION

<br>

Our slope coefficient on fertility is -3.7 and that gets applied to EVERY change in fertility rate we plug in.

- So, move from fertility rate 1 to fertility rate 2 and life expectancy goes down by 3.7 years

- Move from fertility rate 2 to fertility rate 3 and that's ANOTHER 3.7 years lost

- That's 7.4 years for adding two more births per woman on average

<br>

That's NOT the story with the dummy variable

- The dummy can only ever have one of two values: zero or one

- A wealthier than the median state has almost 5 years more life expectancy and that effect can only happen one time

- So, what we see in the prediction plot is the dummy variable raising or lowering the line but not changing the slope

<br>

**Does that make sense?**

- **e.g. one formula for a line, two sets of predictions depending on the values you plug in**

:::


## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Multiple Regression: Add a Dummy Confounder**
:::

<br>

```{r, echo=FALSE}
modelsummary(models = list("Association" = res1, "Causal Argument" = res2), out = "gt", fmt = 2, stars = c("*" = .05), gof_omit = "IC|F|Log", coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", "wealthier_median" = "Wealth > Median", "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body()) |>
  gt::tab_style(style = gt::cell_fill(color = 'orange'), locations = gt::cells_body(columns = 2:3, rows = 1))
```


::: notes

Keep in mind, our primary interest is in estimating the coefficient on fertility while adjusting for the confounder

- In other words, we're trying to make a causal argument

- In those terms, we don't really care about the coefficient on the dummy variable

- We're trying to make a causal argument about the power of fertility rates to move life expectancies around the world

<br>

IFF our identification strategy is convincing then our second model is getting us closer to causality

- Interestingly for this case these results suggest that the association is stronger than the actual causal effect

- Fertility rates definitely appear to matter, but maybe a little less than the simple correlation suggests

<br>

**Make sense?**

:::



## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Multiple Regression: Add a Numeric Confounder**
:::

<br>

**Rescale GDP pc (10k)**
```{r, echo=TRUE}
d$gdp_pc_10k <- d$gdp_per_capita/10000
```

<br>

**Fit a multiple regression using the rescaled confounder**
```{r, echo=TRUE}
res3 <- lm(data = d, life_expectancy_total ~ 
             fertility_rate_per_woman + gdp_pc_10k)
```

::: notes

Our dummy version was just to get us into multiple regression more easily

- In reality, we do not want to throw away variation so cheaply

- We have tons of variation in wealth around the world so we need to see how that impacts the mechanism we are interested in

<br>

Here I have simply rescaled GDP per capita into 10s of thousands of dollars to make it easier to study

- Otherwise I have not changed the variable at all

<br>

Everybody run this code and examine the results!

- (**SLIDE**)

:::




## {background-image="libs/Images/background-slate_v2.png" .center}

<br>

:::: {.columns}
::: {.column width="50%"}
```{r, echo=FALSE}
modelsummary(models = list(res1, res2, res3), out = "gt", fmt = 2, stars = c("*" = .05), gof_omit = "IC|Log", coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", "wealthier_median" = "Wealth > Median", "gdp_pc_10k" = "GDP per capita", "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "large")
  ), locations = gt::cells_body())
```
:::

::: {.column width="50%"}

<br>

```{r, fig.align='center', fig.asp=.85, fig.width=8}
plot(res3, which = 1)
```
:::
::::

::: notes

**Alright, let's start just with the fit of Model 3**

- **What are your takeaways about the fit of the model?**

<br>

Improvements to the fit: 

- Increased R2, 
- significant coefficients, 
- smaller average residuals

<br>

**Any concerns in the residuals?**

- (My sense, obs 115 is definitely an outlier but the red line is really exaggerating the effect of this point.)

<br>

**Just for practice, how do we interpret the coefficient on GDP per capita?**

- (Each increase of $10k in GDP pc is associated, on average, with an increase in life expectancy of approximately one year)

<br>

**Alright, use ggpredict on our Model 3**

- In this case, just include fertility as the only "term" in the function

- (**SLIDE**)

:::



## {background-image="libs/Images/background-slate_v2.png" .center}

```{r, echo=TRUE, fig.align='center', fig.asp=.85, fig.width=8}
ggpredict(res3, terms = "fertility_rate_per_woman")
```

::: notes

What I want you to see here is that with numeric confounders ggpredict will set them at their average value in the dataset

- Here avg GDP pc is approximately $16k

<br>

**Questions on the mutliple regressions we've worked on today?**

- How to fit?

- How to evaluate?

- How to make predictions?

<br>

**SLIDE**: Assignment for next class

:::



## For Next Class {background-image="libs/Images/background-slate_v2.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

**Possible Confounders**: GDP per capita, (1) unemployment rate, (2) tobacco use, (3) compulsory education

::: notes

For next class everybody try fitting, evaluating and interpreting three more multiple regression models

- Regress life expectancy on fertility plus each of these confounders separately

<br>

**Questions on the assignment?**

:::










