---
format: revealjs
---

## Today's Agenda {background-image="Images/background-data_blue_v4.png" .center}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
library(modelr)
library(ggeffects)

# Load US States Data
d <- read_excel("../../DATA/2026-Spring/Pollock_Edwards_Textbook/Pollock_Edwards-US_States.xlsx", na = "NA")

d$voter_id <- if_else(d$voter.id.law != "No document required", 1, 0) |>
  as_factor()
```

<br>

::: {.r-fit-text}

**Ordinary Least Squares (OLS) Regression**

- Extending an OLS regression

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2026)
:::

::: notes
Prep for Class

1. Check Canvas submissions

<br>

**SLIDE**: Let's kick things off with your assignment for today

:::




## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**RQ: Do voter ID laws reduce voter turnout?**
:::

```{r}
# (Factors that predict swing states AND pressure for voter ID
# - Ethnicity: hispanic.percent, black.percent, white.percent
# - Inequality: gini.2016
# - Education: ba.or.more
# - Elite pressure: trump2016
# - Ideology: public.conservative
# - Wealth: hh.income
# - Electoral stakes: battleground2020

#res3 <- lm(data = d, vep20.turnout ~ voter_id + hh.income)
#res7 <- lm(data = d, vep20.turnout ~ voter_id + gini.2016)

res1 <- lm(data = d, vep20.turnout ~ voter_id)
res2 <- lm(data = d, vep20.turnout ~ voter_id + battleground2020)
res3 <- lm(data = d, vep20.turnout ~ voter_id + public.conservative)
res4 <- lm(data = d, vep20.turnout ~ voter_id + trump2016)
res5 <- lm(data = d, vep20.turnout ~ voter_id + ba.or.more)
res6 <- lm(data = d, vep20.turnout ~ voter_id + hispanic.percent + black.percent)
res7 <- lm(data = d, vep20.turnout ~ voter_id + battleground2020 + public.conservative + trump2016 + ba.or.more + hispanic.percent + black.percent)

plot(ggpredict(res7, terms = "voter_id"))

modelsummary(models = list(res1, res2, res3, res4, res5, res6, res7),
             fmt = 2,
             stars = c("*" = .05),
             out = "gt",
             gof_omit = "IC|F|Log") |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())



```

::: notes

Model 5 here is what you fit, but I wanted to show the results for the in-between models as well.

<br>

**Interpret the partial coefficients in Model 5 for me**

- **How does each predictor increase or decrease voter turnout?**

<br>

**Evaluate Model 5 for me**

1. Missing data?

2. Significance?

3. R2?

<br>

**Bottom line, what do we learn from these models?**

- **Do states with voter ID laws see lower turnout in elections or not?**

<br>

**SLIDE**: For next class

:::




