---
format: revealjs
---

# Today's Agenda {background-image="libs/Images/background-data_blue_v3.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
```

<br>

<br>

**Apply multivariate analyses to our research project**

<br>

<br>

::: r-stack
Justin Leinaweaver (Spring 2024)
:::

::: notes
Prep for Class

1. ???

<br>

Let's work today to add our confounders to our project analyses.
:::



## {background-image="libs/Images/background-slate_v2.png" .center}

What do we learn from our bivariate analysis of the first hypothesis?




## {background-image="libs/Images/background-slate_v2.png" .center}

Research project DAG with a confounder

```{r, fig.retina = 3, fig.align = 'center', out.width='85%', fig.asp=.5, fig.width=10}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Education", "FH Status", "GII")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.5, xend = 2.7, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.2, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.2, arrow = arrow())
```

- Scatter plot: GII x Education Spending

- `fh_status`: Add color, shape, OLS line and facets

::: notes

Let's examine our first hypothesis using our tools from Monday.

- Remake the scatter plot but adjust it for Freedom House classifications

- Does our result hold?
:::



## {background-image="libs/Images/background-slate_v2.png" .center}

Let's build tests of these confounded relationships

::: notes

Analyze these results for me.

### What do we learn from these visualizations about our first hypothesis?

### - Does our confidence from the bivariate plot hold that increasing education spending improves gender inequality?

- It appears that the effect differs in not free and partly free states from the free states.

- BUT, these CIs are HUGE
    - Implies the general trend may hold but we really don't have a good sense of the slope in each facet.
    
<br>

**SLIDE**: Let's check the other confounder
:::


## For Next Class {background-image="libs/Images/background-slate_v2.png" .center}

<br>

1. *Linear Regression in R*, "Multiple OLS" section only

2. Wheelan ch 12 "Common Regression Mistakes"

::: notes
Next class I want to introduce you to multiple regression (e.g. OLS regression with more than one predictor variable)
:::



