---
format: revealjs
---

# Today's Agenda {background-image="libs/Images/background-data_blue_v3.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)

d <- read_excel("../Data_in_Class-SP24/World_Development_Indicators/Practice_Simple_OLS/WDI-Practice_Simple_OLS-2024-04-04.xlsx", na = "NA")
```

<br>

::: {.r-fit-text}
**Multivariate Analyses**

- Adding confounders to OLS regressions
:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2024)
:::

::: notes
Prep for Class

1. Assignment for today used the WDI data on life expectancy

2. Make sure merged project data is available

:::


## For Today {background-image="libs/Images/background-slate_v2.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

**Possible Confounders**: GDP per capita, (1) unemployment rate, (2) tobacco use, (3) compulsory education

::: notes

For today I asked you to practice fitting, evaluating and interpreting three multiple regression models

- **How did this go?**

:::


## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Do higher fertility rates lower average life expectancies around the world?**
:::

```{r}
d$gdp_pc_10k <- d$gdp_per_capita/10000

res1 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman)
res2 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman + gdp_pc_10k)
res3 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman + unemployment_pct)
res4 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman + tobacco_use_pct)
res5 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman + compulsory_education_yrs)

modelsummary(models = list(res1, res2, res3, res4, res5), 
             out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", 
                          "gdp_pc_10k" = "GDP pc ($10k)", 
                          "unemployment_pct" = "Unemployment (%)",
                          "tobacco_use_pct" = "Tobacco Use (%)",
                          "compulsory_education_yrs" = "Education (years)",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "15px")
  ), locations = gt::cells_body())

```

::: notes

Quick note, I don't know why modelsummary is omitting some of the F-statistics...

<br>

**Did everybody get these results?**

<br>

**Interpret each coefficient for me**

- **At what level increase does each predictor make a serious impact on life expectancy?**

<br>

**Any missing data concerns?**

<br>

**Any meaningful changes in the R2?**

<br>

**SLIDE**: Let's check the residuals

:::


## {background-image="libs/Images/background-slate_v2.png"}

:::: {.columns}
::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.625, fig.width=8}
plot(res2, which = 1)

plot(res3, which = 1)
```
:::

::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.625, fig.width=8}
plot(res4, which = 1)

plot(res5, which = 1)
```
:::
::::

::: notes

**Takeaways from checking the residuals?**

<br>

Looks to me like only adjusting for wealth deals with the underlying non-linear effect

- I suspect wealth is a big confounder in our causal mechanism and omitting it messes up the estimated relationship between fertility and life expectancy

:::


## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Do higher fertility rates lower average life expectancies around the world?**
:::

:::: {.columns}
::: {.column width="50%"}

<br>

```{r, fig.align = 'center', fig.asp=1, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.4, y = 1, yend = 1, arrow = arrow(), linewidth = 1.4) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.15, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.15, arrow = arrow())
```
:::

::: {.column width="50%"}
```{r}
modelsummary(models = list(res1, res2, res3, res4, res5), 
             out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", 
                          "gdp_pc_10k" = "GDP pc ($10k)", 
                          "unemployment_pct" = "Unemployment (%)",
                          "tobacco_use_pct" = "Tobacco Use (%)",
                          "compulsory_education_yrs" = "Education (years)",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "15px")
  ), locations = gt::cells_body()) |>
  gt::tab_style(style = gt::cell_fill(color = 'orange'), locations = gt::cells_body(columns = 1:6, rows = 1))

```
:::
::::

::: notes

Remember, the goal of this exercise is to use an identification strategy to make a causal argument about fertility and life expectancy

- **Bottom line, what do we learn from these regressions?**

<br>

- Robustness Test: It is reassuring that the coefficient on fertility rate is fairly stable across models

- The direction of effect makes sense on wealth and unemployment (the other two aren't significant or large enough to interpret)

<br>

**Did anybody compare their regression results to the multivariate visualizations we made in class on Monday?**

- (**SLIDE**: Compare to our multivariate visualizations last week)

:::



## {background-image="libs/Images/background-slate_v2.png"}

::: {.r-fit-text}
**Multivariate Analyses: Education as a Confounder**
:::

:::: {.columns}
::: {.column width="70%"}
```{r, echo=FALSE, fig.asp=.85, fig.align='center', fig.width=8, cache=TRUE}
d |>
  mutate(
    education2 = case_when(
      compulsory_education_yrs < quantile(d$compulsory_education_yrs, probs = .25, na.rm = T) ~ "Education (Q1)",
      compulsory_education_yrs < quantile(d$compulsory_education_yrs, probs = .50, na.rm = T) ~ "Education (Q2)",
      compulsory_education_yrs < quantile(d$compulsory_education_yrs, probs = .75, na.rm = T) ~ "Education (Q3)",
      compulsory_education_yrs >= quantile(d$compulsory_education_yrs, probs = .75, na.rm = T) ~ "Education (Q4)"
      )
    ) |>
  select(fertility_rate_per_woman, life_expectancy_total, education2) |>
  na.omit() |>
  ggplot(aes(x = fertility_rate_per_woman, y = life_expectancy_total, color = education2, shape = education2)) +
  geom_point() + 
  geom_smooth(method = "lm", se = T) +
  theme_bw() +
  facet_wrap(~ education2, ncol = 2) +
  guides(color = "none", shape = "none") +
  labs(x = "Fertility Rate (births per woman)", y = "Life Expectancy (years)",
       caption = "Source: WDI (2024) for data from 2020")
```
:::

::: {.column width="30%"}

<br>

```{r}
modelsummary(models = list(res5), 
             out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", 
                          "compulsory_education_yrs" = "Education (years)",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "16px")
  ), locations = gt::cells_body())
```
:::
::::

::: notes

Ok, let's make sure everybody understands the intuitions of our work on multivariate analyses over the last few weeks

<br>

Help me understand what I'm seeing here.

- On the left, I see scatter plots where the negative effect of fertility rate on life expectancy seems consistent across the levels of education

- On the right, the results of a multiple regression showing education is not a significant predictor of life expectancy

- **Are these two pieces of data analysis consistent with each other or not?**

<br>

This is KEY to understanding our work over the last few weeks

- Never forget our focus is to use data analyses **to make a causal argument that changes in fertility rates CHANGE average life expectancy** around the world

- NONE of these tables or pictures tell us about the causal effect of education on life expectancy

- We are interested in the effect of fertility rates WHEN CONTROLLING FOR EDUCATION

- In other words, for this identification strategy we assume that countries with high vs low fertility rates differ in their levels of education

- e.g. we cannot assume they are identical on average

<br>

**If you were interested in estimating the effect of education on life expectancy, what would you have to do?**

- (**SLIDE**: Generate a brand new ID strategy!)

:::



## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**A new research question...**
:::

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Compulsory\nEducation", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

::: {.r-fit-text}
**Requires a new ID strategy!**
:::

::: notes

In this story, I do not think fertility rates are a confounder

<br>

Probably a mediator, no?

- Education changes fertility rates which changes life expectancy

- That would mean a regression with both fertility and education will produce an underestimate of the importance of education!

<br>

**SLIDE**: Let's check!

:::



## {background-image="libs/Images/background-slate_v2.png" .center}

**Do countries that require more years of compulsory education have higher life expectancy?**

<br>

:::: {.columns}
::: {.column width="50%"}

```{r, fig.align = 'center', fig.asp=1, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Compulsory\nEducation", "Fertility\nRate", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.6, xend = 2.4, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1, xend = 1.7, y = 1.15, yend = 1.85, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.15, arrow = arrow())
```
:::

::: {.column width="50%"}

```{r}
res6 <- lm(data = d, life_expectancy_total ~ compulsory_education_yrs)

modelsummary(models = list(res6, res1, res5), 
             out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("compulsory_education_yrs" = "Education (years)",
                          "fertility_rate_per_woman" = "Fertility (births per woman)", 
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "15px")
  ), locations = gt::cells_body())
```
:::
::::

::: notes

Regression results

- Model 1 shows that compulsory education increases life expectancy considerably

- Model 2 shows that fertility rate increases lower life expectancy dramatically

- And Model 3 shows when you include both it looks like fertility rate eats up education as you would expect if it is a mediator

<br>

**Does this make sense?**

:::



## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Do higher fertility rates lower average life expectancies around the world?**
:::

:::: {.columns}
::: {.column width="50%"}

<br>

```{r, fig.align = 'center', fig.asp=1, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.4, y = 1, yend = 1, arrow = arrow(), linewidth = 1.4) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.15, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.15, arrow = arrow())
```
:::

::: {.column width="50%"}
```{r}
modelsummary(models = list(res1, res2, res3, res4, res5), 
             out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", 
                          "gdp_pc_10k" = "GDP pc ($10k)", 
                          "unemployment_pct" = "Unemployment (%)",
                          "tobacco_use_pct" = "Tobacco Use (%)",
                          "compulsory_education_yrs" = "Education (years)",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "15px")
  ), locations = gt::cells_body()) |>
  gt::tab_style(style = gt::cell_fill(color = 'orange'), locations = gt::cells_body(columns = 1:6, rows = 1))

```
:::
::::

::: notes

All of this is just meant to reinforce the key intuition here

- We selected a mechanism of interest,

- We developed an identification strategy for it alone, and now

- We are adjusting our analyses for the confounders

<br>

In either case, this is all a reminder not to become infatuated with your regression tables

- The confounders are there to help us estimate the main relationship, that's it.

- We don't really care about the significance of wealth, unemployment, education, etc

<br>

**Any questions on all of this?**

:::



## {background-image="libs/Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Using Multiple Regressions on our Project**
:::

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=FALSE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("National Material\nCapabailities", "Confounders", "Human Rights and\nthe Rule of Law")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.6, xend = 2.3, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())

## Input project data
d <- read_excel("../Data_in_Class-SP24/Merged_Dataset-2006-2016_with_confounders.xlsx", na = "NA")
```

**Possible Confounders**: 1) Baseline, 2) GDP, 3) Immunizations DPT, 4) Infant mortality & 5) Compulsory education

::: notes

Today we work to apply our new techniques to our ongoing class project!

<br>

I'm asking everyone to work together to fit these five models

- The baseline is the simple regression of human rights on cinc_pct

- Each of the additional models adds a single confounder to the baseline

- Let's use all of the data! Not just 2016.

<br>

**Any questions?**

- We'll check in and see how we did at 11:35a!

- Go!

<br>

**Notes**

- GDP (constant 2015 US$)
    - (Source: World Bank national accounts data, and OECD National Accounts data files.)
    
- Immunization, DPT (% of children ages 12-23 months)
    - (Source: WHO and UNICEF (http://www.who.int/immunization/monitoring_surveillance/en/).)

- Mortality rate, infant (per 1,000 live births)
    - (Source: Estimates developed by the UN Inter-agency Group for Child Mortality Estimation (UNICEF, WHO, World Bank, UN DESA Population Division) at www.childmortality.org.)

- Compulsory education, duration (years)
    - (Source: UNESCO Institute for Statistics (UIS). UIS.Stat Bulk Data Download Service. Accessed September 19, 2023. https://apiportal.uis.unesco.org/bdds.)

:::



## {background-image="libs/Images/background-slate_v2.png" .center}

```{r}
res1 <- lm(data = d, P3_Human_Rts ~ cinc_pct)
res2 <- lm(data = d, P3_Human_Rts ~ cinc_pct + gdp_2015usd_billions)
res3 <- lm(data = d, P3_Human_Rts ~ cinc_pct + immunization_dpt)
res4 <- lm(data = d, P3_Human_Rts ~ cinc_pct + infant_mortality)
res5 <- lm(data = d, P3_Human_Rts ~ cinc_pct + compulsory_education)

# d2016 <- d |> filter(Year == 2016)
# 
# res1 <- lm(data = d2016, P3_Human_Rts ~ cinc_pct)
# res2 <- lm(data = d2016, P3_Human_Rts ~ cinc_pct + gdp_2015usd_billions)
# res3 <- lm(data = d2016, P3_Human_Rts ~ cinc_pct + immunization_dpt)
# res4 <- lm(data = d2016, P3_Human_Rts ~ cinc_pct + infant_mortality)
# res5 <- lm(data = d2016, P3_Human_Rts ~ cinc_pct + compulsory_education)

modelsummary(models = list(res1, res2, res3, res4, res5), out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("cinc_pct" = "CINC (%)",
                          "gdp_2015usd_billions" = "GDP (billions)", 
                          "immunization_dpt" = "Immunization DPT", 
                          "infant_mortality" = "Infant Mortality", 
                          "compulsory_education" = "Compulsory Education", 
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "18px")
  ), locations = gt::cells_body())
```

::: notes

**Did everybody get these results?**

<br>

**SLIDE**: Quick adjustment before we discuss the results

:::




## {background-image="libs/Images/background-slate_v2.png" .center}

```{r}
d$gdp_2015usd_trillions <- d$gdp_2015usd/1e12

res2a <- lm(data = d, P3_Human_Rts ~ cinc_pct + gdp_2015usd_trillions)

modelsummary(models = list(res1, res2a, res3, res4, res5), out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("cinc_pct" = "CINC (%)",
                          "gdp_2015usd_trillions" = "GDP (trillions)", 
                          "immunization_dpt" = "Immunization DPT", 
                          "infant_mortality" = "Infant Mortality", 
                          "compulsory_education" = "Compulsory Education", 
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "18px")
  ), locations = gt::cells_body()) |>
  gt::tab_style(style = gt::cell_fill(color = 'orange'), locations = gt::cells_body(columns = 1:6, rows = 1))
```

::: notes

Never report a coefficient of zero as we see here with GDP.

- It's not actually zero it's just such a small number it's hard to see

- I converted GDP into trillions to ensure I could report the coefficient on this table

- You should always do the same

<br>

**Ok, what do we learn from these regressions?**

- **What happened to our estimate of the effect of CINC on human rights?**

<br>

1. Missing data problems?

2. Are the coefficients significant?

3. What does the R2 indicate?

4. Any problems in the residuals plots? (**SLIDE**)
:::


## {background-image="libs/Images/background-slate_v2.png"}

:::: {.columns}
::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.625, fig.width=8}
plot(res2a, which = 1)

plot(res3, which = 1)
```
:::

::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.625, fig.width=8}
plot(res4, which = 1)

plot(res5, which = 1)
```
:::
::::

::: notes

Two steps forward, a handful of steps back!

- The GDP residuals are a mess

- Infant mortality is DEFINITELY nonlinear

- Immunization shows a funneling effect...

- Education looks ok!

:::


## {background-image="libs/Images/background-slate_v2.png" .center}

```{r}
modelsummary(models = list(res1, res2a, res3, res4, res5), out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("cinc_pct" = "CINC (%)",
                          "gdp_2015usd_trillions" = "GDP (trillions)", 
                          "immunization_dpt" = "Immunization DPT", 
                          "infant_mortality" = "Infant Mortality", 
                          "compulsory_education" = "Compulsory Education", 
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "18px")
  ), locations = gt::cells_body()) |>
  gt::tab_style(style = gt::cell_fill(color = 'orange'), locations = gt::cells_body(columns = 1:6, rows = 1))
```

::: notes

Heading in the right direction!

<br>

**Questions on our work today?**

<br>

Friday we'll keep working on mutliple regressions

:::