---
format: revealjs
---

## Today's Agenda {background-image="Images/background-data_blue_v4.png" .center}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
library(modelr)
library(ggeffects)

# Input practice WDI data
wdi <- read_excel("../../2026-Spring/DATA/Lectures/World_Development_Indicators-2023-Life_Expectancy.xlsx", na = "NA")

# # Input practice WDI data
# wdi <- read_excel("../../Course_History/2025-Spring/Data_in_Class/Class11-1-Simple_OLS-WDI_LifeExpectancy/WDI-Practice_Simple_OLS-2024-04-04.xlsx", na = "NA")
```

<br>

::: {.r-fit-text}

**Ordinary Least Squares (OLS) Regression**

- Improving the fit of your models

- Notes: Regression_Analysis.R

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2026)
:::

::: notes
Prep for Class

1. Check Canvas submissions

2. TRY TO SAVE TIME AT END OF CLASS so they can work on the assignment

3. ARGH, YOU FORGOT THE ASSIGNMENT ON CANVAS SO JUST HAD THEM DO IT IN CLASS

<br>

**SLIDE**: Let's refresh our memories on evaluating the fit of an OLS regression model

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Evaluating the Fit of an OLS Regression**
:::

<br>

1. Is there a problem with missing data?

2. Are the coefficients significant?

3. How much of the outcome is explained by the model?

4. Is there a problem in the residuals?

::: notes

**What is the concern with missing data in your regression?**

- (Missing at random vs missing not at random, aka selection bias)

<br>

**What do significant coefficients tell me about your regression model?**

- (Can we reject the null hypothesis?)

<br>

**How do I evaluate if my $R^2$ value is "good" or not?**

- (It depends on what you are trying to do)

- (If the goal is predicting outcomes, e.g. who wins an election, higher $R^2$ is better)

- (If the goal is developing a precise estimate of a causal effect, $R^2$ kind of doesn't matter)

<br>

**And finally, what am I checking in the residuals?**

- **What should they look like?**

- (**SLIDE**)

:::




## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Assumptions about OLS Residuals**

1. Zero mean

2. Constant variance (homoscedasticity)

3. No autocorrelation

4. Normally distributed

:::

::: notes

IFF your model fits the data well, then your residuals should have these four qualities

<br>

**What does it mean to say your residuals should have zero mean?**

- (Model line should go through the middle of the data and the positive and negative residuals should offset)

<br>

**What does it mean to say our errors should be homoscedastic?**

- (The spread of the error should be the same across all levels of the model)

<br>

**What does it mean to say that there should be no autocorrelation in our errors?**

- (No patterns in the residuals; no correlation between them)

<br>

**And finally, why should our residuals be normally distributed?**

- (Basically, if you check the other three boxes your residuals are random noise and adding together random noise should look like a normal curve)

<br>

**Any questions on our four steps for evaluating a regression?**

<br>

What we didn't get to last class was what do we do if our residuals indicate a serious problem in the fit of our model

- Today we learn two approaches dependent on the problem revealed by the residuals

<br>

Approaches:

1. If the effect you are studying is nonlinear you can add squared and cubed terms to the regression (or more!)

2. If your outcome variable has a strong right skew you can transform it using the log

<br>

**SLIDE**: Let's use your first Canvas assignment to introduce these approaches

:::



## {background-image="Images/background-data_blue_v4.png" .smaller}

**Do countries with more highly educated populations tend to have longer average life expectancies?**
  
::::: {.columns}
:::: {.column width="40%"}

<br>
  
```{r, echo=FALSE}
model1 <- lm(data = wdi, life_expectancy ~ secondary_education_pct)

modelsummary(models = model1, 
             output = "gt",
             fmt = 2,
             stars = c('*' = .05),
             gof_omit = "IC|F|Log",
             coef_map = c("secondary_education_pct" = "Secondary Education (%)", "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

::::
  
:::: {.column width="60%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(ggpredict(model1, terms = "secondary_education_pct")) +
  labs(x = "Secondary Education (%)",
       y = "Average Life Expectancy (years)")
```

### Residuals 1

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model1, which = 2)
```

### Residuals 2

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model1, which = 1)
```

:::
  
::::
:::::

::: notes

**What is your answer to our first homework question?**

<br>

- **In substantive terms, how big is the effect?**

- **How useful is this as policy guidance?**

<br>

**But does the model fit the data well? Evaluate the statistical significance for me**

- (Residuals 1: No serious concerns in the QQ plot)

- (Residuals 2: Clear heteroskedasticity AND autocorrelation)

    - (A definite non-linear pattern to the residuals)

<br>

I cannot emphasize enough the dangers of assuming a regression is a "good" summary of a relationship UNLESS you evaluate the fit of the line

- Here the significance star is especially dangerous

- Too many people see the star and think they've succeeded in finding truth!

- Regression is a tool and ALL tools are imperfect

<br>

**Does anybody remember how to curve the formula for a line in algebra?**
  
- (**SLIDE**)

:::



## {background-image="Images/background-data_blue_v4.png" .smaller}
  
::: {.r-fit-text}
**Adapting OLS regressions to fit non-linear relationships**
:::
  
::: {.panel-tabset}

### Exponential Growth Data

```{r, fig.retina=3, fig.asp=.618, out.width="80%", fig.align='center'}
d3 <- tibble(
  x1 = 1:50,
  x2 = (1:50)^2,
  y1 = x1 + x2 + rnorm(n = 50, mean = 10, sd = 2)
)

res10 <- lm(data = d3, y1 ~ x1)
res11 <- lm(data = d3, y1 ~ x1 + I(x1^2))

d4 <- d3 |>
  add_predictions(res10, var = "lm") |>
  add_predictions(res11, var = "squared")

d3 |>
  ggplot(aes(x = x1, y = y1)) +
  geom_point() +
  theme_bw() +
  labs(x = "Predictor", y = "Outcome")
```


### Version 1

```{r, fig.retina=3, fig.asp=.618, out.width="80%", fig.align='center'}
d3 |>
  ggplot(aes(x = x1, y = y1)) +
  geom_point() +
  geom_line(data = d4, aes(x = x1, y = lm), color = "red") +
  theme_bw() +
  labs(x = "Predictor", y = "Outcome") +
  annotate("text", x = 15, y = 2000, size = 8, color = "red",
           label = expression("Y = " ~ alpha + beta[1] ~ "x"))
```

### Version 2

```{r, fig.retina=3, fig.asp=.618, out.width="80%", fig.align='center'}
d3 |>
  ggplot(aes(x = x1, y = y1)) +
  geom_point() +
  geom_line(data = d4, aes(x = x1, y = lm), color = "red") +
  geom_line(data = d4, aes(x = x1, y = squared), color = "blue") +
  theme_bw() +
  labs(x = "Predictor", y = "Outcome") +
  annotate("text", x = 2, y = 2000, size = 8, color = "red", hjust = 0,
           label = expression("Y = " ~ alpha + beta[1] ~ "x")) +
  annotate("text", x = 2, y = 1500, size = 8, color = "blue", hjust = 0, label = expression("Y = " ~ alpha + beta[1] ~ "x" + beta[2] ~ "x"^2))
```

:::
  
::: notes

Example 1: Exponential growth

- Here I have plotted some toy data that follows an exponential growth curve

- e.g. increases begin slowly but speed up as we move across the levels of the predictor

<br>
  
**Version 1**: We can definitely fit a simple regression to this data

- Heck, it even comes pretty close to the trend!
  
- However, you can see here the pattern in the residuals without even needing to check the residuals plot

<br>
  
**Version 2**: In this form we allow X to have TWO effects on the line

- X impacts Y both in its regular form (e.g. 1, 2, 3, 4) AND also as a square (e.g. 1, 4, 9, 16)

- Early on the squared term has a small impact on the line and that impact gets bigger and bigger

<br>
  
Note that this formula for a line also allows the effect of each version of X to differ from each other!
  
- Each X has a SEPARATE beta term in the regression

- This makes this form VERY adaptable

<br>
  
For example, you could have a strong, positive slope on regular X but a tiny slope on X^2

- That would be a mostly straight line that skews upward very slowly

<br>
  
Alternatively, you could have a small slope coefficient on X and a huge slope on the squared term

- This is a trend that launches upward like a rocket

<br>
  
**Do these intuitions make sense?**
  
<br>
  
**SLIDE**: We can also do this in the other direction

:::
  
  
  
## {background-image="Images/background-data_blue_v4.png" .smaller}
  
::: {.r-fit-text}
**Adapting OLS regressions to fit non-linear relationships**
:::
  
::: {.panel-tabset}

### Exponential Decline Data

```{r, fig.retina=3, fig.asp=.618, out.width="80%", fig.align='center'}
d3 <- tibble(
  x1 = 1:50,
  x2 = (1:50)^2,
  y1 = x1 - x2 + rnorm(n = 50, mean = 10, sd = 2)
)

res10 <- lm(data = d3, y1 ~ x1)
res11 <- lm(data = d3, y1 ~ x1 + I(x1^2))

d4 <- d3 |>
  add_predictions(res10, var = "lm") |>
  add_predictions(res11, var = "squared")

d3 |>
  ggplot(aes(x = x1, y = y1)) +
  geom_point() +
  theme_bw() +
  labs(x = "Predictor", y = "Outcome")
```


### Version 1

```{r, fig.retina=3, fig.asp=.618, out.width="80%", fig.align='center'}
d3 |>
  ggplot(aes(x = x1, y = y1)) +
  geom_point() +
  geom_line(data = d4, aes(x = x1, y = lm), color = "red") +
  theme_bw() +
  labs(x = "Predictor", y = "Outcome") +
  annotate("text", x = 2, y = -1500, size = 8, color = "red", hjust = 0, label = expression("Y = " ~ alpha + beta[1] ~ "x"))
```

### Version 2

```{r, fig.retina=3, fig.asp=.618, out.width="80%", fig.align='center'}
d3 |>
  ggplot(aes(x = x1, y = y1)) +
  geom_point() +
  geom_line(data = d4, aes(x = x1, y = lm), color = "red") +
  geom_line(data = d4, aes(x = x1, y = squared), color = "blue") +
  theme_bw() +
  labs(x = "Predictor", y = "Outcome") +
  annotate("text", x = 2, y = -1500, size = 8, color = "red", hjust = 0,
           label = expression("Y = " ~ alpha + beta[1] ~ "x")) +
  annotate("text", x = 2, y = -2000, size = 8, color = "blue", hjust = 0, label = expression("Y = " ~ alpha + beta[1] ~ "x" - beta[2] ~ "x"^2))
```

:::
  
::: notes

Example 2: Exponential decline

- Decreases begin slowly but speed up as we move across the levels of the predictor

- Here we see fake data in an exponential decline curve

<br>
  
**Version 1**: If we fit a linear model we see a vague approximation but with a definite pattern to the error

<br>
  
**Version 2**: In this form we allow X to have TWO effects on the line

- X impacts Y both in its regular form (e.g. 1, 2, 3, 4) AND THEN by subtracting its squared term (e.g. 1, 4, 9, 16)

- Early on the squared term has a small impact on the line and that impact gets bigger and bigger

<br>
  
**Make sense?**
  
<br>
  
**SLIDE**: Let's apply this to the secondary education regression

:::



## {background-image="Images/background-data_blue_v4.png" .center}
  
::: {.r-fit-text}
**Adapting OLS regressions to fit non-linear relationships**
:::
  
<br>
  
The I() Function$^*$

- Let's you modify a variable inside the lm() formula

- Combine with "^2" to square the predictor variable

<br>

::: {.r-fit-text}

```{r, echo=TRUE, eval=FALSE}
lm(data = wdi, 
   life_expectancy ~ secondary_education_pct + I(secondary_education_pct^2))
```

*That's a capital "i" in the function name

:::

::: notes

Here's the code you need to add a squared term to the education predictor variable inside the lm() function

- **Does everybody understand the code on the slide?**
  
<br>
  
Ok, fit this new model, make the predictions and check the residuals!
  
- (**SLIDE**)

:::




## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Adapting OLS regressions to fit non-linear relationships**
:::
  
::::: {.columns}
:::: {.column width="50%"}
```{r, fig.asp=.85, fig.align='center', fig.width=6}
model1 <- lm(data = wdi, life_expectancy ~ secondary_education_pct)
model1b <- lm(data = wdi, life_expectancy ~ secondary_education_pct + I(secondary_education_pct^2))

plot(model1, which = 1)
```
::::

:::: {.column width="50%"}
```{r, fig.asp=.85, fig.align='center', fig.width=6}
plot(model1b, which = 1)
```
::::
:::::

::: notes

On the left is our original model and on the right is our new model with the squared term

<br>

**Has this addressed our problems of non-constant variance and autocorrelation?**

- (Yes to the autocorrelation)

- (Some of the non-constant variance has been reduced)

<br>

Good, we have a better model fit!

- **SLIDE**: Now let's interpret the substantive results

:::




## {background-image="Images/background-data_blue_v4.png" .smaller}
  
::::: {.columns}
:::: {.column width="40%"}

<br>
  
```{r, echo=FALSE}
modelsummary(list(model1, model1b),
             out = "gt",
             fmt = 2, 
             stars = c("*" = .05), 
             gof_omit = "IC|F|Log",
             coef_rename = c("secondary_education_pct" = "Secondary Education (%)", "I(secondary_education_pct^2)" = "Secondary Education^2", "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
    gt::cell_fill(color = 'white'),
    gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

::::
  
:::: {.column width="60%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(ggpredict(model1b, terms = "secondary_education_pct")) +
  labs(x = "Secondary Education (%)",
       y = "Average Life Expectancy (years)")
```

### Residuals 1

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model1b, which = 2)
```

### Residuals 2

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model1b, which = 1)
```

:::
  
::::
:::::
  
::: notes

**Have we now improved the fit of the line to the data?**
  
- **In other words, is this a better summary of the relationship?**
  
- Added a significant coefficient, increased the R2 by a lot, reduced the error, fixed the nonlinear residuals!
  
<br>
  
**Ok, what is the story of this line?**
  
- **How does secondary education relate to life expectancy in our global sample?**
  
<br>
  
- At low levels of secondary education the effect is strong and positive

    - e.g. Spreading secondary ed is associated with longer lives

- HOWEVER, once you get beyond the 60-ish% the effect flattens out

- So, the effect of adding more education beyond that point no longer translates into higher life expectancies!

<br>
  
**Anybody want to take a stab at explaining why we might see this relationship in the real world?**

- (Life expectancies are caused by many, many factors and education is just one of them!)

- It matters, but can't replace the others!
  
<br>
  
**Questions on adding a squared predictor to curve the line?**

- Checking the residuals led us to a REALLY cool finding that we would have completely missed if we only checked missing data, significance and $R^2$
  
<br>
  
**SLIDE**: Let's keep practicing

:::




## {background-image="Images/background-data_blue_v4.png" .smaller}

::: {.r-fit-text}
**Do wealthier countries tend to have longer average life expectancies?**
:::

::::: {.columns}
:::: {.column width="40%"}

<br>
  
```{r, echo=FALSE}
wdi$gdp_pc_k <- wdi$gdp_pc_current_usd/1000

model2 <- lm(data = wdi, life_expectancy ~ gdp_pc_k)

modelsummary(models = model2, 
             output = "gt",
             fmt = 2,
             stars = c('*' = .05),
             gof_omit = "IC|F|Log",
             coef_map = c("gdp_pc_k" = "GDP per capita (k, USD)", "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

::::
  
:::: {.column width="60%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(ggpredict(model2, terms = "gdp_pc_k")) +
  labs(x = "GDP per capita (k, USD)",
       y = "Average Life Expectancy (years)")
```

### Residuals 1

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model2, which = 2)
```

### Residuals 2

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model2, which = 1)
```

:::
  
::::
:::::

::: notes

**Did everybody rescale GDP pc into thousands?**

<br>

**What is your answer to our second homework question?**

<br>

- **In substantive terms, how big is the effect?**

- **How useful is this as policy guidance?**

<br>

**But does the model fit the data well? Evaluate the statistical significance for me**

- (Residuals 1: Serious deviations in both tails of the QQ plot)

- (Residuals 2: That is a CLEAR nonlinear effect!)

<br>

Ok, you know where this is going!

- Everybody refit this model but add a squared version of GDP and get ready to report back on the new model

:::






## {background-image="Images/background-data_blue_v4.png" .smaller}

::: {.r-fit-text}
**Do wealthier countries tend to have longer average life expectancies?**
:::

::::: {.columns}
:::: {.column width="40%"}

<br>
  
```{r, echo=FALSE}
wdi$gdp_pc_k <- wdi$gdp_pc_current_usd/1000

model2 <- lm(data = wdi, life_expectancy ~ gdp_pc_k)
model2b <- lm(data = wdi, life_expectancy ~ gdp_pc_k + I(gdp_pc_k^2))


modelsummary(models = list(model2, model2b), 
             output = "gt",
             fmt = 3,
             stars = c('*' = .05),
             gof_omit = "IC|F|Log",
             coef_map = c("gdp_pc_k" = "GDP per capita (k, USD)", 
                          "I(gdp_pc_k^2)" = "GDP pc^2",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

::::
  
:::: {.column width="60%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(ggpredict(model2b, terms = "gdp_pc_k")) +
  labs(x = "GDP per capita (k, USD)",
       y = "Average Life Expectancy (years)")
```

### Residuals 1

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model2b, which = 2)
```

### Residuals 2

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model2b, which = 1)
```

:::
  
::::
:::::

::: notes

I'm using three digits after the decimal place because the effect of the squared term is tiny

<br>

**Ok, what have you learned from this new model?**

- (As wealth increases life expectancy increases up to a certain level and then more money doesn't immediately translate to life expectancy)

<br>

**Changes in the substantive story?**

- **Why is the squared terms so small?**

    - (GDP pc values get BIG and squaring those values produces HUGE numbers)

<br>

**Have we fixed the residuals problem?**

- (Much better, but not yet...)

<br>

Ok, everybody do this again but now ADD a cubed term to the model

- e.g. gdp, gdp squared and gdp cubed

- Go!

:::



## {background-image="Images/background-data_blue_v4.png" .smaller}

::: {.r-fit-text}
**Do wealthier countries tend to have longer average life expectancies?**
:::

::::: {.columns}
:::: {.column width="40%"}

<br>
  
```{r, echo=FALSE}
model2c <- lm(data = wdi, life_expectancy ~ gdp_pc_k + I(gdp_pc_k^2) + I(gdp_pc_k^3))


modelsummary(models = list(model2, model2b, model2c), 
             output = "gt",
             fmt = 3,
             stars = c('*' = .05),
             gof_omit = "IC|F|Log",
             coef_map = c("gdp_pc_k" = "GDP per capita (k, USD)", 
                          "I(gdp_pc_k^2)" = "GDP pc^2",
                          "I(gdp_pc_k^3)" = "GDP pc^3",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "large")
  ), locations = gt::cells_body())
```

::::
  
:::: {.column width="60%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(ggpredict(model2c, terms = "gdp_pc_k")) +
  labs(x = "GDP per capita (k, USD)",
       y = "Average Life Expectancy (years)")
```

### Residuals 1

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model2c, which = 2)
```

### Residuals 2

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model2c, which = 1)
```

:::
  
::::
:::::

::: notes

**Ok, what have you learned from this new model?**

<br>

**Changes in the substantive story?**

<br>

**Have we fixed the residuals problem?**

- (Not yet...)

<br>

**SLIDE**: I won't ask you to keep going with this one, but I will show the next version

:::



## {background-image="Images/background-data_blue_v4.png" .smaller}

::: {.r-fit-text}
**Do wealthier countries tend to have longer average life expectancies?**
:::

::::: {.columns}
:::: {.column width="40%"}

<br>
  
```{r, echo=FALSE}
model2d <- lm(data = wdi, life_expectancy ~ gdp_pc_k + I(gdp_pc_k^2) + I(gdp_pc_k^3) + I(gdp_pc_k^4))

modelsummary(models = list(model2, model2b, model2c, model2d), 
             output = "gt",
             fmt = 3,
             stars = c('*' = .05),
             gof_omit = "IC|F|Log",
             coef_map = c("gdp_pc_k" = "GDP per capita (k, USD)", 
                          "I(gdp_pc_k^2)" = "GDP pc^2",
                          "I(gdp_pc_k^3)" = "GDP pc^3",
                          "I(gdp_pc_k^4)" = "GDP pc^4",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "medium")
  ), locations = gt::cells_body())
```

::::
  
:::: {.column width="60%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(ggpredict(model2d, terms = "gdp_pc_k")) +
  labs(x = "GDP per capita (k, USD)",
       y = "Average Life Expectancy (years)")
```

### Residuals 1

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model2d, which = 2)
```

### Residuals 2

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model2d, which = 1)
```

:::
  
::::
:::::

::: notes

You can continuously add additional multiples of your predictors to the regression model and it will let you tune your model to essentially any number of turns of direction

- The danger here is that we are overfitting the sample data

- REMEMBER, the aim of inferential statistics is to learn about the population from the sample

- I don't want a perfect model for this sample, I want my best estimate for how wealth impacts life expectancy

<br>

SINCE the residuals vs fitted values plot suggests we still haven't addressed the non-linearity I have to consider other options!

- **What do we know about the distribution of GDP pc? Is it normally distributed?**

- (GDP pc has a huge right skew and that might be preventing OLS from getting into the middle!)

<br>

SO, let's try our second approach to adjusting an OLS regression for non-linear residuals

<br>

**SLIDE**: Log transformations

:::



## {background-image="Images/background-data_blue_v4.png" .center}

:::: {.columns}
::: {.column width="50%"}
**Original Scale**

```{r, echo=FALSE, fig.align='center', fig.retina=3, fig.asp=.8, fig.width=5}
#summary(wdi$gdp_pc_current_usd)

wdi |>
  ggplot(aes(x = gdp_pc_current_usd)) +
  geom_histogram(color = "white", bins = 20) +
  theme_bw() +
  labs(x = "GDP per capita (USD)", y = "Count")
```
:::

::: {.column width="50%"}
**Log Scale**

```{r, echo=FALSE, fig.align='center', fig.retina=3, fig.asp=.8, fig.width=5}
wdi$gdppc_log <- log(wdi$gdp_pc_current_usd)

#summary(log(wdi$gdppc_log))

wdi |>
  ggplot(aes(x = gdppc_log)) +
  geom_histogram(color = "white", bins = 20) +
  theme_bw() +
  labs(x = "GDP per capita (logged, USD)", y = "Count")
```
:::
::::

::: {.fragment}
```{r, echo=TRUE}
model2_log <- lm(data = wdi, life_expectancy ~ I(log(gdp_pc_current_usd)))
```
:::

::: notes

Back in Week 6 we explored how we could use a log transformation to help us analyze right skewed data

<br>

GDP per capita is clearly right skewed

- Most countries in the world are on the poorer end of the spectrum with a few massive outliers

<br>

Think of this from the perspective of OLS

- How do you draw a straight line through the original distribution that keeps the same amount of error on both sides?

- But so much easier with the logged version!

<br>

**REVEAL**: Everybody run this model and get ready to report back on its substantive and statistical significance

- Now, remember, you have to take the log of the ORIGINAL variable, NOT the thousands version

- The log of 1 is different from the log of 1,000

<br>

(**SLIDE**)

:::



## {background-image="Images/background-data_blue_v4.png" .smaller}

::: {.r-fit-text}
**Do wealthier countries tend to have longer average life expectancies?**
:::

::::: {.columns}
:::: {.column width="40%"}

<br>
  
```{r, echo=FALSE}
model2_log <- lm(data = wdi, life_expectancy ~ I(log(gdp_pc_current_usd)))

modelsummary(models = list(model2, model2b, model2c, model2_log), 
             output = "gt",
             fmt = 3,
             stars = c('*' = .05),
             gof_omit = "IC|F|Log",
             coef_map = c("gdp_pc_k" = "GDP per capita (k, USD)", 
                          "I(gdp_pc_k^2)" = "GDP pc^2",
                          "I(gdp_pc_k^3)" = "GDP pc^3",
                          "I(gdp_pc_k^4)" = "GDP pc^4",
                          "I(log(gdp_pc_current_usd))" = "log(GDP pc)",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "medium")
  ), locations = gt::cells_body())
```

::::
  
:::: {.column width="60%"}

::: {.panel-tabset}

### Predictions

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(ggpredict(model2_log, terms = "gdp_pc_current_usd")) +
  labs(x = "GDP per capita (USD)",
       y = "Average Life Expectancy (years)") +
  scale_x_continuous(labels = scales::dollar_format())
```

### Residuals 1

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model2_log, which = 2)
```

### Residuals 2

```{r, echo=FALSE, fig.retina=3, fig.asp=.9, fig.align='center', out.width='90%', fig.width=7}
plot(model2_log, which = 1)
```

:::
  
::::
:::::

::: notes

**Ok, what have you learned from this new model?**

- **Changes in the substantive story?**

<br>

Our model here suggests increasing the log value of GDP pc by one increases life expectancy by four years

- **Remind me, what does a change of one on the log scale look like in the normal scale?**

- (Multiply the amount by Euler's number, 2.72)

- (Rule of thumb, tripling of the base)

- (SO, triple your GDP pc and add 4 years life expectancy on average)

<br>

The ggpredict function translates this for us back into the normal scale

- At lower ends of the predictions the effect is fast (easy to triple a small GDP number)

- At the high end, tripling is REALLY hard so the effect slows way down!

<br>

**Have we fixed the residuals problem?**

<br>

**Ok, questions on our two methods for addressing nonlinearity in residuals?**

- If you suspect the effect changes direction try adding a squared or cubed term to the predictors

- If the predictor or outcome is right-skewed try taking the log

<br>

**SLIDE**: For next class

:::



## For Next Class {background-image="Images/background-data_blue_v4.png" .center}

<br>

**Do countries with more basic levels of education tend to be wealthier on average?**

1. Fit and evaluate four versions of this model (baseline, squared term, logged DV and both)

2. Submit your "best" model to Canvas

::: notes

Everybody check out the assignment on Canvas and let me know if you have any questions

- Go!

<br>

Canvas Assignment: Developing Better Models of Global Wealth

- Data: World_Development_Indicators-2023-Life_Expectancy.xlsx
- RQ: Do countries with more basic levels of education (primary_education_pct) tend to be wealthier on average (gdp_pc_current_usd)?

Fit the following models and select the one you argue "best" answers our research question:

1. Baseline: Regress gdp_pc_current_usd on primary_education_pct
2. Squared: Regress gdp_pc_current_usd on primary_education_pct and a squared version of the predictor
3. Logged: Regress a logged version of gdp_pc_current_usd on primary_education_pct
4. Both: Regress a logged version of gdp_pc_current_usd on primary_education_pct and a squared version of the predictor

Submit: 

1. A polished regression table with all four models
2. An argument in favor of one of the models above the other three based on their fits to the data (with the evidence you are using to support the argument)
3. An explanation of your answer to the RQ using your chosen model (with a prediction plot)

<br>

```{r}
wdi$gdp_pc_k <- wdi$gdp_pc_current_usd/1000

educ1 <- lm(data = wdi, gdp_pc_k ~ primary_education_pct)
educ2 <- lm(data = wdi, gdp_pc_k ~ primary_education_pct + I(primary_education_pct^2))
educ3 <- lm(data = wdi, log(gdp_pc_current_usd) ~ primary_education_pct)
educ4 <- lm(data = wdi, log(gdp_pc_current_usd) ~ primary_education_pct + I(primary_education_pct^2))

modelsummary(models = list("GDP pc (K)" = educ1, educ2, "log(GDP)" = educ3, educ4), 
             output = "gt",
             fmt = 2,
             stars = c('*' = .05),
             gof_omit = "IC|F|Log")

plot(educ1, which = 1)
plot(educ1, which = 2)

plot(educ3, which = 1)
plot(educ3, which = 2)

plot(ggpredict(educ3, terms = "primary_education_pct"))



















```





:::

