---
format: revealjs
---

# Today's Agenda {background-image="Images/background-data_blue_v3.png"}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)

d <- read_excel("../Data_in_Class/Class11-1-Simple_OLS-WDI_LifeExpectancy/WDI-Practice_Simple_OLS-2024-04-04.xlsx", na = "NA")
```

<br>

::: {.r-fit-text}
**Multivariate Analyses**

- Adding confounders to OLS regressions
:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2025)
:::

::: notes
Prep for Class

1. Assignment for today used the WDI data on life expectancy

2. Make sure merged project data is available

:::


## For Today {background-image="Images/background-slate_v2.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

**Possible Confounders**: GDP per capita, (1) unemployment rate, (2) tobacco use, (3) compulsory education

::: notes

For today I asked you to practice fitting, evaluating and interpreting three multiple regression models

- **How did this go?**

:::


## {background-image="Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Do higher fertility rates lower average life expectancies around the world?**
:::

```{r}
d$gdp_pc_10k <- d$gdp_per_capita/10000

res1 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman)
res2 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman + gdp_pc_10k)
res3 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman + unemployment_pct)
res4 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman + tobacco_use_pct)
res5 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman + compulsory_education_yrs)

modelsummary(models = list(res1, res2, res3, res4, res5), 
             out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", 
                          "gdp_pc_10k" = "GDP pc ($10k)", 
                          "unemployment_pct" = "Unemployment (%)",
                          "tobacco_use_pct" = "Tobacco Use (%)",
                          "compulsory_education_yrs" = "Education (years)",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "15px")
  ), locations = gt::cells_body())

```

::: notes

Quick note, I don't know why modelsummary is omitting some of the F-statistics...

<br>

**Did everybody get these results?**

<br>

**Interpret each coefficient for me**

- **At what level increase does each predictor make a serious impact on life expectancy?**

<br>

**Any missing data concerns?**

<br>

**Any meaningful changes in the R2?**

<br>

**SLIDE**: Let's check the residuals

:::


## {background-image="Images/background-slate_v2.png"}

:::: {.columns}
::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.625, fig.width=8}
plot(res2, which = 1)

plot(res3, which = 1)
```
:::

::: {.column width="50%"}
```{r, fig.align='center', fig.asp=.625, fig.width=8}
plot(res4, which = 1)

plot(res5, which = 1)
```
:::
::::

::: notes

**Takeaways from checking the residuals?**

<br>

Looks to me like only adjusting for wealth deals with the underlying non-linear effect

- I suspect wealth is a big confounder in our causal mechanism and omitting it messes up the estimated relationship between fertility and life expectancy

:::


## {background-image="Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Do higher fertility rates lower average life expectancies around the world?**
:::

:::: {.columns}
::: {.column width="50%"}

<br>

```{r, fig.align = 'center', fig.asp=1, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.4, y = 1, yend = 1, arrow = arrow(), linewidth = 1.4) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.15, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.15, arrow = arrow())
```
:::

::: {.column width="50%"}
```{r}
modelsummary(models = list(res1, res2, res3, res4, res5), 
             out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", 
                          "gdp_pc_10k" = "GDP pc ($10k)", 
                          "unemployment_pct" = "Unemployment (%)",
                          "tobacco_use_pct" = "Tobacco Use (%)",
                          "compulsory_education_yrs" = "Education (years)",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "15px")
  ), locations = gt::cells_body()) |>
  gt::tab_style(style = gt::cell_fill(color = 'orange'), locations = gt::cells_body(columns = 1:6, rows = 1))

```
:::
::::

::: notes

Remember, the goal of this exercise is to use an identification strategy to make a causal argument about fertility and life expectancy

- **Bottom line, what do we learn from these regressions?**

<br>

- Robustness Test: It is reassuring that the coefficient on fertility rate is fairly stable across models

- The direction of effect makes sense on wealth and unemployment (the other two aren't significant or large enough to interpret)

<br>

**Did anybody compare their regression results to the multivariate visualizations we made in class on Monday?**

- (**SLIDE**: Compare to our multivariate visualizations last week)

:::



## {background-image="Images/background-slate_v2.png"}

::: {.r-fit-text}
**Multivariate Analyses: Education as a Confounder**
:::

:::: {.columns}
::: {.column width="70%"}
```{r, echo=FALSE, fig.asp=.85, fig.align='center', fig.width=8, cache=TRUE}
d |>
  mutate(
    education2 = case_when(
      compulsory_education_yrs < quantile(d$compulsory_education_yrs, probs = .25, na.rm = T) ~ "Education (Q1)",
      compulsory_education_yrs < quantile(d$compulsory_education_yrs, probs = .50, na.rm = T) ~ "Education (Q2)",
      compulsory_education_yrs < quantile(d$compulsory_education_yrs, probs = .75, na.rm = T) ~ "Education (Q3)",
      compulsory_education_yrs >= quantile(d$compulsory_education_yrs, probs = .75, na.rm = T) ~ "Education (Q4)"
      )
    ) |>
  select(fertility_rate_per_woman, life_expectancy_total, education2) |>
  na.omit() |>
  ggplot(aes(x = fertility_rate_per_woman, y = life_expectancy_total, color = education2, shape = education2)) +
  geom_point() + 
  geom_smooth(method = "lm", se = T) +
  theme_bw() +
  facet_wrap(~ education2, ncol = 2) +
  guides(color = "none", shape = "none") +
  labs(x = "Fertility Rate (births per woman)", y = "Life Expectancy (years)",
       caption = "Source: WDI (2024) for data from 2020")
```
:::

::: {.column width="30%"}

<br>

```{r}
modelsummary(models = list(res5), 
             out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", 
                          "compulsory_education_yrs" = "Education (years)",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "16px")
  ), locations = gt::cells_body())
```
:::
::::

::: notes

Ok, let's make sure everybody understands the intuitions of our work on multivariate analyses over the last few weeks

<br>

Help me understand what I'm seeing here.

- On the left, I see scatter plots where the negative effect of fertility rate on life expectancy seems consistent across the levels of education

- On the right, the results of a multiple regression showing education is not a significant predictor of life expectancy

- **Are these two pieces of data analysis consistent with each other or not?**

<br>

This is KEY to understanding our work over the last few weeks

- Never forget our focus is to use data analyses **to make a causal argument that changes in fertility rates CHANGE average life expectancy** around the world

- NONE of these tables or pictures tell us about the causal effect of education on life expectancy

- We are interested in the effect of fertility rates WHEN CONTROLLING FOR EDUCATION

- In other words, for this identification strategy we assume that countries with high vs low fertility rates differ in their levels of education

- e.g. we cannot assume they are identical on average

<br>

**If you were interested in estimating the effect of education on life expectancy, what would you have to do?**

- (**SLIDE**: Generate a brand new ID strategy!)

:::



## {background-image="Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**A new research question...**
:::

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Compulsory\nEducation", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

::: {.r-fit-text}
**Requires a new ID strategy!**
:::

::: notes

In this story, I do not think fertility rates are a confounder

<br>

Probably a mediator, no?

- Education changes fertility rates which changes life expectancy

- That would mean a regression with both fertility and education will produce an underestimate of the importance of education!

<br>

**SLIDE**: Let's check!

:::



## {background-image="Images/background-slate_v2.png" .center}

**Do countries that require more years of compulsory education have higher life expectancy?**

<br>

:::: {.columns}
::: {.column width="50%"}

```{r, fig.align = 'center', fig.asp=1, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Compulsory\nEducation", "Fertility\nRate", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.6, xend = 2.4, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1, xend = 1.7, y = 1.15, yend = 1.85, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.15, arrow = arrow())
```
:::

::: {.column width="50%"}

```{r}
res6 <- lm(data = d, life_expectancy_total ~ compulsory_education_yrs)

modelsummary(models = list(res6, res1, res5), 
             out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("compulsory_education_yrs" = "Education (years)",
                          "fertility_rate_per_woman" = "Fertility (births per woman)", 
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "15px")
  ), locations = gt::cells_body())
```
:::
::::

::: notes

Regression results

- Model 1 shows that compulsory education increases life expectancy considerably

- Model 2 shows that fertility rate increases lower life expectancy dramatically

- And Model 3 shows when you include both it looks like fertility rate eats up education as you would expect if it is a mediator

<br>

**Does this make sense?**

:::



## {background-image="Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Do higher fertility rates lower average life expectancies around the world?**
:::

:::: {.columns}
::: {.column width="50%"}

<br>

```{r, fig.align = 'center', fig.asp=1, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.4, y = 1, yend = 1, arrow = arrow(), linewidth = 1.4) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.15, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.15, arrow = arrow())
```
:::

::: {.column width="50%"}
```{r}
modelsummary(models = list(res1, res2, res3, res4, res5), 
             out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", 
                          "gdp_pc_10k" = "GDP pc ($10k)", 
                          "unemployment_pct" = "Unemployment (%)",
                          "tobacco_use_pct" = "Tobacco Use (%)",
                          "compulsory_education_yrs" = "Education (years)",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "15px")
  ), locations = gt::cells_body()) |>
  gt::tab_style(style = gt::cell_fill(color = 'orange'), locations = gt::cells_body(columns = 1:6, rows = 1))

```
:::
::::

::: notes

All of this is just meant to reinforce the key intuition here

- We selected a mechanism of interest,

- We developed an identification strategy for it alone, and now

- We are adjusting our analyses for the confounders

<br>

In either case, this is all a reminder not to become infatuated with your regression tables

- The confounders are there to help us estimate the main relationship, that's it.

- We don't really care about the significance of wealth, unemployment, education, etc

<br>

**Any questions on all of this?**

:::



## Why use OLS regressions? {background-image="Images/background-slate_v2.png" .center}

::: {.r-fit-text}
- Assumes **"regression to the mean"**

- **Quantifies** the relationship between variables

- Uses **ALL** of the data

- Provides estimates of **uncertainty**

- Can include a **confounder**
:::



## Why use multiple regressions? {background-image="Images/background-slate_v2.png" .center}

::: {.r-fit-text}
- Can include more than one **confounder**

- Can be adjusted for **non-linear effects**
:::



## {background-image="Images/background-slate_v2.png" .center}

::: {.r-fit-text}
**Do higher fertility rates lower average life expectancies around the world?**
:::

:::: {.columns}
::: {.column width="50%"}

<br>

```{r, fig.align = 'center', fig.asp=1, fig.width=7, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Fertility\nRate", "Confounders", "Life\nExpectancy")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.4, y = 1, yend = 1, arrow = arrow(), linewidth = 1.4) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.15, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.15, arrow = arrow())
```
:::

::: {.column width="50%"}
```{r}
d <- read_excel("../Data_in_Class/Class11-1-Simple_OLS-WDI_LifeExpectancy/WDI-Practice_Simple_OLS-2024-04-04.xlsx", na = "NA")

d$gdp_pc_10k <- d$gdp_per_capita/10000

res1 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman)
res2 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman + gdp_pc_10k)
res3 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman + unemployment_pct)
res4 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman + tobacco_use_pct)
res5 <- lm(data = d, life_expectancy_total ~ fertility_rate_per_woman + compulsory_education_yrs)

modelsummary(models = list(res1, res2, res3, res4, res5), 
             out = "gt", 
             fmt = 2, stars = c("*" = .05), 
             gof_omit = "IC|Log", 
             coef_map = c("fertility_rate_per_woman" = "Fertility (births per woman)", 
                          "gdp_pc_10k" = "GDP pc ($10k)", 
                          "unemployment_pct" = "Unemployment (%)",
                          "tobacco_use_pct" = "Tobacco Use (%)",
                          "compulsory_education_yrs" = "Education (years)",
                          "(Intercept)" = "Constant")) |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "15px")
  ), locations = gt::cells_body()) |>
  gt::tab_style(style = gt::cell_fill(color = 'orange'), locations = gt::cells_body(columns = 1:6, rows = 1))

```
:::
::::

::: notes

**Alright, from these options what is our identification strategy for this relationship?**

- *ON BOARD*

<br>

Ok, everybody fit the new multiple regression that includes all of the confounders we decided were necessary to identify the relationship.

- *You code this too!*

<br>

*FIRST, class leads you through evaluating the model*

1. Check for missing data (and where is it missing from?)

2. Are the coefficients significant?

3. Is the F test significant?

4. What does the R2 tell us?

5. What do the residuals indicate about the fit?

<br>

*SECOND, ask the class to generate predictions using ggeffects:ggpredict*

:::



## How confident are you that higher fertility rates directly lower average life expectancies around the world? {background-image="Images/background-slate_v2.png" .center}

<br>

Confidence depends on the quality of...

1. The Data

2. The Sample

3. The Regression


::: notes

Ok, data analysts, the big question is ultimately how convinced are you from our work that increasing fertility rates directly reduces average life expectancy around the world?

<br>

Your certainty should depend on the quality of:

1. The data (methodolgy and measurement)

2. The sample (representative of the population or not)

3. The regression

<br>

Let's consider each of those:

**FIRST, what's your level of confidence in the quality of the data pulled form the World Bank?**

<br>

**SECOND, what's your level of confidence that we have a representative sample of the countries in the world in 2022?**

<br>

**THIRD, what's your level of confidence in the fit of the multiple regression model?**

<br>

**Bottom line, have we built a useful causal argument or not?**

::: 



## The Ross (1990) Dataset {background-image="Images/background-slate_v2.png" .center}

<br>

Ross, Catherine E. (1990) *Work, Family and Well-being in the United States*. Inter-university Consortium for Political and Social Research (ICPSR).

<br>

This study measures the effects of various social conditions on individuals' physical and mental health.

::: notes

Let's practice this again!

<br>

Everybody grab the data and the codebook from Canvas (ICPSR_Ross_Survey_1990.xlsx)

<br>

Note: Since 1990, the USD is up 237% so I adjusted the reported income levels for this inflation so we can work with what $ look like for us today

- example: $50k in 1990 is $118k in 2024

```{r, eval=FALSE}
names(d)

lm(data = d, earnk2024 ~ education) |> summary()
lm(data = d, earnk2024 ~ ethnicity) |> summary()
lm(data = d, earnk2024 ~ height) |> summary()
lm(data = d, earnk2024 ~ male) |> summary()
lm(data = d, earnk2024 ~ exercise) |> summary()



lm(data = d, earnk2024 ~ education) |> summary()
lm(data = d, earnk2024 ~ education + male) |> summary()
lm(data = d, earnk2024 ~ education + angry) |> summary()
lm(data = d, earnk2024 ~ education + ethnicity) |> summary()



```

:::


## {background-image="Images/background-slate_v2.png" .center}

```{r, fig.retina = 3, fig.align = 'center', out.width='80%', fig.height=4, fig.width=9, cache=TRUE}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Education", "Confounders", "Personal\nIncome")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 8) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.25, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

**Possible Confounders**: Height, weight, gender, ethnicity, walk, exercise, smokenow, tense, angry, age

::: notes

**Alright, from these options what is our identification strategy for this relationship?**

- *ON BOARD*

<br>

Ok, everybody fit the new multiple regression that includes all of the confounders we decided were necessary to identify the relationship.

- Include the baseline regression too (education -> earnk2024)

- *You code this too!*

<br>

*FIRST, class leads you through evaluating the model*

1. Check for missing data (and where is it missing from?)

2. Are the coefficients significant?

3. Is the F test significant?

4. What does the R2 tell us?

5. What do the residuals indicate about the fit?

<br>

*SECOND, ask the class to generate predictions using ggeffects:ggpredict*

<br>

**Codebook Notes**

"height"

- What is your height without shoes on? (inches)

"weight"

- What is your weight without clothing? (lb)

"male"

- dummy variable = 1 if male and 0 if female

"earn"

- During 1989, what was your personal income from your own wages, salary, or other sources, before taxes?

"earnk"

- As above but rescaled into thousands of dollars

"ethnicity"

"education"

- What is the highest grade or year in school that you have completed?
    - None . . . . . . . . . . . . . . . . . . . .  00
    - Elementary . . . . . . . 01 02 03 04 05 06 07 08
    - High school . . . . . . . .  . . . . 09 10 11 12
    - College . . . . . . . . . .  . . . . 13 14 15 16
    - Some graduate school . . . . .  . . . . . . . 17
    - Graduate or professional degree . . . . . . . 18

"mother_education"

- What is the highest grade or year of school your mother completed?
    - None . . . . . . . . . . . . . . . . . . . .  00
    - Elementary . . . . . . . 01 02 03 04 05 06 07 08
    - High school . . . . . . . .  . . . . 09 10 11 12
    - College . . . . . . . . . .  . . . . 13 14 15 16
    - Some graduate school . . . . .  . . . . . . . 17
    - Graduate or professional degree . . . . . . . 18
    - DON'T KNOW  . . . . . . . . . . . . . . . . . 98

"father_education"

- What is the highest grade or year of school your father completed?
    - None . . . . . . . . . . . . . . . . . . . .  00
    - Elementary . . . . . . . 01 02 03 04 05 06 07 08
    - High school . . . . . . . .  . . . . 09 10 11 12
    - College . . . . . . . . . .  . . . . 13 14 15 16
    - Some graduate school . . . . .  . . . . . . . 17
    - Graduate or professional degree . . . . . . . 18
    - DON'T KNOW  . . . . . . . . . . . . . . . . . 98

"walk"

- How often do you take a walk? (Includes walking to work/train station etc.) Would you say . . . 
    - Never, . . . . . . . . . . .  01
    - Once a month or less, . . . . 02
    - About twice a month, . . . .  03
    - About once a week, . . . . .  04
    - Twice a week, . . . . . . . . 05
    - Three times a week, or . . .  06
    - More than 3 times a week? . . 07
    - EVERY DAY . . . . . . . . . . 08

"exercise"

- How often do you do strenuous exercise such as running, basketball, aerobics, tennis, swimming, biking, and so on? Would you say . . .
    - Never, . . . . . . . . . . .  1
    - Once a month or less, . . . . 2
    - About twice a month, . . . .  3
    - About once a week, . . . . .  4
    - Twice a week, . . . . . . . . 5
    - Three times a week, or . . .  6
    - More than 3 times a week? . . 7

"smokenow"

- Do you currently smoke 7 or more cigarettes a week?
    - Yes . . . . . . . . . . . . . 1
    - No  . . . . . . . . . . . . . 2

"tense"

- On how many of the past 7 days have you felt tense or anxious?

"angry"

- On how many of the past 7 days have you felt angry?

"age"

:::



## How confident are you that higher levels of education directly increase yearly earnings? {background-image="Images/background-slate_v2.png" .center}

<br>

Confidence depends on the quality of...

1. The Data

2. The Sample

3. The Regression

::: notes

Ok, data analysts, the big question is ultimately how convinced are you from our work that pursuing more education will increase your personal earnings?

<br>

Let's consider each of those:

**FIRST, what's your level of confidence in the quality of the dataset built by Ross (1990)?**

<br>

**SECOND, what's your level of confidence that we have a representative sample of the people in the US in 1990?**

- **Is data from 30 years ago still applicable today? Why or why not?**

<br>

**THIRD, what's your level of confidence in the fit of the multiple regression model?**

<br>

**Bottom line, have we built a useful causal argument or not?**


<br>

**Codebook Notes**

"height"

- What is your height without shoes on? (inches)

"weight"

- What is your weight without clothing? (lb)

"male"

- dummy variable = 1 if male and 0 if female

"earn"

- During 1989, what was your personal income from your own wages, salary, or other sources, before taxes?

"earnk"

- As above but rescaled into thousands of dollars

"ethnicity"

"education"

- What is the highest grade or year in school that you have completed?
    - None . . . . . . . . . . . . . . . . . . . .  00
    - Elementary . . . . . . . 01 02 03 04 05 06 07 08
    - High school . . . . . . . .  . . . . 09 10 11 12
    - College . . . . . . . . . .  . . . . 13 14 15 16
    - Some graduate school . . . . .  . . . . . . . 17
    - Graduate or professional degree . . . . . . . 18

"mother_education"

- What is the highest grade or year of school your mother completed?
    - None . . . . . . . . . . . . . . . . . . . .  00
    - Elementary . . . . . . . 01 02 03 04 05 06 07 08
    - High school . . . . . . . .  . . . . 09 10 11 12
    - College . . . . . . . . . .  . . . . 13 14 15 16
    - Some graduate school . . . . .  . . . . . . . 17
    - Graduate or professional degree . . . . . . . 18
    - DON'T KNOW  . . . . . . . . . . . . . . . . . 98

"father_education"

- What is the highest grade or year of school your father completed?
    - None . . . . . . . . . . . . . . . . . . . .  00
    - Elementary . . . . . . . 01 02 03 04 05 06 07 08
    - High school . . . . . . . .  . . . . 09 10 11 12
    - College . . . . . . . . . .  . . . . 13 14 15 16
    - Some graduate school . . . . .  . . . . . . . 17
    - Graduate or professional degree . . . . . . . 18
    - DON'T KNOW  . . . . . . . . . . . . . . . . . 98

"walk"

- How often do you take a walk? (Includes walking to work/train station etc.) Would you say . . . 
    - Never, . . . . . . . . . . .  01
    - Once a month or less, . . . . 02
    - About twice a month, . . . .  03
    - About once a week, . . . . .  04
    - Twice a week, . . . . . . . . 05
    - Three times a week, or . . .  06
    - More than 3 times a week? . . 07
    - EVERY DAY . . . . . . . . . . 08

"exercise"

- How often do you do strenuous exercise such as running, basketball, aerobics, tennis, swimming, biking, and so on? Would you say . . .
    - Never, . . . . . . . . . . .  1
    - Once a month or less, . . . . 2
    - About twice a month, . . . .  3
    - About once a week, . . . . .  4
    - Twice a week, . . . . . . . . 5
    - Three times a week, or . . .  6
    - More than 3 times a week? . . 7

"smokenow"

- Do you currently smoke 7 or more cigarettes a week?
    - Yes . . . . . . . . . . . . . 1
    - No  . . . . . . . . . . . . . 2

"tense"

- On how many of the past 7 days have you felt tense or anxious?

"angry"

- On how many of the past 7 days have you felt angry?

"age"

:::



## Next Week: The Final Report {background-image="Images/background-slate_v2.png" .center}

Are powerful states more likely to respect the human rights of their citizens?

<br>

::: {.r-fit-text}

**IDEA Evaluations**

https://drury.campuslabs.com/eval-home/

:::

::: notes

Next week the final reports!

<br>

Also, please take the time to do the course evals

- I really appreciate all the feedback!

:::







