---
format: revealjs
---

## Today's Agenda {background-image="Images/background-data_blue_v4.png" .center}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
#library(modelsummary)
library(crosstable)

load("../../DATA/2026-Spring/ANES/ANES2020_Data_Extract.RData")

## Notes on modelsummary
# # Install the package ONE TIME
# install.packages("modelsummary")
# 
# # Load the package EVERY time
# library(modelsummary)
# 
# # Cross-tab with row %s
# datasummary_crosstab(data = mpg, drv ~ class, 
#                      statistic = ~ Percent("row"))
# 
# # Cross-tab with column %s
# datasummary_crosstab(data = mpg, drv ~ class, 
#                      statistic = ~ Percent("col"))
# 
# # Cross-tab with counts AND column %s
# datasummary_crosstab(data = mpg, drv ~ class, 
#                      statistic = ~ N + Percent("col"))
```

<br>

::: {.r-fit-text}

**Controlled Comparison Analysis**

- Multivariate descriptive statistics

- New R file: Making_Controlled_Comparisons.R

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2026)
:::

::: notes
Prep for Class

1. Review Canvas submissions

<br>

**SLIDE**: Let's refresh our work thus far on making causal arguments

:::



## Building a Causal Argument {background-image="Images/background-data_blue_v4.png" .center}

<br>

"Causal inference refers to the process of drawing a conclusion that a specific treatment (i.e. intervention) was the 'cause' of the effect (or outcome) that was observed" (Frey 2018). 

<br>

```{r, fig.align = 'center', fig.width = 8, fig.height=1.25}
## Manual DAG
d1 <- tibble(
  x = c(-3, 3),
  y = c(1, 1),
  labels = c("Predictor", "Outcome")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 0, color = "white") +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = -1.9, xend = 2.2, y = 1, yend = 1, arrow = arrow())
```

::: notes

Remember, the aim of any scientific discipline is to generate new knowledge that explains why things happen in the world

- We refer to this as explanatory or causal analysis

- In short, we want to demonstrate that changes in our chosen predictors CAUSE a change in the outcome variable

<br>

**And what is the fundamental problem with causal inference?**

- **In other words, why is establishing causality so hard?**

- (**SLIDE**)

:::



## {.center background-image="Images/background-data_blue_v4.png"}

::: {.r-fit-text}
**The Fundamental Problem of Causal Inference**
:::

<br>

"To measure causal effects, we need to compare the factual outcome with the counter-factual outcome, but **we can never observe the counter-factual outcome**" (Llaudet and Imai 2023).

::: notes

Don't forget, this is a problem for ALL of science, not just the social sciences.

- Whether its gravity, speed, biological reproduction or whatever else, we cannot observe actual counter-factuals

- Even experimental methods do NOT actually solve this problem!

<br>

**And how do experimental methods, e.g. using a randomized controlled trial (RCT), address this problem?**

- (**SLIDE**)

:::



## Randomized Controlled Trials (RCTs) {.center background-image="Images/background-data_blue_v4.png"}

```{r, fig.retina=3, fig.align='center', fig.asp=.6, fig.width=10}
## Manual DAG
d2 <- tibble(
  x = c(2.5, 2, 3, 2, 3),
  y = c(3, 2, 2, 1, 1),
  labels = c("Participants", "Treatment\nGroup", "Control\nGroup", "Outcome", "Outcome")
)

ggplot(data = d2, aes(x = x, y = y)) +
  geom_point(size = 0, color = "white") +
  theme_void() +
  coord_cartesian(xlim = c(1.75, 3.25), ylim = c(0.75, 3.25)) +
  geom_label(aes(label = labels), size = 10) +
  annotate("segment", x = 2.5, xend = 2.5, y = 2.85, yend = 2.7, arrow = arrow(length = unit(0.1, "inches"))) +
  annotate("segment", x = 2.5, xend = 2.2, y = 2.5, yend = 2.2, arrow = arrow()) +
  annotate("segment", x = 2.5, xend = 2.85, y = 2.5, yend = 2.2, arrow = arrow()) +
  annotate("segment", x = 2, xend = 2, y = 1.75, yend = 1.15, arrow = arrow()) +
 annotate("segment", x = 3, xend = 3, y = 1.75, yend = 1.15, arrow = arrow()) +
  annotate("text", x = 2.5, y = 2.6, label = "Random Allocation", color = "red", size = 10)
```

::: notes

**How does an RCT help us address the fundamental problem?**

1. Random assignment means that even though both groups are composed of different individuals, the two groups are comparable to each other **on average** in all respects.

2. Random assignment means that the only thing that distinguishes the treatment group from the control group **BESIDES THE TREATMENT** is chance.

<br>

So, with random assignment we ASSUME the groups are identical on average which lets us pretend the outcome for group 1 represents the counterfactual for group 2

- **Make sense?**

<br>

Last class we explored the conceptual tools you need for dealing with situations where you cannot run an RCT

- **So, what do we do if we are analyzing observational data?**

- (**SLIDE**)

:::



## Making Controlled Comparisons {background-image="Images/background-data_blue_v4.png" .center}

```{r, fig.retina=3, fig.align='center', fig.asp=.6, fig.width=8}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Treatment", "Confounder", "Outcome")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 0, color = "white") +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.5, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.15, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.15, arrow = arrow())
```

::: notes

We develop **identification strategies** that identify the **confounders** in our observational relationships.

- This is what the textbook refers to as a "controlled comparison" (p160)

- The hope is that by controlling for all of the confounders we get to a place where the treatment and control groups are identical on average

<br>

**Any questions on our work from last class on making controlled comparisons?**

<br>

**SLIDE**: Seems like a good time to review the assignment for today!

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**Pollock & Edwards Exercise 4: Types of Relationship**
:::

<br>

Does frequent attendance at religious services reduce tolerance for allowing "radical" Muslim speech when controlling for education?

```{r}
tab1 <- tibble(
  allow = c(121, 44, 226, 162),
  not = c(197, 187, 140, 195)
) |>
  mutate(
    total = allow + not,
    allow2 = allow/total
  )
  
tab1a <- tibble(
  allow = c(121, 44, 226, 162),
  not = c(197, 187, 140, 195)
) |>
  mutate(
    total = allow + not,
    allow2 = str_c(round(allow/total*100, 0), "%"),
    not2 = str_c(round(not/total*100, 0), "%"),
    allow = str_c("(", allow, ")"),
    not = str_c("(", not, ")")
  ) |>
  select("Allow Speech" = allow2, allow, "Forbid Speech" = not2, not) |>
  t() 

row.names(tab1a) <- c("Allow Speech", "", "Forbid Speech", "")

tab1a |>
  kable(digits = 2, align = c("l", rep("c", 4)), col.names = c("", "Low Religion", "High Religion", "Low Religion", "High Religion")) |>
  add_header_above(c("", "Education: Low" = 2, "Education: High" = 2)) |>
  kable_styling(font_size = 25)
```

::: notes

**Did everybody calculate these percentages?**

<br>

**So, does the data support this hypothesis?**

- (Yes)

- Higher religious attendance, less support for free speech of radical clerics even when controlling for education

<br>

Now tell me about the partial relationship here

- **Is this a spurious, additive, or interaction relationship?**

- (**SLIDE**: Line plot)

:::



## {background-image="Images/background-data_blue_v4.png" .center}

Does frequent attendance at religious services reduce tolerance for allowing "radical" Muslim speech when controlling for education?

```{r, fig.align='center', fig.asp=.75, fig.width=7}
tab1 |>
  mutate(
    education = c("Low", "Low", "High", "High"),
    religion = c(1, 2, 1, 2)
  ) |>
  ggplot(aes(x = religion, y = allow2, color = education)) +
  geom_line(linewidth = 1.2) +
  theme_bw() +
  labs(x = "Frequency of Religious Attendance", y = "Support for 'Radical' Muslim Speech", color = "Education") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_continuous(limits = c(.9, 2.1), breaks = c(1,2), label = c("Low\nAttendance", "High\nAttendance")) +
  theme(legend.position = "top")
```

::: notes

**Was everybody able to translate the table into the line chart?**

<br>

The visualization is SO much easier to interpret, right?

- The pattern is additive (e.g. lines fairly parallel)

- The relationship between religious attendance and speech opinions remains consistent across both levels of education. 

<br>
            
**SLIDE**: Exercise 5
            
:::



## Pollock & Edwards Exercise 5 {background-image="Images/background-data_blue_v4.png" .center}

<br>

Do countries with proportional representation electoral systems have higher levels of voter turnout (controlling for level of development)?

```{r}
tab5 <- tibble(
  cat1 = c("Not PR", "PR", "Overall"),
  overall = c(63.1, 68.7, NA_integer_),
  low_gdp = c(56.1, 69.8, 61.6),
  high_gdp = c(66, 68.5, 67.4)
)

tab5 |>
  kable(col.names = c("", "Overall", "Low GDP", "High GDP"), align = "c") |>
  #add_header_above(c("", "Zero-Order", "Partial Effect" = 2)) |>
  kable_styling(font_size = 30)
```

::: notes

**Everybody make this table (or something like it)?**

<br>

**Does the data support this hypothesis?**

- (Yes!)

<br>

This version I've made displays both the zero-order effect and the partial effect.

- **What on this table represents the zero-order effect and what the partial effect?**

- (**SLIDE**)

:::



## Pollock & Edwards Exercise 5 {background-image="Images/background-data_blue_v4.png" .center}

<br>

Do countries with proportional representation electoral systems have higher levels of voter turnout (controlling for level of development)?

```{r}
tab5 <- tibble(
  cat1 = c("Not PR", "PR", "Overall"),
  overall = c(63.1, 68.7, NA_integer_),
  low_gdp = c(56.1, 69.8, 61.6),
  high_gdp = c(66, 68.5, 67.4)
)

tab5 |>
  kable(col.names = c("", "Overall", "Low GDP", "High GDP"), align = "c") |>
  add_header_above(c("", "Zero-Order", "Partial Effect" = 2)) |>
  kable_styling(font_size = 30)
```

::: notes

The zero-order effect is the results of a simple bivariate analysis

- e.g. a description of the relationship between two variables

<br>

The partial effect is the result of a controlled comparison

- e.g. a description of the relationship between two variables while controlling for a third variable

<br>

**Which of these should you present when testing a hypothesis?**

- (Both!)

- Generally speaking you want to be transparent in your methods

<br>

**So, is this a spuriousness, additive, or interactive relationship?**

- (**SLIDE**: Line plot)

:::



## Pollock & Edwards Exercise 5 {background-image="Images/background-data_blue_v4.png" .center}

<br>

Do countries with proportional representation electoral systems have higher levels of voter turnout (controlling for level of development)?

```{r, fig.align='center', fig.asp=.75, fig.width=7}
tab5 |>
  filter(cat1 != "Overall") |>
  mutate(
    x1 = c(1, 2)
  ) |>
  pivot_longer(cols = low_gdp:high_gdp, names_to = "gdp", values_to = "value") |>
  ggplot(aes(x = x1, y = value, color = gdp)) +
  geom_line(linewidth = 1.2) +
  theme_bw() +
  labs(x = "Electoral System", y = "Voter Turnout (mean %)", color = "") +
  #scale_y_continuous(labels = scales::percent_format()) +
  scale_x_continuous(limits = c(.9, 2.1), breaks = c(1,2), label = c("Not\nProportional\nRepresentation", "Proportional\nRepresentation")) +
  theme(legend.position = "top") +
  scale_color_discrete(labels = c("High GDP", "Low GDP"))
```

::: notes

The relationship between electoral systems (PR vs. non-PR) and turnout appears interactive. 

- PR systems are associated with higher turnout across both low and high GDP countries, BUT

- The magnitude of the effect is WAY stronger in poorer countries than in rich ones!

<br>

This also shows why the claim in Part D is so wrong!

- Instituting PR in poorer countries may have substantially larger effects on voter turnout!

<br>

**Overall, how did we do with these exercises?**

<br>

**SLIDE**: Alright let's learn the code we need to perform controlled comparisons in R

:::



## Making Controlled Comparisons {background-image="Images/background-data_blue_v4.png"}

<br>

Add to your notes:

```{r, echo=TRUE}
# Step 1: Split your sample by levels of the control variable
# New tool: subset()

# Step 2: Generate bivariate statistics for each subset
# Cross-tabs or descriptive statistics by group

# Step 3: Present the results in tables and line plots
# new geom in ggplot: geom_line()

# Access the built-in data
library(tidyverse)
```

::: notes

Everybody copy these notes and let's introduce the approach with our built-in data

<br>

Let's learn the process!

- **SLIDE**: We'll start with the relationship between two categorical variables

:::



## {background-image="Images/background-data_blue_v4.png"}

**Making Controlled Comparisons: Categorical x Categorical**

::: {.panel-tabset}

### Question

<br>

**Does the class of a car predict its drive train controlling for the year of manufacture?**

- `class` is a nominal variable

- `drv` is a nominal variable

- `year` is an interval variable


### Cross-tab

**Does the class of a car predict its drive train? (zero-order effect)**

```{r, echo=TRUE, eval=TRUE}
proportions(table(mpg$drv, mpg$class), margin = 2)
```

:::

::: notes

I know the question isn't particularly interesting, but, as ever, I want us to learn the tools using built-in datasets

<br>

So, imagine our causal model that we want to test is that car class causes drive train designs

- e.g. bigger car classes force automakers to use certain kinds of drive train to support them

<br>

My proposed confounder is time

- I feel confident in saying that time is a predictor of what kinds of car classes manufacturer's are building, AND

- That time is a predictor of the kinds of drive trains available for those cars

<br>

Now, we always begin our analyses with the zero-order effects

- **Remind me, what is a zero-order effect?**

- (Bivariate analysis: Simple description of the relationship between two variables)

<br>

Let's first check the zero-order effect

- Everybody make a cross-tab of `class` and `drive` 

- Use proportions in the cross-tab and make sure the outome variable (`drv`) is on the rows

<br>

*Open next tab*

**And what is the answer to the first part of this question?**

- **Does class predict drive train?**

- (Quite strongly!)

<br>

Now, to this point you've been making tables by hand

- The key for me is that you are getting good with table() and proportions() as those are base R functions that should always work

- HOWEVER, there are MANY different ways to automate this process in R

- I'm going to show you one of them now, BUT I make ZERO promises that this approach will work into the future

- **SLIDE**

:::



## {background-image="Images/background-data_blue_v4.png"}

```{r, echo=TRUE, eval=FALSE}
# Install the package ONE TIME
install.packages("crosstable")

# Load the package when you want to use it
library(crosstable)
```

<br>

crosstable(data, cols, by)

- `data` is the dataset

- `cols` is the row variable

- `by` is the column variable

as_flextable()

- Polishes a crosstable object for printing

::: notes

Everybody install the crosstable package

<br>

This gives us two useful functions for our work making controlled comparisons

- *Talk through the details*

<br>

**SLIDE**: Let's put this to work!

:::



## {background-image="Images/background-data_blue_v4.png"}

```{r, echo=TRUE, eval=FALSE}
# Univariate table
crosstable(mpg, cols = drv) |> 
  as_flextable()
```

<br>

```{r}
crosstable(mpg, cols = drv) |> 
  as_flextable(fontsizes = list(body=25, header=25))
```

::: notes

Crosstable applied to one categorical variable gives you a univariate count AND the proportions

<br>

One new piece of R code here is called the pipe

- e.g. the '|>'

- The pipe is an AWESOME piece of code that says take whatever is produced on one line and send it to the next function

- Here we see crosstable creating a new object and then sending it to as_flextable to print it

- **Make sense?**

<br>

**SLIDE**: The bivariate case - a cross-tab

:::



## {background-image="Images/background-data_blue_v4.png"}

```{r, echo=TRUE, eval=FALSE}
# Bivariate table (row %)
crosstable(mpg, cols = drv, by = class) |> 
  as_flextable()
```

<br>

```{r}
# Bivariate table (row %)
crosstable(mpg, cols = drv, by = class) |> 
  as_flextable(fontsizes = list(body=15, header=15))
```

::: notes

By default crosstable calculates %s by row, NOT column

<br>

**SLIDE**: The code to change this to column is annoying

:::



## {background-image="Images/background-data_blue_v4.png"}

```{r, echo=TRUE, eval=FALSE}
# Bivariate table (column %)
crosstable(mpg, cols = drv, by = class, 
           percent_pattern = "{p_col} ({n})") |> 
  as_flextable()
```

<br>

```{r}
# Bivariate table (column %)
crosstable(mpg, cols = drv, by = class, percent_pattern = "{p_col} ({n})") |> 
  as_flextable(fontsizes = list(body=15, header=15))
```

::: notes


:::



## {background-image="Images/background-data_blue_v4.png"}

```{r, echo=TRUE, eval=FALSE}
# Bivariate table (column %), change % decimals
crosstable(mpg, cols = drv, by = class, 
           percent_pattern = "{p_col} ({n})",
           percent_digits = 0) |> 
  as_flextable()
```

<br>

```{r}
# Bivariate table (column %)
crosstable(mpg, cols = drv, by = class, 
           percent_pattern = "{p_col} ({n})",
           percent_digits = 0) |> 
  as_flextable(fontsizes = list(body=15, header=15))
```

::: notes

I appreciate how easy this code is and that it produces a table with BOTH the counts and %s per cell

- **Questions on this code?**

<br>

**SLIDE**: Back to our research question about car class and drive trains

:::



## {background-image="Images/background-data_blue_v4.png"}

**Making Controlled Comparisons: Categorical x Categorical**

::: {.panel-tabset}

### Question

<br>

**Does the class of a car predict its drive train controlling for the year of manufacture?**

- `class` is a nominal variable

- `drv` is a nominal variable

- `year` is an interval variable


### Step 1

<br>

```{r, echo=TRUE}
# Split the sample into subsets by year

# Subset 1: 1999 cars
mpg1999 <- subset(mpg, year == "1999")

# Subset 2: 2008 cars
mpg2008 <- subset(mpg, year == "2008")
```

### Step 2

```{r, echo=TRUE}
proportions(table(mpg1999$drv, mpg1999$class), margin = 2)
proportions(table(mpg2008$drv, mpg2008$class), margin = 2)
```

### Polished Table

```{r, echo=TRUE, eval=TRUE}
# Partial effects cross-tab
crosstable(mpg, cols = drv, by = c(year, class),
           percent_pattern = "{p_col} ({n})",
           percent_digits = 0) |> 
  as_flextable()

crosstable(mpg, cols = drv, by = c(class, year),
           percent_pattern = "{p_col} ({n})",
           percent_digits = 0) |> 
  as_flextable()


# How do car manufacturers change their fleets over time? (e.g. market demands, new tech, government regulations, etc)
crosstable(mpg, cols = class, by = year,
           percent_pattern = "{p_col} ({n})",
           percent_digits = 0) |> 
  as_flextable()

crosstable(mpg, cols = class, by = c(year, drv),
           percent_pattern = "{p_col} ({n})",
           percent_digits = 0) |> 
  as_flextable()


```

:::

::: notes

BUT now it is time for a controlled comparison

- In other words, has the car market changed from 1999 to 2008?

<br>

Step 1 is to create the subsets

- We will give the subset function TWO arguments

- FIRST, you tell it the dataset you want to split into pieces

- SECOND, you tell it how to decide which rows to keep and which to remove

- *Step through arguments on slide*

<br>

In short, our controlled comparison will involve running our analyses on two different groups

- In this case we want to describe the relationship between drv and class one time in 1999 and one time in 2008

<br>

Step 2 we use the exact same tools as we used in bivariate analyses

- Since these variables are both categorical, we make cross-tabs!

- Now we just do it twice, once for each year

- The top table shows the relationship in 1999 and the bottom table in 2008

- These are the partial effects

<br>

The crosstable package can make cross-tabs with a simple control variable pretty darn easy

- **Any differences between the tables stand out for you?**

- (Only differences in compact and subcompact classes)
    
- (More compact cars are getting 4 wheel drive and subcompacts are losing 4 wheel drive)

<br>

**Questions on making controlled comparisons between two categorical variables?**

<br>

**SLIDE**: Let's practice!

:::



## {background-image="Images/background-data_blue_v4.png" .center}

**Does party ID predict policy preferences on the rules for purchasing firearms?**

<br>

```{r, fig.align = 'center', fig.width = 9, fig.height=1.25}
## Manual DAG
d1 <- tibble(
  x = c(-3, 3),
  y = c(1, 1),
  labels = c("Party ID", "Strict Gun\nPurchasing Rules")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 0, color = "white") +
  theme_void() +
  coord_cartesian(xlim = c(-4, 4)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = -1.9, xend = 1.6, y = 1, yend = 1, arrow = arrow())
```

::: notes

There are a lot of claims out there regarding the viewpoints of Americans on the rules around buying guns

- I'd like to know if the divide across the parties is as sharp as the conventional wisdom indicates

<br>

**SLIDE**: BUT, I think there is a clear confounder in this relationship

:::




## {background-image="Images/background-data_blue_v4.png" .center}

**Does party ID predict policy preferences on the rules for purchasing firearms?**

```{r, fig.retina=3, fig.align='center', fig.asp=.6, fig.width=8}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Party ID\n(partyid3)", "Gender\n(gender)", "Strict Gun\nPurchasing Rules\n(gun.buying)")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 0, color = "white") +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.5, xend = 2.25, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.15, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.25, arrow = arrow())
```

::: notes

Alright research team, here is your task

- Use the ANES (2020) data to answer my question

- You should perform BOTH a zero-order test AND a partial effects test

- Since Party ID is categorical just make me cross-tabs

- **Questions?**

- Go!

<br>

**SLIDE**: Results

:::



## {background-image="Images/background-data_blue_v4.png"}

**Does party ID predict policy preferences on the rules for purchasing firearms?**

::: {.panel-tabset}

### Zero-Order

<br>

```{r, fig.align='center', fig.asp=.618}
## Example 1: Does party ID predict policy preferences on the rules for purchasing firearms?
anes2 <- na.omit(select(anes, gun.buying, partyid3, gender))

crosstable(anes2, cols = gun.buying, by = partyid3, percent_digits = 0, percent_pattern = "{p_col} ({n})") |>
  as_flextable(fontsizes = list(body = 14, header = 14))
```

### Partial Table

<br>

```{r, fig.align='center', fig.asp=.618}
crosstable(anes2, cols = gun.buying, by = c(partyid3, gender),
           percent_digits = 0, percent_pattern = "{p_col} ({n})") |>
  as_flextable(fontsizes = list(body = 14, header = 14))
```

:::

::: notes

**Is there a difference in preferences regarding gun purchasing rules across the parties?**

<br>

**Is this a spurious, an additive, or an interactive effect when controlling for gender?**

- (Appears additive to me)

- Slope similar direction and magnitude in each control group

<br>

**Questions on making a controlled comparison with two categorical variables?**

<br>

**SLIDE**: Let's now switch to an example with an interval outcome variable

:::


## {background-image="Images/background-data_blue_v4.png"}

**Making Controlled Comparisons: Interval x Categorical**

::: {.panel-tabset}

### Question

<br>

**Does the cut of a diamond predict its value when controlling for its size?**

- `price` is an interval variable

- `cut` is an ordinal variable

- `carat` is an interval variable


### Group Statistics

**Does the cut of a diamond predict its value?**

```{r, echo=TRUE, eval=TRUE}
aggregate(data = diamonds, price ~ cut, mean)

aggregate(data = diamonds, price ~ cut, median)
```

:::

::: notes

We've investigated the zero-order version of this question a few times now

<br>

**What do we keep seeing? Does cut matter for price?**

- (*Tab*: Nope! Even looks backwards!)

<br>

**SLIDE**: Partial effects time

:::



## {background-image="Images/background-data_blue_v4.png"}

**Making Controlled Comparisons: Interval x Categorical**

::: {.panel-tabset}

### Subset Method

```{r, echo=TRUE, eval=FALSE}
# Subset 1: Smallest diamonds (bottom 50%)
diamonds_smaller <- subset(diamonds, carat <= .7)

# Subset 2: Largest diamonds (top 50%)
diamonds_larger <- subset(diamonds, carat > .7)

# Group descriptive statistics on each subset
aggregate(data = diamonds_smaller, price ~ cut, mean)

aggregate(data = diamonds_larger, price ~ cut, mean)
```

```{r}
diamonds$size <- if_else(diamonds$carat <= .7, "Smallest 50%", "Largest 50%")

aggregate(data = diamonds, price ~ cut + size, mean) |>
  pivot_wider(names_from = size, values_from = price) |>
  select("Diamond Cut" = cut, `Smallest 50%`, `Largest 50%`) |>
  kable() |>
  kable_styling(font_size = 20)
```

### Catgeorical Method

```{r, echo=TRUE}
# Create a dummy variable to split the sample
diamonds$size <- if_else(diamonds$carat <= .7, true = "Smallest 50%", 
                         false = "Largest 50%")

# Give the control variable to aggregate
aggregate(data = diamonds, price ~ cut + size, mean)
```
:::

::: notes

Two methods for the case with an interval outcome and a categorical predictor

<br>

The subset method looks a lot like what we just did, right?

- The disadvantage is that you will have to combine these results manually into a table

- ADVANTAGE 1 is that this method works no matter how many levels of the control variable there are!

- So, if we wanted to split diamond weight by quartiles we would make four subsets and run aggregate four times

- ADVANTAGE 2 this method works if the control variable is nominal, ordinal and interval!

- **Make sense?**

<br>

IFF the control variable is categorical, then there's any easier approach!

- *Categorical method tab*

- Aggregate will let you add a control variable and then it will make a single, long-form table

- **Everybody got this code working?**

<br>

**SLIDE**: The real advantage of this is that a single table means we can more easily visualize the results!

:::



## {background-image="Images/background-data_blue_v4.png"}

**Making Controlled Comparisons: Interval x Categorical**

```{r, echo=TRUE, fig.align='center', fig.asp=.85, out.width="35%", fig.width=5}
# Save the aggregate result as a new object
effect1 <- aggregate(data = diamonds, price ~ cut + size, mean)

# Convert the cut variable levels into numbers
effect1$x1 <- as.numeric(effect1$cut)

# Use the new object for a line plot
ggplot(effect1, aes(x = x1, y = price, color = size)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(labels = levels(effect1$cut))
```

::: notes

*Step through the code*

- Key here is that IFF the categorical predictor is ordinal we can visualize the results as a line plot

<br>

**Questions on visualizing the aggregate results?**

<br>

**SLIDE**: Our new tools

:::



## {background-image="Images/background-data_blue_v4.png"}

**Making Controlled Comparisons**

**If categorical x categorical**

- Step 1: Split your sample by levels of the control variable
- Step 2: Generate cross-tabs for each subset
- Step 3: Polish tables with crosstable()

**If interval x categorical**

- Step 1: Split your sample by levels of the control variable
- Step 2: Generate statistics by group for each subset 
- Step 3: Visualize partial effects with line plots

::: notes

**Does the subset approach to making controlled comparisons make sense?**

<br>

**SLIDE**: Let's practice!

:::



## For Next Class {background-image="Images/background-data_blue_v4.png" .center}

Do higher levels of frustration with the current state of the economy (econ.current) predict presidential approval (ft.trump.pre) when controlling for party ID (partyid3)?

::: {.incremental}

1. Table 1: Group means of presidential approval by each respondents' view of the economy (zero-order)

2. Table 2: Redo Table 1 BUT control for party ID (partial effects)

3. Figure 1: Make a line plot of the partial effects calculated in Step 2 above

:::

::: notes



Canvas Assignment: Practice Making Controlled Comparisons (interval x categorical)

Do higher levels of frustration with the current state of the economy (econ.current) predict presidential approval (ft.trump.pre) when controlling for party ID (partyid3)?

1. Table 1: Submit a table of group means of presidential approval by each respondents' view of the economy (zero-order effects)

2. Table 2: Submit a table of group means of presidential approval by each respondents' view of the economy while controlling for party ID (partial effects)

3. Figure 1: Visualize the results in Step 2 above with a line plot. Be sure to categorize it as a spurious, additive or interactive effect. You should have three lines on this plot (one per party identification).



:::




## {background-image="Images/background-data_blue_v4.png"}

```{r, eval=FALSE}
names(anes)

# Do higher levels of frustration with the current state of the economy (econ.current) predict presidential approval (ft.trump.pre)?






aggregate(anes, ft.trump.pre ~ econ.current, mean)
aggregate(anes, ft.trump.pre ~ econ.current + partyid3, mean)
aggregate(anes, ft.trump.pre ~ econ.current + gender, mean)

aggregate(anes, ft.trump.post ~ econ.current, mean)
aggregate(anes, ft.trump.post ~ econ.current + partyid3, mean)
aggregate(anes, ft.trump.post ~ econ.current + gender, mean)

aggregate(anes, ft.trump.pre ~ partyid3, mean)
aggregate(anes, ft.trump.pre ~ partyid3 + social.class, mean)




# Subsets
anes_male <- subset(anes, gender == "1. Male")
anes_female <- subset(anes, gender == "2. Female")

anes_dem <- subset(anes, partyid3 == "1. Democrat")
anes_ind <- subset(anes, partyid3 == "2. Independent")
anes_rep <- subset(anes, partyid3 == "3. Republican")

summary(anes$age)
anes_age1 <- subset(anes, age < 37)
anes_age2 <- subset(anes, age > 37 & age < 66)
anes_age3 <- subset(anes, age >= 66)

# class -> economy approval
proportions(table(anes$econ.current, anes$social.class), margin = 2)
proportions(table(anes$econ.current, anes$social.class, anes$gender), margin = c(2,3))

proportions(table(anes$econ.current, anes$partyid3), margin = 2)
proportions(table(anes$econ.current, anes$partyid3, anes$gender), margin = c(2,3))





# OOH this is cool!
# Social class x party x age
proportions(table(anes$partyid3, anes$social.class), margin = 2)
proportions(table(anes_age1$partyid3, anes_age1$social.class), margin = 2)
proportions(table(anes_age2$partyid3, anes_age2$social.class), margin = 2)
proportions(table(anes_age3$partyid3, anes_age3$social.class), margin = 2)

# Social class x donation to candidate x party
proportions(table(anes$polact.givecand, anes$social.class), margin = 2)
proportions(table(anes_dem$polact.givecand, anes_dem$social.class), margin = 2)
proportions(table(anes_ind$polact.givecand, anes_ind$social.class), margin = 2)
proportions(table(anes_rep$polact.givecand, anes_rep$social.class), margin = 2)

# Abortion x party x gender
proportions(table(anes$abortion.legal, anes$partyid3), margin = 2)
proportions(table(anes_male$abortion.legal, anes_male$partyid3), margin = 2)
proportions(table(anes_female$abortion.legal, anes_female$partyid3), margin = 2)

# Death penalty x party x gender
proportions(table(anes$death.penalty, anes$partyid3), margin = 2)
proportions(table(anes_male$death.penalty, anes_male$partyid3), margin = 2)
proportions(table(anes_female$death.penalty, anes_female$partyid3), margin = 2)




proportions(table(anes$death.penalty, anes$partyid3), margin = 2)
proportions(table(anes$death.penalty, anes$partyid3, anes$gender), margin = c(2,3))

proportions(table(anes$gun.buying, anes$partyid3), margin = 2)
proportions(table(anes$gun.buying, anes$partyid3, anes$gender), margin = c(2,3))





```
