---
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
fontsize: 14pt
classoption: "aspectratio=169"
header-includes:
- \usepackage{caption, xcolor}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "./09_1-Models.Rmd", output_file = "./09_1-Models.pdf")

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```


## Today's Agenda

\begin{Large}
1. What is a model?

\vspace{.25in}

2. Brainstorming models for our research project
\end{Large}

\vspace{.5in}

\begin{center}
Justin Leinaweaver (Fall 2021)
\end{center}


## {.plain}
\center
\includegraphics[keepaspectratio=true,width=.95\paperwidth]{Images/08_2-wing_wind1.png}


## {.plain}
\center
\includegraphics[keepaspectratio=true,width=.95\paperwidth]{Images/08_2-wing_wind2.png}



## {.plain}
\center
\includegraphics[keepaspectratio=true,width=.95\paperwidth]{Images/08_2-wing_wind3.png}



## {.plain}
\center
\includegraphics[keepaspectratio=true,width=.95\paperwidth]{Images/08_2-boston_t1.png}



## {.plain}
\center
\includegraphics[keepaspectratio=true,width=.95\paperwidth]{Images/08_2-boston_t2.png}



##
\begin{columns}
\begin{column}{0.4\textwidth}
\center
\includegraphics[keepaspectratio=true,width=.4\paperwidth]{Images/08_2-mbta_map1.png}

```
\end{column}
\begin{column}{0.6\textwidth}
\begin{large}
\hspace{.1in} Maps are:
\begin{itemize}
\item Neither true nor false
\item Limited in their accuracy
\item Partial representations
\item Useful for only some uses
\item A reflection of the interests of the designer
\end{itemize}
\end{large}
\end{column}
\end{columns}


##
\begin{columns}
\begin{column}{0.4\textwidth}
\center
\includegraphics[keepaspectratio=true,width=.4\paperwidth]{Images/08_2-mbta_map1.png}

```
\end{column}
\begin{column}{0.6\textwidth}
\begin{large}
\hspace{.1in}\textbf{Scientific models} are:
\begin{itemize}
\item Neither true nor false
\item Limited in their accuracy
\item Partial representations
\item Useful for only some uses
\item A reflection of the interests of the designer
\end{itemize}
\end{large}
\end{column}
\end{columns}




## Assignment 3
\Large
\begin{center}
\textbf{Answer a research question with data}
\end{center}

\vspace{.2in}

\large

Dataset 1: Freedom House (civil liberties and political rights)

\vspace{.1in}

Dataset 2: CoW (National Material Capabilities)

\vspace{.2in}

\Large
\begin{center}
\textbf{Let's brainstorm research questions!}
\end{center}




## Assignment 3: Brainstorm a Model
\Large

What's the story that connects these predictors to these outcomes? 

+ Do we need a time lag?

```{r, echo = FALSE, fig.asp = .4, fig.width = 5, fig.align = 'center', out.width = '90%'}
tribble(
    ~Name, ~x, ~y,
    "Predictor", 1, 1,
    "Outcome", 3, 1
) %>%
    ggplot(aes(x = x, y = y)) +
    geom_point(size = .1) +
    geom_label(aes(label = Name), size = 6) +
    coord_cartesian(xlim = c(0.8, 3.3), ylim = c(.93, 1.04)) +
    ggthemes::theme_tufte() +
    annotate("segment", x = 1.4, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
    scale_x_continuous(breaks = NULL) +
    scale_y_continuous(breaks = NULL) +
    labs(x = "", y = "")

```



## Assignment 3: The Research Report {.plain}

1. Introduction
   + What is the question, why is it important and what is your argument?

\vspace{.1in}

2. Bivariate Analyses
   + What is the answer to the question using only bivariate visualizations and simple OLS regressions?

\vspace{.1in}

3. Multivariate Analyses
   + What is the answer to the question using multivariate visualizations and multiple OLS regressions?

\vspace{.1in}

4. Conclusion
   + What did you find and how confident should we be in your results?
   
   
  
## Assignment 3: The Research Report {.plain}
\LARGE

1. Introduction

	+ \Large What is the question?

	+ \Large Why is it important?

	+ \Large What is your argument?


## For Tuesday
\large

Read Wheelan ch4 on correlation

Check out the following recipes in Chang (2018)

+ Recipes 3.2, 3.3, 3.7, 3.8 on stacked bar plots

+ Recipes in ch 5 on scatter plots

+ Recipe 15.14 on recoding a Continuous Variable to a Categorical Variable






