---
format: revealjs
---

## Today's Agenda {background-image="Images/background-data_blue_v4.png" .center}

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(modelsummary)
library(modelr)
library(ggeffects)
library(crosstable)

# Load US States Data
d <- read_excel("../../DATA/2026-Spring/Pollock_Edwards_Textbook/Pollock_Edwards-US_States.xlsx", na = "NA")
```

<br>

::: {.r-fit-text}

**Ordinary Least Squares (OLS) Regression**

- Introduction to multiple regression

:::

<br>

::: r-stack
Justin Leinaweaver (Spring 2026)
:::

::: notes
Prep for Class

1. Check Canvas submissions

2. Don't go super slow through this. There's nothing wrong with assuming they can do basic analyses at this point.

<br>

Last class we got in a ton of practice fitting, interpreting and evaluating OLS regressions

- Today we segue into multiple regression

- **SLIDE**: We'll do that by focusing on answering a research question with data

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}

**Do voter ID laws reduce voter turnout?**

<br>

1. Load the Pollock_Edwards-US_States.xlsx

2. Univariate analyses

    - Outcome: `vep20.turnout`
    
    - Predictor: `voter.id.law`

:::

::: notes

Our research question: Do voter ID laws reduce voter turnout?

<br>

As with all projects we begin with univariate analyses

- Load the data and then summarize the variation in the two key measures we will be analyzing

:::



## {background-image="Images/background-data_blue_v4.png"}

::: {.panel-tabset}

### VEP

```{r, echo=TRUE}
summary(d$vep20.turnout)
```

```{r, fig.align='center', fig.asp=.618, out.width="65%", fig.width=6}
d |>
  ggplot(aes(x = vep20.turnout)) +
  geom_histogram(color = "white", bins = 10) +
  labs(x = "Voting Eligible Population (% in 2020)", y = "Count") +
  scale_x_continuous(breaks = seq(50,80,5), labels = scales::percent_format(scale = 1))
```


### Voter ID

```{r, echo=FALSE}
crosstable(data = d, cols = voter.id.law) |>
  as_flextable()
```

```{r, fig.align='center', fig.asp=.618, out.width="55%", fig.width=6}
d |>
  ggplot(aes(x = voter.id.law)) +
  geom_bar() +
  labs(x = "", y = "Count") +
  theme_bw()
```

:::

::: notes

**Takeaways from your univariate analysis of the VEP in 2020?**

- **How much variation is there in voter turnout across the 50 US states?**

<br>

**Takeaways from your univariate analysis of the voter ID regimes in 2020?**

- **How much variation is there in voter ID policies across the 50 US states?**

<br>

Alright, let's explore the research question

- **What statistical tool should I use to examine the relationship between an interval variable and a categorical variable?**

- (difference of means)

- Everybody do it!

<br>

**SLIDE**: Result

:::




## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**RQ: Do voter ID laws reduce voter turnout?**
:::

```{r, echo=TRUE}
# Calculate group means
aggregate(data = d, vep20.turnout ~ voter.id.law, mean)
```

<br>

::: {.fragment}

```{r, echo=TRUE}
# Null hypothesis test using ANOVA
aov(data = d, vep20.turnout ~ voter.id.law) |> 
  summary()
```

:::

::: notes

**What do you see in these group means?**

- **Are voter ID policies associated with lower voter turnout?**

<br>

Next step, we need a test to determine if these groups are different from what the null would predict

- **What function do we need to run a null hypothesis test for a difference of means with more than two levels?**

- (ANOVA)

- Ok, do that!

<br>

**REVEAL: So, what do we learn from these ANOVA results?**

- (We cannot reject the NULL hypothesis of no difference between the groups)

<br>

Remember, the F statistic is a ratio of between-group error divided by within-group error

- This high p-value is almost certainly because so many of the group means are so similar

- HOWEVER, there is definitely something going on when comparing the no id group to all the different flavors of ID group

- **SLIDE**: So, let's attack this differently

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**RQ: Do voter ID laws reduce voter turnout?**
:::

<br>

```{r, echo=TRUE}
# Create a dummy variable
d$no_voter_id <- 
  if_else(d$voter.id.law == "No document required", 1, 0)
```

<br>

```{r, echo=TRUE}
# Check the new variable
table(d$no_voter_id)
```

::: notes

Everybody create this dummy variable so we can split our sample by whether or not a state has any kind of voter ID policy in place

<br>

Once you create this variable I want you to replicate our analyses

- **Remind me, what is the appropriate null hypothesis test for using this new variable?**

- (A t-test! Only two levels in the categorical variable so prefer a t-test for its simplicity and ease of interpretation)

<br>

Ok, go!

- Calculate the group means and then run a null hypothesis test on the results


:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**RQ: Do voter ID laws reduce voter turnout?**
:::

```{r, echo=TRUE}
# Null hypothesis test using a t-test
t.test(data = d, vep20.turnout ~ no_voter_id)
```

::: notes

**What do you see in these group means?**

- **Are voter ID policies associated with lower voter turnout?**

- (Seems clear from this that states without voter ID laws see 3% higher turnout on average)

<br>

**So, what do we learn from these t-test results?**

- **Can we reject the null?**

- (Narrowly insignificant at p-value of .08)

- **SLIDE**: Let's look at it

:::



## {background-image="Images/background-data_blue_v4.png" .center .smaller}

```{r, echo=TRUE}
# Null hypothesis test using a t-test
t.test(data = d, vep20.turnout ~ no_voter_id)
```

```{r, fig.align='center', fig.asp=.618, out.width="50%", fig.width=6}
x1 <- rnorm(n = 1000000, mean = 0, sd = 1)

sim1 <- tibble(
  x = density(x1)$x,
  y = density(x1)$y
)

sim2 <- sim1 |>
  filter(x <= 1.96 & x >= -1.96)

# sim2 <- sim1 |>
#   filter(x <= -1.96 | x >= 1.96)

ggplot(sim1, aes(x = x, y = y)) +
  geom_line() +
  geom_area(data = sim2, aes(x = x, y = y), fill = "lightblue") +
  theme_bw() +
  labs(x = "Null Hypothesis", title = "Not all p-values are created equal") +
  scale_x_continuous(limits = c(-3.5,3.5), breaks = -3:3) +
  geom_vline(xintercept = 1.78, color = "red") # t value from t.test
```

::: notes

Key to remember, statistical significance is not the end-all, be-all test of usefulness

- It is just one piece of how we evaluate our statistical analyses

<br>

Long story short, I'm not willing to ignore this finding due just to this p-value

- Given the small sample size, a p-value of .08 is relatively good

- Let's call it strongly suggestive rather than conclusive support

<br>

Ok, set aside the statistical significance debate for a moment

- **What is missing from our work so far if the aim is to make a causal argument about this linkage?**

- (**SLIDE**: Confounders!)

:::




## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**RQ: Do voter ID laws reduce voter turnout?**
:::

```{r, fig.retina=3, fig.align='center', out.width="90%", fig.width=7}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Voter ID", "Confounders", "Voter\nTurnout")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 0, color = "white") +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 5) +
  annotate("segment", x = 1.5, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.15, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.15, arrow = arrow())
```

::: notes

With observational data we can be confident that our comparison groups are not identical on average

- Specifically, the characteristics of states that have voter ID laws are likely quite different from those that don't have voter ID laws

- Also, it stands to reason that the characteristics of states with very high voter turnout are likely different from those with low turnout

- Our job is to identify the confounders that, when we control for them, will allow us to isolate the effect of voter ID laws on turnout

<br>

**Help me brainstorm, what possible confounders are in this relationship?**

- **Everybody take a look at the codebook for variables in our dataset we can consider!**

- (*ON BOARD*)

- (Factors that predict swing states AND Republican pressure for voter ID
    - foreign.born
    - hispanic.percent
    - gini2016 (inequality)
    - ba.or.more
    - term.limits
    - trump16

<br>

We'll come back to this list at the end, but for now let's focus on one story

- **SLIDE**: Battleground states

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**RQ: Do voter ID laws reduce voter turnout?**
:::

```{r, fig.align='center'}
library(usmap)

test1 <- tibble(
  state = c("AZ", "FL", "GA", "IA", "MI", "MN", "NV", "NH", "NC", "OH", "PA", "TX", "WI"),
  values = rep("Battleground", 13)
)

plot_usmap(data = test1) +
  labs(title = "Battleground States (Cook Political Report 2020)") +
  scale_fill_manual(values = "#AA4499", na.value = "#DDCC77") + 
  theme(legend.position = "none")

# plot_usmap(regions = "states", include = c("AZ", "FL", "GA", "IA", "MI", "MN", "NV", "NH", "NC", "OH", "PA", "TX", "WI"), fill = "purple2") +
#   labs(title = "Battleground States (Cook Political Report 2020)")
```

::: notes

By "battleground state" I am adopting the Cook Political Report's definition

- Includes states that have shifted between Democratic and Republican victories in recent presidential election cycles, AND

- States rated as highly competitive, often too close to call until the final days of an election.

<br>

In the dataset this includes 13 states:

- AZ, FL, GA, IA, MI, MN, NV, NH, NC, OH, PA, TX, WI

<br>

My argument here is that:

1. Being a battleground state increases the pressure on Republicans in the legislature to enact voter ID

2. Having a close election in your state increases the incentive for individual voters to vote (better chance your vote will matter!)

<br>

Let's test it

- Everybody calculate the group means but now include both our voter ID dummy AND the battleground state dummy

- Go!

:::



## {background-image="Images/background-data_blue_v4.png" .center}

```{r, echo=TRUE}
# Difference of means
aggregate(data = d, 
          vep20.turnout ~ no_voter_id + battleground2020, mean)
```

<br>

::: {.fragment}

```{r, echo=TRUE}
# Two-factor ANOVA
aov(data = d, vep20.turnout ~ no_voter_id + battleground2020) |> 
  summary()
```

:::

::: notes

**Remind me, are these zero-order or partial effects?**

- (Partial!)

<br>

**What do you see in these group means?**

- **Are voter ID policies associated with lower voter turnout?**

- (Seems clear that states without voter ID laws see higher turnout in BOTH battleground and non-battleground states on average)

<br>

BUT, is this a significant difference?

- **What function do we need to run a null hypothesis test of these group means?**

- (two-factor ANOVA)

<br>

**REVEAL: So, what do we learn from these ANOVA results?**

- (Battleground state effect is significant so we can reject the null that says no difference in turnout between battleground and non-battleground states)

- (Voter ID effect has come down to .07, but is still not quite at the .05 threshold)

- **BUT** this doesn't tell us the size or direction of either effect!

<br>

**SLIDE**: Summing up in order to move forward

:::



## Null Hypothesis Testing {background-image="Images/background-data_blue_v4.png" .center}

<br>

**Two Groups**

- Difference of means and a t-test

- Difference of proportions and a proportions test

**More Groups**

- Difference of means and an ANOVA (aov)

- Cross-tab and a $\chi^2$ test

::: notes

In sum, these methods have allowed us to accomplish a great deal in terms of inferential statistics

- e.g. describing a sample in order to learn about a population

<br>

Moving forward, we need to embrace regression in order to push our tools further

- OLS will let us add more confounders to each test

- OLS will let us include interval variables as controls much more easily

- OLS will let us address nonlinear relationships in our data

- And more!

<br>

So, let's take a few steps back and shift to OLS in order to move ourselves forward in answering our research question!

- **SLIDE**: We'll start by focusing on the relationship WITHOUT a confounder

:::



## {background-image="Images/background-data_blue_v4.png" .center .smaller}

::: {.r-fit-text}
**RQ: Do voter ID laws reduce voter turnout?**
:::

<br>

:::: {.columns}

::: {.column width="45%"}
**Difference of Means and T-Test**

```{r, echo=TRUE}
t.test(data = d, 
       vep20.turnout ~ no_voter_id)
```
:::

::: {.column width="55%"}
**Simple OLS Regression**

```{r, echo=TRUE}
lm(data = d, vep20.turnout ~ no_voter_id) |> 
  summary()
```
:::

::::

::: notes

On the left is the difference of means and t-test, on the right is the OLS

- Compare and contrast these results for me

- **What do you see?**

<br>

Let's use the simple OLS regression to make predictions

- **What is the predicted voter turnout in states with no voter ID policies in place?**

    - (The intercept: 66.85%)

- **What is the predicted voter turnout in states with voter ID policies in place?**

    - (66.85 + 3.05 = 69.9%)

- **Does that look or sound like anything on the screen?**

    - (The group means!)

<br>

In the case of a binary predictor the OLS generates the group means!

<br>

**OK, how else are these two approaches the same?**

- Both offer null hypothesis testing in p-values (and the p-values are VERY similar)

- Both offer null hypothesis testing in CIs

    - OLS needs an extra function, but the SE is right there for us

<br>

**How are these approaches different?**

- BIG difference in terms of the null hypothesis testing!

- Difference of means t-test is asks IF the groups are actually different from the null

    - Doesn't tell you direction or magnitude

- OLS significance test asks if your SPECIFIC, measured difference is a better fit to the data than the null

    - This estimates both direction and magnitude!

- PLUS OLS gives us other useful ways to think about the fit (SEs, R2, Adjusted R2, shape and size of the residuals, etc)

<br>

**SLIDE**: In order to add our confounder in we need to change our formula for a line

:::



## Making Controlled Comparisons {background-image="Images/background-data_blue_v4.png" .center}

<br>

**Simple OLS Regression**

$$ y = \alpha + \beta(x) + \epsilon $$

**Multiple OLS Regression**

$$ y = \alpha + \beta_1(x_1) + \beta_2(x_2) + ... + \beta_k(x_k) + \epsilon $$

::: notes

A simple OLS regression is a bivariate test

- One predictor used to model one outcome

- We then interpret the beta coefficient (e.g. the slope of the line) as the effect on the outcome of increasing X by one

<br>

Multiple OLS regression is a multivariate test

- We can now include as many predictor variables as we want in order to model an outcome

- e.g. all the confounders we need to establish a causal argument!

<br>

**Per the textbook, how is multiple regression like the controlled comparisons we have been doing so far?**

- **In other words, how do we interpret the partial regression coefficients in the multiple regression?**

- (You CANNOT interpret ANY single coefficient without reference to the others!)

    - e.g. these are partial effects, not zero-order effects
    
- To be clear, BOTH of these formulas describe a SINGLE line

- In the multiple regression version the slope of the line is determined by ALL of these partial regression coefficients

- That means the meaning of any one partial regression coefficient depends on holding all other predictor variables constant

- This is the ceteris paribus condition

<br>

**Questions on this intuition from the textbook?**

<br>

Ok, textbook Exercise 5 asks why you shouldn't select variables to maximize the adjusted R2.

- **Why is this a terrible idea?**

- (**SLIDE**)

:::



## Making Causal Arguments {background-image="Images/background-data_blue_v4.png" .center}

```{r, fig.retina=3, fig.align='center', fig.asp=.6, fig.width=8}
## Manual DAG
d1 <- tibble(
  x = c(1, 2, 3),
  y = c(1, 2, 1),
  labels = c("Treatment", "Confounder", "Outcome")
)

ggplot(data = d1, aes(x = x, y = y)) +
  geom_point(size = 0, color = "white") +
  theme_void() +
  coord_cartesian(xlim = c(0, 4), ylim = c(.75, 2.25)) +
  geom_label(aes(label = labels), size = 7) +
  annotate("segment", x = 1.5, xend = 2.5, y = 1, yend = 1, arrow = arrow()) +
  annotate("segment", x = 1.7, xend = 1, y = 1.85, yend = 1.15, arrow = arrow()) +
  annotate("segment", x = 2.3, xend = 3, y = 1.85, yend = 1.15, arrow = arrow())
```

$$ y = \alpha + \beta_1(x_1) + \beta_2(x_2) + ... + \beta_k(x_k) + \epsilon $$

::: notes

Because we're making causal arguments and not garbage can models

- We're not economists!

<br>

Maximizing adjusted R-square leads to overfitting your data

- e.g. describing a line that so perfectly mirrors your sample it stops representing the population we are interested in!

- Remember, the goal of inferential statistics is to learn about the population NOT perfectly describe the sample

<br>

Bottom line, if it isn't your primary mechanism or a confounder then leave it out!

<br>

**SLIDE**: Ok, let's take this back to our RQ for today

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**RQ: Do voter ID laws reduce voter turnout?**
:::

<br>

```{r, echo=TRUE}
# Fit the multiple regression
lm(data = d, vep20.turnout ~ no_voter_id + battleground2020) |>
  summary()
```

::: notes

Everybody make sure the code runs

- **Does everybody see that all we've done is added a second variable to the right side of the formula in the lm function?**

<br>

**SLIDE**: Let's evaluate the fit of this model

:::




## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**RQ: Do voter ID laws reduce voter turnout?**
:::

```{r}
lm(data = d, vep20.turnout ~ no_voter_id + battleground2020) |> 
  modelsummary(out = "gt",
             fmt = 2,
             stars = c("*" = .05),
             gof_omit = "IC|F|Log") |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "x-large")
  ), locations = gt::cells_body())
```

::: notes

Ok, talk me through the fit of this model to the data using our four evaluation steps

<br>

**Any concerns with missing data?**

- (Nope. Includes all 50 states!)

<br>

**Is the model statistically significant?**

- (Battleground state: yes)

- (No voter ID is close: .08)

<br>

**What do we learn from the Adjusted R2?**

- (Only explains 10% of the variation in voter turnout)

- BUT, the magnitude of each effect is quite substantial; 3-4% easily moves most elections!

<br>

Step 4 is to evaluate the residuals but we'll hit those next week

<br>

**SLIDE**: Let's reinforce our intuitions on this approach by connecting it back to what we've done previously

:::




## {background-image="Images/background-data_blue_v4.png" .center .smaller}

::: {.r-fit-text}
**RQ: Do voter ID laws reduce voter turnout?**
:::

<br>

:::: {.columns}

::: {.column width="63%"}
**Difference of Means and ANOVA**

```{r, echo=TRUE}
aggregate(data = d, 
          vep20.turnout ~ no_voter_id + battleground2020, 
          mean)

aov(data = d, 
    vep20.turnout ~ no_voter_id + battleground2020) |> 
  summary()
```
:::

::: {.column width="37%"}
**Multiple OLS Regression**

```{r}
lm(data = d, vep20.turnout ~ no_voter_id + battleground2020) |> 
  modelsummary(out = "gt",
             fmt = 2,
             stars = c("*" = .05),
             gof_omit = "IC|F|Log") |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "medium")
  ), locations = gt::cells_body())
```
:::

::::

::: notes

On the left is our difference of means test with two-factor ANOVA

<br>

On the right is the polished multiple regression using the modelsummary function

<br>

**SLIDE**: To connect these dots I want you to translate this multiple regression into the formula for a line

:::




## {background-image="Images/background-data_blue_v4.png" .center .smaller}

:::: {.columns}
::: {.column width="63%"}
**Difference of Means and ANOVA**

```{r, echo=TRUE}
aggregate(data = d, 
          vep20.turnout ~ no_voter_id + battleground2020, 
          mean)

aov(data = d, 
    vep20.turnout ~ no_voter_id + battleground2020) |> 
  summary()
```
:::

::: {.column width="37%"}
**Multiple OLS Regression**

```{r}
lm(data = d, vep20.turnout ~ no_voter_id + battleground2020) |> 
  modelsummary(out = "gt",
             fmt = 2,
             stars = c("*" = .05),
             gof_omit = "IC|F|Log") |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "medium")
  ), locations = gt::cells_body())
```
:::
::::

<br>

$$ y = \alpha + \beta_1(x_1) + \beta_2(x_2) + ... + \beta_k(x_k) + \epsilon $$

::: notes

Ok, everybody write down the formula for a line but replace the variable letters with the actual results from the table

- Go!

<br>

(**SLIDE**)

:::




## {background-image="Images/background-data_blue_v4.png" .center .smaller}

:::: {.columns}
::: {.column width="60%"}
**Difference of Means and ANOVA**

```{r, echo=TRUE}
aggregate(data = d, 
          vep20.turnout ~ no_voter_id + battleground2020, 
          mean)

aov(data = d, 
    vep20.turnout ~ no_voter_id + battleground2020) |> 
  summary()
```
:::

::: {.column width="40%"}
**Multiple OLS Regression**

```{r}
lm(data = d, vep20.turnout ~ no_voter_id + battleground2020) |> 
  modelsummary(out = "gt",
             fmt = 2,
             stars = c("*" = .05),
             gof_omit = "IC|F|Log") |>
  gt::tab_style(style = list(
                  gt::cell_fill(color = 'white'),
                  gt::cell_text(size = "medium")
  ), locations = gt::cells_body())
```
:::
::::

$$ y = \alpha + \beta_1(x_1) + \beta_2(x_2) + ... + \beta_k(x_k) + \epsilon $$

$$ \text{Voter Turnout} = 65.93 + 2.95(\text{No ID}) + 3.69(\text{Battleground})$$

::: notes

**Everybody get this?**

<br>

Now I want you to use the formula for a line to make the same four predictions as the group means on the left

<br>

**What is the model's prediction for voter turnout if...**

1. Your state requires ID and is not a battleground?

    - (65.93)
    
2. Your state does not require ID and is not a battleground?

    - (65.93 + 2.95 = 68.88)

3. Your state requires ID and is a battleground?

    - (65.93 + 3.69 = 69.62)
    
4. Your state does not require ID and is a battleground?

    - (65.93 + 2.95 + 3.69 = 72.57)

<br>

**Any idea why our estimates from the multiple OLS no longer match the group means?**

- (Because they answer fundamentally different questions)

1. Regression adjusts for all predictors

    - The group mean is the average ignoring all other factors
    - Regression produces partial effects where each value is adjusted for the others!

2. If the group sizes are different, regression tries to treat each group more equally (up-weighting the small groups, down-weighting the large ones)

    - Our voter ID variable is 18 vs 32 so the regression will adjust to treat them as comparable

3. Regression allows us to consider what if different types of voter ID impact turnout differently? 

    - In that case, the means get adjusted
    
<br>

**Questions on what we've done so far?**

- **e.g. the intuitions of multiple OLS from the textbook?**

<br>

**SLIDE**: Practice time

:::



## {background-image="Images/background-data_blue_v4.png" .center}

::: {.r-fit-text}
**RQ: Do voter ID laws reduce voter turnout?**
:::

<br>

Use a multiple OLS regression to estimate the CAUSAL effect of voter ID laws (`vep20.turnout`) on voter turnout (`no_voter_id`)

- Confounders: Battleground state (`battleground2020`), proportion of the state with a college degree (`ba.or.more`) and the proportion of the state identifying as hispanic (`hispanic.percent`)

::: notes

Which of the following THREE models do you ARGUE is the closest to a causal argument about the effect of voter ID on voter turnout?


```{r}
# Test other confounders?
# foreign.born
# hispanic.percent
# gini2016 (inequality)
# ba.or.more
# term.limits
# trump16

lm(data = d, vep20.turnout ~ no_voter_id + battleground2020) |> summary()

lm(data = d, vep20.turnout ~ no_voter_id + battleground2020 + ba.or.more) |> summary()

lm(data = d, vep20.turnout ~ no_voter_id + battleground2020 + hispanic.percent) |> summary()

lm(data = d, vep20.turnout ~ no_voter_id + battleground2020 + ba.or.more + hispanic.percent) |> summary()





```

:::




































